### 2017-09-22 inst/documents/rotations1.Rmd
tweaks and add first Rmd (currently fails)

---
title: "Rotations and Insecticide Resistance"
output:
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex

#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r,echo=FALSE,results='hide',warning=FALSE}
  library(rotations)
  library(ggplot2)
```

### Rotation interval 50 generations, migration 0.01, steps through insecticides

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1"}

dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration_rate_intervention=0.01)

```

### decreasing migration to 0.001

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1"}

dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration_rate_intervention=0.001)

```

### Rotate when resistant, migration 0.001

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1"}

dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration_rate_intervention=0.001)

```

### 2017-09-22 R/array_named.r

@@ -20,7 +20,7 @@
 
 array_named <- function(...)
   {
-   array(0, dim = lengths(list(...)), dimnames = list(...))
+   array(NA, dim = lengths(list(...)), dimnames = list(...))
   }
 
 

### 2017-09-22 R/fitness_single_locus.r

@@ -106,6 +106,7 @@ fitness_single_locus <- function ( n_insecticides = NULL,
 
 #' @param n_insecticides number of insecticides 
 #' @param a_fitloc array of single locus fitnesses to fill
+#' @param same_insecticides whether to just set fitnesses for all insecticides the same
 #' @param plot whether to plot fitness
 #' 
 #' @examples 
@@ -117,6 +118,7 @@ fitness_single_locus <- function ( n_insecticides = NULL,
 
 fitness_single_locus_test <- function ( n_insecticides = NULL,
                                    a_fitloc = NULL,
+                                   same_insecticides = FALSE,
                                    plot = FALSE)
 {
   
@@ -124,42 +126,46 @@ fitness_single_locus_test <- function ( n_insecticides = NULL,
   if ( is.null(n_insecticides)) n_insecticides <- 4
 
   # create array
-  a_fitloc <-  array_named(insecticide=1:n_insecticides, genotype=c('SS','SR', 'RR'), amount=c('no','lo', 'hi'))
+  a_fitloc <-  array_named(insecticide=1:n_insecticides, genotype=c('SS','RS', 'RR'), amount=c('no','lo', 'hi'))
   
   
   # andy looking to set fitnesses for all insecticides to be same
   # but it seemed to mess up simulation
-  # so disabled for now
-  # a_fitloc[, 'SS', 'no']=0.3; a_fitloc[, 'SS', 'lo']=0.3; a_fitloc[, 'SS', 'hi']=0.3;
-  # a_fitloc[, 'SR', 'no']=0.3; a_fitloc[, 'SR', 'lo']=0.7; a_fitloc[, 'SR', 'hi']=0.7;
[…]

### 2017-09-22 R/rot_plots.r

@@ -22,8 +22,8 @@ rot_plot_resistance <- function(df_res2) {
     mutate( i_in_use = ifelse(stringr::str_detect(region,paste0(insecticide,"_active")),1.05,NA))    
   
   gg <- ggplot( df_res2, aes_string(x='generation',y='resistance') ) + 
-    geom_point(shape=1, colour='blue') +
-    #geom_line( colour='blue') +  
+    #geom_point(shape=1, colour='blue') +
+    geom_line( colour='blue') +  
     facet_wrap('region', ncol=2) +
     #theme(axis.text.x = element_blank()) +
     #add insecticide use indication

### 2017-09-22 R/run_rot.r

@@ -12,11 +12,14 @@
 #' @param plot whether to plot results
 #' @param start_insecticide which insecticide to start with
 #' @param diagnostics whether to output running info
+#' @param same_insecticides whether to just set fitnesses for all insecticides the same
+#' 
 #' 
 #' @examples 
 #' run_rot(rotation_interval=100)
 #' dfr <- run_rot(rotation_interval=50, max_generations = 300)
-#' dfr <- run_rot()
+#' dfr <- run_rot(rotation_interval=0, max_generations = 300)
+#' dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration_rate_intervention=0.01)
 #' 
 #' @import tidyverse 
 #' @return dataframe of results
@@ -33,7 +36,13 @@ run_rot <- function( max_generations = 500, #the maximum number of mosquito gene
                           coverage = 0.8, # "coverage" of the intervention is defined as the proportion of mosquitoes that are covered by the intervention (and 1-C is the proportion of the population in the untreated refugia).
                           plot = TRUE,
                           start_insecticide = 1,
-                          diagnostics = FALSE
+                          diagnostics = FALSE,
+                          hardcode_fitness = TRUE,
+                          same_insecticides = TRUE,
+                          hardcode_exposure = TRUE,
+                          expo_hi = 0.8,
+                          expo_lo = 0,                              
+                          male_expo_prop = 1
                           )
   {
   
@@ -68,13 +77,18 @@ run_rot <- function( max_generations = 500, #the maximum number of mosquito gene
   
   ### set exposures from hardcoded test function or based on other inputs
   exposure <- set_exposure_rot_test( n_insecticides=n_insecticides )
-  #a_expo <- set_exposure_rot()
+  #exposure <- set_exposure_rot()
   
   ### set fitnesses from hardcoded test function or based on other inputs
-  fitness <- fitness_single_locus_test( n_insecticides=n_insecticides )
-  
-  #4 same insecticides
-  #a_fitloc <- fitness_single_locus(n_insecticides=4, eff=0.5, dom=0.5, rr=0.5, cost=0, fitSS=1)
[…]

### 2017-09-22 R/set_exposure_rot.r

@@ -92,8 +92,11 @@ set_exposure_rot_test <- function( n_insecticides = NULL,
   a_expo <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), exposure=c('no','lo','hi'))
   
   #exposure patterns for insecticide 1
-  a_expo[1, 'm', 'lo'] =0.1; a_expo[1, 'm', 'hi'] =0.5;
-  a_expo[1, 'f', 'lo'] =0.1; a_expo[1, 'f', 'hi'] =0.6;
+  #ians initial values
+  #a_expo[1, 'm', 'lo'] =0.1; a_expo[1, 'm', 'hi'] =0.5;
+  #a_expo[1, 'f', 'lo'] =0.1; a_expo[1, 'f', 'hi'] =0.6;
+  a_expo[1, 'm', 'lo'] =0.1; a_expo[1, 'm', 'hi'] =0.1;
+  a_expo[1, 'f', 'lo'] =0.1; a_expo[1, 'f', 'hi'] =0.1;  
   #exposure patterns for insecticide 2
   if(n_insecticides>=2){ #need to avoid exceeding size of the array
     a_expo[2, 'm', 'lo'] =0.1; a_expo[2, 'm', 'hi'] =0.1;

### 2017-10-05 inst/documents/rotations_notes.Rmd
flexibility of n_insecticides 1-5 working

#rotations_notes.Rmd

Just rough notes.

*4/10/2017*

How to make the dataframe that saves results in run_rot flexible to the number of insecticides ?

df_results <- data.frame(generation=1:max_generations,
                         insecticide=NA,
                         r1_active=NA,
                         r1_refuge=NA,
                         r2_active=NA,
                         r2_refuge=NA,
                         r3_active=NA,
                         r3_refuge=NA,
                         r4_active=NA,
                         r4_refuge=NA, stringsAsFactors = FALSE)

Can I adapt this from array_named() in resistance ?
#' n_insecticides <- 3
#' # list(c()) is important below to create one list for each dim
#' l_dimnames <- rep(list(c('no','lo','hi')), n_insecticides)
#' names(l_dimnames) <- paste0("niche", 1:n_insecticides)
#' do.call(array_named,l_dimnames)

# this feels hacky, but it does work & I can put in a function

# create the flexible r* columns
n_insecticides <- 3
l_ <- rep(list(NA), n_insecticides*2) #2 because active & refuge 
names(l_) <- c( paste0("r", 1:n_insecticides, "_active"),
                paste0("r", 1:n_insecticides, "_refuge") )
df_r <- do.call(data.frame,l_)

# create inflexible columns ready to add r* columns on
max_generations <- 10
df_results <- data.frame(generation=1:max_generations,
                         insecticide=NA)

df_results <- data.frame(df_results, df_r)

Note this can only cope with one insceticide being active at a time ? (I could put a comma delimited string in later if I did want to record multiple insecticides).

Does this fit with tidy data ?

Each variable you measure should be in one column.
Each different observation of that variable should be in a different row.

Yes I think it does. Although maybe not ?
Currently I record df_results and gather to df_r2 to allow plotting.
To get :
 generation insecticide    region  resistance
1          1          NA   r1_refuge 0.001000000
2          2           1   r1_refuge 0.001061247

So perhaps I can just save as that to start ? It has multiple rows per generation.
More difficult to see, but easier to plot ?

This allows me to facet by region.
BUT if I wanted to put the lines for active & region into the same subplots (which would make easier to view) I would probably want columns for active and refuge or would I ??

gen active_insecticide resistance_gene active_or_refuge resistance

I can use tidyr::separate() to get from r1_refuge to r1 & refuge in different columns.

separate(region, into=c("resistance_gene","active_or_refuge"))

### 2017-10-05 R/run_rot.r

@@ -50,19 +50,29 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
     
   exposure <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), amount=c('no','lo', 'hi'))
   
-  #andy try to store results in data frame might make easier
-  df_results <- data.frame(generation=1:max_generations,
-                      insecticide=NA,
-                      r1_active=NA,
-                      r1_refuge=NA,
-                      r2_active=NA,
-                      r2_refuge=NA,
-                      r3_active=NA,
-                      r3_refuge=NA,
-                      r4_active=NA,
-                      r4_refuge=NA,
-                      r5_active=NA,
-                      r5_refuge=NA, stringsAsFactors = FALSE)
+  # df_results <- data.frame(generation=1:max_generations,
+  #                     insecticide=NA,
+  #                     r1_active=NA,
+  #                     r1_refuge=NA,
+  #                     r2_active=NA,
+  #                     r2_refuge=NA,
+  #                     r3_active=NA,
+  #                     r3_refuge=NA,
+  #                     r4_active=NA,
+  #                     r4_refuge=NA, stringsAsFactors = FALSE)
+  
+  # setup dataframe to store results, tricky to cope with variable number insecticides
+  l_gene_plus_activity <- rep(list(rep(NA,max_generations)), n_insecticides*2) #2 because active & refuge 
+  
+  names(l_gene_plus_activity) <- c( paste0("r", 1:n_insecticides, "_active"),
+                                    paste0("r", 1:n_insecticides, "_refuge") )
+  
+  df_results <- do.call(data.frame, list(generation = 1:max_generations,
+                                         insecticide = NA,
+                                         stringsAsFactors = FALSE,
+                                         l_gene_plus_activity))  
+  
+  
   #experimental
   # df_res_active <- data.frame(generation=rep(1:max_generations,n_insecticides),
   #                       insecticide=NA,
@@ -329,15 +339,15 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
   break #breaks out of looping generations and terminates the simulation
   }    
     
-   } #end of cycle running the gens up to max_generations
+   } #end of max_generations loop
     
-  #****************************************************************  
-   # NOW collate the data and draw plots
+  #####################################  
[…]

### 2017-10-05 R/rot_plots.r
active and refuge plotted on same sub-plot

@@ -14,21 +14,28 @@
 #' 
 rot_plot_resistance <- function(df_res2) {
   
+  # column names of input dataframe
+  # "generation"  "insecticide"     "resist_gene"  "active_or_refuge" "resistance"
+  
   # to allow plotting of insecticide in use
   # add column which has a value if insecticide in use & NA if not
   # the value (currently 1.05) determines where the line appears on the y axis
   df_res2 <- df_res2 %>%
     #mutate( i_in_use = ifelse(insecticide==1,1,NA))
-    mutate( i_in_use = ifelse(stringr::str_detect(region,paste0(insecticide,"_active")),1.05,NA))    
+    #mutate( i_in_use = ifelse(stringr::str_detect(region,paste0(insecticide,"_active")),1.05,NA))    
+    #now that active & refuge on same plot
+    mutate( i_in_use = ifelse(resist_gene==paste0('r',insecticide),1.05,NA))    
   
-  gg <- ggplot( df_res2, aes_string(x='generation',y='resistance') ) + 
+    
+  gg <- ggplot( df_res2, aes_string(x='generation',y='resistance',colour='active_or_refuge') ) + 
     #geom_point(shape=1, colour='blue') +
-    geom_line( colour='blue') +  
-    facet_wrap('region', ncol=2) +
+    #geom_line( colour='blue') +  
+    geom_line( alpha=0.5 ) + 
+    facet_wrap('resist_gene', ncol=1) +
     #theme(axis.text.x = element_blank()) +
     #add insecticide use indication
-    geom_line( aes_string(x='generation',y='i_in_use'), colour='red', lwd=2) +
-    annotate("text", x = 0, y = 1.15, label = "deployment", size = 2.5, hjust='left') +
+    geom_line( aes_string(x='generation',y='i_in_use'), colour='green4', lwd=2) +
+    #annotate("text", x = 0, y = 1.15, label = "deployment", size = 2.5, hjust='left') +
     #scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1)) +
     scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1)) +
     #theme_bw()

### 2017-10-05 R/run_rot.r

@@ -372,12 +372,12 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
     gather(names(l_gene_plus_activity),
            key=region, value=resistance)
   
-  #use tidyr::separate() to get from r1_refuge to r1 & refuge in different columns.
-  #df_res2 <- separate(df_res2, region, into=c("resistance_gene","active_or_refuge"))
+  # use tidyr::separate() to get from r1_refuge to r1 & refuge in different columns.
+  # to get active & refuge into the same subplot
+  df_res2 <- separate(df_res2, region, into=c("resist_gene","active_or_refuge"))
   
   # do the plots
-  #rot_plot_use(df_res2)
-  rot_plot_resistance(df_res2)
+  if (plot) rot_plot_resistance(df_res2)
   
   invisible(df_res2)
   

### 2017-10-05 inst/documents/rotations1.Rmd
build doc

@@
 ---
-title: "Rotations and Insecticide Resistance"
+title: "Rotations and Insecticide Resistance initial plots"
 output:
   pdf_document:
-    keep_tex: true
+    keep_tex: false
     fig_caption: true
@@
 
-```{r,echo=FALSE,results='hide',warning=FALSE}
+```{r, eval=TRUE, include=FALSE}
+
+library(knitr)
+opts_chunk$set(fig.height=7, fig.width=7)
+# opts_chunk$set(dev="tiff",
+#                dev.args=list(compression="lzw"),
+#                dpi=300,
+#                cache=FALSE,
+#                fig.path='pap2figs/')
+```
+
+```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
   library(rotations)
@@
 
-### Rotation interval 50 generations, migration 0.01, steps through insecticides
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotation interval 50, migration 0.01"}
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1"}
+# Interval 50, migration 0.01
 
@@
 
-### decreasing migration to 0.001
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotation interval 50, migration 0.001"}
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1"}
+# Decrease migration to 0.001
 
@@
 
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotate when resistant, migration 0.001"}
+
 ### Rotate when resistant, migration 0.001
-
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1"}
 

### 2017-10-05 inst/documents/rotations_notes.Rmd
active and refuge plotted on same sub-plot
build doc
allow starting frequencies to be set

@@
 separate(region, into=c("resistance_gene","active_or_refuge"))
+
+I now have run_rot() working for flexible numbers of insecticides (1-5).
+
+Done : Next, get active and refuge onto the same plots.
+
+# TODO
+~ test setting inputs correctly (move away from hardcoded test functions)
+~ find what went wrong when I set fitnesses
+~ get Ian to pull changes
+
+# state of play
+~ rotations1.Rmd : example plots

### 2017-10-05 R/run_rot.r

@@ -4,7 +4,8 @@
 #' but at present, input will only allow a maximumum of 5
 #' 
 #' @param max_generations maximum number of mosquito generations to run the simulation
-#' @param n_insecticides MAX is 5<<<<the number of insecticides (and hence loci) in the simuation MAX IS 5<<<
+#' @param n_insecticides number of insecticides (and hence loci), current max is 5
+#' @param start_freqs starting frequencies of resistance either one per insecticide or same for all
 #' @param rotation_interval frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
 #' @param rotation_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
 #' @param migration_rate_intervention migration rate into and out-of the treated area. It is the proportion of the treated population that migrates. We assume that immigration=emigration.
@@ -12,7 +13,7 @@
 #' @param plot whether to plot results
 #' @param start_insecticide which insecticide to start with
 #' @param diagnostics whether to output running info
-#' @param same_insecticides whether to just set fitnesses for all insecticides the same
+#' @param same_insecticides only used with hardcode_fitness, whether to just set fitnesses for all insecticides the same
 #' 
 #' 
 #' @examples 
@@ -28,7 +29,7 @@
 
 run_rot <- function( max_generations = 200, #the maximum number of mosquito generations to run the simulation
                           n_insecticides = 4, #MAX is 5<<<<the number of insecticides (and hence loci) in the simuation MAX IS 5<<<
-                          
+                          start_freqs = 0.001,
                           rotation_interval = 0, #frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
                           rotation_criterion = 0.5, #resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
                           
@@ -82,8 +83,12 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
   
   
   ### set starting allele frequencies from hardcoded test function or based on other inputs
-  RAF <- set_start_freqs_test( n_insecticides=n_insecticides, max_generations=max_generations )  
-  
+  #RAF <- set_start_freqs_test( n_insecticides=n_insecticides, max_generations=max_generations )  
+  #todo add checks thats start_freqs is either length 1 or n_insecticides
+  RAF <- set_start_freqs( n_insecticides=n_insecticides, 
+                          max_generations=max_generations, 
[…]

### 2017-10-05 R/set_start_freqs.r

@@ -1,3 +1,50 @@
+#' set_start_freqs allows starting allele frequencies to be set
+#' 
+#' 
+#' @param n_insecticides number of insecticides, optional can just be specified by number of items in vector expo
+#' @param freqs starting allele frequencies either one per insecticide or same for all
+#'              in this version same for m,f,intervention & refuge    
+#' 
+#' @examples
+#' #frequencies the same for all insecticides
+#' RAF <- set_start_freqs(n=3, freqs=0.001)
+#' RAF[,,,1] # to view generation 1
+#' #frequencies different for each insecticide
+#' RAF <- set_start_freqs(n=3, freqs=c(0.1,0.01,0.001))
+#' RAF[,,,1] # to view generation 1
+#' 
+#' #allowing array to be viewed differently
+#' as.data.frame(RAF)
+#' 
+#' @return array of allele frequencies with just first record filled
+#' @export
+#' 
+set_start_freqs <- function( n_insecticides = NULL,
+                             max_generations = 200,
+                             freqs = 0.001 )
+{
+  
+  #get n_insecticides if it is not specified
+  #todo add checks, allow single
+  if ( is.null(n_insecticides)) n_insecticides <- length(freqs)
+  
+  # create empty array
+  # RAF stands for resistance allele frequency
+  RAF <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), site=c('intervention','refugia'), gen=1:max_generations)
+
+  # fill generation 1
+  # in this version same for m,f,intervention & refuge
+  RAF[, 'm', 'intervention', 1] = freqs
+  # dim1 :n_insecticides
+  # dim2 :sex: m,f
+  # dim3 :site: intervention,refuge
+  # dim4 : generations
+  # I checked this fills correctly for dims2 & 3 when freqs is a vector
+  RAF[ , , , 1] = freqs
+   
+  return(RAF)  
+}  
+
 #' set_start_freqs_test allows starting allele frequencies to be hardcoded
 #' 
 #' 
@@ -48,8 +95,6 @@ set_start_freqs_test <- function( n_insecticides = NULL,
     RAF[5, 'm', 'refugia',1]=0.001;      RAF[5, 'f', 'refugia',1]=0.001
   } 
   
-  
-  #if (plot) plot_exposure(a_expo)
   
   return(RAF)
 }

### 2017-10-05 inst/documents/rotations_notes.Rmd
fixed setting exposures differently for each insecticide

--- 2017-10-05 inst/documents/rotations_notes.Rmd
@@
 separate(region, into=c("resistance_gene","active_or_refuge"))
+
+I now have run_rot() working for flexible numbers of insecticides (1-5).
+
+done : Next, get active and refuge onto the same plots.  
+done : allowed starting frequencies to be set
+
+implemented set_expsoure_rot() but currently changing exposure seems to make no difference ?
+dfr <- run_rot(rotation_interval=50, expo_hi=0.9)
+dfr <- run_rot(rotation_interval=50, expo_hi=0.2)
+# these produce identical frequency curves for r1 & r2
+dfr <- run_rot(n=2, rotation_interval=50, expo_hi=c(1,0.1))
+
+# TODO
+~ implement set_exposure_rot() to replace test version
+
+~ test setting inputs correctly (move away from hardcoded test functions)
+~ find what went wrong when I set fitnesses
+~ get Ian to pull changes
+
+# state of play
+~ rotations1.Rmd : example plots

### 2017-10-05 R/run_rot.r

@@ -14,7 +14,9 @@
 #' @param start_insecticide which insecticide to start with
 #' @param diagnostics whether to output running info
 #' @param same_insecticides only used with hardcode_fitness, whether to just set fitnesses for all insecticides the same
-#' 
+#' @param expo_hi exposure to insecticide in hi niche, either single or vector of 1 per insecticide
+#' @param expo_lo exposure to insecticide in lo niche, either single or vector of 1 per insecticide
+#' @param male_expo_prop proportion tht males are exposed relative to f, default 1, likely to be <1 (could possibly be a vector per insecticide)
 #' 
 #' @examples 
 #' run_rot(rotation_interval=100)
@@ -90,8 +92,11 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
                           freqs = start_freqs )  
     
   ### set exposures from hardcoded test function or based on other inputs
-  exposure <- set_exposure_rot_test( n_insecticides=n_insecticides )
-  #exposure <- set_exposure_rot()
+  #exposure <- set_exposure_rot_test( n_insecticides=n_insecticides )
+  exposure <- set_exposure_rot( n_insecticides=n_insecticides,
+                                expo_hi = expo_hi,
+                                expo_lo = expo_lo,                              
+                                male_expo_prop = male_expo_prop)
   
   ### set fitnesses from hardcoded test function or based on other inputs
   if (hardcode_fitness)

### 2017-10-05 R/set_exposure_rot.r

@@ -10,8 +10,9 @@
 #' fills an array of exposure values
 #' 
 #' @param n_insecticides number of insecticides, optional can just be specified by number of items in vector expo
-#' @param expo exposure to the insecticides
-#' @param male_expo_prop proportion tht males are exposed relative to f, default 1, likely to be <1
+#' @param expo_hi exposure to insecticide in hi niche, either single or vector of 1 per insecticide
+#' @param expo_lo exposure to insecticide in lo niche, either single or vector of 1 per insecticide
+#' @param male_expo_prop proportion tht males are exposed relative to f, default 1, likely to be <1 (could possibly be a vector per insecticide)
 #' @param plot whether to plot exposure    
 #' 
 #' @examples
@@ -32,8 +33,11 @@ set_exposure_rot <- function( n_insecticides = NULL,
 {
   
   #get n_insecticides if it is not specified
-  #todo add checks, allow single
   if ( is.null(n_insecticides)) n_insecticides <- length(expo_hi)
+  
+  if ( ( length(expo_hi) != 1 & length(expo_hi) != n_insecticides ) |
+       ( length(expo_lo) != 1 & length(expo_lo) != n_insecticides ) )
+    stop("expo_hi & lo need to be length 1 or n_insecticides")
   
   #exposure to insecticide
   #exposure array initialise with 0s 

### 2017-10-05 inst/documents/rotations1.Rmd
fixed setting exposures differently for each insecticide
starting to allow fitnesses to be set properly

@@
 
+# Current state of play
+
+1. Flexible number of insecticides between 1 & 5
+1. Exposure can be set for each insecticide or same for all.
+1. Starting frequency can be set for each insecticide or same for all.
+
+Some example plots follow, these act as a visual test that the model is doing what we expect.
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="exposure set to 1 for r1 and 0.1 for r2"}
+
+#expo_hi=c(1,0.1)
+dfr <- run_rot(n=2, rotation_interval=100, expo_hi=c(1,0.1))
+
+```
+
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotation interval 50, migration 0.01"}
@@
 
+# TODO remove hardcode fitnesses from here
 dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration_rate_intervention=0.01)
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotation interval 50, migration 0.001"}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="decrease migration to 0.001"}
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotate when resistant, migration 0.001"}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="cost 0.1 effectiveness 1"}
+dfr <- run_rot(n=3, rotation_interval=50, eff=1, cost=0.1)
+```
 
-### Rotate when resistant, migration 0.001
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotate when resistant, exposure 0.5 migration 0.01"}
 
-dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration_rate_intervention=0.001)
+### Rotate when resistant, migration 0.01
+
+dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration_rate_intervention=0.01)
 

### 2017-10-05 inst/documents/rotations_notes.Rmd
starting to allow fitnesses to be set properly

--- 2017-10-05 inst/documents/rotations_notes.Rmd
@@
 separate(region, into=c("resistance_gene","active_or_refuge"))
+
+I now have run_rot() working for flexible numbers of insecticides (1-5).
+
+done : Next, get active and refuge onto the same plots.  
+done : allowed starting frequencies to be set
+
+implemented set_expsoure_rot() now working
+dfr <- run_rot(rotation_interval=50, expo_hi=0.9)
+dfr <- run_rot(rotation_interval=50, expo_hi=0.2)
+# diff frequency curves for r1 & r2
+dfr <- run_rot(n=2, rotation_interval=50, expo_hi=c(1,0.1))
+
+started fitness_single_locus() from run_rot()
+
+dfr <- run_rot(n=2, rotation_interval=50, eff=c(1,0.1))
+
+# TODO
+~ test fitness_single_locus() from run_rot() and add to Rmd
+
+~ test setting inputs correctly (move away from hardcoded test functions)
+~ find what went wrong when I set fitnesses
+~ get Ian to pull changes
+
+# state of play
+~ rotations1.Rmd : example plots

### 2017-10-09 inst/documents/rotations_notes.Rmd
added more params to UI, and output of code

@@
 
+*9/10/17*
+~ started first shiny UI : rotresist1
+~ add rot interval and generations to rotresist
+
 # TODO
+
+~ seems that resistance evolves more slowly in later insecticides despite same params. Why ?
+
+e.g. :
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 50 , eff = 0.7 , dom = 0.5 , rr = 0.5 , expo_hi = 0.7 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
+
+*Ian said first question is to assess whether there is any benefit to rotations when there are no costs and no emigration*
+
 ~ test fitness_single_locus() from run_rot() and add to Rmd

### 2017-10-09 R/run_rot.r

@@ -225,7 +225,7 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
       norm_coeff=F_female_r_intervention+F_female_s_intervention
       RAF[insecticide, 'f', 'intervention', gen]=F_female_r_intervention/norm_coeff
       
-      if(diagnostics) message(sprintf("generation %d: just completed insecticide selection for locus/insecticide %d\n", gen, insecticide))
+      if(diagnostics) message(sprintf("generation %d: completed insecticide selection for locus/insecticide %d\n", gen, insecticide))
       
       
     } #end of loop that deals with this insecticide if it is being deployed
@@ -254,7 +254,7 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
     #same allele frequencies in both sexes if no differential exposure so 
        RAF[insecticide, 'f', 'intervention', gen]=RAF[insecticide, 'm', 'intervention', gen]
        
-      if(diagnostics) message(sprintf("generation %d: just completed natural selection against locus %d in intervention site\n", gen, insecticide))   
+      if(diagnostics) message(sprintf("generation %d: completed selection against locus %d in intervention site\n", gen, insecticide))   
           
      
        } #end of code for insecticides that are not being deployed  in the intervention site
@@ -279,7 +279,7 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
       #same allele frequencies in both sexes if no differential selection so
       RAF[insecticide, 'f', 'refugia', gen]=RAF[insecticide, 'm', 'refugia', gen]
       
-      if(diagnostics) message(sprintf("generation %d: just completed natural selection against locus %d in refugia\n", gen, insecticide))
+      if(diagnostics) message(sprintf("generation %d: completed selection against locus %d in refugia\n", gen, insecticide))
       
     }   #end of cycling insecticides
     
@@ -337,7 +337,7 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
   for(temp_int in 1:n_insecticides){
     if(candidate==n_insecticides) candidate=1 else candidate=candidate+1 
     if(RAF[candidate, 'f','intervention', gen]<rotation_criterion){
-      message(sprintf("generation %d, switching from insecticide %d to insecticide %d; RAF are %f and %f respectively\n",
+      message(sprintf("generation %d, switch from insecticide %d to %d; frequencies = %f and %f\n",
                       gen, current_insecticide, candidate,
[…]

### 2017-10-09 inst/shiny/rotresist1/server.r

@@ -22,44 +22,33 @@ shinyServer(function(input, output) {
         
         #cat("running resistSimple with these inputs:", input$P_1, input$P_2*input$P_1, input$h.RS1_00, input$h.RS2_00,"\n")
         
-        run_rot(n_insecticides = input$n_A, 
-                rotation_interval = 50, 
-                max_generations = 200,
-                eff = input$effectiveness_A1,
-                dom = input$dominance_A1,
-                rr = input$advantage_A1,
-                cost = input$cost_A1)
-        
-        # runcurtis_f2( max_gen = 500,
-        #               P_1 = input$frequency_A1, 
-        #               P_2 = input$frequency_A2,
-        #               #P_2 = input$P_2*input$P_1, #if doing 2 as a proportion of 1, i need to put a limit of 1 on this
-        #               h.RS1_A0 = input$dominance_A1, 
-        #               h.RS2_0B = input$dominance_A2,
-        #               exposure = input$exposure_A,
-        #               phi.SS1_A0 = input$effectiveness_A1,
-        #               phi.SS2_0B = input$effectiveness_A2,
-        #               rr_restoration_ins1 = input$advantage_A1,
-        #               rr_restoration_ins2 = input$advantage_A2,
-        #               cex.axis = 1,
-        #               addCombinedStrategy = FALSE )
+        run_rot(n_insecticides =    input$n_A, 
+                max_generations =   input$max_generations_A,
+                start_freqs =       input$frequency_A,
+                rotation_interval = input$rotation_interval_A, 
+                eff =               input$effectiveness_A,
+                dom =               input$dominance_A,
+                rr =                input$advantage_A,
+                expo_hi =           input$exposure_A,
+                coverage =          input$coverage_A,
+                migration_rate_intervention = input$migration_A, 
+                cost =              input$cost_A)
 
                 
-        #a hack to output the inputs
-        # cat("A:\n")
-        # cat("runcurtis_f2( max_gen=500, ", 
-        #               "P_1 =",input$frequency_A1,",", 
-        #               "P_2 =",input$frequency_A2,",", 
-        #               "h.RS1_A0 =",input$dominance_A1,",", 
-        #               "h.RS2_0B =",input$dominance_A2,",",
-        #               "exposure =",input$exposure_A,",",
-        #               "phi.SS1_A0 =",input$effectiveness_A1,",",
-        #               "phi.SS2_0B =",input$effectiveness_A2,",",
[…]

### 2017-10-10 inst/documents/rotations1.Rmd
improve UI and plots add log scale option

@@
 ```
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Showing how with cost (0.1) an initial decline in the freq of the later insecticides when they are not being used can lead them to behave differently than earlier ones."}
+
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
+
+```
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="When cost is set to 0 all insecticides give the same curves."}
+
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0 )

### 2017-10-10 inst/documents/rotations_notes.Rmd
improve UI and plots add log scale option

@@
 ~ add rot interval and generations to rotresist
+~ looking good
+
+*10/10/17*
+
+~ seems that resistance evolves more slowly in later insecticides despite same params. Why ?
+
+~ aha! the new UI helped me to see that this can be a result of the decline in resistance frequency for the later insecticides when they are not being used at the start. e.g.
+
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
+
+When cost is set to 0 all 3 insecticides give the same curves :
+
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0 )
+
+done ~ in plots change r1 to insecticide_1 etc.
+done ~ in plots add the deployment symbol to legend
+
+done ~ can I offer option to have plots on log y axis to make it clearer whats going on at lower frequencies ?
+
+Now from the UI ticking log y axis influences the next plot, (but is isolated so allows you to compare log & non log plots).
+
+Starting to think about sensitivity analysis.
+For Levick et al the inputs and ranges were hardcoded into resistance::sensiAnPaperPart()
+
+start_freq |    0.0001 - 0.1 log uniform
+exposure |      0.1 - 0.9
+male_exposure | 0 - 1
+effectiveness | 0.3 - 1
+dominance     | 0 - 1
+rr_restoration |  0.2 - 1
+correct_mix   | 0.5 - 1
+
+To add for rotations
+
+n_insecticides    | 
+max_generations   | 
+rotation_interval | 
+coverage          | 
+migration         |  
+cost              |
 
@@
 
-~ seems that resistance evolves more slowly in later insecticides despite same params. Why ?
+~ meet with Ian PM to prepare table for sensitivity analysis, we'll need to think about num insecticides, num generations & coverage too
 
-e.g. :
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 50 , eff = 0.7 , dom = 0.5 , rr = 0.5 , expo_hi = 0.7 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
+~ what were the ranges for the Levick paper
+
+~ Ian said we may want to 
+~ Ian will send me grant app to assess single v multi locus insecticide resistance
+~ Ian said we may want to put something in to stop Frequency going below initial starting value when not in use (because this could represent background balance between benefits & cost)
 
@@
 
-~ test setting inputs correctly (move away from hardcoded test functions)
-~ find what went wrong when I set fitnesses
 ~ get Ian to pull changes

### 2017-10-10 R/rot_plots.r

@@ -1,6 +1,7 @@
 #' rot_plot_resistance plot rotation simulation resistance results
 #'
 #' @param df_res2 dataframe of resistance results from run_rot()
+#' @param logy whether to use log scale for y axis
 #'
 #'
 #' @import ggplot2
@@ -12,7 +13,8 @@
 #' df_res2 <- run_rot()
 #' rot_plot_resistance(df_res2)
 #' 
-rot_plot_resistance <- function(df_res2) {
+rot_plot_resistance <- function(df_res2,
+                                logy = TRUE) {
   
   # column names of input dataframe
   # "generation"  "insecticide"     "resist_gene"  "active_or_refuge" "resistance"
@@ -24,7 +26,7 @@ rot_plot_resistance <- function(df_res2) {
     #mutate( i_in_use = ifelse(insecticide==1,1,NA))
     #mutate( i_in_use = ifelse(stringr::str_detect(region,paste0(insecticide,"_active")),1.05,NA))    
     #now that active & refuge on same plot
-    mutate( i_in_use = ifelse(resist_gene==paste0('r',insecticide),1.05,NA))    
+    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide),1.05,NA))    
   
     
   gg <- ggplot( df_res2, aes_string(x='generation',y='resistance',colour='active_or_refuge') ) + 
@@ -34,12 +36,25 @@ rot_plot_resistance <- function(df_res2) {
     facet_wrap('resist_gene', ncol=1) +
     #theme(axis.text.x = element_blank()) +
     #add insecticide use indication
-    geom_line( aes_string(x='generation',y='i_in_use'), colour='green4', lwd=2) +
+    #geom_line( aes_string(x='generation',y='i_in_use'), colour='green4', lwd=2) +
+    #trying to get 2nd legend (this puts i_in_use in middle OK start)
+    geom_line( aes_string(x='generation',y='i_in_use', colour=factor('i_in_use')), lwd=2) +
+    #the two below don't work to remove title
+    #guide_legend(title=NULL) +
+    #scale_fill_discrete(guide = guide_legend(title = NULL)) +
     #annotate("text", x = 0, y = 1.15, label = "deployment", size = 2.5, hjust='left') +
     #scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1)) +
[…]

### 2017-10-10 inst/shiny/rotresist1/server.r

@@ -32,7 +32,8 @@ shinyServer(function(input, output) {
                 expo_hi =           input$exposure_A,
                 coverage =          input$coverage_A,
                 migration_rate_intervention = input$migration_A, 
-                cost =              input$cost_A)
+                cost =              input$cost_A,
+                logy =              input$logy ) #inefficient that it reruns when changing to log
 
                 
         # a hack to output the inputs, so it can be run from the console
@@ -47,7 +48,8 @@ shinyServer(function(input, output) {
                       "expo_hi =",input$exposure_A,",",
                       "coverage =",input$coverage_A,",",
                       "migration_rate_intervention =",input$migration_A,",",
-                      "cost =",input$cost_A,
+                      "cost =",input$cost_A,",",
+                      "logy =",input$logy,
                       ")\n" )
         
                
@@ -73,7 +75,9 @@ shinyServer(function(input, output) {
                 expo_hi =           input$exposure_B,
                 coverage =          input$coverage_B,
                 migration_rate_intervention = input$migration_B, 
-                cost =              input$cost_B)
+                cost =              input$cost_B,
+                logy =              input$logy ) #inefficient that it reruns when changing to log
+        
         
         
         # a hack to output the inputs, so it can be run from the console
@@ -88,7 +92,8 @@ shinyServer(function(input, output) {
             "expo_hi =",input$exposure_B,",",
             "coverage =",input$coverage_B,",",
             "migration_rate_intervention =",input$migration_B,",",
-            "cost =",input$cost_B,
+            "cost =",input$cost_B,",",
+            "logy =",input$logy,
             ")\n" )
         
       }) #end isolate  

### 2017-10-11 inst/documents/rotations_notes.Rmd
add comment about getting out of rotation if frequency exceeded

@@
 
-To add for rotations
+For rotations
 
-n_insecticides    | 
+n_insecticides    | 3,4,5
 max_generations   | 
-rotation_interval | 
-coverage          | 
+rotation_interval | 5 - 50 (integer)
+male_expo_prop    | 0 - 1
+coverage          | 0.1 - 0.9
 migration         |  
-cost              |
+cost              | 0 - 0.05
+
+dominance of selection  |
+dominance of cost       | 0 - 1
+
+Sudo manuscript rotations are always a single generation.
+
+How do we deal with the problem that resistance frequenciesin the rotation scenario may go above 50% and this makes time-to-resistance not a good measure of how good the intervention is (because the rotation).
+
+Ian wants it to be a fair comparison.
+
+I'm nervous about making the rotation repsonsive. We both said that i unlikely to happen in the field.
+
+Could we have number of generations at < 50% ? Not perfect either.
 
@@
 
-~ meet with Ian PM to prepare table for sensitivity analysis, we'll need to think about num insecticides, num generations & coverage too
+~ time to 50% resistance.
 
-~ what were the ranges for the Levick paper
+~ sort outputs which need to be time to resistance for planned rotations and RwR rotate when resistant. Sequence.
+
+~ fix migration rate to be between 0 & 1, where 1 is random movement.
+~ read Ians document
+
+~ set dominance of cost and dominance of selection to be different in fitness_single_locus().
+
+~ Ian & I talked about using fitness to detremine changes.
+and that fitness could be used as an index of population, but Ian was sceptical, he said to look up fitness determinants of population on wikipedia, about soft & hard selection.
 

### 2017-10-11 R/run_rot.r

@@ -328,7 +328,9 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
     #(a) it will continue to be use if it is the only one below the threshold 
     #(b) it will be replaced if one of the other insecticides is now below the threshold due to fitness effects or migration
     #(c) the simulation will terminate if all insecticides exceed the threshold
-      
+  
+  # TODO this keeps the rotation going, if I want to add in responsive that stops the rotation if 
+  # the resistance frequency is reached then I can add it in here with &
   if (rotation_count!=rotation_interval) rotation_count=rotation_count+1
   else change_insecticide=1 #i.e. its time to rotate so need to identify the next insecticide in the rotation
   }

### 2017-10-11 inst/documents/rotations_notes.Rmd
separate dominance of selection and dominance of cost

@@
 
+*11/10/17*
+
 # TODO
+
+~ set dominance of cost and dominance of selection to be different in fitness_single_locus().
 
@@
 ~ read Ians document
-
-~ set dominance of cost and dominance of selection to be different in fitness_single_locus().
 
@@
 
-~ get Ian to pull changes
+~ check in resistance::fitnessSingleLocus() on separation of dominance of cost & selection. Was probably OK for Levick et al because we didn't have cost in. 
+But I think that dominance of cost may be set to 0 which means that the RS experiences no effect of cost.
+
+    a_dom       <- array_named(locusNum=c(1,2), exposure=c('no','lo','hi')) 
+    a_dom[1, 'hi'] <- dom1
+    a_dom[2, 'hi'] <- dom2
+
+    #exposure 0 'no'
+    a_fitloc[ paste0('RS',locusNum), 'no'] <- 1 - (a_dom[locusNum, 'no'] * a_cost[locusNum])
+    a_fitloc[ paste0('RR',locusNum), 'no'] <- 1 - a_cost[locusNum]
 

### 2017-10-11 R/fitnessSingleLocus.r

@@ -1,4 +1,5 @@
 #' calculate single locus fitness for 2 loci from insecticide exposure
+#' VERSION FROM RESISTANCE NOT USED IN ROTATIONS
 #' 
 #' can be used in 2 ways
 #' 1) by passing arrays a_dom etc. as done from runModel2()

### 2017-10-11 R/fitness_single_locus.r

@@ -1,13 +1,13 @@
 #' calculate single locus fitness for a flexible number of insecticides
 #' 
 #' can be used in 2 ways
-#' 1) by passing arrays a_dom etc. as done from runModel2()
-#' 2) pass a single value and n_insecticides to give all insecticides the same value
+#' 1) pass a single value and n_insecticides to give all insecticides the same value
 #' 2) pass vectors - new way for rotations
 
 #' @param n_insecticides number of insecticides (not needed if vectors or arrays are passed)
 #' @param eff effectiveness, for all insecticides or individually
-#' @param dom dominance, for all insecticides or individually
+#' @param dom_sel dominance of selection, for all insecticides or individually
+#' @param dom_cos dominance of cost, for all insecticides or individually
 #' @param rr resistance restoration, for all insecticides or individually 
 #' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
 #' @param fitSS fitness of SS if no insecticide, for all insecticides or individually
@@ -23,18 +23,19 @@
 #' #defaults
 #' fitness_single_locus()
 #' #2 different insecticides
-#' fitness_single_locus(eff=c(0.5,1), dom=c(0.5,0.5), rr=c(0.5,0.5), cost=c(0,0.1), fitSS=c(1,1))
+#' fitness_single_locus(eff=c(0.5,1), dom_sel=c(0.5,0.5), dom_cost=c(0.5,0.5), rr=c(0.5,0.5), cost=c(0,0.1), fitSS=c(1,1))
 #' #4 same insecticides
-#' fitness_single_locus(n_insecticides=4, eff=0.5, dom=0.5, rr=0.5, cost=0, fitSS=1)
+#' fitness_single_locus(n_insecticides=4, eff=0.5, dom_sel=0.5, dom_cost=0.5, rr=0.5, cost=0, fitSS=1)
 #' #4 pairs of the same insecticides
-#' fitness_single_locus(n_insecticides=8, eff=c(0.5,1), dom=c(0.5,0.5), rr=c(0.5,0.5), cost=c(0,0.1), fitSS=c(1,1))
+#' fitness_single_locus(n_insecticides=8, eff=c(0.5,1), dom_sel=c(0.5,0.5), dom_cost=c(0.5,0.5), rr=c(0.5,0.5), cost=c(0,0.1), fitSS=c(1,1))
 
 #' @return fitness values
 #' @export
 
 fitness_single_locus <- function ( n_insecticides = NULL,
[…]

### 2017-10-11 R/run_rot.r

@@ -18,7 +18,8 @@
 #' @param expo_lo exposure to insecticide in lo niche, either single or vector of 1 per insecticide
 #' @param male_expo_prop proportion tht males are exposed relative to f, default 1, likely to be <1 (could possibly be a vector per insecticide)
 #' @param eff effectiveness, for all insecticides or individually
-#' @param dom dominance, for all insecticides or individually
+#' @param dom_sel dominance of selection, for all insecticides or individually
+#' @param dom_cos dominance of cost, for all insecticides or individually
 #' @param rr resistance restoration, for all insecticides or individually 
 #' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
 #' @param fitSS fitness of SS if no insecticide, for all insecticides or individually
@@ -52,7 +53,8 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
                       expo_lo = 0,                              
                       male_expo_prop = 1,
                      eff = 0.5, #c(0.5, 0.7, 0.9),
-                     dom = 0.5, #c(0.5, 0.5, 0.5),
+                     dom_sel = 0.5, #c(0.5, 0.5, 0.5),
+                     dom_cos = 0.5, #c(0.5, 0.5, 0.5),
                      rr = 0.5, #c(0.5, 0.5, 0.5),
                      cost = 0.5, #c(0,0,0),
                      fitSS = 1,
@@ -114,8 +116,13 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
     fitness <- fitness_single_locus_test( n_insecticides=n_insecticides, same_insecticides = same_insecticides )
   } else
   {
-    #fitness <- fitness_single_locus(n_insecticides=n_insecticides, eff=0.5, dom=0.5, rr=0.5, cost=0, fitSS=1)
-    fitness <- fitness_single_locus(n_insecticides=n_insecticides, eff=eff, dom=dom, rr=rr, cost=cost, fitSS=fitSS)
+    fitness <- fitness_single_locus(n_insecticides=n_insecticides, 
+                                    eff=eff, 
+                                    dom_sel=dom_sel, 
+                                    dom_cos=dom_cos, 
+                                    rr=rr, 
+                                    cost=cost, 
+                                    fitSS=fitSS)
     
   }
   

### 2017-10-31 inst/documents/rotations_notes.Rmd
get UI to cope with split dominance

@@
 
-I'm nervous about making the rotation repsonsive. We both said that i unlikely to happen in the field.
+I'm nervous about making the rotation responsive. We both said that i unlikely to happen in the field.
 
@@
 
-*11/10/17*
+~ setting dominance of cost and dominance of selection to be different in fitness_single_locus().
+
+*31/10/17*
+
+UI :
+Warning: Error in run_rot: argument 6 matches multiple formal arguments
+
+probably due to recent splitting of dominance
+
+Yes, fixed now.
+
+Increasing dominance of cost decreases resistance, 
+increasing dominance of selection increases resistance. (both as expected)
 
 # TODO
-
-~ set dominance of cost and dominance of selection to be different in fitness_single_locus().
 
@@
 ~ rotations1.Rmd : example plots
+~ UI : rotations/shiny/rotresist1/server.r
+
+# funding opportunities
+https://lstmed.sharepoint.com/research/RBPS/Pages/Funding%20Information/Funding-Opportunities.aspx     
+https://www.researchprofessional.com/funding/simple?q=software#1509451534514

### 2017-10-31 R/fitness_single_locus.r

@@ -23,11 +23,11 @@
 #' #defaults
 #' fitness_single_locus()
 #' #2 different insecticides
-#' fitness_single_locus(eff=c(0.5,1), dom_sel=c(0.5,0.5), dom_cost=c(0.5,0.5), rr=c(0.5,0.5), cost=c(0,0.1), fitSS=c(1,1))
+#' fitness_single_locus(eff=c(0.5,1), dom_sel=c(0.5,0.5), dom_cos=c(0.5,0.5), rr=c(0.5,0.5), cost=c(0,0.1), fitSS=c(1,1))
 #' #4 same insecticides
-#' fitness_single_locus(n_insecticides=4, eff=0.5, dom_sel=0.5, dom_cost=0.5, rr=0.5, cost=0, fitSS=1)
+#' fitness_single_locus(n_insecticides=4, eff=0.5, dom_sel=0.5, dom_cos=0.5, rr=0.5, cost=0, fitSS=1)
 #' #4 pairs of the same insecticides
-#' fitness_single_locus(n_insecticides=8, eff=c(0.5,1), dom_sel=c(0.5,0.5), dom_cost=c(0.5,0.5), rr=c(0.5,0.5), cost=c(0,0.1), fitSS=c(1,1))
+#' fitness_single_locus(n_insecticides=8, eff=c(0.5,1), dom_sel=c(0.5,0.5), dom_cos=c(0.5,0.5), rr=c(0.5,0.5), cost=c(0,0.1), fitSS=c(1,1))
 
 #' @return fitness values
 #' @export

### 2017-10-31 inst/shiny/rotresist1/ui.r

@@ -57,9 +57,10 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
   fluidRow(
     column(2, p()),
     column(2, actionButton('aButtonRunA', 'Run Scenario A')),
-    column(1, p()),   
     #column(3, h6("(takes a few seconds)")),
-    column(3, checkboxInput("logy", "log y axis", FALSE)),
+    column(1, sliderInput("max_generations", "generations", val=150, min = 100, max = 500, step = 50, ticks=FALSE)),
+    column(1, p()), 
+    column(1, checkboxInput("logy", "log y axis", FALSE)),
     column(2, actionButton('aButtonRunB', 'Run Scenario B')),
     column(2, p())
   ), #end fixed row
@@ -80,14 +81,15 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
            hr(), #hr(),hr(),
            sliderInput("n_B", NULL, val=3, min = 1, max = 5, step = 1, ticks=FALSE),
            hr()
-    ),   
-    column(1, offset = 0,
-           h5("generations"),
-           sliderInput("max_generations_A", NULL, val=150, min = 100, max = 500, step = 50, ticks=FALSE),
-           hr(), #hr(),hr(),
-           sliderInput("max_generations_B", NULL, val=150, min = 100, max = 500, step = 50, ticks=FALSE),
-           hr()
-    ),  
+    ), 
+    # moved up to one slider for both scenarios
+    # column(1, offset = 0,
+    #        h5("generations"),
+    #        sliderInput("max_generations_A", NULL, val=150, min = 100, max = 500, step = 50, ticks=FALSE),
+    #        hr(), #hr(),hr(),
+    #        sliderInput("max_generations_B", NULL, val=150, min = 100, max = 500, step = 50, ticks=FALSE),
+    #        hr()
+    # ),  
     column(1, offset = 0,
            h5("interval"),
            sliderInput("rotation_interval_A", NULL, val=50, min = 0, max = 100, step = 10, ticks=FALSE),
@@ -103,10 +105,17 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
            hr()          
     ),    
     column(1, offset = 0,
-           h5("dominance"),
-           sliderInput("dominance_A", NULL, val=0.5, min = 0, max = 1, step = 0.1, ticks=FALSE),
+           h5("dominance, selection"),
[…]

### 2017-11-01 inst/documents/rotations1.Rmd
rescale migration 0-1, don't plot refuge if migration=0

@@
 # TODO remove hardcode fitnesses from here
-dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration_rate_intervention=0.01)
+dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.01)
 
@@
 
-dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration_rate_intervention=0.001)
+dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.001)
 
@@
 
-dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration_rate_intervention=0.01)
+dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration=0.01)
 
@@
 
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.1 )
 
@@
 
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0 )
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0 )

### 2017-11-01 inst/documents/rotations_notes.Rmd
rescale migration 0-1, don't plot refuge if migration=0

@@
 
+Ians document : manuscript notes.doc
+
+We will compare 4 strategies
+•	RwR (rotate when resistant.. really a sequential use policy but the term RwR emphasises that can return to insecticides used earlier in the sequence if they are now effective
+•	Periodically rotate every 10 generations (provided there replacement insecticide has resistance levels below the threshold. 10 generations is about every year
+•	Rotate every 20 generations (~ 2 years)
+•	Rotate every 50 generations (~ 5 years).
+
+Threshold for a policy change: resistance allele frequency > 50% (may also want to investigate a lower level e.g. 25% and/or a fitness criterion)
+
+End of policy life: when resistance levels to all insecticides exceed the threshold for policy change.
+
+Sensitivity analysis to test which of the policies is best will explore the parameter space defined on table 1.
+However, there are a few test cases we will examine first to test our understanding and illustrate the basic principles.
+
+Test 1. No refugia (i.e. Coverage=100%) or fitness costs.
+Would expect all policies to be equivalent.
+
+Test 2. Dominance of resistance and fitness costs both set to 0.5.
+I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles.
+I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same.
+I’m probably wrong but it’s a nice test…..
+
+Things still to consider
+RAF can go below the original defined level  due to natural selection…de we want to stop this e.g. by saying this minimum value is mutation/selection
+
+What should we do about planned rotations where RAF>50% for several generation i.e. is an ineffective deployment. This gives it an ‘unfair’ advantage over RwR
+
+We still need to set a trap so can’t rotate on the basis of a few generations e.g. every 2 gens indefinitely as selection brings RAF>0.5 and natural selection brings RAF<50%
+
+Notes:
+List of research questions (draft keep adding)
+
+•	What are the relative impacts of fitness costs and size of refugia. Presumably we find lifespan of policy then do PRC (partial rank correlation) to quantify their impact.
+
+•	Routine rotation may be better than a “rotate-when-resistant” (RwR) policy if resistance is recessive i.e.
+rapid rotation ensures resistance allele frequency in treated populations remains so low that resistant homozygotes are rare.
+
+•	Can contrast to mixtures.
+Would have to use the full model for mixtures and this one for rotation.
+Unfortunately, we can’t have a separate refugia in the full model at present but will build this when simulating macro-mosaics.
+Once that is done we can use the full model for rotations so don’t need the simple model described here, although the algebra above may be faster than running the full model.
+
+We are not necessarily recommending rotate when resistant (RwR) policy, but it serves a good baseline to compare reactive policies (such as RwR) with pro-active polices such as rotations.
+
+RwR may be good if there is an exceptionally long lived insecticide in the armoury as will find and use it.
+If that long-lived insecticide is part of a rotation sequence, you may never find it (until it has outlasted all the others).
+
+When RwR you know how many insecticides you have left. In a strict rotation, they may all start to exhibit resistance at roughly the same time with little fore-warning.
+
+~ looking at fixing migration rate to be between 0 & 1, where 1 is random movement.
+
+~~ currently suggestion is that it is set to (0.1-0.9)*(1- Coverage)
+
+So I want to remove dependence on the 1-coverage bit, by putting that into the code.
+
+migration_rate_intervention = 0.01,
+
+migration rate into and out-of the treated area. It is the proportion of the treated population that migrates. We assume that immigration=emigration.
+
+First line of run_rot()
+
+migration_rate_refugia = migration_rate_intervention * coverage/(1-coverage)
+
+This is done for all insecticides
+
+     fem_intervention=
+       (1-migration_rate_intervention)*RAF[temp_int, 'f', 'intervention', gen]+
+       migration_rate_intervention*RAF[temp_int, 'f', 'refugia', gen]
+
+     male_intervention=
+       (1-migration_rate_intervention)*RAF[temp_int, 'm', 'intervention', gen]+
+       migration_rate_intervention*RAF[temp_int, 'm', 'refugia', gen]
+
+     fem_refugia=
+       (1-migration_rate_refugia)*RAF[temp_int, 'f', 'refugia', gen]+
+       migration_rate_refugia*RAF[temp_int, 'f', 'intervention', gen]
+
+     male_refugia=
+       (1-migration_rate_refugia)*RAF[temp_int, 'm', 'refugia', gen]+
+       migration_rate_refugia*RAF[temp_int, 'm', 'intervention', gen]
+
+    RAF[temp_int, 'f', 'intervention', gen]=fem_intervention
+    RAF[temp_int, 'm', 'intervention', gen]=male_intervention 
+    RAF[temp_int, 'f', 'refugia', gen]=fem_refugia
+    RAF[temp_int, 'm', 'refugia', gen]=male_refugia
+
+Perhaps I can put this into a little spreadsheet to test ?
+Or into a little test function.
+e.g. that when migration set to 1 mixing is random ?
+Just for 1 gender.
+
+coverage
+
+migration_refugia = migration_intervention * coverage/(1-coverage)
+
+freq_new <- (1-migration)*
+
+Nearly there, testing Ians migration :
+
+Ah yes 1-c is max migration, it gives equal frequencies in one generation
+
+> migration_test(migration=0.4, startfreqs=c(0.8,0.2), coverage = 0.6)
+              gen
+site             1    2    3    4    5    6    7    8    9   10
+  intervention 0.8 0.56 0.56 0.56 0.56 0.56 0.56 0.56 0.56 0.56
+  refugia      0.2 0.56 0.56 0.56 0.56 0.56 0.56 0.56 0.56 0.56
+
 # TODO
+
+~ fix migration rate to be between 0 & 1, where 1 is random movement.
 
@@
 
+~ improve plot legend to show when insecticide in use
+
 ~ sort outputs which need to be time to resistance for planned rotations and RwR rotate when resistant. Sequence.
 
-~ fix migration rate to be between 0 & 1, where 1 is random movement.
-~ read Ians document
-
-~ Ian & I talked about using fitness to detremine changes.
+~ Ian & I talked about using fitness to determine changes.
 and that fitness could be used as an index of population, but Ian was sceptical, he said to look up fitness determinants of population on wikipedia, about soft & hard selection.
 
-~ Ian said we may want to 
 ~ Ian will send me grant app to assess single v multi locus insecticide resistance

### 2017-11-01 inst/shiny/rotresist1/server.r

@@ -32,7 +32,7 @@ shinyServer(function(input, output) {
                 rr =                input$advantage_A,
                 expo_hi =           input$exposure_A,
                 coverage =          input$coverage_A,
-                migration_rate_intervention = input$migration_A, 
+                migration =         input$migration_A, 
                 cost =              input$cost_A,
                 logy =              input$logy ) #inefficient that it reruns when changing to log
 
@@ -49,7 +49,7 @@ shinyServer(function(input, output) {
                       "rr =",input$advantage_A,",",
                       "expo_hi =",input$exposure_A,",",
                       "coverage =",input$coverage_A,",",
-                      "migration_rate_intervention =",input$migration_A,",",
+                      "migration =",input$migration_A,",",
                       "cost =",input$cost_A,",",
                       "logy =",input$logy,
                       ")\n" )
@@ -77,7 +77,7 @@ shinyServer(function(input, output) {
                 rr =                input$advantage_B,
                 expo_hi =           input$exposure_B,
                 coverage =          input$coverage_B,
-                migration_rate_intervention = input$migration_B, 
+                migration =         input$migration_B, 
                 cost =              input$cost_B,
                 logy =              input$logy ) #inefficient that it reruns when changing to log
         
@@ -95,7 +95,7 @@ shinyServer(function(input, output) {
             "rr =",input$advantage_B,",",
             "expo_hi =",input$exposure_B,",",
             "coverage =",input$coverage_B,",",
-            "migration_rate_intervention =",input$migration_B,",",
+            "migration =",input$migration_B,",",
             "cost =",input$cost_B,",",
             "logy =",input$logy,
             ")\n" )

### 2017-11-01 inst/documents/rotations_notes.Rmd
improve plot legends

@@
 
+done ~ fix migration rate to be between 0 & 1, where 1 is random movement.
+
+~ allowed plots not to show refuge
+
 # TODO
 
-~ fix migration rate to be between 0 & 1, where 1 is random movement.
+~ improve plot legend to show when insecticide in use
+~ do a few plots for Dave Malone
+#high cost, no migration
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0.1 , logy = FALSE )
+#no cost
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
+#no cost and rotate when resistant
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 0 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
 
 ~ time to 50% resistance.
-
-~ improve plot legend to show when insecticide in use
 

### 2017-11-20 inst/documents/rotations_notes.Rmd
notes

--- 2017-11-01 inst/documents/rotations_notes.Rmd
@@
 
+done ~ fix migration rate to be between 0 & 1, where 1 is random movement.
+
+~ allowed plots not to show refuge
+
+#high cost, no migration
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0.1 , logy = FALSE )
+#no cost, no migration
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
+#no cost and rotate when resistant, no migration
+run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 0 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
+
+done ~ improve plot legend to show when insecticide in use
+
+*20/11/17*
+
+~ looking at refactoring the migration code in run_rot()
+
+Can do this to get the allele frequencies for generation 2 (by )
+
+RAF[,,,2]
+
+~ currently it does 4 very similar calcs for m|f, refuge|active.
+
+~ i can replace with single, except need to be acreful not to change values before they have been used in the other calculations.
+
+FROM :
+    for(temp_int in 1:n_insecticides){
+
+     fem_intervention=
+       (1-migration_rate_intervention)*RAF[temp_int, 'f', 'intervention', gen]+
+       migration_rate_intervention*RAF[temp_int, 'f', 'refugia', gen]
+
+     male_intervention=
+       (1-migration_rate_intervention)*RAF[temp_int, 'm', 'intervention', gen]+
+       migration_rate_intervention*RAF[temp_int, 'm', 'refugia', gen]
+
+     fem_refugia=
+       (1-migration_rate_refugia)*RAF[temp_int, 'f', 'refugia', gen]+
+       migration_rate_refugia*RAF[temp_int, 'f', 'intervention', gen]
+
+     male_refugia=
+       (1-migration_rate_refugia)*RAF[temp_int, 'm', 'refugia', gen]+
+       migration_rate_refugia*RAF[temp_int, 'm', 'intervention', gen]
+
+    RAF[temp_int, 'f', 'intervention', gen]=fem_intervention
+    RAF[temp_int, 'm', 'intervention', gen]=male_intervention 
+    RAF[temp_int, 'f', 'refugia', gen]=fem_refugia
+    RAF[temp_int, 'm', 'refugia', gen]=male_refugia
+
+   }
+
+TO :
+
+    mig_intervention <- (1-migration_rate_intervention)*RAF[,, 'intervention', gen]+
+                           migration_rate_intervention *RAF[,, 'refugia', gen]
+
+    mig_refugia      <- (1-migration_rate_refugia)*RAF[,, 'refugia', gen]+
+                           migration_rate_refugia *RAF[,, 'intervention', gen]
+
+    RAF[,, 'intervention', gen] <- mig_intervention
+    RAF[,, 'refugia', gen] <- mig_refugia
+
+# testing multiplying RAF arrays
+# to allow making code more concise
+
+RAF <- set_start_freqs(n=3, max_generations = 1, freqs=c(0.1,0.01,0.001))
+RAF2 <- set_start_freqs(n=3, max_generations = 1, freqs=c(1,2,3))
+
+RAF * RAF2
+           sex
+insecticide     m     f
+          1 0.100 0.100
+          2 0.020 0.020
+          3 0.003 0.003
+
 # TODO
 
-~ fix migration rate to be between 0 & 1, where 1 is random movement.
-
-~ time to 50% resistance.
-
-~ improve plot legend to show when insecticide in use
+~ put migration code into a function and add a test
 
 ~ sort outputs which need to be time to resistance for planned rotations and RwR rotate when resistant. Sequence.
+
+~ time under resitance threshold, e.g. 50%
+
+* rotations understand code better
+* rotations add catch to stop freq declining
+* rotate when resistant add catch to stop switch back to insecticide used < x years ago
 
@@
 
+~ RAF[,,,2] to get allele frequencies for generation 2 (by insecticide,sex and active|refuge)
+
 # funding opportunities

### 2017-11-21 inst/documents/rotations_notes.Rmd
tweaks to pass CHECK

@@
 
+Which is more concise and shows more clearly what is going on.
+
 # testing multiplying RAF arrays
@@
 
+done ~ put migration code into a function 
+started ~ add a test that checks migration for diff insecticides & mf
+
+*21/11/17*
+
 # TODO
 
-~ put migration code into a function and add a test
+~ is it worth creating a get_freq() or getraf() function similar to in rtsetse but not as complex ?
+~ pass RAF
+~ get_freq(insecticide=1, site='intervention', sex='f', gen=2 )
+~ assume all for any not specified
+~ have an as_df arg to return as a dataframe
+
+~ maybe have a set_raf() function too
+
+~ rdpeng suggests not having non standard evaluation in package code :
+If you have any non-standard evaluation in your package code (which you’ll notice because of the “no visible bindings” warnings you’ll get when you check the package), go through and change any instances to use standard evaluation alternatives.
+This change prevents these warnings when you check your package and will also ensure that the functions behave like you expect them to when they are run by other users in their own R sessions.
+~ https://bookdown.org/rdpeng/RProgDA/non-standard-evaluation.html
 

### 2017-11-21 R/run_rot.r

@@ -13,7 +13,9 @@
 #' @param plot whether to plot results
 #' @param start_insecticide which insecticide to start with
 #' @param diagnostics whether to output running info
+#' @param hardcode_fitness whether to use hardcoded fitness, default FALSE
 #' @param same_insecticides only used with hardcode_fitness, whether to just set fitnesses for all insecticides the same
+#' @param hardcode_exposure whether to use hardcoded exposure, default FALSE
 #' @param expo_hi exposure to insecticide in hi niche, either single or vector of 1 per insecticide
 #' @param expo_lo exposure to insecticide in lo niche, either single or vector of 1 per insecticide
 #' @param male_expo_prop proportion tht males are exposed relative to f, default 1, likely to be <1 (could possibly be a vector per insecticide)
@@ -24,13 +26,14 @@
 #' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
 #' @param fitSS fitness of SS if no insecticide, for all insecticides or individually
 #' @param logy whether to use log scale for y axis
-#' @param add_gens_under50 whether to add a label of num generations under 50% resistance
+#' @param add_gens_under50 whether to add a label of num generations under 50 pcent resistance
 #' 
 #' @examples 
 #' run_rot(rotation_interval=100)
 #' dfr <- run_rot(rotation_interval=50, max_generations = 300)
 #' dfr <- run_rot(rotation_interval=0, max_generations = 300)
-#' dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.01)
+#' dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, 
+#'                same_insecticides =TRUE, migration=0.01)
 #' 
 #' @import tidyverse 
 #' @return dataframe of results
@@ -43,7 +46,7 @@ run_rot <- function( max_generations = 200, #the maximum number of mosquito gene
                       rotation_interval = 10, #frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
                       rotation_criterion = 0.5, #resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
                       migration = 0.01, 
-                      #migration_rate_intervention = 0.01, # migration rate into and out-of the treated area. It is the proportion of the treated population that migrates. We assume that immigration=emigration.
[…]

### 2017-11-21 R/set_start_freqs.r

@@ -2,6 +2,7 @@
 #' 
 #' 
 #' @param n_insecticides number of insecticides, optional can just be specified by number of items in vector expo
+#' @param max_generations maximum number of mosquito generations to run the simulation
 #' @param freqs starting allele frequencies either one per insecticide or same for all
 #'              in this version same for m,f,intervention & refuge    
 #' 
@@ -13,8 +14,8 @@
 #' RAF <- set_start_freqs(n=3, freqs=c(0.1,0.01,0.001))
 #' RAF[,,,1] # to view generation 1
 #' #testing mutliplying frequency arrays
-#' RAF <- set_start_freqs(n=3, max_generations = 1, freqs=c(0.1,0.01,0.001))
-#' RAF2 <- set_start_freqs(n=3, max_generations = 1, freqs=c(1,2,3))
+#' RAF <- set_start_freqs(max_generations = 1, freqs=c(0.1,0.01,0.001))
+#' RAF2 <- set_start_freqs(max_generations = 1, freqs=c(1,2,3))
 #' RAF*RAF2
 #' 
 #' #allowing array to be viewed differently
@@ -53,6 +54,7 @@ set_start_freqs <- function( n_insecticides = NULL,
 #' 
 #' 
 #' @param n_insecticides number of insecticides, optional can just be specified by number of items in vector expo
+#' @param max_generations maximum number of mosquito generations to run the simulation
 #' @param plot whether to plot exposure    
 #' 
 #' @examples

### 2017-11-21 inst/documents/rotations1.Rmd
max_generations -> max_gen

@@
 # TODO remove hardcode fitnesses from here
-dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.01)
+dfr <- run_rot(rotation_interval=50, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.01)
 
@@
 
-dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.001)
+dfr <- run_rot(rotation_interval=50, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.001)
 
@@
 
-dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration=0.01)
+dfr <- run_rot(rotation_interval=0, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration=0.01)
 
@@
 
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.1 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.1 )
 
@@
 
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0 )

### 2017-11-21 inst/documents/rotations_notes.Rmd
max_generations -> max_gen

@@
 
-df_results <- data.frame(generation=1:max_generations,
+df_results <- data.frame(generation=1:max_gen,
                          insecticide=NA,
@@
 # create inflexible columns ready to add r* columns on
-max_generations <- 10
-df_results <- data.frame(generation=1:max_generations,
+max_gen <- 10
+df_results <- data.frame(generation=1:max_gen,
                          insecticide=NA)
@@
 
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
 
@@
 
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0 )
 
@@
 n_insecticides    | 3,4,5
-max_generations   | 
+max_gen   | 
 rotation_interval | 5 - 50 (integer)
@@
 #high cost, no migration
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0.1 , logy = FALSE )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0.1 , logy = FALSE )
 #no cost, no migration
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
 #no cost and rotate when resistant, no migration
-run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.01 , rotation_interval = 0 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 0 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
 
@@
 
-RAF <- set_start_freqs(n=3, max_generations = 1, freqs=c(0.1,0.01,0.001))
-RAF2 <- set_start_freqs(n=3, max_generations = 1, freqs=c(1,2,3))
+RAF <- set_start_freqs(n=3, max_gen = 1, freqs=c(0.1,0.01,0.001))
+RAF2 <- set_start_freqs(n=3, max_gen = 1, freqs=c(1,2,3))
 

### 2017-11-21 R/migration_test.r

@@ -3,7 +3,7 @@
 #' @param migration migration between 0-1
 #' @param coverage proportion of popn in treated area
 #' @param startfreqs vector of 2 starting resistance frequencies, first in treated area, 2nd in refuge
-#' @param max_generations maximum number of mosquito generations to run the simulation
+#' @param max_gen maximum number of mosquito generations to run the simulation
 #'
 # @import
 # @importFrom
@@ -16,19 +16,19 @@
 migration_test <- function(migration = 0.5,
                            coverage = 0.6,
                            startfreqs = c(0.2,0.8),
-                           max_generations = 10) {
+                           max_gen = 10) {
   
 #convert migration 0-1 to proportion leaving intervention
 m_int <- migration*(1-coverage)  
 m_ref <- m_int*coverage/(1-coverage)  
   
 afreq <- array_named(site=c('intervention','refugia'), 
-                      gen=1:max_generations)
+                      gen=1:max_gen)
 
 # set frequencies in gen1 for intervention & refuge  
 afreq[,1] <- startfreqs
 
-for( gen in 2:max_generations)
+for( gen in 2:max_gen)
 {
 
   # sort of worked, kept sum(f) as 1, but went to 0.5 in first gen

### 2017-11-21 R/rot_migrate.r

@@ -12,8 +12,8 @@
 #' RAF <- set_start_freqs(freqs=c(0.1,0.01,0.001))
 #' RAF[,,,1] # to view generation 1
 #' #testing mutliplying frequency arrays
-#' RAF <- set_start_freqs(max_generations = 1, freqs=c(0.1,0.01,0.001))
-#' RAF2 <- set_start_freqs(max_generations = 1, freqs=c(1,2,3))
+#' RAF <- set_start_freqs(max_gen = 1, freqs=c(0.1,0.01,0.001))
+#' RAF2 <- set_start_freqs(max_gen = 1, freqs=c(1,2,3))
 #' RAF*RAF2
 #' 
 #' 
@@ -34,7 +34,7 @@ rot_migrate <- function( RAF1gen,
   migrate_intervention <- migration*(1-coverage)
   migrate_refugia <- migrate_intervention*coverage/(1-coverage)  
   
-  #RAF <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), site=c('intervention','refugia'), gen=1:max_generations)
+  #RAF <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), site=c('intervention','refugia'), gen=1:max_gen)
   
   # RAF[,, ensures calc is repeated for each insecticide and sex 
   # more concise version of original code

### 2017-11-21 R/set_start_freqs.r

@@ -2,7 +2,7 @@
 #' 
 #' 
 #' @param n_insecticides number of insecticides, optional can just be specified by number of items in vector expo
-#' @param max_generations maximum number of mosquito generations to run the simulation
+#' @param max_gen maximum number of mosquito generations to run the simulation
 #' @param freqs starting allele frequencies either one per insecticide or same for all
 #'              in this version same for m,f,intervention & refuge    
 #' 
@@ -14,8 +14,8 @@
 #' RAF <- set_start_freqs(n=3, freqs=c(0.1,0.01,0.001))
 #' RAF[,,,1] # to view generation 1
 #' #testing mutliplying frequency arrays
-#' RAF <- set_start_freqs(max_generations = 1, freqs=c(0.1,0.01,0.001))
-#' RAF2 <- set_start_freqs(max_generations = 1, freqs=c(1,2,3))
+#' RAF <- set_start_freqs(max_gen = 1, freqs=c(0.1,0.01,0.001))
+#' RAF2 <- set_start_freqs(max_gen = 1, freqs=c(1,2,3))
 #' RAF*RAF2
 #' 
 #' #allowing array to be viewed differently
@@ -25,7 +25,7 @@
 #' @export
 #' 
 set_start_freqs <- function( n_insecticides = NULL,
-                             max_generations = 200,
+                             max_gen = 200,
                              freqs = 0.001 )
 {
   
@@ -35,7 +35,7 @@ set_start_freqs <- function( n_insecticides = NULL,
   
   # create empty array
   # RAF stands for resistance allele frequency
-  RAF <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), site=c('intervention','refugia'), gen=1:max_generations)
+  RAF <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), site=c('intervention','refugia'), gen=1:max_gen)
 
   # fill generation 1
   # in this version same for m,f,intervention & refuge
@@ -54,7 +54,7 @@ set_start_freqs <- function( n_insecticides = NULL,
 #' 
 #' 
 #' @param n_insecticides number of insecticides, optional can just be specified by number of items in vector expo
-#' @param max_generations maximum number of mosquito generations to run the simulation
+#' @param max_gen maximum number of mosquito generations to run the simulation
 #' @param plot whether to plot exposure    
 #' 
 #' @examples
@@ -67,7 +67,7 @@ set_start_freqs <- function( n_insecticides = NULL,
 #' @export
 #' 
[…]

### 2017-11-21 inst/shiny/rotresist1/server.r

@@ -23,7 +23,7 @@ shinyServer(function(input, output) {
         #cat("running resistSimple with these inputs:", input$P_1, input$P_2*input$P_1, input$h.RS1_00, input$h.RS2_00,"\n")
         
         run_rot(n_insecticides =    input$n_A, 
-                max_generations =   input$max_generations,
+                max_gen =   input$max_gen,
                 start_freqs =       input$frequency_A,
                 rotation_interval = input$rotation_interval_A, 
                 eff =               input$effectiveness_A,
@@ -40,7 +40,7 @@ shinyServer(function(input, output) {
         # a hack to output the inputs, so it can be run from the console
         cat("A:\n")
         cat("run_rot( n_insecticides =",input$n_A,",", 
-                      "max_generations =",input$max_generations,",", 
+                      "max_gen =",input$max_gen,",", 
                       "start_freqs =",input$frequency_A,",", 
                       "rotation_interval =",input$rotation_interval_A,",",
                       "eff =",input$effectiveness_A,",",
@@ -68,7 +68,7 @@ shinyServer(function(input, output) {
       isolate({
         
         run_rot(n_insecticides =    input$n_B, 
-                max_generations =   input$max_generations,
+                max_gen =   input$max_gen,
                 start_freqs =       input$frequency_B,
                 rotation_interval = input$rotation_interval_B, 
                 eff =               input$effectiveness_B,
@@ -86,7 +86,7 @@ shinyServer(function(input, output) {
         # a hack to output the inputs, so it can be run from the console
         cat("B:\n")
         cat("run_rot( n_insecticides =",input$n_B,",", 
-            "max_generations =",input$max_generations,",", 
+            "max_gen =",input$max_gen,",", 
             "start_freqs =",input$frequency_B,",", 
             "rotation_interval =",input$rotation_interval_B,",",
             "eff =",input$effectiveness_B,",",

### 2017-11-21 tests/testthat/test_rot_migrate.r

@@ -8,7 +8,7 @@ context("migration")
 # test that tot freqs for sites are same before as after
 # maybe test that sum of whole array before and after is the same
 
-RAF2 <- set_start_freqs(n=3, max_generations=1, freqs=c(0.1,0.2,0.3))
+RAF2 <- set_start_freqs(n=3, max_gen=1, freqs=c(0.1,0.2,0.3))
 # set freqs in refugia to 0 to see effects of migration
 RAF2[,,"refugia",] <- 0
 bef <- RAF2[,,,1]
@@ -24,7 +24,7 @@ sum(bef[1,,]) == sum(aft[1,,])
 
 test_that("migration works for m & f ...", {
   
-  RAF2 <- set_start_freqs(n=3, max_generations=1, freqs=c(0.1,0.2,0.3))
+  RAF2 <- set_start_freqs(n=3, max_gen=1, freqs=c(0.1,0.2,0.3))
   
   # set freqs in refugia to 0 to see effects of migration
   RAF2[,,"refugia",] <- 0

### 2017-11-21 inst/documents/rotations_notes.Rmd
finishing migration but seem to have introduced bug, run_rot() failing

@@
 run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 0 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
+#no cost, some migration
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.05 , cost = 0 , logy = FALSE )
 
@@
 *21/11/17*
-
-# TODO
-
 ~ is it worth creating a get_freq() or getraf() function similar to in rtsetse but not as complex ?
@@
 ~ have an as_df arg to return as a dataframe
+
+# TODO
+
+Fix this :
+Error in filter(., active_or_refuge == "active") : 
+  object 'active_or_refuge' not found
 

### 2017-11-21 R/rot_migrate.r

@@ -1,8 +1,9 @@
 #' migration between intervention and refugia
 #' 
-#' @param RAF array of resistance allele frequencies
-#' @param migrate_intervention migration from intervention
-#' @param migrate_refugia migration from refugia
+#' @param RAF1gen single generation array of resistance allele frequencies
+#' @param migration migration rate between treated & untreated areas 0-1. We assume that immigration=emigration.
+#' @param coverage proportion of mosquitoes that are covered by the intervention (and 1-C is the proportion of the population in the untreated refugia).
+#' @param verbose whether to output before and after migration
 #' 
 #' @examples
 #' #frequencies the same for all insecticides
@@ -25,9 +26,6 @@ rot_migrate <- function( RAF1gen,
                          coverage = 0.6,
                          verbose = FALSE)
 {
-  
-  # TODO I think I should get this to work
-  # on RAF without the generation dimension
   
   # to allow migration to be on a scale from 0-1 (1-coverage is the max)
   # calculating this in rot_migrate function allows easier testing of function

### 2017-11-21 inst/documents/rotations_notes.Rmd
commenting out add_gens_under50 calc because of NSE error

@@
 
+ARG! got migration seeming like it's working
+
+BUT now have come across non-standard-evaluation error in run_rot()
+
+Fix this :
+Error in filter(., active_or_refuge == "active") : 
+  object 'active_or_refuge' not found
+
+Draft that is sposed to help explain :
+http://rpubs.com/lionel-/programming-draft
+and posts in here :
+https://community.rstudio.com/t/should-tidyeval-be-abandoned/2238/37
+
+I could remove the generations under 50 bit for now to try to get run_rot() working again.
+
+removed add_gens_under50 and now run_rot() does run.
+
 # TODO
@@
 ~ https://bookdown.org/rdpeng/RProgDA/non-standard-evaluation.html
+
+# programing with dplyr
+http://dplyr.tidyverse.org/articles/programming.html
 

### 2017-11-21 R/run_rot.r

@@ -36,6 +36,7 @@
 #'                same_insecticides =TRUE, migration=0.01)
 #' 
 #' @import tidyverse 
+#' @importFrom rlang .data #to help with standard evaluatoin of dplyr
 #' @return dataframe of results
 #' @export
 
@@ -64,7 +65,7 @@ run_rot <- function( max_gen = 200, #the maximum number of mosquito generations 
                      cost = 0.1, #c(0,0,0),
                      fitSS = 1,
                      logy = FALSE,
-                     add_gens_under50 = TRUE) 
+                     add_gens_under50 = FALSE) 
   {
   
     
@@ -359,26 +360,30 @@ run_rot <- function( max_gen = 200, #the maximum number of mosquito generations 
   
   # use tidyr::separate() to get from r1_refuge to r1 & refuge in different columns.
   # to get active & refuge into the same subplot
-  df_res2 <- separate(df_res2, region, into=c("resist_gene","active_or_refuge"))
+  df_res2 <- tidyr::separate(df_res2, region, into=c("resist_gene","active_or_refuge"))
   
   # calculate number generations under 50% resistance to be used in plotting 
-  # probably should be somewhere else !
-  # ? just for active area
+  # probably should be somewhere else ! ? just for active area
   # this does give the answer, but only 1 per insecticide
   # so all the results per generation are lost
-  df_res2 <- df_res2 %>%
-    filter(active_or_refuge=='active') %>%
-    group_by(resist_gene) %>%
-    summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
-    ungroup() %>%
-    left_join(df_res2, by='resist_gene')
+  # ARG! struggling with replacing NSE which started causing problems
+  # commented out for now
+  # df_res2 <- df_res2 %>%
+  #   filter(.data$active_or_refuge=='active') %>%
+  #   group_by(.data$resist_gene) %>%
+  #   summarise(gens_under50 = sum(.data$resistance < 0.5, na.rm=TRUE)) %>%
+  #   ungroup() %>%
+  #   left_join(df_res2, by='resist_gene')
   
   # if migration is set to 0 don't show refuge in plots
   plot_refuge <- ifelse(migration==0,FALSE,TRUE)
   
   # do the plots
+  # if (plot) rot_plot_resistance(df_res2, plot_refuge=plot_refuge, 
+  #                               logy=logy, add_gens_under50=add_gens_under50)
[…]

### 2017-11-21 inst/documents/rotations_notes.Rmd
fixing few glitches, tidying
tidy insecticide switching

@@
 
+~ rdpeng suggests not having non standard evaluation in package code :
+If you have any non-standard evaluation in your package code (which you’ll notice because of the “no visible bindings” warnings you’ll get when you check the package), go through and change any instances to use standard evaluation alternatives.
+This change prevents these warnings when you check your package and will also ensure that the functions behave like you expect them to when they are run by other users in their own R sessions.
+~ https://bookdown.org/rdpeng/RProgDA/non-standard-evaluation.html
+
+# programing with dplyr
+http://dplyr.tidyverse.org/articles/programming.html
+
 *21/11/17*
@@
 
+Ians text removed from the insecticide switching portion of code
+
+  #NOTE. At the moment if all but one of the insecticides are over the threshold frequency, the one under the threshold
+  #will be repeatedly deployed. For example, if there are 3 insecticides and rotation is every 10 generations. 
+  #if #3 is being deployed and due to rotate out at generetion 100: if only #3 is under the threshold
+  #it will not rotate out, but will continue to be used util the next scheduled rotation at which point
+    #(a) it will continue to be use if it is the only one below the threshold 
+    #(b) it will be replaced if one of the other insecticides is now below the threshold due to fitness effects or migration
+    #(c) the simulation will terminate if all insecticides exceed the threshold
+
 # TODO
 
-Fix this :
-Error in filter(., active_or_refuge == "active") : 
-  object 'active_or_refuge' not found
+~ take the insecticide switching out into its own function(s)
+
+~ create a new version of run_rot2() allowing me to refactor the code and check that it still generates same results as run_rot()
 
 ~ maybe have a set_raf() function too
-
-~ rdpeng suggests not having non standard evaluation in package code :
-If you have any non-standard evaluation in your package code (which you’ll notice because of the “no visible bindings” warnings you’ll get when you check the package), go through and change any instances to use standard evaluation alternatives.
-This change prevents these warnings when you check your package and will also ensure that the functions behave like you expect them to when they are run by other users in their own R sessions.
-~ https://bookdown.org/rdpeng/RProgDA/non-standard-evaluation.html
-
-# programing with dplyr
-http://dplyr.tidyverse.org/articles/programming.html
 

### 2017-11-21 R/run_rot.r

@@ -272,60 +272,86 @@ run_rot <- function( max_gen = 200, #the maximum number of mosquito generations 
 
 
     
+    ##########################################
+    # checking if insecticide switch is needed
     
-  #do we need to switch current insecticide ?
-    if(rotation_interval==0){  #i.e.  its a RwR policy:
-          if(RAF[current_insecticide, 'f','intervention', gen] > rotation_criterion) change_insecticide=1
-          #message(sprintf("confirm. RAF=%f, change_insecticide=%d \n", RAF[current_insecticide, 'f','intervention', gen], change_insecticide))
-          }  
-    
-  if(rotation_interval!=0){ #i.e. if freq_rotation>0 then its a policy of routine, periodic rotation
-  #NOTE. At the moment if all but one of the insecticides are over the thereshold frequency, the one under the threshold
-  #will be repeatedly deployed. For example, if there are 3 insecticides and rotation is every 10 generations. 
-  #if #3 is being deployed and due to rotate out at generetion 100: if only #3 is under the threshold
-  #it will not rotate out, but will continue to be used util the next scheduled rotation at which point
-    #(a) it will continue to be use if it is the only one below the threshold 
-    #(b) it will be replaced if one of the other insecticides is now below the threshold due to fitness effects or migration
-    #(c) the simulation will terminate if all insecticides exceed the threshold
-  
-  # TODO this keeps the rotation going, if I want to add in responsive that stops the rotation if 
-  # the resistance frequency is reached then I can add it in here with &
-  if (rotation_count!=rotation_interval) rotation_count=rotation_count+1
-  else change_insecticide=1 #i.e. its time to rotate so need to identify the next insecticide in the rotation
-  }
-    
-  if(change_insecticide==1){
-  rotation_count=1; next_insecticide_found=0; candidate=current_insecticide 
+    # if rotate-when resistant
+    if (rotation_interval==0)
+    {  
+      #TOCHECK switch criterion is female only
+      if (RAF[current_insecticide, 'f','intervention', gen] > rotation_criterion)
+      {
+        change_insecticide <- 1        
+      }
+      #message(sprintf("confirm. RAF=%f, change_insecticide=%d \n", RAF[current_insecticide, 'f','intervention', gen], change_insecticide))
+
+    } else if (rotation_interval != 0)
+    # if periodic rotation  
+    {
+      if (rotation_count != rotation_interval) 
+      {
+        # keeps the rotation going
[…]

### 2017-11-22 inst/documents/rotations1.Rmd
rename rotation_ vars to rot_, add sentence to UI about rotate when resistant at interval 0

@@
 #expo_hi=c(1,0.1)
-dfr <- run_rot(n=2, rotation_interval=100, expo_hi=c(1,0.1))
+dfr <- run_rot(n=2, rot_interval=100, expo_hi=c(1,0.1))
 
@@
 # TODO remove hardcode fitnesses from here
-dfr <- run_rot(rotation_interval=50, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.01)
+dfr <- run_rot(rot_interval=50, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.01)
 
@@
 
-dfr <- run_rot(rotation_interval=50, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.001)
+dfr <- run_rot(rot_interval=50, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.001)
 
@@
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="cost 0.1 effectiveness 1"}
-dfr <- run_rot(n=3, rotation_interval=50, eff=1, cost=0.1)
+dfr <- run_rot(n=3, rot_interval=50, eff=1, cost=0.1)
 ```
@@
 
-dfr <- run_rot(rotation_interval=0, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration=0.01)
+dfr <- run_rot(rot_interval=0, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration=0.01)
 
@@
 
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.1 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rot_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.1 )
 
@@
 
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rot_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0 )

### 2017-11-22 inst/documents/rotations_notes.Rmd
rename rotation_ vars to rot_, add sentence to UI about rotate when resistant at interval 0

@@
 implemented set_expsoure_rot() now working
-dfr <- run_rot(rotation_interval=50, expo_hi=0.9)
-dfr <- run_rot(rotation_interval=50, expo_hi=0.2)
+dfr <- run_rot(rot_interval=50, expo_hi=0.9)
+dfr <- run_rot(rot_interval=50, expo_hi=0.2)
 # diff frequency curves for r1 & r2
-dfr <- run_rot(n=2, rotation_interval=50, expo_hi=c(1,0.1))
+dfr <- run_rot(n=2, rot_interval=50, expo_hi=c(1,0.1))
 
@@
 
-dfr <- run_rot(n=2, rotation_interval=50, eff=c(1,0.1))
+dfr <- run_rot(n=2, rot_interval=50, eff=c(1,0.1))
 
@@
 
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rot_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0.1 )
 
@@
 
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rot_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration_rate_intervention = 0.01 , cost = 0 )
 
@@
 max_gen   | 
-rotation_interval | 5 - 50 (integer)
+rot_interval | 5 - 50 (integer)
 male_expo_prop    | 0 - 1
@@
 #high cost, no migration
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0.1 , logy = FALSE )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0.1 , logy = FALSE )
 #no cost, no migration
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
 #no cost and rotate when resistant, no migration
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 0 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 0 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0 , cost = 0 , logy = FALSE )
 #no cost, some migration
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rotation_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.05 , cost = 0 , logy = FALSE )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 10 , eff = 1 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.05 , cost = 0 , logy = FALSE )
 
@@
 
+*22/11/17*
+
+# check with Ian
+~ insecticide switch criterion is female only
+
 # TODO
@@
 ~ take the insecticide switching out into its own function(s)
+insecticide_check()
+insecticide_switch()
 

### 2017-11-22 inst/shiny/rotresist1/server.r

@@ -25,7 +25,7 @@ shinyServer(function(input, output) {
         run_rot(n_insecticides =    input$n_A, 
                 max_gen =   input$max_gen,
                 start_freqs =       input$frequency_A,
-                rotation_interval = input$rotation_interval_A, 
+                rot_interval = input$rot_interval_A, 
                 eff =               input$effectiveness_A,
                 dom_sel =           input$dom_sel_A,
                 dom_cos =           input$dom_cos_A,
@@ -42,7 +42,7 @@ shinyServer(function(input, output) {
         cat("run_rot( n_insecticides =",input$n_A,",", 
                       "max_gen =",input$max_gen,",", 
                       "start_freqs =",input$frequency_A,",", 
-                      "rotation_interval =",input$rotation_interval_A,",",
+                      "rot_interval =",input$rot_interval_A,",",
                       "eff =",input$effectiveness_A,",",
                       "dom_sel =",input$dom_sel_A,",",
                       "dom_cos =",input$dom_cos_A,",",
@@ -70,7 +70,7 @@ shinyServer(function(input, output) {
         run_rot(n_insecticides =    input$n_B, 
                 max_gen =   input$max_gen,
                 start_freqs =       input$frequency_B,
-                rotation_interval = input$rotation_interval_B, 
+                rot_interval = input$rot_interval_B, 
                 eff =               input$effectiveness_B,
                 dom_sel =           input$dom_sel_B,
                 dom_cos =           input$dom_cos_B,
@@ -88,7 +88,7 @@ shinyServer(function(input, output) {
         cat("run_rot( n_insecticides =",input$n_B,",", 
             "max_gen =",input$max_gen,",", 
             "start_freqs =",input$frequency_B,",", 
-            "rotation_interval =",input$rotation_interval_B,",",
+            "rot_interval =",input$rot_interval_B,",",
             "eff =",input$effectiveness_B,",",
             "dom_sel =",input$dom_sel_B,",",
             "dom_cos =",input$dom_cos_B,",",

### 2017-11-22 inst/shiny/rotresist1/ui.r

@@ -52,7 +52,7 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
     column(6, plotOutput('plotA')),
     column(6, plotOutput('plotB'))
     #column(1)    
-  ), #end fixed row
+  ), 
   
   fluidRow(
     column(2, p()),
@@ -63,8 +63,7 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
     column(1, checkboxInput("logy", "log y axis", FALSE)),
     column(2, actionButton('aButtonRunB', 'Run Scenario B')),
     column(2, p())
-  ), #end fixed row
-  
+  ), 
   
   hr(),
   
@@ -91,10 +90,10 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
     #        hr()
     # ),  
     column(1, offset = 0,
-           h5("interval"),
-           sliderInput("rotation_interval_A", NULL, val=50, min = 0, max = 100, step = 10, ticks=FALSE),
+           h5("rotation interval*"),
+           sliderInput("rot_interval_A", NULL, val=50, min = 0, max = 100, step = 10, ticks=FALSE),
            hr(), #hr(),hr(),
-           sliderInput("rotation_interval_B", NULL, val=50, min = 0, max = 100, step = 10, ticks=FALSE),
+           sliderInput("rot_interval_B", NULL, val=50, min = 0, max = 100, step = 10, ticks=FALSE),
            hr()
     ),     
     column(1,
@@ -159,12 +158,14 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
            hr(),
            sliderInput("cost_B", NULL, val=0.1, min = 0, max = 0.2, step = 0.01, ticks=FALSE),
            hr()          
-    )    
-    # column(2, offset = 0,
-    #        h5("Extra params not included in Curtis"),
-    #        numericInput("correct_mix_deploy", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1),
-    #        checkboxInput("addCombinedStrategy", "plot adaptive strategy",FALSE)
-    # )
-  ) #end fixed row
+    )
+  ), #end fluid row
+
+  fluidRow(
+    column(1, NULL),
+    column(11,h6("* setting rotation interval to 0 causes insecticde to be kept in use until resistance threshold is reached (rotate-when-resistant)"))
+  ) #end fluid row
+
[…]

### 2017-11-22 inst/documents/rotations_notes.Rmd
move checking and switching of insecticides into 2 functions

@@
 *22/11/17*
+~ added sentence about interval 0 RwR to the UI
+
+~ removed this commented out alternative to df_results
+  #experimental
+  # df_res_active <- data.frame(generation=rep(1:max_gen,n_insecticides),
+  #                       insecticide=NA,
+  #                       region=NA,
+  #                       resistance=NA, stringsAsFactors = FALSE)
+  # df_res_refuge <- df_res_active
+
+~ I can redesign the insecticide switch function slightly :
+
+to pass :
+current_insecticide
+n_insecticides
+RAF1gen
+df_results : to allow assessment of which insecticides have been used previously
+
+return :
+current_insecticide : if it has failed to find one then return 0
+
+BUT i need to allow ability to assess when insecticides have been used previously so may need to pass df_results too
+
+      rot_count <- 1 
+      next_insecticide_found <- 0
+      candidate <- current_insecticide
+
+      for(temp_int in 1:n_insecticides)
+      {
+        #search through insecticides and go back to start if reach end
+        candidate <- ifelse(candidate==n_insecticides, yes=1, no=candidate+1)
+
+        if (RAF[candidate, 'f','intervention', gen] < rot_criterion)
+        {
+          next_insecticide_found <- 1 
+          current_insecticide <- candidate 
+          change_insecticide <- 0
+
+          message(sprintf("generation %d, switch from insecticide %d to %d; frequencies = %f and %f\n",
+                          gen, current_insecticide, candidate,
+                          RAF[current_insecticide, 'f','intervention', gen], RAF[candidate, 'f','intervention', gen]))
+        }
+
+        if (next_insecticide_found==1) break
+
+      } # end of loop checking each insecticide
 
@@
 
+~ rename rot_count to gens_this_insecticide
+
 ~ create a new version of run_rot2() allowing me to refactor the code and check that it still generates same results as run_rot()
@@
 ~ maybe have a set_raf() function too
+
+~ get an online version up on shinyapps
 

### 2017-11-22 R/insecticide_switch.r

@@ -1,0 +1,57 @@
+#' insecticide_switch choose a new insecticide to switch to
+#'
+#' 
+#' 
+#' @param RAF array of resistance allele frequencies
+#' @param current_insecticide id number of current insecticide
+#' @param n_insecticides number of insecticides (and hence loci)
+#' @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
+#' @param gen generation number
+#' @param df_results results of sim so far
+# @param rot_interval frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
+# @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
+# @param rot_count num generations this insecticide has been used for
+#' 
+# @examples 
+#' 
+#' @return integer : insecticide number to switch to, if 0 none found 
+#' @export
+
+
+insecticide_switch <- function( RAF,
+                                current_insecticide,
+                                n_insecticides,
+                                rot_criterion,
+                                gen,
+                                df_results )
+{
+  next_insecticide_found <- 0
+  candidate <- current_insecticide 
+  
+  #TODO use df_results to stop switching to an insecticide that has been used recently
+  
+  for(temp_int in 1:n_insecticides)
+  {
+    #search through insecticides and go back to start if reach end
+    candidate <- ifelse(candidate==n_insecticides, yes=1, no=candidate+1)
+    
+    if (RAF[candidate, 'f','intervention', gen] < rot_criterion)
+    {
+      message(sprintf("generation %d, switch from insecticide %d to %d; frequencies = %f and %f",
+                      gen, current_insecticide, candidate,
+                      RAF[current_insecticide, 'f','intervention', gen], RAF[candidate, 'f','intervention', gen]))
+      
+      next_insecticide_found <- 1 
+      current_insecticide <- candidate 
+      #change_insecticide <- 0 
+    }
+    
+    if (next_insecticide_found==1) break   
+    
+  } # end of loop checking each insecticide
+ 
+# return 0 if no insecticide to use next found  
+if (next_insecticide_found==0) current_insecticide <- 0 
+   
+return(current_insecticide)  
+}

### 2017-11-22 R/rot_migrate.r

@@ -32,6 +32,11 @@ rot_migrate <- function( RAF1gen,
   migrate_intervention <- migration*(1-coverage)
   migrate_refugia <- migrate_intervention*coverage/(1-coverage)  
   
+  # not needed because now calculated
+  # if(migrate_intervention>(1-coverage)){
+  # message(sprintf("warning from calibration: migration rate in/out of intervenation exceed 1 minus coverage\n"))   
+  # }
+  
   #RAF <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), site=c('intervention','refugia'), gen=1:max_gen)
   
   # RAF[,, ensures calc is repeated for each insecticide and sex 

### 2017-11-22 R/run_rot.r

@@ -1,10 +1,9 @@
 #' run_rot to run simulation of the effect of rotations on the spread of resistance
 #'
 #' can run any number of insecticides/loci 
-#' but at present, input will only allow a maximumum of 5
 #' 
 #' @param max_gen maximum number of mosquito generations to run the simulation
-#' @param n_insecticides number of insecticides (and hence loci), current max is 5
+#' @param n_insecticides number of insecticides (and hence loci)
 #' @param start_freqs starting frequencies of resistance either one per insecticide or same for all
 #' @param rot_interval frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
 #' @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
@@ -32,8 +31,7 @@
 #' run_rot(rot_interval=100)
 #' dfr <- run_rot(rot_interval=50, max_gen = 300)
 #' dfr <- run_rot(rot_interval=0, max_gen = 300)
-#' dfr <- run_rot(rot_interval=0, max_gen = 300, hardcode_fitness = TRUE, 
-#'                same_insecticides =TRUE, migration=0.01)
+#' dfr <- run_rot(rot_interval=0, max_gen = 300, migration=0.01)
 #' 
 #' @import tidyverse 
 #to try help with standard evaluation of dplyr couldn't get to work
@@ -69,8 +67,6 @@ run_rot <- function( max_gen = 200, #the maximum number of mosquito generations 
                      add_gens_under50 = FALSE) 
   {
   
-    
-  exposure <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), amount=c('no','lo', 'hi'))
   
   # setup dataframe to store results, tricky to cope with variable number insecticides
   l_gene_plus_activity <- rep(list(rep(NA,max_gen)), n_insecticides*2) #2 because active & refuge 
@@ -84,27 +80,21 @@ run_rot <- function( max_gen = 200, #the maximum number of mosquito generations 
                                          l_gene_plus_activity))  
   
   
-  #experimental
-  # df_res_active <- data.frame(generation=rep(1:max_gen,n_insecticides),
-  #                       insecticide=NA,
-  #                       region=NA,
-  #                       resistance=NA, stringsAsFactors = FALSE)
-  # df_res_refuge <- df_res_active
-  
-  
-  ### set starting allele frequencies from hardcoded test function or based on other inputs
[…]

### 2017-11-22 inst/documents/rotations_notes.Rmd
add min_rwr_interval

--- 2017-11-22 inst/documents/rotations_notes.Rmd
@@
 *22/11/17*
+~ added sentence about interval 0 RwR to the UI
+
+~ removed this commented out alternative to df_results
+  #experimental
+  # df_res_active <- data.frame(generation=rep(1:max_gen,n_insecticides),
+  #                       insecticide=NA,
+  #                       region=NA,
+  #                       resistance=NA, stringsAsFactors = FALSE)
+  # df_res_refuge <- df_res_active
+
+~ take the insecticide switching out into its own function(s)
+insecticide_check()
+insecticide_switch()
+
+~ rename rot_count to gens_this_insecticide
 
@@
 
-~ take the insecticide switching out into its own function(s)
-insecticide_check()
-insecticide_switch()
+~ add in an option to limit switch back to previous insecticides
+~ could add something to put a lower limit on length of rotations
 

### 2017-11-22 R/insecticide_check.r

@@ -7,7 +7,7 @@
 #' @param rot_interval frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
 #' @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
 #' @param gens_this_insecticide num generations this insecticide has been used for
-#' 
+#' @param min_rwr_interval minimum rotate-when-resistant interval to stop short switches, only used when rot_interval==0. set to 0 to have no effect.
 # @examples 
 #' 
 #' @return TRUE/FALSE whether to change 
@@ -18,7 +18,8 @@ insecticide_check <- function( RAF1gen,
                                current_insecticide,
                                rot_interval, 
                                rot_criterion,
-                               gens_this_insecticide
+                               gens_this_insecticide,
+                               min_rwr_interval
                               ) 
 {
   change_insecticide <- FALSE
@@ -27,7 +28,9 @@ insecticide_check <- function( RAF1gen,
   if (rot_interval == 0)
   {  
     #TODO check with Ian that switch criterion is female only
-    if (RAF1gen[current_insecticide, 'f','intervention'] > rot_criterion)
+    if ( RAF1gen[current_insecticide, 'f','intervention'] > rot_criterion &
+         #to add a min interval to stop lots short switches
+         gens_this_insecticide > min_rwr_interval )
     {
       change_insecticide <- TRUE        
     }

### 2017-11-22 inst/documents/rotations_notes.Rmd
added initial no_r_below_start to stop resistance declining below start vals

@@
 
+~ add in an option to limit switch back to previous insecticides
+~ added something to put a lower limit on length of rotations for rwr
+
 # check with Ian
@@
 
-~ add in an option to limit switch back to previous insecticides
-~ could add something to put a lower limit on length of rotations
+~ add option to stop resistance frequencies going below starting value (the thinking being that they could be maintained at this value by selection).
+It happens in the simulation when there is a cost and the insecticide is not in use.
 

### 2017-11-23 inst/documents/sensi_an_rotations1.Rmd
starting on sensi analysis

---
title: "sensi_an_rotations1 : start of sensitivity analysis for rotations"
output: pdf_document:
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
  library(rotations)
  library(ggplot2)

  outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

Starting question : is there any benefit to rotations when there are no costs and no migration to an untreated area ?

Test 1. No refugia (i.e. Coverage=100%) or fitness costs. Would expect all policies to be equivalent.

Test 2. Dominance of resistance and fitness costs both set to 0.5.
I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles. I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same. I’m probably wrong but it’s a nice test…..

```{r, eval=TRUE}

#start small & build it up
#maybe create an equiv. of resistance::sensiAnPaperPart.r later

# example for 1 run
# linputs <- list(max_gen = 50, n_insecticides = 2)
# do.call(run_rot, linputs)
# dfinputs <- as.data.frame(linputs)

# So one way of doing is to create an linputs list for each run & pass to run_rot()
# so run_rot() only ever runs for 1 scenario
# in contrast resistance::runModel2() accepted an input matrix & would run for multiple scenarios
# not sure which is better

n_scenarios <- 5

df_res_all <- data.frame()

for(i in 1:n_scenarios)
{
  # expo_hi 0.1 – 0.9
  # eff 0.3 - 1
  linputs <- list()

  linputs$expo_hi <-  runif(1, min=0.1, max=0.9)
  linputs$eff <-      runif(1, min=0.3, max=1)

  # run one scenario
  dfres <- do.call(run_rot, linputs)

  # summarise results per scenario
  res <- dfres %>%
  dplyr::filter(active_or_refuge=='active') %>%
  group_by(resist_gene) %>%
  summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
  summarise(mean_gens_under50 = mean(gens_under50)) %>%
  unlist()

  # add results onto inputs list
  linputs <- c(linputs, res)

  # change list to dataframe
  df_in_out <- as.data.frame(linputs)

  # rbind this scenario inputs & results onto previous
  # not a good speed or memory way of doing ... but a start
  df_res_all <- rbind(df_res_all, df_in_out)

}

```

<!-- eval=FALSE to stop slow code being executed after 1st time --> 
```{r, eval=FALSE}

  # this is how it was done for resistance paper1
  # nScenarios <- 10  
  # ## do model runs
  # listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
  # ## save results objects as rda for re-use
  # save(listOutMix, file=paste0(outFolder,'listOutMix_rr_10.rda'))

```

### 2017-11-23 inst/documents/rotations_notes.Rmd
tidying

@@
 
+~ added option to stop resistance frequencies going below starting value (the thinking being that they could be maintained at this value by selection).
+It happens in the simulation when there is a cost and the insecticide is not in use.
+
+~ deployed to shinyapps : working well now
+https://andysouth.shinyapps.io/rotresist1/
+
+*23/11/17*
+~ sent draft model-data gap workshop invite to IVCC
+
+~ tweaked UI input defaults to show a comparison of fixed rotation versus sequence (rwr)
+
+start protocol for sensitivity analysis.
+
+Currently run_rot() outputs :
+'data.frame':	1600 obs. of  5 variables:
+ $ generation      : int  1 2 3 4 5 6 7 8 9 10 ...
+ $ insecticide     : num  NA 1 1 1 1 1 1 1 1 1 ...
+ $ resist_gene     : chr  "insecticide1" "insecticide1" "insecticide1" "insecticide1" ...
+ $ active_or_refuge: chr  "active" "active" "active" "active" ...
+ $ resistance      : num  0.001 0.00141 0.002 0.00283 0.00399 ...
+
+ insecticide : indicates which is in use
+ resist_gene : which gene does this frequency relate to
+
+ from run_rot() now commented out
+
+  # calculate number generations under 50% resistance
+
+  df_res2 <- df_res2 %>%
+    dplyr::filter(active_or_refuge=='active') %>%
+    group_by(resist_gene) %>%
+    summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
+    ungroup() %>%
+    left_join(df_res2, by='resist_gene')
+
+tst <- run_rot()
+
+tst %>%
+dplyr::filter(active_or_refuge=='active') %>%
+group_by(resist_gene) %>%
+summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE))
+
+This gives generations under 50% resistance for each insecticide in the active area
+
+1 insecticide1          134
+2 insecticide2          144
+3 insecticide3          154
+4 insecticide4          164
+
+I could then calculate the mean gens_under50 across all the insecticides.
+
+tst %>%
+dplyr::filter(active_or_refuge=='active') %>%
+group_by(resist_gene) %>%
+summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
+summarise(mean_gens_under50 = mean(gens_under50))
+
+  mean_gens_under50
+              <dbl>
+1               149
+
+Look back at the resistance sensitivity code and see how much of it I can use.
+
+# resistance files to do runs
+1. sensiAnPaper1All.Rmd : code to run whole sensitivity analysis & produce plots 
+1. sensiAnPaperPart.r   : sets up param values and runs for one treatment (e.g. mixture, insecticide1)
+1. setInputOneScenario.r: creates input object for each run
+1. runModel2.r          : runs model for the passed scenarios
+
+I would like to modify setInputOneScenario.r so I'm not referencing inputs by there location in the matrix which seems dangerous, but I might not be able to get around ?
+
+Can I store the inputs in a dataframe with a name column, or with rownames ?
+Or probably it should be a dataframe with a column name for each input, this will make it easy to add multiple runs together.
+
+resistance::runModel2() just accepts an input matrix (i.e. not individually specified params) you need to call setInputOneScenario() first.
+input <- setInputOneScenario()
+listOut <- runModel2(input)
+
+rotations::run_rot() does currently accept individually specified params
+
+I want to make rotations reasonably similar to resistance but make improvements where I can.
+
+I can possibly 
+1. create a set_input_one_scenario() function that returns a dataframe (or list ?) of the inputs.
+1. continue to allow run_rot() to accept individual inputs with an optional list
+
+maybe a diff way of doing too using ... operator, if I have as a list can I pass directly without having to write them all out ?
+
+in resistance this function passes all it's arguments onto another.
+
+resistSimple <- function(...)
+{
+  input <- setInputOneScenario(...)
+
+can I create a list and then pass that to run_rot() (yes using do.call) and also save as a dataframe ?
+This use of do.call given in Hadleys advanced R.
+
+linputs <- list(max_gen = 50, n_insecticides = 2)
+do.call(run_rot, linputs)
+dfinputs <- as.data.frame(linputs)
+
+This is good in that I don't have to specify all the inputs.
+Also easy afterwards to rbind dfinputs together.
+
+Can I make the rotations workflow with fewer files than the resistance one ? :
+rotations workflow :
+1. sensi_an_rotation1.Rmd : run whole sensi analysis including setting input ranges, & analysing outputs
+1. run_rot.r              : a single model run
+
+resistance workflow
+1. sensiAnPaper1All.Rmd : code to run whole sensitivity analysis & produce plots 
+1. sensiAnPaperPart.r   : sets up param values and runs for one treatment (e.g. mixture, insecticide1)
+1. setInputOneScenario.r: creates input object for each run
+1. runModel2.r          : runs model for the passed scenarios
+
 # check with Ian
@@
 
-~ add option to stop resistance frequencies going below starting value (the thinking being that they could be maintained at this value by selection).
-It happens in the simulation when there is a cost and the insecticide is not in use.
+Look back at the resistance sensitivity code and see how much of it I can use.
+
+start thinking about how best to run sensitivty analysis & collect or post-process to get num gens under 50% threshold.
+
+*Ian said first question is to assess whether there is any benefit to rotations when there are no costs and no emigration*
+
+*protocol for sensitivity analysis*
+~ store input ranges
+~ generate inputs for each run
+~ run
+~ store outputs
+~ calculate output summary stats per run
+~ concatenate inputs & outputs
+~ relate outputs to inputs
 
@@
 
-~ get an online version up on shinyapps
-
 ~ sort outputs which need to be time to resistance for planned rotations and RwR rotate when resistant. Sequence.
@@
 
-* rotations understand code better
-* rotations add catch to stop freq declining
-* rotate when resistant add catch to stop switch back to insecticide used < x years ago
+Test 1. No refugia (i.e. Coverage=100%) or fitness costs.
+Would expect all policies to be equivalent.
+
+Test 2. Dominance of resistance and fitness costs both set to 0.5.
+I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles.
+I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same.
+I’m probably wrong but it’s a nice test…..
 
@@
 and that fitness could be used as an index of population, but Ian was sceptical, he said to look up fitness determinants of population on wikipedia, about soft & hard selection.
-
-~ Ian will send me grant app to assess single v multi locus insecticide resistance
-~ Ian said we may want to put something in to stop Frequency going below initial starting value when not in use (because this could represent background balance between benefits & cost)
-
-*Ian said first question is to assess whether there is any benefit to rotations when there are no costs and no emigration*
 

### 2017-11-23 R/run_rot.r

@@ -9,12 +9,7 @@
 #' @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
 #' @param migration migration rate between treated & untreated areas 0-1. We assume that immigration=emigration.
 #' @param coverage proportion of mosquitoes that are covered by the intervention (and 1-C is the proportion of the population in the untreated refugia).
-#' @param plot whether to plot results
 #' @param start_insecticide which insecticide to start with
-#' @param diagnostics whether to output running info
-#' @param hardcode_fitness whether to use hardcoded fitness, default FALSE
-#' @param same_insecticides only used with hardcode_fitness, whether to just set fitnesses for all insecticides the same
-#' @param hardcode_exposure whether to use hardcoded exposure, default FALSE
 #' @param expo_hi exposure to insecticide in hi niche, either single or vector of 1 per insecticide
 #' @param expo_lo exposure to insecticide in lo niche, either single or vector of 1 per insecticide
 #' @param male_expo_prop proportion tht males are exposed relative to f, default 1, likely to be <1 (could possibly be a vector per insecticide)
@@ -24,10 +19,16 @@
 #' @param rr resistance restoration, for all insecticides or individually 
 #' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
 #' @param fitSS fitness of SS if no insecticide, for all insecticides or individually
-#' @param logy whether to use log scale for y axis
-#' @param add_gens_under50 whether to add a label of num generations under 50 pcent resistance
 #' @param min_rwr_interval minimum rotate-when-resistant interval to stop short switches, only used when rot_interval==0. set to 0 to have no effect.
 #' @param no_r_below_start to stop resistance frequencies going below starting values TRUE or FALSE
+#' 
+#' @param plot whether to plot results
+#' @param diagnostics whether to output running info
+# @param hardcode_fitness whether to use hardcoded fitness, default FALSE
+# @param same_insecticides only used with hardcode_fitness, whether to just set fitnesses for all insecticides the same
+# @param hardcode_exposure whether to use hardcoded exposure, default FALSE
+#' @param logy whether to use log scale for y axis
+#' @param add_gens_under50 whether to add a label of num generations under 50 pcent resistance
 #' 
 #' @examples 
[…]

### 2018-01-08 inst/documents/sensi_an_rotations1.Rmd
struggling with dplyr

@@
   # summarise results per scenario
+  # NOTE this is also currently calc at end of run_rot() prob better to calc there
+  # and put in the output dfres, but its then repeated per 
+  # TODO first AG fix this maybe go for non dplyr solution
+  # if I do calc in run_rot then just find value at first row for each gene
+  # Error in dplyr::filter(active_or_refuge == "active") : 
+  # object 'active_or_refuge' not found
   res <- dfres %>%
+  # only assess in control areas not in refugia  
   dplyr::filter(active_or_refuge=='active') %>%
   group_by(resist_gene) %>%
-  summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
-  summarise(mean_gens_under50 = mean(gens_under50)) %>%
+  # for all insecticides in all generations  
+  # summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
+  # summarise(mean_gens_under50 = mean(gens_under50)) %>%
+  # just for deployed insecticides 
+  summarise(gens_dep_under50 = sum(resistance < 0.5 &
+                                   #finds insecticide in use = this one
+                                   #TODO in_use is also calculated in rot_plots, but set to 1.05 for the plots
+                                   resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
+  summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
   unlist()

### 2018-01-19 inst/documents/rotations1.Rmd
working for coverage=1, new documents

@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="exposure set to 1 for r1 and 0.1 for r2"}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="exposure set to 1 for r1 and 0.4 for r2"}
 
 #expo_hi=c(1,0.1)
-dfr <- run_rot(n=2, rot_interval=100, expo_hi=c(1,0.1))
+dfr <- run_rot(n_insecticides=2, rot_interval=100, expo_hi=c(1,0.4))
 
@@
 
-# TODO remove hardcode fitnesses from here
-dfr <- run_rot(rot_interval=50, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.01)
+dfr <- run_rot(rot_interval=50, max_gen = 300, migration=0.01)
 
@@
 
-dfr <- run_rot(rot_interval=50, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.001)
+dfr <- run_rot(rot_interval=50, max_gen = 300, migration=0.001)
 
-```
-
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="cost 0.1 effectiveness 1"}
-dfr <- run_rot(n=3, rot_interval=50, eff=1, cost=0.1)
 ```
@@
 
-dfr <- run_rot(rot_interval=0, max_gen = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration=0.01)
+dfr <- run_rot(rot_interval=0, max_gen = 300, expo_hi=0.5, migration=0.01)
 
@@
 
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rot_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.1 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rot_interval = 50 , eff = 0.6 , dom_sel = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.1, no_r_below_start =FALSE )
 
@@
 
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rot_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0 )
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.1 , rot_interval = 50 , eff = 0.6 , dom_sel = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0 )

### 2018-01-19 inst/documents/rotations_notes.Rmd
working for coverage=1, new documents

@@
 
-I would like to modify setInputOneScenario.r so I'm not referencing inputs by there location in the matrix which seems dangerous, but I might not be able to get around ?
+I would like to modify setInputOneScenario.r so I'm not referencing inputs by their location in the matrix which seems dangerous, but I might not be able to get around ?
 
@@
 
-# check with Ian
-~ insecticide switch criterion is female only
+But may make sense to keep closer to the resistance workflow.
+
+*27/11/17*
+
+~ talked to Ian fine to send out invite for IVCC workshop now
+~ ask Hilary & Martin if they have any other staff that can come
+~ do a doodle poll for dates
+
+~ checked with Ian, fine that insecticide switch criterion is female only
+
+#Q for Ian :
+Is it mean generations under 50 for all insecticides that we want ?
+
+#discussion
+Probably just for the currently active insecticide. Might be called total effectively controlled generations.
+
+but then how do we compare multiple insecticides.
+
+We might also be interested in the other insecticides recovering.
+
+So we might want more than one output.
+
+Output1 : Total well controlled generations.
+For each generation count whether the active insecticide is below the resistance threshold of 50%.
+
+Output2 : (may not be as useful) Mean number of generations below resistance threshold across all insecticides at all times.
+
+Scenarios.
+
+Ian not that keen on my minimum timespan for rotate-when-resistant, because that potentially forces continued use of an insecticide that is ineffective.
+He would rather that something that stops switching back to an insecticide which has been used less than e.g.
+5 generations ago.
+But I asked whether testing for resistance is likely to be tht frequent ?
+If testing is only once per year than changes can perhaps only be made every 10 generations.
+
+We agreed that we could come back to this later, because it only really becomes relevant when there are fairly high costs that being resistance down when insecticides are not in use.
+
+gens_dep_under50 : is the 
+tot_gens_dep_under50 : summed across all insecticides
+
+Good meeting with Rosemary about spray data.
+
+*28/11/17*
+Planning IVCC workshop.
+potential dates :
+8,9,11,12 and 29th January are best.
+
+8th is a monday I'd like to miss.
+
+2 hour or half day ? I'm inclined to 2 hours. More likely to get people.
+
+Sent invite.
+
+started a word doc with outline
+
+Decisions to make before : 
+~ Focus on mix v sequence, or rotations, or both
+
+*29/11/17*
+From WHO World Malaria report 2017 Q&A.
+About insecticide resistance & malria burden :
+http://www.who.int/malaria/media/world-malaria-report-2017-qa/en/
+A large multicountry evaluation coordinated by WHO between 2011 and 2016 showed that insecticides continue to be an effective tool for malaria prevention, even in areas where mosquitoes have developed resistance to pyrethroids.
+Further, there is no clear correlation between insecticide resistance and trends in malaria burden: some countries with resistance to pyrethroids have shown reductions in disease burden; others with less resistance have shown an increase.
+Ultimately, we need more data on the public health impact of insecticide resistance and further investments in this area.
+
+Also IRS coverage gone down probably because switching to more expensive insecticides :
+
+The proportion of the population at risk protected by indoor residual spraying (IRS), an intervention that involves spraying the inside walls of dwellings with insecticides, declined globally from a peak of 5.8% in 2010 to 2.9% in 2016, with decreases seen in all WHO regions.
+The number of people protected in 2010 was 180 million globally, reducing to about 100 million in 2016.
+While there is yet no clear evidence of why this is happening, the decreases in coverage with IRS insecticides are occurring as countries change or rotate the chemicals they are using to try to prevent the spread of mosquitoes resistant to pyrethroids.
+Because alternative insecticides are more expensive than pyrethroids, increased funding for IRS is needed to maintain high coverage levels.
+
+*8/1/2018*
+assessing where I am with rotations
+
+sensi_an_rotations1.Rmd is starting to do what I want.
+
+When I set coverage to 1 and cost to 0 I get this :
+
+run_rot(coverage=1)
+Error in if (RAF[insecticide, sex, site, gen] < start_freqs[insecticide]) RAF[insecticide, : missing value where TRUE/FALSE needed
+
+I suspect because when coverage=1 site='refugia' is not populated.
+
+BEWARE I've added a few changes with major implications in to try to fix this ...
+
+Might want to just put a condition in to not run migration when coverage=1
+
+    # migration between refugia and intervention site
+    RAF[,,,gen] <- rot_migrate(RAF[,,,gen], migration=migration, coverage=coverage)
+
+*9/1/2018*
+
+getting there with model working with no refugia
+this works 
+run_rot(coverage=1)
+
+But knitting sensi_an_rotations1.Rmd
+gives error in RAF1gen[] incorrect number of dimension calls
+from insecticide_check() and I suspect because the site dimension has been dropped because it is just length 1
+
+added drop=FALSE in insecticide_check()
+but still seem to get error
+
+may need to add drop=FALSE at other places in 
+or maybe the build hasn't worked yet
+
+I think working now.
+
+*12/1/2018*
+
+run_rot(coverage=1, rot_interval=0)
+
+Error in RAF1gen[current_insecticide, "f", "intervention"] : 
+  incorrect number of dimensions
+Called from: insecticide_check(RAF1gen = RAF[, , , gen, drop = FALSE]
+
+*19/1/2018*
+
+model is now working with a coverage of 1
+
+From GPIRM :
+Idea that cost of resistance likely to decline when ferquency of resistance gene has been high. (We don't have this in our models).
+Insecticide resistance management strategies must, however, be implemented before the resistance gene becomes common and stable in the population; otherwise, the resistant gene will not recede even if use of the insecticide causing selection pressure is discontinued.
+As demonstrated by a study of blowflies by McKenzie and Whitten in 1982 (3), fitness cost is not an intrinsic property of the gene.
+Therefore, if that gene is allowed sufficient time to become common in a population, the rest of the genome will adapt to incorporate it without a significant fitness cost.
+At this point, even if the selection pressure is removed, the resistance gene will remain in the population.
+
+Example of sequential use and cost from GPIRM
+Example in Colombia: resistance genes disappeared from the vector population when selection pressure was removed.
+A report submitted by WHO AMRO regional entomologists showed that, in 2005–2006, resistance to pyrethroids and DDT was identified in An.
+darlingi.
+A decision was quickly made to change to fenitrothion, an organophosphate with a different mode of action, for IRS.
+Rapid implementation of this alternative, which thereby removed the selection pressure, reduced the
+frequency of resistance.
+In 2010, susceptibility tests showed that the frequency of resistance genes in the vector population had dropped below the level of detection, and pyrethroids were once again introduced into the IRS programme, albeit on a more limited scale.
+
+Working on rotations_story.Rmd to create visuals about rotations we can use in upcoming talks.
 
@@
 
-Look back at the resistance sensitivity code and see how much of it I can use.
+~ get sensi_an_rotations1.Rmd working for Ians Q1
 
-start thinking about how best to run sensitivty analysis & collect or post-process to get num gens under 50% threshold.
+gens_dep_under50 is ugly and I need to rewrite in sensi_an_rotations !
+
+BEWARE active in active_or_refuge is potentially confusing, does not mean that insecticide is active/deployed, just means its is the active area (a different insecticide may be deployed)
 
@@
 # state of play
+~ sensi_an_rotations1.Rmd : starting on sensitivity analysis
+~~  produces an output dataframe containing both inputs and outputs
+
 ~ rotations1.Rmd : example plots
 ~ UI : rotations/shiny/rotresist1/server.r
+      https://andysouth.shinyapps.io/rotresist1/
+~ RAF[,,,2] to get allele frequencies for generation 2 (by insecticide,sex and active|refuge)
 
-~ RAF[,,,2] to get allele frequencies for generation 2 (by insecticide,sex and active|refuge)
+~ c 15 generations per year for mosquitos, 25 day generation time.
+~ 3 years = 45 generations
+~ https://www.cdc.gov/malaria/about/biology/mosquitoes/
+~
 

### 2018-01-19 inst/documents/rotations_story.Rmd
working for coverage=1, new documents

---
title: "Rotations and Insecticide Resistance the story"
output:
  pdf_document:
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex

#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r, eval=TRUE, include=FALSE}

library(knitr)
#https://stackoverflow.com/a/36234023/1718356 for fig.pos = 'h'
opts_chunk$set(fig.height=3, fig.width=7, fig.pos = 'h', echo=FALSE, warning=FALSE, message=FALSE)
# opts_chunk$set(dev="tiff",
#                dev.args=list(compression="lzw"),
#                dpi=300,
#                cache=FALSE,
#                fig.path='pap2figs/')
```

```{r, results='hide'}
  library(rotations)
  library(ggplot2)
```

In **sequential use** an insecticide is changed when it reaches a resistance threshold, in contrast in a **rotation** an insecticide is changed before it reaches a resistant threshold.

If there are no costs of resistance or migration from untreated areas then rotations and sequences are likely to result in a similar effect on resistance frequencies as seen in the following plots.

*Question to myself* why do rotations last c 5 generations longer in all comparisons ?

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=2.5, fig.width=7, fig.align='left', fig.cap="A sequence, insecticide changed when resistance threshold reached."}

dfr <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0, coverage=1)

```

```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A rotation, insecticide changed every 15 generations (c 1 year)."}

dfr <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0, coverage=1)

```

```{r, eval=FALSE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A rotation, insecticide changed every 45 generations (c 3 years)."}
#eval=FALSE
dfr <- run_rot(n_insecticides=2, rot_interval=45, expo_hi=0.5, eff=0.5, cost=0, coverage=1)

```

\newpage
If there are costs of resistance then we would expect resistance frequencies to decline when an insecticide is not in use. However this may not make much difference between the effect of a sequence and a rotation. In a sequence the resistance frequency builds up once and then gets a long period to decline when not in use. In a rotation resistance frequency increases as a series of small steps with short declines in between (see following plots).

(This is assuming that costs don't change according to the frequency of the resistance allele - there is some evidence to suggest that costs may decline when a resistance allele has been at a high frequency for a long time).

```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with resistance cost of 0.01."}

dfr <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0.01, coverage=1)

```

```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with resistance cost of 0.01."}

dfr <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0.01, coverage=1)

```

\newpage
Even without costs of resistance we might expect resistance frequencies to decline when an insecticide is not in use due to immigration from nearby areas that were not treated originally. This is shown in the following two plots where a neighbouring untreated area and exchange of migrants is represented.

```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with dispersal between treated and untreated areas (no resistance costs)."}

dfr <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0, coverage=0.8, migration = 0.01)

```

```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with with dispersal between treated and untreated areas (no resistance costs)."}

dfr <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0, coverage=0.8, migration = 0.01)

```

### 2018-01-19 inst/documents/sensi_an_rotations1.Rmd
working for coverage=1, new documents

@@
 ---
-title: "sensi_an_rotations1 : start of sensitivity analysis for rotations"
-output: pdf_document:
+title: "sensi_an_rotations1. Start of sensitivity analysis for rotations."
+output: pdf_document
 #output: word_document
@@
 
-```{r, eval=TRUE}
+Testing sensitvity analysis to answer question 1.
+Are sequences and rotations equivalent when coverage 100% and no fitness costs.
+On each page should see a run with 1 set of inputs first for sequence then for rotation.
+
+\pagebreak
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=4}
+#results='hide',
 
@@
 {
-  # expo_hi 0.1 – 0.9
-  # eff 0.3 - 1
+  # inputs (could also be saved to a dataframe first & extracted one row at a time)
   linputs <- list()
-
+  # constant inputs
+  # for Ians Q1 set coverage to 1 & cost to 0 
+  linputs$coverage <- 1
+  linputs$cost <- 0  
+  # variable inputs
   linputs$expo_hi <-  runif(1, min=0.1, max=0.9)
@@
 
+  # because I want to compare sequence rot_when_r to rotation
+  # I could run all the scenarios above for both seq & rot
+  # (and filter out later)
+
+  for(rot_or_not in 0:1)
+  {  
+    #TODO look up about generating random integers in R
+
+    #beware currently interval generated one more time than it is needed
+    #random rot interval
+    linputs$rot_interval <- round( runif(1,min=5,max=50), 0 )
+
+    #multiply by rot_or_not to get rot_int of 0 every other run
+    linputs$rot_interval <- linputs$rot_interval * rot_or_not
+
+    #print scenario num and inputs to doc for exploration
+    cat("scenario",i,
+        " expo_hi",linputs$expo_hi,
+        " eff",linputs$eff,
+        " rot_interval",linputs$rot_interval,
+        "\n")
+
+    ##################
     # run one scenario
@@
 
-  # add results onto inputs list
-  linputs <- c(linputs, res)
+    #print results in doc for diagnostics
+    cat("tot gens deployed under freq 0.5 =", res,"\n\n")
 
-  # change list to dataframe
-  df_in_out <- as.data.frame(linputs)
+    # create dataframe from inputs and outputs
+    df_in_out <- as.data.frame(c(linputs, res))
 
@@
 
-}
+  } #end rot_or_not loop
+} #end scenario loop
+
+# save object containing inputs and results as rda for analysis
+save(df_res_all, file='listOutMix_rr_10.rda') #paste0(outFolder,'*.rda'))
 
 ```
-
-<!-- eval=FALSE to stop slow code being executed after 1st time --> 
-```{r, eval=FALSE}
-
-  # this is how it was done for resistance paper1
-  # nScenarios <- 10  
-  # ## do model runs
-  # listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
-  # ## save results objects as rda for re-use
-  # save(listOutMix, file=paste0(outFolder,'listOutMix_rr_10.rda'))
-
-```

### 2018-01-19 R/insecticide_check.r

@@ -31,7 +31,9 @@ insecticide_check <- function( RAF1gen,
     
     #TODO later alternative stop switching back to insecticide that has been used within 5 generations
     
-    if ( RAF1gen[current_insecticide, 'f','intervention'] > rot_criterion &
+    #if ( RAF1gen[current_insecticide, 'f','intervention'] > rot_criterion &
+    # 12/1/18 because have drop=false to preserve site dimension the gen dimension is preserved too
+    if ( RAF1gen[current_insecticide, 'f','intervention',1] > rot_criterion &
          #to add a min interval to stop lots short switches
          gens_this_insecticide > min_rwr_interval )
     {

### 2018-01-19 R/run_rot.r

@@ -91,7 +91,9 @@ run_rot <- function(max_gen = 200,
   #todo add checks thats start_freqs is either length 1 or n_insecticides
   RAF <- set_start_freqs( n_insecticides=n_insecticides, 
                           max_gen=max_gen, 
-                          freqs = start_freqs )
+                          freqs = start_freqs,
+                          coverage = coverage )
+  
   #old hardcoded test function
   #RAF <- set_start_freqs_test( n_insecticides=n_insecticides, max_gen=max_gen )    
     
@@ -226,42 +228,51 @@ run_rot <- function(max_gen = 200,
       ######  
       # refugia
       
-      # RS coefficient common to equations 2 and 3
-      temp_coeff <- (RAF[insecticide, 'm', 'refugia',gen-1]*(1-RAF[insecticide, 'f', 'refugia',gen-1])+
-                     RAF[insecticide, 'f', 'refugia',gen-1]*(1-RAF[insecticide, 'm', 'refugia',gen-1]))*
-                     0.5*fitness[insecticide, 'RS', 'no']
+      # andy adding a condition, refugia not needed if coverage=1
+      if ( coverage < 1 )
+      {
+        # RS coefficient common to equations 2 and 3
+        temp_coeff <- (RAF[insecticide, 'm', 'refugia',gen-1]*(1-RAF[insecticide, 'f', 'refugia',gen-1])+
+                         RAF[insecticide, 'f', 'refugia',gen-1]*(1-RAF[insecticide, 'm', 'refugia',gen-1]))*
+          0.5*fitness[insecticide, 'RS', 'no']
+        
+        # male RR   
+        F_male_r_refugia <- RAF[insecticide, 'm', 'refugia',gen-1]*
+          RAF[insecticide, 'f', 'refugia',gen-1]*
+          fitness[insecticide, 'RR', 'no']+temp_coeff
+        # male SS
+        F_male_s_refugia  <- (1-RAF[insecticide, 'm', 'refugia',gen-1])*
+          (1-RAF[insecticide, 'f', 'refugia',gen-1])*
+          fitness[insecticide, 'SS', 'no']+temp_coeff 
+        
+        #normalise and store results
+        norm_coeff <- F_male_r_refugia + F_male_s_refugia
+        RAF[insecticide, 'm', 'refugia', gen] <- F_male_r_refugia/norm_coeff
+        
+        # no insecticides in use so same frequencies for both sexes
[…]

### 2018-01-26 inst/documents/rotations_notes.Rmd
improve vis of insecticide use

@@
 From GPIRM :
-Idea that cost of resistance likely to decline when ferquency of resistance gene has been high. (We don't have this in our models).
+Idea that cost of resistance likely to decline when frequency of resistance gene has been high. (We don't have this in our models).
 Insecticide resistance management strategies must, however, be implemented before the resistance gene becomes common and stable in the population; otherwise, the resistant gene will not recede even if use of the insecticide causing selection pressure is discontinued.
@@
 At this point, even if the selection pressure is removed, the resistance gene will remain in the population.
+
+But I looked at the 1982 paper and not sure it supports that ?
+Seems to me that McKenzie1982 shows that relative fitness of the susceptible genotypes increase over time as would be expected by a decline in the concentration of the active ingredient ?
+(and similar to our WoS paper)
 
@@
 
+*22/1/2018*
+
+Why do rotations last longer than sequences ? (even when coverage=1 & cost=0)
+
+I think it's just because the rotation interval continues to its end.
+
+e.g. added an option to plot a 2nd scenario to rot_plot_resistance()
+
+dfrot0 <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0, coverage=1)
+dfrot15 <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0, coverage=1)
+rot_plot_resistance(dfrot0, df_resanother=dfrot15, plot_refuge=FALSE)
+
+So maybe we should have assessment of the resistance frequency in the rotation strategy too ?
+I could add it as a flag.
+
+done, now works for simple scenario
+
+BUT rotations still lasting a few generations longer in later scenarios ?
+dfseq <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0.01, coverage=1)
+dfrot <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0.01, coverage=1)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)
+
+rotation
+insecticide2 192 2 active 0.46672149
+insecticide2 193 2 active 0.48524932
+insecticide2 194 NA active 0.50379224
+
+sequence
+insecticide1 187 0.4710576
+insecticide1 188 0.4895936
+insecticide1 189 0.5081326
+
+Could it be because an extra generation is slipped in at each rotation switchover ?
+Or might be that more decline happens in rotations
+
+Actually I think it's due to min_rwr_interval which was set at 5 by default and in the current code is implemented for sequence but not for rotation.
+
+hmmm but even when I set min_rwr_interval to 1 there seems to be the difference ?
+is there any info in the console messages ?
+
+###rotation :
+generation 166, switch from insecticide 1 to 2; frequencies = 0.402193 and 0.178446
+generation 181, switch from insecticide 2 to 1; frequencies = 0.402193 and 0.384221
+generation 188, switch from insecticide 1 to 2; frequencies = 0.512362 and 0.393770
+
+simulation terminating at generation 194 because all RAFs above threshold of 0.500000
+
+frequency of resistance in females to insecticide 1 is 0.504826
+frequency of resistance in females to insecticide 2 is 0.503792
+
+###sequence :
+generation 92, switch from insecticide 1 to 2; frequencies = 0.510798 and 0.001000
+generation 183, switch from insecticide 2 to 1; frequencies = 0.510798 and 0.397957
+
+simulation terminating at generation 189 because all RAFs above threshold of 0.500000
+
+frequency of resistance in females to insecticide 1 is 0.508133
+frequency of resistance in females to insecticide 2 is 0.503261
+
+Hi Ian,
+
+I'm looking into rotations and mostly I find very little difference from sequences even with costs or refuges.
+In a sequence resistance builds up once and then declines, in rotations it builds up in steps and declines in between but the time-to-resistance is the same for both.
+
+Then I refound this idea in GPIRM that costs of resistance are likely to decline when frequency of resistance gene has been high.
+
+"Insecticide resistance management strategies must, however, be implemented before the resistance gene becomes common and stable in the population; otherwise, the resistant gene will not recede even if use of the insecticide causing selection pressure is discontinued.
+As demonstrated by a study of blowflies by McKenzie and Whitten in 1982 (3), fitness cost is not an intrinsic property of the gene.
+Therefore, if that gene is allowed sufficient time to become common in a population, the rest of the genome will adapt to incorporate it without a significant fitness cost.
+At this point, even if the selection pressure is removed, the resistance gene will remain in the population."
+
+If this was the case it would probably favour rotations over sequences.
+
+Do you have any thoughts ? Do you think it likely that costs will decline over time at high resistance frequencies ?
+
+I looked at the 1982 paper (attached) and I don't think it supports the idea that costs decline over time (but it does seem to have data on relative fitness of RR,RS,SS that we could use in the WoS paper, gives evidence for dominance changing over time and could be used to derive our model parameters.)
+
+Andy
+
+Thanks Andy
+
+That results should put the cat amongst the pigeons i.e. little difference between sequences and rotations.
+
+There are examples of “compensatory” mutations offsetting the fitness cost of a mutation.
+Often from bacteria where populations can be very high and the population is asexual so that the mutation and modifier are always linked.
+It is less well studied in larger animals.
+There may be examples from IR but I am unaware.
+Often it is a post hoc explanation plucked out the air to explain why there is no apparent fitness effect i.e.
+“there were fitness effect, but they must have been modified”.
+
+I like the 1982 paper… I have put it in the WoS folder with a note to discuss it.
+
+*25/1/2018*
+
+noticed that rotation plots don't have a log y axis where resistance plots do.
+There is a logy option in run_rot() should probably set that TRUE by default.
+yes fixed
+
+Rowland(1981) has some good stuff on costs of resistance and selection coefficients and putting these into a model and testing against field data :
+
+Withdrawal of dieldrin in two regions of India led to the frequency of SS increasing from 0.31 to 0.62 within 7 months (Rao et al., 1960) and from 0.11 to 0.85 within 9 months (Bhatia & Deobhankar, 1963).
+In Fig.
+9 the computed reversions in An.garnbiae and Anstephemi have been redrawn, expressed in terms of SS rather than R frequency.
+Superimposed on these are the An.
+culicifacies field reversions.
+Assuming that one generation takes 1 month (cf.
+Curtis et al., 1978), the computer
+predictions seem to be consistent with the field reversions.
+
+Having established why and to what extent resistance should revert when cyclodiene usage is ended, the question remains could yHCH be effectively re-introduced at a later date as part of a insecticide-rotation resistance management strategy?
+Because RR is the most deleterious genotype, the rate of reversion should be rapid initially.
+Only when R is predominantly carried by heterozygotes would the rate of reversion begin to slow down.
+For this reason resistance cannot be expected to revert completely, but it should nevertheless revert sufficiently to warrant the useful re-introduction of yHCH for several generations at a time.
+An unresolved question, potentially damagingto the long-term future of the rotation strategy,is would modifiers of fitness evolve if cyclodieneusage was prolonged?
+
+It seems, then, that rotation strategies might prolong dieldrin efficacy indefinitely, not just against anophelines but against many species of insect.
+
+[because the combination of selection coefficient, cost and dominance means that resistance doesn't rise too fast when used and declines when not in use].
+
+Rowland2013 data on hut trials of actellic - pirimphos-methyl - an OP.
+Includes monthly change and at diff concentrations so could be useful for WoS, but doesn't have resistant & susceptible (does have resistance frequencies).
+
+Matowo2015 selection of insecticide resistance, seems to show changing insecticide resistance in response to interventions. but only shows mortality change not resistance frequency change.
+
+Kim2016 interesting risk paper :
+The main policy implication of this analysis is that evolutionary speed of insecticide resistance and
+maximal larviciding efficacy strongly affect the preferred actions implied by our decision analysis tool and that these parameters jointly offer the largest potential benefits from reduced uncertainty
+
+*26/1/2018*
+
+what to call new UI resistrot consistent with resistmixseq
+
 # TODO
 
+Explore idea from Rowland (1981) that resistance declines faster when it is mostly held by the RR (i.e. at higher allele frequencies)
+
+Look into data on changing allele frequencies over time (e.g. for Kdr)...
+
+It occurred to me that maybe a benefit of rotations is that if it is difficult to measure resistance allele frequencies they provide a lower risk of reaching resistance thresholds for any insecticide.
+i.e.
+are they a better strategy in light of uncertainty in resistance allele frequencies ?
+
+~ check again on current small differences between rotations & sequences
+
+~ maybe I can write a function to plot 2 scenarios on the same graph (in the same way that curtis fig2 plots mix & seq on same graph) ?
+~ I could have the insecticide use still along the top but with diff colours showing the two alternatives
+
 ~ get sensi_an_rotations1.Rmd working for Ians Q1
+
+~ can we get cost to decrease over time at high frequency as suggested in GPIRM ?
 

### 2018-01-26 inst/documents/rotations_story.Rmd
improve vis of insecticide use

@@
 
-If there are no costs of resistance or migration from untreated areas then rotations and sequences are likely to result in a similar effect on resistance frequencies as seen in the following plots.
+If there are no costs of resistance or migration from untreated areas then rotations and sequences are likely to result in very similar effect on resistance frequencies as seen in the following plots.
 
-*Question to myself* why do rotations last c 5 generations longer in all comparisons ?
+These plots assume that an insecticide is changed when it reaches the resistance frequency threshold even if the rotation interval has yet to be reached.
+Without this assumption then rotations always take a little longer time than sequences but at the expense of reaching slightly higher resistance frequencies.
+
+The plots here are for 2 insecticides to keep things simple however the model can be run for any number of insecticides. Effectiveness of insecticides at killing SS is set to 0.8 unless indicated.
 
@@
 
-dfr <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0, coverage=1)
+dfrot0 <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0, coverage=1)
 
@@
 
-```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A rotation, insecticide changed every 15 generations (c 1 year)."}
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A rotation, insecticide changed every 10 generations (c 1 year)."}
 
-dfr <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0, coverage=1)
+dfrot15 <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0, coverage=1)
 
@@
 
-```{r, eval=FALSE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A rotation, insecticide changed every 45 generations (c 3 years)."}
+```{r, eval=FALSE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A rotation, insecticide changed every 30 generations (c 3 years)."}
 #eval=FALSE
-dfr <- run_rot(n_insecticides=2, rot_interval=45, expo_hi=0.5, eff=0.5, cost=0, coverage=1)
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=30, expo_hi=0.5, eff=0.8, cost=0, coverage=1)
 
@@
 \newpage
+###costs of resistance
+
 If there are costs of resistance then we would expect resistance frequencies to decline when an insecticide is not in use.
@@
 In a rotation resistance frequency increases as a series of small steps with short declines in between (see following plots).
+The time it takes for both insecticides to reach the resistance threshold of 50% is still very similar for a sequence and a rotation.
 
-(This is assuming that costs don't change according to the frequency of the resistance allele - there is some evidence to suggest that costs may decline when a resistance allele has been at a high frequency for a long time).
+(This is assuming that costs don't change according to the frequency of the resistance allele - there is some evidence to suggest that costs may decline when a resistance allele has been at a high frequency for a long time (mentioned in GPIRM), if this is the case then it could indeed provide an advantage for rotations over sequences).
 
@@
 
-dfr <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0.01, coverage=1)
+dfseq <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.01, coverage=1)
 
@@
 
-dfr <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0.01, coverage=1)
+dfrot <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.01, coverage=1)
+
+#to compare 2 previous scenarios
+#rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)
 
@@
 \newpage
-Even without costs of resistance we might expect resistance frequencies to decline when an insecticide is not in use due to immigration from nearby areas that were not treated originally.
-This is shown in the following two plots where a neighbouring untreated area and exchange of migrants is represented.
+###costs of resistance
 
-```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with dispersal between treated and untreated areas (no resistance costs)."}
+Costs increased to 5%.
 
-dfr <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0, coverage=0.8, migration = 0.01)
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with resistance cost of 0.05."}
+
+dfseq <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.05, coverage=1)
 
@@
 
-```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with with dispersal between treated and untreated areas (no resistance costs)."}
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with resistance cost of 0.05."}
 
-dfr <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0, coverage=0.8, migration = 0.01)
+dfrot <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.05, coverage=1)
+
+#to compare 2 previous scenarios
+#rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)
 
 ```
+
+\newpage
+###immigration from untreated areas
+
+Even without costs of resistance we might expect resistance frequencies to decline when an insecticide is not in use due to immigration from nearby areas that were not treated originally.
+This is shown in the following two plots where a neighbouring untreated area ('refuge') and exchange of migrants is represented.
+In this example the treated and untreated areas are set to the same size and migration at 1%.
+The time it takes for both insecticides to reach the resistance threshold of 50% is still very similar for a sequence and a rotation.
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with 1 percent dispersal between treated and untreated areas of equal size (no resistance costs)."}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0, coverage=0.5, migration = 0.01)
+
+```
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with 1 percent dispersal between treated and untreated areas of equal size (no resistance costs)."}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0, coverage=0.5, migration = 0.01)
+
+```
+
+\newpage
+###costs and immigration
+
+Including both costs of resistance and immigration from an untreated area, still results in little difference between a sequence and rotation in terms of time-to-resistance to both insecticides.
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with dispersal between treated and untreated areas and resistance cost of 0.005."}
+
+#upped max_gen because migration and cost decrease rise in resistance
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.005, coverage=0.5, migration = 0.01)
+
+```
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with with dispersal between treated and untreated areas and resistance cost of 0.005."}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=0.5, migration = 0.01)
+
+```
+
+\newpage
+###dominance
+
+In all the preceeding plots dominance of resistance and dominance of cost were set to 0.5.
+Here we check whether changing the dominance of cost between 0 and 1 has an effect on rotations.
+It does not appear to change time to the 50% resistance threshold, on the following page we try increasing the threshold.
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with a resistance cost of 0.005 and dominance of cost set to 0 (recessive)."}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 0)
+
+```
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with a resistance cost of 0.005 and dominance of cost set to 1 (dominant)."}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 1)
+
+```
+
+\newpage
+###dominance of cost effects with increased resistance threshold (90%)
+
+Still little difference between sequence and rotation.
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="dominance of cost 0"}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 0, rot_criterion = 0.9, start_freqs = 0.001)
+
+```
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="dominance of cost 0."}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 0, rot_criterion = 0.9, start_freqs = 0.001)
+
+```
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="dominance of cost 1."}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 1, rot_criterion = 0.9, start_freqs = 0.001)
+
+```
+
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="dominance of cost 1."}
+
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 1, rot_criterion = 0.9, start_freqs = 0.001)
+
+```

### 2018-01-26 R/insecticide_check.r

@@ -8,6 +8,8 @@
 #' @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
 #' @param gens_this_insecticide num generations this insecticide has been used for
 #' @param min_rwr_interval minimum rotate-when-resistant interval to stop short switches, only used when rot_interval==0. set to 0 to have no effect.
+#' @param exit_rot whether to exit rotation interval if rot_criterion is reached
+
 # @examples 
 #' 
 #' @return TRUE/FALSE whether to change 
@@ -19,15 +21,16 @@ insecticide_check <- function( RAF1gen,
                                rot_interval, 
                                rot_criterion,
                                gens_this_insecticide,
-                               min_rwr_interval
+                               min_rwr_interval,
+                               exit_rot
                               ) 
 {
   change_insecticide <- FALSE
   
-  # if rotate-when resistant
-  if (rot_interval == 0)
+  # if rotate-when resistant OR exit_rot==TRUE so rotation can be exited
+  if (rot_interval == 0 | exit_rot == TRUE)
   {  
-    #TODO check with Ian that switch criterion is female only
+    #BEWARE that switch criterion is female only
     
     #TODO later alternative stop switching back to insecticide that has been used within 5 generations
     
@@ -40,9 +43,11 @@ insecticide_check <- function( RAF1gen,
       change_insecticide <- TRUE        
     }
     #message(sprintf("confirm. RAF=%f, change_insecticide=%d \n", RAF1gen[current_insecticide, 'f','intervention'], change_insecticide))
-    
-  } else if (rot_interval != 0)
-    # if periodic rotation  
+  } 
+
+  # even if in a rotation it my have assessed in previous loop that insecticide needs to be changed  
+  if (rot_interval != 0)
+  # if a rotation  
   {
     if (gens_this_insecticide != rot_interval) 
     {

### 2018-01-26 R/rot_plots.r

@@ -4,6 +4,7 @@
 #' @param plot_refuge whether to plot refuge as well as intervention
 #' @param logy whether to use log scale for y axis
 #' @param add_gens_under50 whether to add a label of num generations under 50percent resistance
+#' @param df_resanother exploratory option to plot results of another scenario on the same graph
 #'
 # check said that namespace dependencies not required
 # @import ggplot2
@@ -15,10 +16,13 @@
 #' df_res2 <- run_rot()
 #' rot_plot_resistance(df_res2)
 #' 
+
+#' 
 rot_plot_resistance <- function(df_res2,
                                 plot_refuge = TRUE,
-                                logy = FALSE,
-                                add_gens_under50 = TRUE) {
+                                logy = TRUE,
+                                add_gens_under50 = TRUE,
+                                df_resanother = NULL) {
   
   # column names of input dataframe
   # "generation"  "insecticide"     "resist_gene"  "active_or_refuge" "resistance"
@@ -29,27 +33,35 @@ rot_plot_resistance <- function(df_res2,
   
   # to allow plotting of insecticide in use
   # add column which has a value if insecticide in use & NA if not
-  # the value (currently 1.05) determines where the line appears on the y axis
+  # the value determines where the line appears on the y axis
+  to_plot_in_use <- 1
+  
   df_res2 <- df_res2 %>%
-    #mutate( i_in_use = ifelse(insecticide==1,1,NA))
     #mutate( i_in_use = ifelse(stringr::str_detect(region,paste0(insecticide,"_active")),1.05,NA))    
     #now that active & refuge on same plot
-    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide),1.05,NA))    
+    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide), to_plot_in_use, NA))    
   
-  # only colour by active_or_refuge if their is a refuge
+  # only colour by active_or_refuge if there is a refuge
   if (plot_refuge) {
     gg <- ggplot( df_res2, aes_string(x='generation',y='resistance',colour='active_or_refuge') ) +
-    geom_line( alpha=0.5, lwd=1.5 )
+    geom_line( alpha=0.5, lwd=1.5 ) +
+    #legend for the lines, allows me to set title & labels  
+    scale_colour_manual("areas connected\nby migration",values=c("red3","navy"), labels=c("treated","untreated"))
   }   else
   {
[…]

### 2018-02-07 inst/documents/rotations_notes.Rmd
to pass check plus doc and add warning to shiny

@@
 *26/1/2018*
+new UI resistrot, similar name to resistmixseq
 
-what to call new UI resistrot consistent with resistmixseq
+*1/2/2018*
+
+I want to create a figure that shows effect of untreated area, refugia
+Can I do this in ggplot2, or do it in powerpoint ?
+see : resistance-dispersal-dilution-figure.pptx
+
+How to create a greater effect of dispersal, reduce the coverage.
+
+run_rot(n_insecticides = 1)
+Error in RAF1gen[, , "intervention"] : incorrect number of dimensions
+Called from: rot_migrate(RAF[, , , gen], migration = migration, coverage = coverage)
+
+but this is OK
+run_rot(n_insecticides = 1, coverage=1)
+
+I suspect problem is with dropping dimensions in rot_migrate when there's just one insecticide
+For now just force n_insecticides > 1 from the UI
 
@@
 
-Explore idea from Rowland (1981) that resistance declines faster when it is mostly held by the RR (i.e. at higher allele frequencies)
+~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ?
+
+~ something not right, when run with coverage=0 still get increase in resistance in treated areas, even though technically there is no treated area, maybe I just need to force coverage to be > 0 :
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 0 , eff = 0.8 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0 , migration = 0.01 , cost = 0.01 , logy = TRUE no_r_below_start = TRUE min_rwr_interval = 5 )
+
+~ add error-check to run_rot if n.insecticides=1 & migration > 0
+
+Maybe explore idea from Rowland (1981) that resistance declines faster when it is mostly held by the RR (i.e. at higher allele frequencies)
 

### 2018-02-07 inst/documents/rotations_story.Rmd
to pass check plus doc and add warning to shiny

@@
 #https://stackoverflow.com/a/36234023/1718356 for fig.pos = 'h'
+
 opts_chunk$set(fig.height=3, fig.width=7, fig.pos = 'h', echo=FALSE, warning=FALSE, message=FALSE)
-# opts_chunk$set(dev="tiff",
-#                dev.args=list(compression="lzw"),
-#                dpi=300,
-#                cache=FALSE,
-#                fig.path='pap2figs/')
+
+# #these options make the pdf fail but do create the tiffs
+opts_chunk$set(fig.height=3, fig.width=7, fig.pos = 'h', echo=FALSE, warning=FALSE, message=FALSE,
+               dev="tiff",
+               dev.args=list(compression="lzw"),
+               dpi=300,
+               cache=FALSE,
+               fig.path='figs/')
 ```
@@
 
-```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with 1 percent dispersal between treated and untreated areas of equal size (no resistance costs)."}
+```{r, eval=TRUE, fig.height=2.5, fig.width=5, fig.show='hold', fig.align='left', fig.cap="A sequence with 2 percent dispersal between treated and untreated areas of equal size (no resistance costs)."}
 
-dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0, coverage=0.5, migration = 0.01)
+dfr <- run_rot(n_insecticides=2, max_gen = 130, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0, coverage=0.5, migration = 0.02)
 
@@
 
-```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with 1 percent dispersal between treated and untreated areas of equal size (no resistance costs)."}
+```{r, eval=TRUE, fig.height=2.5, fig.width=5, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with 2 percent dispersal between treated and untreated areas of equal size (no resistance costs)."}
 
-dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0, coverage=0.5, migration = 0.01)
+dfr <- run_rot(n_insecticides=2, max_gen = 130, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0, coverage=0.5, migration = 0.02)
+
+```
+
+\newpage
+### more insecticides
+
+```{r, eval=TRUE, fig.height=5, fig.width=5, fig.show='hold', fig.align='left', fig.cap="A sequence with 10 percent coverage and 5 perecnt dispersal."}
+
+dfr <- run_rot(n_insecticides=4, max_gen = 130, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0, coverage=0.1, migration = 0.05)
+
+```
+
+```{r, eval=TRUE, fig.height=5, fig.width=5, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with 1 percent dispersal between treated and untreated areas of equal size (no resistance costs)."}
+
+dfr <- run_rot(n_insecticides=4, max_gen = 130, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0, coverage=0.1, migration = 0.05)
 
@@
 
-```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with dispersal between treated and untreated areas and resistance cost of 0.005."}
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A sequence with dispersal between treated and untreated areas and resistance cost of 0.05."}
 
-#upped max_gen because migration and cost decrease rise in resistance
-dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.005, coverage=0.5, migration = 0.01)
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.05, coverage=0.5, migration = 0.01)
 
@@
 
-```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with with dispersal between treated and untreated areas and resistance cost of 0.005."}
+```{r, eval=TRUE, fig.height=2.5, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A 1 year rotation with with dispersal between treated and untreated areas and resistance cost of 0.05."}
 
-dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=0.5, migration = 0.01)
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.05, coverage=0.5, migration = 0.01)
 
 ```
+
+knitr::knit_exit() 
+### plots after this don't show much interesting
 

### 2018-02-07 R/rot_plots.r

@@ -46,7 +46,7 @@ rot_plot_resistance <- function(df_res2,
     gg <- ggplot( df_res2, aes_string(x='generation',y='resistance',colour='active_or_refuge') ) +
     geom_line( alpha=0.5, lwd=1.5 ) +
     #legend for the lines, allows me to set title & labels  
-    scale_colour_manual("areas connected\nby migration",values=c("red3","navy"), labels=c("treated","untreated"))
+    scale_colour_manual("areas connected\nby migration",values=c("red3","navy"), labels=c("treated","untreated refugia"))
   }   else
   {
     gg <- ggplot( df_res2, aes_string(x='generation',y='resistance')) +

### 2018-02-07 inst/shiny/rotresist1/ui.r

@@ -1,6 +1,7 @@
 #rotations/shiny/rotresist1/ui.r
 #andy south 9/10/17
 
+#original UI with development options
 
 library(shiny)
 library(markdown)
@@ -45,12 +46,12 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
   title = "insecticide rotations effects on resistance",
 
   #navbarPage sets up navbar, title appears on left
-  navbarPage("insecticide resistance evolution with rotations and sequences", id="selectedTab",
+  navbarPage("insecticide resistance evolution with rotations and sequences DEVELOPMENT VERSION", id="selectedTab",
            
            # tab UI
            tabPanel("UI", 
   
-  h6("Needs to be run on a widescreen. Modify inputs and compare 2 scenarios. Select About for outline."),  
+  h6("Warning! This is a research tool and not for making operational decisions.  Needs to be run on a widescreen. Modify inputs and compare 2 scenarios. Select About for outline."),  
   
   #fixedRow(
   fluidRow(
@@ -82,9 +83,9 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
     ),
     column(1, offset = 0,
            h5("n. insecticides"),
-           sliderInput("n_A", NULL, val=3, min = 1, max = 5, step = 1, ticks=FALSE),
+           sliderInput("n_A", NULL, val=3, min = 2, max = 5, step = 1, ticks=FALSE),
            hr(), #hr(),hr(),
-           sliderInput("n_B", NULL, val=3, min = 1, max = 5, step = 1, ticks=FALSE),
+           sliderInput("n_B", NULL, val=3, min = 2, max = 5, step = 1, ticks=FALSE),
            hr()
     ), 
     # moved up to one slider for both scenarios
@@ -153,23 +154,24 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
     ),
     column(1, offset = 0,
            h5("coverage"),
-           sliderInput("coverage_A", NULL, val=0.8, min = 0, max = 1, step = 0.1, ticks=FALSE),
+           sliderInput("coverage_A", NULL, val=0.8, min = 0.01, max = 1, step = 0.1, ticks=FALSE),
            hr(),#hr(),hr(),
-           sliderInput("coverage_B", NULL, val=0.8, min = 0, max = 1, step = 0.1, ticks=FALSE),
+           sliderInput("coverage_B", NULL, val=0.8, min = 0.01, max = 1, step = 0.1, ticks=FALSE),
[…]

### 2018-02-13 inst/documents/rotations_notes.Rmd
notes

@@
 
+re-read this 
+Sternberg & Thomas 2017
+Insights from agriculture for the management of insecticide resistance in disease vectors
+Because of the situation-specific nature of insecticide resistance management plans, public health may have an advantage over agriculture. Agricultural pest management is concerned with hundreds
+of species, and in many cases, detailed data are not available for each pest species (and potential nontarget organisms).
+Compared to agriculture, public health is concerned with a small number of insect species.
+This should enable a strongly data-driven approach to selecting resistance management strategies, but in reality, very little is currently known.
+
+• The success of an IRM strategy is highly dependent on ecological and biological specifics; there is not yet a solid evidence base of recommending one strategy over another.
+• If control is achieved via the use of multiple tools with diverse modes of action, not only will selection for insecticide resistance be reduced but also, the impact of resistance will likely be less severe.
+The pending resistance crisis creates an urgent need to develop and implement integrated, multitactic IVM strategies that parallel IPM in agriculture.
+
+To our knowledge, there is only one field trial aimed specifically at testing resistance management strategies against malaria mosquitoes (Penilla et al., 2007).
+Perhaps worryingly, this trial revealed no clear benefit of either insecticide rotations or mosaics in slowing the spread of resistance when compared to monotherapy.
+However, these results do not necessarily mean that rotations or mosaics cannot be effective, but rather highlights a critical gap in understanding when, how, and why such strategies might work.
+
+Download papers on modelling oxitech
+Insectcide resistance
+https://doi.org/10.1093/jee/100.5.1642
+https://doi.org/10.1603/029.102.0233
+
+These are not about insecticide resistance (they are about resistance to the gene itself but may have useful pointers)
+Alphey2011. Modeling resistance to genetic control of insects
+https://doi.org/10.1016/j.jtbi.2010.11.016
+
+Watkinson-Powell2017 Resistance to genetic insect control: Modelling the effects of space
+https://doi.org/10.1016/j.jtbi.2016.10.014
+
+Type of fitness cost influences the rate of evolution of resistance to transgenic Bt crops
+Looks like could be very interesting :
+DOI: 10.1111/1365-2664.12680
+
+Combining the high-dose/refuge strategy and self-limiting transgenic
+insects in resistance management- a test in experimental mesocosms 
+Zhou2017
+
+submit links to UIs and paper once it is published to :
+vectorlearningxchange.com
+
+send Rowland rotation paper to Ian.
+
 ~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ?

### 2018-03-16 inst/documents/rotations_notes.Rmd
sensi analysis

@@
 *1/2/2018*
-
 I want to create a figure that shows effect of untreated area, refugia
@@
 
-# TODO
+~ c 15 generations per year for mosquitos, 25 day generation time.
+~ 3 years = 45 generations
+~ https://www.cdc.gov/malaria/about/biology/mosquitoes/
 
-re-read this 
+*14/3/2018*
+re-red this 
 Sternberg & Thomas 2017
@@
 
+*15/3/2018* back from Peru and Geneva
+~ get trial run of sensitivity analysis going
+~ sensi_an_rotations1.Rmd
+
+~ now have 500 runs working, max gens up to 400
+~ seems that generations are identical for sequence and rotations
+~ results saved in df_res_all_rot.rda
+
+Setting coverage to <1 (0.5) and migration to 0.01 does result in rotations lasting longer than sequences in some scenarios.
+Particularly shorter ones.
+Greatest slope is when rot_interval set to 5, one scenario gives 300 for sequence and 350 for rotation.
+
+So maybe difference is created by the changeover, and more frequent changeovers create greater difference.
+
+Also when coverage set to 1 again and cost > 0 there was 1 scenario with a very short rot-interval where the rotation lasts fair bit longer than the sequence.
+
+## scenario 79 expo_hi 0.79 eff 0.72 rot_interval 0
+## tot gens deployed under freq 0.5 = 199
+
+## scenario 79 expo_hi 0.79 eff 0.72 rot_interval 6
+## tot gens deployed under freq 0.5 = 251
+
+run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
+run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
+
+It could perhaps be just a rounding thing, because the number of generations is an integer there will always be some overshooting, when you have more switches there is the potential for more overshooting and the runs will last for longer.
+
+~ checking on coverage=0, currently resistance frequencies increase when they shouldn't do 
+run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=0, add_gens_under50=TRUE)
+
+in run_rot()
+
+coverage is passed to :
+set_start_freqs() but only to determine whether the
+
+RAF 3rd dimension site is set to : c('intervention','refugia') or just 'intervention'
+
+actually if coverage=0 perhaps dimension should just be set to refuge.
+
+So maybe it doesn't point to an error that resistance rises when coverage=0 just a property of the model structure.
+
+# TODO
+
+~ in run_rot finish code condensing around :
+      raf_m <- RAF[insecticide, 'm', 'intervention',gen-1]
+      raf_f <- RAF[insecticide, 'f', 'intervention',gen-1]
+
+~ sensi_an_rotations1.Rmd
+
+~ Maybe don't allow coverage to be set to 0 ? or put in some kind of warning.
+
+~ how to deal with 'overshooting' leading to more generations under 50 when rotation intervals are shorter ? 
+run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
+run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
+
+~ can I create an easy function to compare two scenarios (seq & rotate) with otherwise same params ?
+
 Download papers on modelling oxitech
-Insectcide resistance
+Insecticide resistance
 https://doi.org/10.1093/jee/100.5.1642
@@
 vectorlearningxchange.com
-
-send Rowland rotation paper to Ian.
 
@@
 
-~ c 15 generations per year for mosquitos, 25 day generation time.
-~ 3 years = 45 generations
-~ https://www.cdc.gov/malaria/about/biology/mosquitoes/
-~
-
 # funding opportunities

### 2018-03-16 inst/documents/sensi_an_rotations1.Rmd
sensi analysis

--- 2017-11-23 inst/documents/sensi_an_rotations1.Rmd
@@
 
-n_scenarios <- 5
+n_scenarios <- 100
 
@@
 {
-  # expo_hi 0.1 – 0.9
-  # eff 0.3 - 1
+  # inputs (could also be saved to a dataframe first & extracted one row at a time)
   linputs <- list()
-
+  # constant inputs
+  # for Ians Q1 set coverage to 1 & cost to 0 
+  linputs$coverage <- 1
+  #linputs$migration <- 0.1
+  linputs$cost <- 0.1 #0
+  linputs$max_gen <- 400
+  # variable inputs
   linputs$expo_hi <-  runif(1, min=0.1, max=0.9)
@@
 
+  # because I want to compare sequence rot_when_r to rotation
+  # I could run all the scenarios above for both seq & rot
+  # (and filter out later)
+
+  #random rot interval integer
+  linputs$rot_interval <- sample( 5:50, 1 )  #beware set replace=T if more than 1
+
+  for(rot_or_not in 0:1)
+  {  
+    #to get rot_int of 0 every other run
+    if (rot_or_not == 0) linputs$rot_interval <- 0
+
+    ##################
     # run one scenario
@@
     # summarise results per scenario
+    # NOTE this is also currently calc at end of run_rot() prob better to calc there
+    # and put in the output dfres, but its then repeated per 
+    # TODO first AG fix this maybe go for non dplyr solution
+    # if I do calc in run_rot then just find value at first row for each gene
+    # Error in dplyr::filter(active_or_refuge == "active") : 
+    # object 'active_or_refuge' not found
     res <- dfres %>%
+    # only assess in control areas not in refugia  
     dplyr::filter(active_or_refuge=='active') %>%
     group_by(resist_gene) %>%
-  summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
-  summarise(mean_gens_under50 = mean(gens_under50)) %>%
+    # for all insecticides in all generations  
+    # summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
+    # summarise(mean_gens_under50 = mean(gens_under50)) %>%
+    # just for deployed insecticides 
+    summarise(gens_dep_under50 = sum(resistance < 0.5 &
+                                     #finds insecticide in use = this one
+                                     resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
+                                 summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
                                  unlist()
 
-  # add results onto inputs list
-  linputs <- c(linputs, res)
+    #print scenario num and inputs to doc for exploration
+    cat("scenario",i,
+        " expo_hi",signif(linputs$expo_hi,2),
+        " eff",signif(linputs$eff,2),
+        " rot_interval",linputs$rot_interval,
+        "\n")
 
-  # change list to dataframe
-  df_in_out <- as.data.frame(linputs)
+    #print results in doc for diagnostics
+    cat("tot gens deployed under freq 0.5 =", res,"\n\n")
+
+    # create dataframe from inputs and outputs
+    # first add scenario column onto results
+    res$scenario <- i
+    df_in_out <- as.data.frame(c(linputs, res))
 
@@
 
-}
+  } #end rot_or_not loop
+} #end scenario loop
+
+# save object containing inputs and results as rda for analysis
+save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))
 
@@
 
-<!-- eval=FALSE to stop slow code being executed after 1st time --> 
-```{r, eval=FALSE}
+\pagebreak
 
-  # this is how it was done for resistance paper1
-  # nScenarios <- 10  
-  # ## do model runs
-  # listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
-  # ## save results objects as rda for re-use
-  # save(listOutMix, file=paste0(outFolder,'listOutMix_rr_10.rda'))
+# summary plot of all runs, would show any differences between sequence & rotations as non horizontal lines
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
+
+# want to plot line between sequence and rotation value
+# any deviation from horizontal will be obvious
+# could use rotation interval as the x axis, so that longer intervals will be clearer
+
+# maybe not for publication
+# but useful for quick view
+
+    gg <- ggplot( df_res_all, aes_string(x='rot_interval',
+                                         y='tot_gens_dep_under50',
+                                         group='scenario',
+                                         colour='scenario') ) +
+          geom_line( alpha=0.5, lwd=1 )
+
+    plot(gg)
 

### 2018-03-16 inst/documents/rotations_notes.Rmd
notes

@@
 
+*16/3/18*
+~ code condensing in run_rot()
+
+~ added stop in run_rot() for if coverage=0
+
+~ looking at comparing two scenarios (rot & seq) on same plot
+
+I tried this before with : df_resanother in rot_plots. It puts on the resistance curves for the 2 scenarios, but doesn't show the insecticide use for each
+dfseq <- run_rot(n_insecticides=2, rot_interval=0, expo_hi=0.5, eff=0.5, cost=0.01, coverage=1)
+dfrot <- run_rot(n_insecticides=2, rot_interval=15, expo_hi=0.5, eff=0.5, cost=0.01, coverage=1)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)
+
+dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)
+
+I have improved plotting of 2 scenarios together.
+Seems that overshooting problem occurs in rotations.
+Possibly because rotations are not stopped when the threshold is reached.
+See dfrot from prev e.g.
+whichj keeps going up to 0.62 at gen 127.
+This may not matter if the counting of gens only includes those < 0.5.
+The other option is to stop rotations early,m which I thought I was already doing with exit_rot = TRUE ??
+
+in insecticide_check() exit_rot should prompt exit from rotation when 0.5 threshold is exceeded.
+
+dfrot2 <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, diagnostics=TRUE)
+
+Aha! It could be a problem with min_rwr_interval, which is 5 by default and so may force staying within rotations and thus cause overshooting of threshold.
+
+But i thought that only gens under thresh are counted ???
+
 # TODO
 
-~ in run_rot finish code condensing around :
-      raf_m <- RAF[insecticide, 'm', 'intervention',gen-1]
-      raf_f <- RAF[insecticide, 'f', 'intervention',gen-1]
-
-~ sensi_an_rotations1.Rmd
-
-~ Maybe don't allow coverage to be set to 0 ? or put in some kind of warning.
+~ check how gens_under_thresh are counted in sensi_an_rotations1
+    summarise(gens_dep_under50 = sum(resistance < 0.5 &
+                                     #finds insecticide in use = this one
+                                     resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
+                                 summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
+                                 unlist()
 
@@
 
-~ can I create an easy function to compare two scenarios (seq & rotate) with otherwise same params ?
+~ move github repo back to mine
+~~ From Github I merged changes from Ians with a pull request after comparing across the forks.
+~~ Now I can possibly change RStudio to point to the repo on my Github (do when I have a bit more time)
+~~ update R
+
+~ sensi_an_rotations1.Rmd
 

### 2018-03-20 inst/documents/phaps-rotations-better.Rmd
thorough check

---
title: "perhaps rotations are better than sequences ?"
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

  ```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(rotations)
library(ggplot2)

outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

resistance frequencies seem to take longer to rise for rotations than sequences once cost (or rufugia) are included. 
particularly when the rotation interval is short

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=9}

#for above
#\pagebreak

linputs <- list()

linputs$n_insecticides <- 2
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0.1 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$plot <-     FALSE

linputs$rot_interval=0
dfseq <- do.call(run_rot, linputs)
linputs$rot_interval=6
dfrot <- do.call(run_rot, linputs)

ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

```

### 2018-03-20 inst/documents/rotations_notes.Rmd
thorough check

@@
 
+*19/3/2018* on train to liverpool
+
+trying to understand overshooting issue
+
+dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)
+
+but doesen't look like much time over the 0.5 thresh for the rotations can explain why they take longer. The short rotations only get to 0.5 towards the end.
+
+Can I refine this example to understand what may be happening ?
+
+turn off log ?
+
+dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
+
+2 insecticides
+dfseq <- run_rot(rot_interval=0, max_gen = 70, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, n_insecticides=2)
+dfrot <- run_rot(rot_interval=6, max_gen = 70, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, n_insecticides=2)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
+gens_under_thresh(dfrot)
+gens_under_thresh(dfseq)
+
+This does go over the 0.5 threshold ? can I count this example in sensi_an_rotations ?
+
+gens_under_thresh(dfrot) : 49
+gens_under_thresh(dfseq) : 46
+
+When there is no cost they are identical (39) :
+dfseq <- run_rot(rot_interval=0, max_gen = 70, migration=0.01, expo_hi=0.8, eff=0.8, cost=0, coverage=1, n_insecticides=2)
+dfrot <- run_rot(rot_interval=6, max_gen = 70, migration=0.01, expo_hi=0.8, eff=0.8, cost=0, coverage=1, n_insecticides=2)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
+gens_under_thresh(dfrot)
+gens_under_thresh(dfseq)
+
+but they still look diff in plots, i think because stoppping criteria are different in plots for seq & rot.
+
+# weirdly when I went to 3 insecticides, & cost 0.1
+# gen calc gives same result for each at 69 and plots finish at same point ?
+# aha! that was because max_gen=70 so max result could be 69 i.e. no insectidides in either scenario reached threshold
+
+dfseq <- run_rot(rot_interval=0, max_gen = 100, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, n_insecticides=3)
+dfrot <- run_rot(rot_interval=6, max_gen = 100, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, n_insecticides=3)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
+gens_under_thresh(dfrot)
+gens_under_thresh(dfseq)
+
+1 20-20 = 0
+2 insecticides 46 - 49 = 6
+3 insecticides difference is 83-79 = 4
+4 : 147 - 124 = 23
+5 : 218 - 200 = 18
+6 : 437 - 391 = 46
+
+? no obvious pattern
+
+done ~ write function to count gens under thresh
+done ~ add thresh line to plot
+done ~ add warning to run_rot() for if threshold not reached in time
+
+exit_rot seems not quite to be doing what I expect. e.g. if I set to just 1 insecticide sequence keeps going longer than rot (although this is a somewhat weird example)
+
+Aha! it is min_rwr_interval that did this. It stops a change early in a rotation but this can stop the simulation from ending.
+
+(however it shouldn't effect the gens_under_threshold)
+
+Try seeing if the gens_under_thresh are the same when min_rwr_interval is disabled.
+
+(also note that min_rwr_interval isn't really sposed to apply to rotations, partly a result of the structure )
+
+setting min_rwr_interval to 1 now results if diff of 86-77 = 9 for the 3 insecticides ??
+
+*20/3/18*
+Go back to the scenarios that was very different in sensi_an_rotations_test num79
+
+dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
+gens_under_thresh(dfrot)
+gens_under_thresh(dfseq)
+
+To enable side by side comparison
+
+library(gridExtra)
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE)
+grid.arrange(ggrot, ggseq, ncol = 1)
+
+I think maybe need to go back to two insecticides to try to find small difference :
+
+ni <- 2
+cost <- 0.1
+gens <- 70 
+dfseq <- run_rot(rot_interval=0, max_gen=gens, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
+dfrot <- run_rot(rot_interval=6, max_gen=gens, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE)
+grid.arrange(ggrot, ggseq, ncol = 1)
+gens_under_thresh(dfrot)
+gens_under_thresh(dfseq)
+
+rot:48, seq:46
+
+Note that in the above plot other strange issues, which may be plotting probs.
+~ insecticide use boxes don't join up, seem to be gaps
+~ resistance keeps increasing after insecticide box has finished
+~ first rotation seems shorter than later, because insecticide is classed as NA in firs tstep, is that right ?
+
+Difference is for i2 which has 4 rots of 6, 24, only reaches threshold at end of final rot
+In seq, 12 last for 2 generations less.
+
+Trying too look at plot in more detail
+dfrot2 <- run_rot(rot_interval=6, max_gen=20, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
+
+Not sure if I'm going to be able to fix the ribbons joining ? can I +1 to the x upper limit ?
+
+  to_plot_in_use <- 1
+
+  dfrot2 <- dfrot2 %>%
+    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide), to_plot_in_use, NA))
+
+  gg <- ggplot( dfrot2, aes_string(x='generation',y='resistance')) +
+    geom_line( alpha=0.5, lwd=1, colour='red3' ) +
+    facet_wrap('resist_gene', ncol=1) +
+    geom_ribbon( aes_string(x='generation', ymin=0, ymax='i_in_use'))
+
+  plot(gg)
+
+I think ribbons not meeting and curve increasing after are just plotting issues that I don't know how to fix.
+
+If a rotation is 10 generations, the first finishes at 10 and the next is shown as starting at 11.
+The effect of the first continues until 11.
+Setting i_in_use to 1 for the following generation would fix the plotting issue.
+
+Do I need to try to add a generation 0 to store starting conditions ? NO
+
+Generations
+The conditions within a generation influence resistance which is stored in the following generation.
+Therefore it's good to set insecticide for time 1, and this is consistent with later generations.
+
+So I've fixed saving of insecticide for t1
+
+ni <- 2
+cost <- 0.1
+gens <- 70 
+dfseq <- run_rot(rot_interval=0, max_gen=gens, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
+dfrot <- run_rot(rot_interval=6, max_gen=gens, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE)
+grid.arrange(ggrot, ggseq, ncol = 1)
+gens_under_thresh(dfrot)
+gens_under_thresh(dfseq)
+
+Still have difference between 24 & 22 for i2 in above example
+
+# add in_use columns  
+  dfrot <- dfrot %>%
+    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide), 1, NA)) 
+  dfseq <- dfseq %>%
+    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide), 1, NA)) 
+#remove NA refuge rows    
+dfrot <- dfrot[dfrot$active_or_refuge=='active',]
+dfseq <- dfseq[dfseq$active_or_refuge=='active',]
+
+write_csv(dfrot,'rot.csv')
+write_csv(dfseq,'seq.csv')
+
+see : rot_seq_compare_march2018.xlsx
+
+With cost it seems that rotation may be able to last longer than a sequence.
+
+Could this be because decline in resistance freq is more rapid at higher frequencies.
+
+Both increases and decreases in resistance frequency are higher at higher resistance frequencies.
+
+But does this explain ?
+
+Is it something to do with more frequent changes ?
+
+Could it be because :
+~ frequency changes are bigger as frequencies get higher, 
+~ allowing in a sequence to reach high frequencies straight away makes it slightly quicker than see-sawing at lower frequencies
+~ is their some maths to prove this ?
+~ ask Ian
+~ can I print an example ?
+
+Ian said that at higher dominance the difference in change between low & high frequencies will be greater, so might expect the difference between rot and seq to be greater. Try it and see.
+
+linputs <- list()
+
+linputs$n_insecticides <- 2
+linputs$coverage <- 1
+#linputs$migration <- 0.1
+linputs$cost <-     0.1 
+linputs$max_gen <-  200
+linputs$expo_hi <-  0.6
+linputs$eff <-      0.6
+linputs$dom_sel <-  1
+linputs$dom_cos <-  1
+
+linputs$rot_interval=0
+dfseq <- do.call(run_rot, linputs)
+linputs$rot_interval=6
+dfrot <- do.call(run_rot, linputs)
+
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE)
+library(gridExtra)
+grid.arrange(ggrot, ggseq, ncol=1)
+gens_under_thresh(dfrot)
+gens_under_thresh(dfseq)
+
+with higher dominance there is only 1 generation difference (31:30) may be partly because it is faster
+
+when I reduced exposure and effectiveness then a largish difference between rot & seq did return.
+
 # TODO
+
+~ compare dfrot & dfseq for above example
+
+~ seems that min_rwr_interval explains the plots looking different but not the measured gens_under_thresh
+
+~ maybe come back to i_in_use ribbons not meeting issue
+
+~ rename rot_criterion to rot_threshold
+
+~ add in some unit-tests ?
 
@@
 ~ how to deal with 'overshooting' leading to more generations under 50 when rotation intervals are shorter ? 
-run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
-run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
+dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
+rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)
 
@@
 ~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ?
-
-~ something not right, when run with coverage=0 still get increase in resistance in treated areas, even though technically there is no treated area, maybe I just need to force coverage to be > 0 :
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 0 , eff = 0.8 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0 , migration = 0.01 , cost = 0.01 , logy = TRUE no_r_below_start = TRUE min_rwr_interval = 5 )
 

### 2018-03-20 inst/documents/sensi_an_rotations1.Rmd
thorough check

@@
         " rot_interval",linputs$rot_interval,
+        " cost", linputs$cost,
         "\n")

### 2018-03-20 R/insecticide_switch.r

@@ -27,7 +27,7 @@ insecticide_switch <- function( RAF,
                                 df_results,
                                 diagnostics)
 {
-  next_insecticide_found <- 0
+  next_insecticide_found <- FALSE
   candidate <- current_insecticide 
   
   #TODO use df_results to stop switching to an insecticide that has been used recently
@@ -44,17 +44,17 @@ insecticide_switch <- function( RAF,
                       gen, current_insecticide, candidate,
                       RAF[current_insecticide, 'f','intervention', gen], RAF[candidate, 'f','intervention', gen]))
       
-      next_insecticide_found <- 1 
+      next_insecticide_found <- TRUE 
       current_insecticide <- candidate 
       #change_insecticide <- 0 
     }
     
-    if (next_insecticide_found==1) break   
+    if (next_insecticide_found) break   
     
   } # end of loop checking each insecticide
  
 # return 0 if no insecticide to use next found  
-if (next_insecticide_found==0) current_insecticide <- 0 
+if (!next_insecticide_found) current_insecticide <- 0 
    
 return(current_insecticide)  
 }

### 2018-03-20 R/run_rot.r

@@ -62,7 +62,7 @@ run_rot <- function(max_gen = 200,
                     rr = 0.5, #c(0.5, 0.5, 0.5),
                     cost = 0.1, #c(0,0,0),
                     fitSS = 1,
-                    min_rwr_interval = 5,
+                    min_rwr_interval = 1,
                     no_r_below_start = TRUE,
                     exit_rot = TRUE,
                     
@@ -86,7 +86,7 @@ run_rot <- function(max_gen = 200,
   names(l_gene_plus_activity) <- c( paste0('insecticide', 1:n_insecticides, "_active"),
                                     paste0('insecticide', 1:n_insecticides, "_refuge") )
   
-  df_results <- do.call(data.frame, list(generation = 1:max_gen,
+  df_results <- do.call(data.frame, list(generation = 1:max_gen,                                         
                                          insecticide = NA,
                                          stringsAsFactors = FALSE,
                                          l_gene_plus_activity))  
@@ -132,11 +132,17 @@ run_rot <- function(max_gen = 200,
   
   # usually start the rotation sequence at #1 but can specify any one start
   current_insecticide <- start_insecticide 
+  # store the insecticide for generation1 although its effects will only be seen in following gen
+  # this is consistent with later generations.
+  df_results$insecticide[1] <- current_insecticide
+  
   gens_this_insecticide <- 1
   
-  #start at generation 2 because generation 1 holds the user-defined initial allele frequencies
+  ## start at generation 2 because generation 1 holds the user-defined initial allele frequencies
+  ## The conditions within a generation influence resistance which is stored in the following generation.
+  ## in the plots there is gap between rotation that stops e.g. at gen 10 and next starts at gen 11
   for(gen in 2:max_gen)
-  { 
+    { 
     for(insecticide in 1:n_insecticides)
     {
       # extract resistance allele freqs (raf) for previous timestep
@@ -311,6 +317,9 @@ run_rot <- function(max_gen = 200,
                                                 diagnostics=diagnostics)
     } 
   
+    #message("gen", gen, " curr_ins=", current_insecticide, " change=", change_insecticide)
+    
+    
     # if no suitable insecticide left, break out of generations loop
     if (current_insecticide == 0) #(next_insecticide_found==0)
     {
@@ -323,11 +332,12 @@ run_rot <- function(max_gen = 200,
     }    
  
     # recording the insecticide that's going to be used in next timestep
[…]

### 2018-03-22 inst/documents/phaps-rotations-better.Rmd
finding when rotations and sequences favoured

@@
 
+```{r, eval=TRUE, include=FALSE}
+
+library(knitr)
+#https://stackoverflow.com/a/36234023/1718356 for fig.pos = 'h'
+
+opts_chunk$set(fig.height=7, fig.width=9, fig.pos = 'h', echo=FALSE, warning=FALSE, message=FALSE)
+
+```
+
 ```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
@@
 
-resistance frequencies seem to take longer to rise for rotations than sequences once cost (or rufugia) are included. 
-particularly when the rotation interval is short
+Resistance frequencies can take longer to rise for rotations than sequences once cost (or refugia) are included. 
+Particularly noticeable when the rotation interval is short, costs (0.1) and dominance (1) high.
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=9}
+Can this be a simple mathematical result of:
+
+1. slower rate of resistance frequency change at lower resistance frequencies. 
+2. rotation initially keeps frequencies lower where this rate of change is less.
+
+```{r, eval=TRUE}
 
@@
 grid.arrange(ggrot, ggseq, ncol=1)
-gens_under_thresh(dfrot)
-gens_under_thresh(dfseq)
+grot <- as.numeric(gens_under_thresh(dfrot))
+gseq <- as.numeric(gens_under_thresh(dfseq))
+cat('sum generations deployed insecticide under threshold rotation:', grot,' sequence:',gseq)
+
+print(unlist(linputs))
 
 ```
+
+\pagebreak
+
+Cost reduced from 0.1 to 0.02. Radically cuts difference between rotation and sequence.
+
+```{r, eval=TRUE}
+
+#for above
+#\pagebreak
+
+linputs <- list()
+
+linputs$n_insecticides <- 2
+linputs$coverage <- 1
+#linputs$migration <- 0.1
+linputs$cost <-     0.02 
+linputs$max_gen <-  200
+linputs$expo_hi <-  0.6
+linputs$eff <-      0.6
+linputs$dom_sel <-  1
+linputs$dom_cos <-  1
+linputs$plot <-     FALSE
+
+linputs$rot_interval=0
+dfseq <- do.call(run_rot, linputs)
+linputs$rot_interval=6
+dfrot <- do.call(run_rot, linputs)
+
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+library(gridExtra)
+grid.arrange(ggrot, ggseq, ncol=1)
+grot <- as.numeric(gens_under_thresh(dfrot))
+gseq <- as.numeric(gens_under_thresh(dfseq))
+cat('sum generations deployed insecticide under threshold rotation:', grot,' sequence:',gseq)
+
+print(unlist(linputs))
+
+```
+
+\pagebreak
+
+Decreasing dominance of resistance and cost from 1 to 0 (had to increase exposure, effectiveness and rr to get similar times-to-resistance). Then the difference between rotation and sequence is less.
+
+```{r, eval=TRUE}
+
+linputs <- list()
+
+linputs$n_insecticides <- 2
+linputs$coverage <- 1
+#linputs$migration <- 0.1
+linputs$cost <-     0.02 
+linputs$max_gen <-  200
+linputs$expo_hi <-  0.8
+linputs$eff <-      0.8
+linputs$rr  <-      0.8
+linputs$dom_sel <-  0
+linputs$dom_cos <-  0
+linputs$plot <-     FALSE
+
+linputs$rot_interval=0
+dfseq <- do.call(run_rot, linputs)
+linputs$rot_interval=6
+dfrot <- do.call(run_rot, linputs)
+
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+library(gridExtra)
+grid.arrange(ggrot, ggseq, ncol=1)
+grot <- as.numeric(gens_under_thresh(dfrot))
+gseq <- as.numeric(gens_under_thresh(dfseq))
+cat('sum generations deployed insecticide under threshold rotation:', grot,' sequence:',gseq)
+
+print(unlist(linputs))
+
+```
+
+\pagebreak
+
+Dominance of cost 0 and dom_sel 1.
+
+Benefit of rotation is lost, sequence lasts a tiny bit longer. Suggests it is a high dominance of cost that favours sequence over rotation.
+
+```{r, eval=TRUE}
+
+linputs <- list()
+
+linputs$n_insecticides <- 2
+linputs$coverage <- 1
+#linputs$migration <- 0.1
+linputs$cost <-     0.1 
+linputs$max_gen <-  200
+linputs$expo_hi <-  0.6
+linputs$eff <-      0.6
+linputs$dom_sel <-  1
+linputs$dom_cos <-  0
+linputs$plot <-     FALSE
+
+linputs$rot_interval=0
+dfseq <- do.call(run_rot, linputs)
+linputs$rot_interval=6
+dfrot <- do.call(run_rot, linputs)
+
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+library(gridExtra)
+grid.arrange(ggrot, ggseq, ncol=1)
+grot <- as.numeric(gens_under_thresh(dfrot))
+gseq <- as.numeric(gens_under_thresh(dfseq))
+cat('sum generations deployed insecticide under threshold rotation:', grot,' sequence:',gseq)
+
+print(unlist(linputs))
+
+```
+
+\pagebreak
+dom_sel and dom_cost 0.1, small favour of rotation
+
+```{r, eval=TRUE}
+
+linputs <- list()
+
+linputs$n_insecticides <- 2
+linputs$coverage <- 1
+#linputs$migration <- 0.1
+linputs$cost <-     0.1 
+linputs$max_gen <-  200
+linputs$expo_hi <-  0.8
+linputs$eff <-      0.8
+linputs$dom_sel <-  0.1
+linputs$dom_cos <-  0.1
+linputs$plot <-     FALSE
+
+linputs$rot_interval=0
+dfseq <- do.call(run_rot, linputs)
+linputs$rot_interval=6
+dfrot <- do.call(run_rot, linputs)
+
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+library(gridExtra)
+grid.arrange(ggrot, ggseq, ncol=1)
+grot <- as.numeric(gens_under_thresh(dfrot))
+gseq <- as.numeric(gens_under_thresh(dfseq))
+cat('sum generations deployed insecticide under threshold rotation:', grot,' sequence:',gseq)
+
+print(unlist(linputs))
+
+```

### 2018-03-22 inst/documents/rotations_notes.Rmd
finding when rotations and sequences favoured

@@
 
+What value of cost did Rowland1991 use ? Seems that he used high costs of 0.46 and 0.35, see table 8.
+
 Rowland2013 data on hut trials of actellic - pirimphos-methyl - an OP.
@@
 
+eek
+      0 [main] us 0 init_cheap: VirtualAlloc pointer is null, Win32 error 487
+AllocationBase 0x0, BaseAddress 0x68570000, RegionSize 0xE0000, State 0x10000
+C:\Program Files (x86)\Git\bin\ssh.exe: *** Couldn't reserve space for cygwin's heap, Win32 error 0
+fatal: Could not read from remote repository.
+
+Please make sure you have the correct access rights
+and the repository exists.
+
 # TODO
 
-~ compare dfrot & dfseq for above example
+~ look more at phaps-rotations-better.Rmd
 

### 2018-03-22 R/run_rot.r

@@ -46,7 +46,7 @@
 
 run_rot <- function(max_gen = 200, 
                     n_insecticides = 4, 
-                    start_freqs = 0.001,
+                    start_freqs = 0.01,
                     rot_interval = 10, 
                     rot_criterion = 0.5, 
                     migration = 0.01, 

### 2018-04-04 inst/documents/sensi_an_rotations1.Rmd
add extra outputs

--- 2018-03-16 inst/documents/sensi_an_rotations1.Rmd
@@
   library(ggplot2)
+  library(gridExtra)
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=4}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
 #results='hide',
+#comment=NA removes ## before text output
 
@@
 
-n_scenarios <- 100
+n_scenarios <- 200
 
@@
   linputs <- list()
-  # constant inputs
+
   # for Ians Q1 set coverage to 1 & cost to 0 
+  # April 2018, having go at including variable cost, may want to do some cost=0 runs later
+
+  ## constant inputs ##
   linputs$coverage <- 1
-  #linputs$migration <- 0.1
-  linputs$cost <- 0.1 #0
-  linputs$max_gen <- 400
-  # variable inputs
-  linputs$expo_hi <-  runif(1, min=0.1, max=0.9)
-  linputs$eff <-      runif(1, min=0.3, max=1)
+  linputs$migration <- 0
+  #linputs$cost <- 0 #variable cost below
+  linputs$max_gen <- 500
 
-  # because I want to compare sequence rot_when_r to rotation
-  # I could run all the scenarios above for both seq & rot
-  # (and filter out later)
+  #beware min_rwr_interval, it's to stop repeated small rotations
+  linputs$min_rwr_interval = 5
+  #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
+  linputs$plot <-     FALSE
+
+  ## variable inputs ##
+  linputs$cost <-           runif(1, min=0,   max=0.1) 
+  linputs$expo_hi <-        runif(1, min=0.4, max=0.9)  #min=0.1, max=0.9
+  linputs$male_expo_prop <- runif(1, min=0,   max=1)
+  linputs$eff <-            runif(1, min=0.5, max=1) # min=0.3, max=1
+  linputs$rr  <-            runif(1, min=0.1, max=0.9)
+  linputs$dom_sel <-        runif(1, min=0,   max=1)
+  linputs$dom_cos <-        runif(1, min=0,   max=1) 
+  linputs$start_freqs <-    runif(1, min=0.005,   max=0.1) #note not log yet   
+  linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
 
@@
 
-  for(rot_or_not in 0:1)
+  # create null objects used plater for plotting
+  ggrot <- ggseq <- NULL
+
+  # to compare sequence to rotation run all the scenarios above for both seq & rot 
+  # beware has to be 1:0 because interval is overwritten for seq
+  for(rot_or_not in 1:0)
   {  
@@
     # summarise results per scenario
-    # NOTE this is also currently calc at end of run_rot() prob better to calc there
-    # and put in the output dfres, but its then repeated per 
-    # TODO first AG fix this maybe go for non dplyr solution
-    # if I do calc in run_rot then just find value at first row for each gene
-    # Error in dplyr::filter(active_or_refuge == "active") : 
-    # object 'active_or_refuge' not found
+    # NOTE this is also currently calc at end of run_rot() here or other post-process is probably better
+    # and put in the output dfres, but the value has to be rpeated for every generation 
     res <- dfres %>%
@@
 
-    #print scenario num and inputs to doc for exploration
-    cat("scenario",i,
-        " expo_hi",signif(linputs$expo_hi,2),
-        " eff",signif(linputs$eff,2),
-        " rot_interval",linputs$rot_interval,
-        "\n")
-
-    #print results in doc for diagnostics
-    cat("tot gens deployed under freq 0.5 =", res,"\n\n")
-
     # create dataframe from inputs and outputs
@@
 
+    ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+    gens_res <- as.numeric(gens_under_thresh(dfres))
+
+    # to plot results for checking
+    if ( linputs$rot_interval == 0 )
+    {
+      ggseq <- ggres
+      gens_seq <- gens_res
+    } else
+    {
+      ggrot <- ggres
+      gens_rot <- gens_res
+    }
+
   } #end rot_or_not loop
+
+  #printing input values
+  #print(unlist(linputs))
+
+  # try to plot input vals instead
+  # extract out those between 0 & 1
+  dfi <- data.frame(t(as.data.frame(linputs)))
+  names(dfi) <- 'value'
+  dfi$inputs <- rownames(dfi)
+
+  #take out columns not between 0 & 1
+  dfi <- filter(dfi, !inputs %in% c('max_gen','rot_interval','n_insecticides','plot','min_rwr_interval'))
+  #also remove coverage & migration both set to 0 for now
+  dfi <- filter(dfi, !inputs %in% c('coverage','migration'))
+
+  ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
+            geom_point() +
+            theme_minimal() +
+            ylim(0,1) +
+            theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))
+
+  grid.arrange(ggrot, ggseq, ggins, ncol=1)
+
+  cat('scenario',i,'sum generations deployed insecticide under thresh rotation:', gens_rot,' sequence:',gens_seq,"\n\n")
+
+  #can I give the code to rerun this scenario
+  cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
+              "cost=",signif(linputs$cost,2),", ",
+              "start_freqs=",signif(linputs$start_freqs,2),", ", 
+              "rot_interval=",linputs$rot_interval,", ",
+              "eff=",signif(linputs$eff,2),",\n",
+              "dom_sel=",signif(linputs$dom_sel,2),", ",
+              "dom_cos=",signif(linputs$dom_cos,2),", ",
+              "rr=",signif(linputs$rr,2),", ",
+              "expo_hi=",signif(linputs$expo_hi,2),", ",
+              "coverage=",signif(linputs$coverage,2),",\n",
+              "migration=",signif(linputs$migration,2),", ",
+              "max_gen=",linputs$max_gen,", ", 
+              "min_rwr_interval=",linputs$min_rwr_interval,            
+              ")\n") )
+
 } #end scenario loop
@@
 
+# check proportion of scenarios not reaching thresholds by max_runs
+
+    nruns_fail <- df_res_all %>% 
+                  filter(tot_gens_dep_under50 == linputs$max_gen) %>%
+                  summarise(n=n()) %>% 
+                  unlist()
+
+    cat('proportion runs not reaching thresholds : ', nruns_fail/(2*n_scenarios) ) #2* beacuse eeach scenario run for seq & rot
+
+# plot gens againts exposure and effectiveness
+# as a quick test to see if I can set the starting values of these higher
+
+    gg <- ggplot( df_res_all, aes_string(x='eff',
+                                         y='tot_gens_dep_under50') ) +
+          geom_point( alpha=0.5 )
+    plot(gg)
+
+    gg <- ggplot( df_res_all, aes_string(x='expo_hi',
+                                         y='tot_gens_dep_under50') ) +
+          geom_point( alpha=0.5 )
+    plot(gg)
+
+    gg <- ggplot( df_res_all, aes_string(x='expo_hi*eff*male_expo_prop',
+                                         y='tot_gens_dep_under50') ) +
+          geom_point( alpha=0.5 )
+    plot(gg)
+
 ```

### 2018-04-27 R/sensi_an_param_call_list.Rmd
starting paper3 and developing sensi analysis

---
  title: "sensi_an_param_call_list"
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

To create a list of function calls one for each scenario in the sensitivity analysis.

To allow any run from the sensitivity analysis to be rerun and checked.

Relies on the seed being the same as in sensi_an_rotations.

A better option might be to read a line from df_res_all to extract the input params to rerun the model.

df_res_all[624,1:15]

### 2018-04-27 inst/documents/paper3/paper3_rotations.Rmd
starting paper3 and developing sensi analysis

---
title: "Insecticide rotations delay evolution of resistance in a minority of model runs when compared to sequential use."
# output: 
#   pdf_document:
#     keep_tex: true
#     fig_caption: true
#     latex_engine: pdflatex

output: word_document
#author: "Andy South and Ian Hastings"
#date: "`r Sys.Date()`"
bibliography: library.bib
csl: biomed-central.csl
fontsize: 11pt
spacing: double
# below worked in test_linenumbers.Rmd
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
---

Andy South * and Ian M. Hastings

Department of Parasitology, Liverpool School of Tropical Medicine, Liverpool L3 5QA, UK.

southandy@gmail.com, andy.south@lstmed.ac.uk

ian.hastings@lstmed.ac.uk

\* Corresponding author

```{r, eval=TRUE, include=FALSE}
#!!!! PAPER PROTOCOL
#!!!! submission MS, knit to word_document 
#!!!!  in WORD: delete figs, set double space, add line numbers 
#!!!!  (spacing & line numbers work for pdf output but not for word)
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='pap2figs/')
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
  library(rotations)
  library(ggplot2)
  library(gridExtra)

  outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

## Abstract

Insecticide resistance threatens the control of the vectors of dangerous diseases including malaria, dengue and zika. Recent increases in insecticide resistance in public health are an evolutionary process caused by sustained exposure of insect populations to a small number of available insecticides. Resistance is a particular problem in public health as compared to agriculture because insecticides need to be longer lasting to provide affordable protection and there are fewer available active ingredients. Efforts to limit the development of insecticide resistance are grouped under the term Insecticide Resistance Management (IRM). The main strategies advocated to reduce the development of resistance are 1) rotations, 2) sequences, 3) mixtures  and 4) mosaics. Rotations regularly switch between the use of different insecticides with a short time step (one or a few years) irrespective of resistance levels. Sequences, in contrast, switch from one insecticide to another only when resistance levels have reached a critical, defined threshold usually over a longer timescale.

Rotations are often advocated as one of the best options for limiting the development of resistance.

Testing different IRM strategies in the field is difficult and there has been little recent modelling work on insecticide rotations.

In this paper a model is described allowing rotations and sequences to be compared in terms of their effect on the evolution of insecticide resistance. The model is used to develop a mechanistic understanding of when and why insecticide resistance is likely to evolve faster with rotations or sequences.

The results suggest that under more likely circumstances than not the evolution of insecticide resistance will reach resistance thresholds at very similar times for rotations and sequences.

However under a less common, but still plausible, set of circumstances there are predicted to be large adavnatages to using a rotations as compared to a sequence.

The advantages to a rotation only occur when there are costs of resistance or dispersal from untreated areas, and there is high dominance of cost and low dominance of selection.

The mechanisms for these results are explained.

Other operational factors will favour rotations or sequences aside from the implications for the evolution of resistance. Developing an understanding of the evolutionary implications allows a more explicit consideration of these other factors on their own merits.

\pagebreak

## Introduction
Despite much field and modelling work on the evolution of resistance under different insecticide or drug strategies [@Consortium2013] there is a small evidence base to support decisions about the use of insecticides in public health [@Sternberg2017].

Rotations are advocated as one of the most favoured approaches for Insecticide Resistance Management (IRM)[@WHO2012].

Here a modern modelling approach is described to allow an assessment of the potential benefits, in terms of the evolution of resistance, of rotations over sequences.

## Methods

A population genetic model was developed simulating changes over time in the frequencies of resistant and susceptible alleles in response to fitness differences. The model uses a similar approach to that described in [@Levick2017,@South2018] to compare insecticide mixtures and sequences. The implementation is simpler here because for rotations and sequences there is only a need to follow one insecticide at a time. The simpler implementation allows the model to be run for an unlimited number of insecticides.

Sequences : one insecticide is used until resistance threshold frequency (0.5 in this case) is reached, then switch to new insecticide. 
Rotations : use insecticide for set time, switch to other.

In both cases continue until no more insecticides below the resistance threshold remain.

Model inputs effect fitness as shown in Figure 1.

The model was run under four main scenarios :
1. no resistance fitness costs or dispersal link to untreated areas
2. fitness costs of resistance added
3. dispersal link to untreated areas (no costs)
4. costs and dispersal

To generate a decline in resistance over time either fitness costs or a dispersal link to an untreated area are required.

*talk about model outputs, i.e. generations deployed below threshold*
*try to come up with brief name to make it easier to refer to it later*

## Results

### no resistance fitness costs or dispersal link to untreated refugia

When there are no resistance fitness costs or dispersal links to untreated refugia resistance frequencies do not decline when an insecticide is not in use. Thus for sequences once resistance thresholds are reached the resistance frequencies remain at that level and the insecticide cannot be re-used. For rotations resistance frequencies step upwards when insecticides are in use and remain at a plateau when they are not. This results in

### fitness costs of resistance added

## Discussion

\pagebreak

## Figures

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1. [todo need to modify caption from MJ paper, start by saying that only in left panel during application of the insecticide] The effect of model inputs on the fitness of genotypes for a single insecticide. Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis. Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed). For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of resistance determines how the fitness for SR lies between that of SS and RR. For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR. In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of resistance=0.7 which sets the fitness of the SR closer to RR than SS. Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}

library(resistance) 
plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.7, dominance_cost=0.7, cost=0.3 )

```

```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 2. Example of rotation and sequence when no resistance fitness costs or dispersal link to untreated refugia."}

# revisit how best to do model runs from within the paper

linputs <- list()

linputs$n_insecticides <- 2
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$plot <-     FALSE

linputs$rot_interval=0
dfseq <- do.call(run_rot, linputs)
linputs$rot_interval=10
dfrot <- do.call(run_rot, linputs)

ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)

```

## References
<!-- now auto created here-->

### 2018-04-27 inst/documents/rotations_notes.Rmd
starting paper3 and developing sensi analysis

@@
 
+stackoverflow says this is a cygwin memory error and rebooting should fix it.
 eek
@@
 
-Please make sure you have the correct access rights
-and the repository exists.
+Just rebooted and that fixed it.
+
+*Seems that rotations favoured by higher costs, higher dominance of cost, lower start freq (longer to act), shorter rot interval (more changes)*
+
+*3/4/18*
+
+~ set up initial sensi analysis of 1000 runs
+~ 200 runs take < 30 mins
+
+~ seems that a lot of runs don't reach resistance thresholds by 400 generations, and this is with start-freqs set to 0.01, if I set lower start freqs then even more runs will be excluded.
+
+done ~ add ranges for variable params
+done ~ copy across grid.arrange from phaps-rotations-better
+
+around 79% of runs failing to make thresholds
+
+upped max gens to 500 and the lower exposure val to try to improve number making thresholds
+
+*4/4/2018*
+
+ooo with new ranges getting a few runs where there is a big advantage to rotation. Seems to be assoc with high dom_cos > 0.8
+Only c 1% of scenarios, 2 out of 200
+
+# scenario 137 sum generations deployed insecticide under thresh rotation: 500 sequence: 258
+run_rot(n_insecticides=3, cost=0.026, start_freqs=0.032, rot_interval=10, eff=0.5,
+dom_sel=0.36, dom_cos=0.82, rr=0.32, expo_hi=0.73, coverage=1,
+migration=0, max_gen=500, min_rwr_interval=5)
+
+# scenario 143 sum generations deployed insecticide under thresh rotation: 500 sequence: 126
+run_rot(n_insecticides=2, cost=0.049, start_freqs=0.064, rot_interval=0, eff=0.71,
+dom_sel=0.089, dom_cos=0.87, rr=0.9, expo_hi=0.65, coverage=1,
+migration=0, max_gen=500, min_rwr_interval=5)
+
+In 137 even the rounding makes the rotation last slightly less long (e.g. c400 rather than 500)
+
+start writing code to calc difference between sequence and rotation ready to put that into PRCC.
+
+currently i have repeated following rows for sequence and rotation with identical input params
+
+i probably want to replace this with one row and columns for gens_rot and gens_seq
+
+new version one row per scenario :
+
+#advantage to rotation despite dom_cos=0.23, cost=0.02
+#scenario 153 sum generations deployed insecticide under thresh rotation: 406 sequence: 289
+run_rot(n_insecticides=4, cost=0.023, start_freqs=0.019, rot_interval=28, eff=0.92,
+dom_sel=0.019, dom_cos=0.23, rr=0.63, expo_hi=0.75, coverage=1,
+migration=0, max_gen=500, min_rwr_interval=5)
+
+Looking at some more of literature :
+
+From Mallet 1989 :
+Since a naive strategy - of using insecticides in turn until resistance evolves to each - gives almost the same total control period as a planned rotation23, rotations can hardly even be defined as management.
+However, in some cases rotations may be the only practical way of limiting resistance to a single favoured compound 23 27.
+
+23 is curtis 1987 a book chapter I couldn't find.
+
+Curtis 1993 has some good quotes and references about rotations.
+It covers the oncheriasis control program which is taken by some to be evidence in favour of rotations
+Fig 3 & 4 is lab data comparing sequence & rotation, very similar to our plots, I should definitely cite it.
+
+"A programme such as the OCP differs from these simple models in that more than two insecticides are available for use and a given compound can therefore be out of use for more generations than it is in use."
+
+I hadn't thought of that. Having more than 2 insecticides allows each to be out of use for more time than it is in use. 
+So if the increase in resistance while in use was the same as the decrease when not in use then a rotation could be sustainable in the long term.
+However it is likely that resistance will increase faster in presence than it will decrease in absence (I think Curtis says that too).
+
+Curtis cites how his 1987 work has already included effect of modifier genes, and shows it here in Fig 6.
+"Supporters of prearranged rotations usually tacitly or explicitly assume that the general fitness due to resistance genes will improve if they are exposed to selection for many generations.
+Curtis (1987) simulated the behaviour of a modifer gene which could have this effect and found that it would only have an appreciable influence if one allowed resistance gene frequencies to become very high before switching insecticides.""
+
+"One possible objection to the reactive system is that it requires an efficient and
+comprehensive system for testing for resistance to avoid ‘overshooting’ and
+getting stuck as in the dotted line in Fig. 6. However, running a prearranged
+rotation without a monitoring system would mean that one would have no way
+of knowing if the system was still working."
+
+on p16 there is also interesting consideration of what expected in a mixture if one insecticide decays before another.
+
+It's negative about rotations :
+'Except where there are seasonal/environmental reasons for a pre-arranged rotation (as in the OCP), we do not find that such rotations are likely to have any advantage over well-managed reactive rotation, and we do not consider that the latter can validly be called a 'strategy' for resistance management.
+The use of a suitable insecticide mixture seems more promising in the light of the limited data so far available on resistance in relation to insecticide impregnated bednets.'
+
+I would like to get the Curtis 1987 paper
+Curtis CF. 1987. Genetic aspects of selection for resistance. In Ford MG, Holloman DW, Khambay BPS,
+Sawicki RM, eds. Combating resistance to xenobiotics. Chichester, U.K.: Ellis Horwood, 15e161.
+
+but couldn't find and probably adds little
+
+*5/4/2018*
+added progress counter to sensi_an_rotations after 2000 runs failed to finish overnight
+also now save the output every 200 runs
+
+100 scenarios take < 10 mins
+2000 scenarios should take 20*10 mins, 200 mins < 2 hours
+but seems that RAM fills up that may slow it down
+
+got 1400 runs, PC ground to a halt.
+
+No very clear pattern for high difference runs.
+Small positice effect with cost & dom_cos, small -ve effect of dom_sel
+
+*9/4/2018*
+turned off plotting in sensi_an_rotations
+10,000 runs took c 100 minutes (1hr20)
+
+*27/4/2018*
+
+First draft of paper3
+
+trying to get references to work without having to copy the .bib file from resistance
+
+C:/rsprojects/resistance/inst/documents/paper2/
+
+Getting filter pandoc-citeproc' had status 83 Execution halted
+https://stackoverflow.com/questions/36826916/pandoc-citeproc-error-83-with-rmarkdown-file
+
+removed all '-' from citation keys in mendeley, didn't fix
+
+Temporary solution of cutting out all refs except 2 from the .bib file fixed it.
+Need to go throught the bib file and look for error.
+Can I reduce size of the bib file to just the IR papers ? What did I do for ecocrop ?
 
@@
 
-~ look more at phaps-rotations-better.Rmd
-
-~ seems that min_rwr_interval explains the plots looking different but not the measured gens_under_thresh
+~ can I add figures from specific sensi_an runs ? or safer to run the model for set runs from the paper rmd.
+
+~ how to sort library.bib ? Have a much reduced copy in rotations folder now.
+
+~ change random seed and run another 1400 runs
+~ start PRCC. 
+~ can I try a predictive model to predict rot_minus_seq from inputs ?
+
+think about killer plot to display difference between rot & seq generations
+
+be careful that min_rwr_interval doesn't come back to bite me. Ian wanted it different, i.e. not going back to an insecticide that has been used recently ...
+
+~ sensi_an_rotations1.Rmd
+
+~ *decide on min_rwr_interval for sensi analysis
+~ decide on max_gen, what was it for paper1 ?
+~ one generation per month, 12 generations per year, 500 gens is ~40 years
 
@@
 ~~ update R
-
-~ sensi_an_rotations1.Rmd
 
@@
 
+# good refs for paper
+resistance : malERA2017_updated-research-agenda-resistance-in-malaria.pdf
+
+"What are the benefits of insecticide rotations, mixtures, or spatial mosaics of different compounds?
+What is the impact of adding nonpyrethroid IRS where LLINs are already deployed at high coverage and quality?
+When should new insecticides be adopted?
+What is the ideal rotation period or mosaic configuration?
+How many insecticide classes are needed for effective rotation or mosaic strategies?
+Despite the absence of data to answer these questions, some countries have already developed operational frameworks for resistance management that could be adopted by other programmes.""
+
+The malERA Refresh Consultative Panel on Insecticide and Drug Resistance (2017) malERA: An updated research agenda for insecticide and drug resistance in malaria elimination and eradication.
+PLoS Med 14(11): e1002450.
+https://doi.org/10.1371/journal.pmed.1002450
+
 # state of play
@@
 ~~  produces an output dataframe containing both inputs and outputs
+~~ saves outputs to df_res_all_rot.rda
 
@@
 
+# useful
+~ one generation per month, 12 generations per year, 500 gens is ~40 years
+
 # funding opportunities

### 2018-04-27 inst/documents/sensi_an_rotations1.Rmd
starting paper3 and developing sensi analysis

@@
 
-Starting question : is there any benefit to rotations when there are no costs and no migration to an untreated area ?
+10,000 runs comparing rotations and sequences.
 
-Test 1. No refugia (i.e. Coverage=100%) or fitness costs. Would expect all policies to be equivalent.
+```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
+#eval=FALSE just to temporarily hide this text
 
-Test 2. Dominance of resistance and fitness costs both set to 0.5.
-I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles.
+# Starting question : is there any benefit to rotations when there are no costs and no migration to an untreated area ?
+# 
+# Test 1. No refugia (i.e. Coverage=100%) or fitness costs. Would expect all policies to be equivalent.
+# 
+# Test 2. Dominance of resistance and fitness costs both set to 0.5.
+# I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles.
 I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same.
 I’m probably wrong but it’s a nice test…..
-
-Testing sensitvity analysis to answer question 1.
+# 
+# 
+# Testing sensitvity analysis to answer question 1.
 Are sequences and rotations equivalent when coverage 100% and no fitness costs.
@@
 
-\pagebreak
+#\pagebreak
+```
 
@@
 
-n_scenarios <- 200
+# set random seed so that results can be reproduced
+#set.seed(1)
+set.seed(2)
+max_gen         <- 500
+plot_scenarios  <- FALSE
+n_scenarios     <- 10000
+already_ran     <- TRUE
 
+# if already done runs, just load data for plots
+if (already_ran)
+{
+  #C:/rsprojects/rotations/inst/documents/
+
+  load("df_res_all_rot10000.rda")
+
+}else
+{
   df_res_all <- data.frame()
@@
   {
+    #message should output to the R markdown console
+    if (i%%500 == 1) message("scenario ",i," of ",n_scenarios," ",Sys.time())
+
     # inputs (could also be saved to a dataframe first & extracted one row at a time)
@@
     #linputs$cost <- 0 #variable cost below
-  linputs$max_gen <- 500
+    linputs$max_gen <- max_gen
 
@@
 
-  #random rot interval integer
-  linputs$rot_interval <-   sample( 5:50, 1 )  #beware set replace=T if more than 1
-
     # create null objects used plater for plotting
@@
     # to compare sequence to rotation run all the scenarios above for both seq & rot 
-  # beware has to be 1:0 because interval is overwritten for seq
-  for(rot_or_not in 1:0)
+    for(rot_or_not in 0:1)
     {  
@@
       if (rot_or_not == 0) linputs$rot_interval <- 0
+      else   #random rot interval integer
+          linputs$rot_interval <-   sample( 5:50, 1 )  #beware set replace=T if more than 1
 
@@
       # summarise results per scenario
-    # NOTE this is also currently calc at end of run_rot() here or other post-process is probably better
-    # and put in the output dfres, but the value has to be rpeated for every generation 
-    res <- dfres %>%
-      # only assess in control areas not in refugia  
-      dplyr::filter(active_or_refuge=='active') %>%
-      group_by(resist_gene) %>%
-      # for all insecticides in all generations  
-      # summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
-      # summarise(mean_gens_under50 = mean(gens_under50)) %>%
-      # just for deployed insecticides 
-      summarise(gens_dep_under50 = sum(resistance < 0.5 &
-                                       #finds insecticide in use = this one
-                                       resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
-      summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
-      unlist()
+      # note this is also currently calc at end of run_rot() here is probably better
 
-    # create dataframe from inputs and outputs
-    # first add scenario column onto results
+      res <- gens_under_thresh(dfres)
+
+      # to plot results for checking
+      if ( linputs$rot_interval == 0 )
+      {
+        gens_seq <- as.numeric(res)
+      } else
+      {
+        gens_rot <- as.numeric(res)
+      }
+
+      if (plot_scenarios)
+      {
+        # make plot
+        ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+
+        # to plot results for checking
+        if ( linputs$rot_interval == 0 )
+        {
+          ggseq <- ggres
+          gens_seq <- as.numeric(res)
+        } else
+        {
+          ggrot <- ggres
+          gens_rot <- as.numeric(res)
+        }
+      }
+
+    # TODO I may want to concat all dfres into one object (adda column for scenario num) to save for later analysis
+    # e.g. if I wanted to calculate average resistance frequencies over time in addition to gens below threshold
+
+    } #end rot_or_not loop
+
+    # make an output row for each scenario with inputs, gens_rot, gens_seq
+    # need to make sure I retain the rot_interval for rot
+
+    # add scenario id column onto results
     res$scenario <- i
-    df_in_out <- as.data.frame(c(linputs, res))
+    df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq))
+    df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
 
@@
 
-    ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
-    gens_res <- as.numeric(gens_under_thresh(dfres))
-
-    # to plot results for checking
-    if ( linputs$rot_interval == 0 )
+    if (plot_scenarios)
     {  
-      ggseq <- ggres
-      gens_seq <- gens_res
-    } else
-    {
-      ggrot <- ggres
-      gens_rot <- gens_res
-    }
-
-  } #end rot_or_not loop
-
-  #printing input values
-  #print(unlist(linputs))
-
-  # try to plot input vals instead
+      # plot input vals
       # extract out those between 0 & 1
@@
 
-} #end scenario loop
+    }
+
+    #I could save partial outputs every 200 scenarios
+    #just in case doesn't make to end
+    if (i%%1000 == 0)
+    {
+      #df_res_all$rot_minus_seq <- df_res_all$gens_rot - df_res_all$gens_seq
 
@@
       save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))    
+    }
+
+  } #end scenario loop
+
+}
+
+```
+
+# proportion of scenarios not reaching resistance thresholds
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+# check proportion of scenarios not reaching thresholds by max_runs for both seq & rot
+
+    nruns_fail <- df_res_all %>% 
+                  filter(gens_seq == max_gen & gens_rot == max_gen) %>%
+                  summarise(n=n()) %>% 
+                  unlist()
+
+    cat('proportion runs not reaching thresholds : ', nruns_fail/nrow(df_res_all) )
+
+    # remove runs where thresholds not reached for seq & rot 
+    df_res_all <- df_res_all %>% 
+                  filter(gens_seq != max_gen | gens_rot != max_gen)
+
+    nruns_same <- df_res_all %>% 
+                  filter(gens_seq == gens_rot) %>%
+                  summarise(n=n()) %>% 
+                  unlist()
+
+    cat('proportion runs (that did reach threshold) where rotations and sequence the same  : ', nruns_same/nrow(df_res_all) )
+
+    nruns_rot_20pc_better <- df_res_all %>% 
+                  filter(rot_minus_seq/gens_seq >= 0.2 ) %>%
+                  summarise(n=n()) %>% 
+                  unlist()
+
+    cat('proportion runs (that did reach threshold) where rotations >20 % better than sequence  : ', nruns_rot_20pc_better/nrow(df_res_all) )
+
+```
+
+# difference between rotations and sequences across all runs
+
+## mostly (~67%) no difference but infrequent where rotations better
+## ~ 7% of runs, rotation at least 20% better than sequence
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+    print( ggplot(df_res_all, aes(x=rot_minus_seq)) +
+
+             geom_histogram(binwidth=10) +
+             xlab('rot_minus_seq : rotation generations minus sequence') +
+             ylab('count (sqrt scale to show low values)') +
+             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+             theme_minimal() +
+             theme(panel.grid.minor.y=element_blank()))
 
@@
 
+# Examples of runs where rotations > 20% better than sequences
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=3.5}
+
+scenarios_to_plot <- c(624,2167,602,4028,5771)
+
+for(scenario in scenarios_to_plot)
+{
+  cat(scenario,"\n")
+
+  linputs <- as.list( df_res_all[scenario, 1:15] )
+
+  # run one scenario
+  dfres <- do.call(run_rot, linputs)
+
+  rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=TRUE)
+
+  #now run as sequence
+  linputs$rot_interval <- 0
+  dfres <- do.call(run_rot, linputs)
+
+  rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=TRUE)
+
+  dfi <- data.frame(t(as.data.frame(linputs)))
+  names(dfi) <- 'value'
+  dfi$inputs <- rownames(dfi)
+
+  #take out columns not between 0 & 1
+  dfi <- filter(dfi, !inputs %in% c('max_gen','rot_interval','n_insecticides','plot','min_rwr_interval'))
+  #also remove coverage & migration both set to 0 for now
+  dfi <- filter(dfi, !inputs %in% c('coverage','migration'))
+
+  ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
+            geom_point() +
+            theme_minimal() +
+            ylim(0,1) +
+            theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))
+  plot(ggins)
+
+}
+
+```
+
+#How does the difference between rotations and sequences respond to model inputs ?
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
+
+# go through selected inputs and plot them on x with rot_minus_seq on y 
+# initially with 200 reps doesn't show anything obvious, try it with 2000 overnight runs
+in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')
+
+y <- 'rot_minus_seq'
+
+for(i in in_cols)
+{
+
+    #plotnum <- plotnum + 1
+
+    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
+    #plotlist[[plotnum]] <- 
+    print( ggplot(df_res_all, aes_string(x=i, y=y)) +
+
+             geom_point(shape=1, size=1, alpha=0.5, show.legend=FALSE) +
+             theme_minimal() +
+             geom_smooth(linetype='dashed',size=0.5) )
+
+}
+
+```
+
+# PRCC
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+  library(sensitivity)
+
+# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
+# 
+# Arguments
+# X a data frame containing model input variables
+# y a vector containing the responses 
+# rank logical. If TRUE, the analysis is done on the ranks.
+# nboot the number of bootstrap replicates.
+# conf the confidence level of the bootstrap confidence intervals.
+
+#same as above
+in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')
+
+    x <- df_res_all[,in_cols]
+    y <- df_res_all['rot_minus_seq']
+
+    pcc_res <- pcc(x, y, rank=TRUE)
+
+    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
+    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
+    #plot(pcc_res)
+
+    #results are here I can probably rbind them together into a df that I can ggplot
+    #pcc_res$PRCC
+
+    to_plot <- pcc_res$PRCC
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    to_plot$inputs <- rownames(to_plot)
+
+    axis_text_x <- element_text(angle = 45,hjust = 1, vjust = 1)
+    height <- 0.9
+
+    gg <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
+             geom_point(shape=1, colour='red') +
+             theme(axis.text.x = axis_text_x) +
+             geom_hline(yintercept = 0, linetype=3) +
+             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
+             ylim(-1,1) +
+             #ggtitle(paste(strat_name[strategy_num])) +
+             xlab(NULL)
+
+    plot(gg)
+
+  #to put back in some ref gridlines that are removed by cowplot
+  #theme_set(theme_gray())
+  #theme_set(theme_bw())
+
+  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
+  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.6), labels='AUTO')
+
+```
+
+# Runs where rotations are >20% better (in red) are associated with higher dominance of cost and lower dominance of selection
+
+# but that does not guarantee, must be interaction with other inputs too.
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
+
+#'cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs'
+
+#'cost','dom_cos','dom_sel'
+
+y <- 'rot_minus_seq'
+
+ggplot(df_res_all, aes(x=dom_sel, y=dom_cos)) + 
+     geom_point(shape=3, size=1, alpha=0.2, show.legend=FALSE) +
+     theme_minimal() +
+     geom_point(data=filter(df_res_all, rot_minus_seq/gens_seq >= 0.3 ), shape=1, size=1, col='red', alpha=0.8)
+
+# looking at modelling it
+# library(mgcv)
+# mod_lm2 <- gam(rot_minus_seq ~ cost + dom_cos + dom_sel, data=df_res_all)
+# summary(mod_lm2)  
+# 
+# mod_gam2 <- gam(rot_minus_seq ~ s(cost) + s(dom_cos) + s(dom_sel) + s(rr) + s(expo_hi) + s(male_expo_prop) + s(eff) + s(start_freqs), data=df_res_all)
+# 
+# summary(mod_gam2)  
+# #Deviance explained = 19.5%
+# 
+# mod_gam3 <- gam(rot_minus_seq ~ te(cost,dom_cos,dom_sel), data=df_res_all)
+# summary(mod_gam3)
+
+#actually decision trees may be useful
+
+```
+
+```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
+#eval=FALSE tmp because changes above will break this
 # summary plot of all runs, would show any differences between sequence & rotations as non horizontal lines
-
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
 
@@
 
+#it was useful for me but now that I've moved to one row per scenario it won't work !!
+
     gg <- ggplot( df_res_all, aes_string(x='rot_interval',
@@
     plot(gg)
-
-# check proportion of scenarios not reaching thresholds by max_runs
-
-    nruns_fail <- df_res_all %>% 
-                  filter(tot_gens_dep_under50 == linputs$max_gen) %>%
-                  summarise(n=n()) %>% 
-                  unlist()
-
-    cat('proportion runs not reaching thresholds : ', nruns_fail/(2*n_scenarios) ) #2* beacuse eeach scenario run for seq & rot
 

### 2018-04-30 inst/documents/paper3/paper3_rotations.Rmd
running sensi analyses writing paper3

@@
   outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
+
+  # load simulation outputs for some plots
+  # beware that each object is called df_res_all
+  # some plots are created by directly running the model
+  load("df_res_all_rot_nocost_nomig1000.rda") #nocost nomig  
+  df_nocost_nomig <- df_res_all
+  load("df_res_all_rot10000.rda") #cost
+  df_cost <- df_res_all
+  load("df_res_all_rot_mig1000.rda") #mig
+  df_mig <- df_res_all
+
 ```
@@
 The simpler implementation allows the model to be run for an unlimited number of insecticides.
+The algebra behind the model is described in the supplementary information.
 
@@
 
-Model inputs effect fitness as shown in Figure 1.
+Within the model inputs effect fitness as shown in Figure 1, differences in fitness lead to the changes in allele frequencies over time.
 
@@
 
-To generate a decline in resistance over time either fitness costs or a dispersal link to an untreated area are required.
+To generate a decline in resistance over time when an insecticide is not being used either fitness costs or a dispersal link to an untreated area are required.
 
-*talk about model outputs, i.e. generations deployed below threshold*
-*try to come up with brief name to make it easier to refer to it later*
+#### Assessing relative performance of insecticide use strategies
+
+It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative.
+For this we summed the number of generations when an insecticide was deployed and the resistance allele frequency for that insecticide was below 0.5.
+This allowed comparison between strategies using a variable number of insecticides.
+Previously time-to-resistance, namely the number of generations it takes to reach a defined resistance threshold (usually a resistance allele frequency of 0.5), was used.
+Using time-to-resistance worked when considering just two insecticides, but when the number of insecticides is a variable it is not sufficient.
 
@@
 
-### no resistance fitness costs or dispersal link to untreated refugia
+#### Results1. no resistance fitness costs or dispersal from untreated refugia
 
@@
 For rotations resistance frequencies step upwards when insecticides are in use and remain at a plateau when they are not.
-This results in
 
-### fitness costs of resistance added
+When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold.
+
+#### Results2. fitness costs of resistance added
+
+With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold.
+
+In contrast to the previous section 7% of model runs generated an advantage of greater than 20% for rotations over sequences.
+
+#### Results3. dispersal from untreated refugia added
 
 ## Discussion
+
+Comparing time below resistance thresholds for rotations and sequences, the model shows :
+
+1. rotations and sequences the same when no costs or dispersal
+2. with resistance costs, 67% of runs still the same for rotations and sequences, but 7% of runs have > 20% advantage for rotations
+3. with dispersal, ...
+
+In very rare circumstances rotations can lead to the resistance frequencies for a group of insecticides being kept below resistance thresholds for a very long period of time (and substantially longer than for sequential use).
+Is it possible that these rare circumstances where parameter values are 'just-right' could be implemented operationally ?
+
+### Caveats
+
+The model does not (although it could) include the potential effects of modifier genes ameliorating the costs of resistance alleles.
+It has been suggested that rotations could be favoured as a strategy because they keep resistance frequences lower and restrict selection pressure for modifier genes [@WHO2012].
+However, there is little evidence to support the existence of such modifier genes in insectide resistance [].
+Even with the existence of modifier genes it is unlikely that they would provide much of an advantage to rotations in the scenarios we describe.
+In the modelled scenarios the use of an insecticide is stopped when the resistance frequency for that insecticide reaches 0.5.
+Thus in all scenarios resistance frequencies do not remain at high levels for long.
+In this situation it seems unlikely that modifiers would be selected for.
+
+### Conclusions
+
+In their effect on the evolution of insecticide resistance, rotations and sequences are most often the same.
+Infrequently, in less than 10% of model runs set between plausible limits, there are substantial benefits to the rotation strategy over a sequence.
+There was naver a substantial benefit of a sequence over a rotation.
+Whether or not insecticide rotations are predicted to offer sizeable benefits over sequences is dependent on coarse and fine scale issues about the nature of insecticide resistance that are yet to be agreed upon.
+The model shows that resistance costs and/or dispersal from untreated refugia are necessary to provide an evolutionary advantage to rotations.
+However even when these requirements are satisfied particular combinations of inputs are required to generate the advantage of rotations.
+
+The model only compares rotations and sequences in terms of their effect on the evolution of resistance. There are other operational reasons for why a sequence or a rotation may be favoured.
+
+## TODO
+*try to come up with brief name for model output to make it easier to refer to it later*
 
@@
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1.
-[todo need to modify caption from MJ paper, start by saying that only in left panel during application of the insecticide] The effect of model inputs on the fitness of genotypes for a single insecticide.
-Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
-Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of resistance determines how the fitness for SR lies between that of SS and RR.
-For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of resistance=0.7 which sets the fitness of the SR closer to RR than SS.
-Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}
-
-library(resistance) 
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.7, dominance_cost=0.7, cost=0.3 )
-
-```
-
-```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 2.
-Example of rotation and sequence when no resistance fitness costs or dispersal link to untreated refugia."}
+```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1.
+Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia.
+The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached.
+In this example the rotation and the sequence are stopped at the same time.
+For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again.
+For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and remains at the constant level while not.
+There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use.
+The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold.
+All other inputs are kept constant between the rotation and the sequence."}
 
@@
 
-linputs$n_insecticides <- 2
+linputs$n_insecticides <- 3
 linputs$coverage <- 1
@@
 
-ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
-ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
 library(gridExtra)
@@
 
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2.
+Comparing the increase in resistance frequencies over time for a rotation and a sequence when there are resistance fitness costs.
+The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached.
+In this example the rotation lasts longer than the sequence.
+For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step but in contrast to the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again.
+For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines  while not.
+There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use.
+The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold.
+All other inputs are kept constant between the rotation and the sequence."}
+
+# revisit how best to do model runs from within the paper
+
+linputs <- list()
+
+linputs$n_insecticides <- 3
+linputs$coverage <- 1
+#linputs$migration <- 0.1
+linputs$cost <-     0.05 
+linputs$max_gen <-  200
+linputs$expo_hi <-  0.6
+linputs$eff <-      0.6
+linputs$dom_sel <-  1
+linputs$dom_cos <-  1
+linputs$plot <-     FALSE
+
+linputs$rot_interval=0
+dfseq <- do.call(run_rot, linputs)
+linputs$rot_interval=10
+dfrot <- do.call(run_rot, linputs)
+
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+library(gridExtra)
+grid.arrange(ggrot, ggseq, ncol=1)
+
+```
+
+```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3.
+Comparing the increase in resistance frequencies over time for a rotation and a sequence with dispersal from untreated refugia.
+The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached.
+In this example the rotation and the sequence last the same time.
+In both cases the resistance frequencies in the treated areas (red lines) increase when an insecticide is in use.
+Resistance frequencies in the untreated areas (blue lines) increase after those in the treated areas as a result of dispersal from the treated areas.
+When an insecticide stops being used the resistance frequency in the treated area declines due to dispersal from the untreated area.
+For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and consistent with the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again.
+For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines while not.
+There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use.
+The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold.
+All other inputs are kept constant between the rotation and the sequence."}
+
+# revisit how best to do model runs from within the paper
+
+linputs <- list()
+
+linputs$n_insecticides <- 3
+linputs$coverage <- 0.5
+linputs$migration <- 0.1
+linputs$cost <-     0 
+linputs$max_gen <-  200
+linputs$expo_hi <-  0.6
+linputs$eff <-      0.6
+linputs$dom_sel <-  1
+linputs$dom_cos <-  1
+linputs$plot <-     FALSE
+
+linputs$rot_interval=0
+dfseq <- do.call(run_rot, linputs)
+linputs$rot_interval=10
+dfrot <- do.call(run_rot, linputs)
+
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=TRUE, logy=TRUE, plot=FALSE)
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=TRUE, logy=TRUE, plot=FALSE)
+library(gridExtra)
+grid.arrange(ggrot, ggseq, ncol=1)
+
+```
+
+```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4.
+Effect of cost and dispersal on the difference between rotations and sequences"}
+
+  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=rot_minus_seq)) +
+
+             geom_histogram(binwidth=10) +
+             xlim(-20,400) +
+             xlab('no cost or dispersal : rotation generations minus sequence') +
+             ylab('count (sqrt scale to show low values)') +
+             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+             theme_minimal() +
+             theme(panel.grid.minor.y=element_blank())
+
+  gg_cost <- ggplot( df_cost, aes(x=rot_minus_seq)) +
+
+             geom_histogram(binwidth=10) +
+             xlim(-20,400) +
+             xlab('cost : rotation generations minus sequence') +
+             ylab('count (sqrt scale to show low values)') +
+             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+             theme_minimal() +
+             theme(panel.grid.minor.y=element_blank())
+
+  gg_mig <- ggplot( df_mig, aes(x=rot_minus_seq)) +
+
+             geom_histogram(binwidth=10) +
+             xlim(-20,400) +
+             xlab('dispersal : rotation generations minus sequence') +
+             ylab('count (sqrt scale to show low values)') +
+             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+             theme_minimal() +
+             theme(panel.grid.minor.y=element_blank())
+
+  grid.arrange(gg_nocost_nomig, gg_cost, gg_mig, ncol=1)
+
+```
+
+```{r FigX, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure X.
+[todo need to modify caption from MJ paper, start by saying that only in left panel during application of the insecticide] The effect of model inputs on the fitness of genotypes for a single insecticide.
+Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
+Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
+For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of resistance determines how the fitness for SR lies between that of SS and RR.
+For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
+In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of resistance=0.7 which sets the fitness of the SR closer to RR than SS.
+Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}
+
+library(resistance) 
+plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.7, dominance_cost=0.7, cost=0.3 )
+
+```
+
+\pagebreak
+
 ## References

### 2018-04-30 inst/documents/rotations_notes.Rmd
running sensi analyses writing paper3

@@
 
+# 30/4/18
+
+100 runs with no cost all had identical times for rotations and sequences
+
+from Rmd got error when tried 1000
+worked from console
+
+With no cost and no migration, majority of runs the same. Just 3/1000 seq better than rot by < 3 generations.
+tst <- filter(df_res_all, rot_minus_seq!=0)
+
+Now trying 1000 runs with coverage and dispersal set between 0.1 and 0.9
+
+Seems that less benefit to rotations with dispersal rather than costs.
+
 # TODO
 
+~ run 10k runs of dispersal overnight
+~ try histograms on log scale ?
+~ try decision tree to predict rot_minus_seq from inputs
+
+~ send MS to Ian before 11
+
+~ modify sensi_an_rotations so that I have a reproduceable record of a) no cost or dispersal b)cost only c)dispersal only, d) cost & dispersal
+
 ~ can I add figures from specific sensi_an runs ? or safer to run the model for set runs from the paper rmd.
@@
 ~ how to sort library.bib ? Have a much reduced copy in rotations folder now.
-
-~ change random seed and run another 1400 runs
-~ start PRCC. 
-~ can I try a predictive model to predict rot_minus_seq from inputs ?
 

### 2018-04-30 inst/documents/sensi_an_rotations1.Rmd
running sensi analyses writing paper3

@@
 plot_scenarios  <- FALSE
-n_scenarios     <- 10000
+n_scenarios     <- 1000 # try 1000 runs with no cost
+#already_ran     <- FALSE
 already_ran     <- TRUE
@@
 
-  load("df_res_all_rot10000.rda")
+  #load("df_res_all_rot10000.rda")
+  load("df_res_all_rot_mig1000.rda")
 
@@
     #message should output to the R markdown console
-    if (i%%500 == 1) message("scenario ",i," of ",n_scenarios," ",Sys.time())
+    if (i%%200 == 1) message("scenario ",i," of ",n_scenarios," ",Sys.time())
 
@@
     ## constant inputs ##
-    linputs$coverage <- 1
-    linputs$migration <- 0
+    #linputs$coverage <- 1
+    #linputs$migration <- 0
     #linputs$cost <- 0 #variable cost below
@@
     ## variable inputs ##
-    linputs$cost <-           runif(1, min=0,   max=0.1) 
+    #linputs$cost <-           runif(1, min=0,   max=0.1) 
+    # try 1000 runs with no cost
+    linputs$cost <-           0 
     linputs$expo_hi <-        runif(1, min=0.4, max=0.9)  #min=0.1, max=0.9
@@
     linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
+    # dispersal (initially with cost set to 0)
+    linputs$coverage <-       runif(1, min=0.1, max=0.9)
+    linputs$migration <-      runif(1, min=0.1, max=0.9)
 
-    # create null objects used plater for plotting
+    # create null objects used later for plotting
     ggrot <- ggseq <- NULL
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=3.5}
+```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=3.5}
+#eval false because smaller runs don't have these
 

### 2018-05-02 inst/documents/paper3/paper3_rotations.Rmd
paper3 first draft

@@
   df_cost <- df_res_all
-  load("df_res_all_rot_mig1000.rda") #mig
+  load("df_res_all_rot_mig10000.rda") #mig 10k
   df_mig <- df_res_all
+  load("df_res_all_rot_cost_mig10000.rda") #costmig 10k
+  df_cost_mig <- df_res_all
 
@@
 
+In a rotation the insecticide used is changed on a set time interval irrespective of resistance levels.
+Sequential use, in contrast, denotes the practice of using an insecticide until resistance to it reaches a certain threshold and then switching to another.
+Different rotation time intervals can be considered, e.g.
+in vector control for Malaria, Indoor Residual Spraying may be repeated every year (around 10 anopheles mosquito generations) and Insecticide Treated Nets may be replaced every two years (around 20 generations).
+Multiples of these intervals could be used as the rotation interval.
+Multiple generational rotations like this contrast with the situation in agriculture where a rotation is sometimes more tightly defined as having an interval of a single generation [@Sudo2017].
+
 Here a modern modelling approach is described to allow an assessment of the potential benefits, in terms of the evolution of resistance, of rotations over sequences.
@@
 A population genetic model was developed simulating changes over time in the frequencies of resistant and susceptible alleles in response to fitness differences.
-The model uses a similar approach to that described in [@Levick2017,@South2018] to compare insecticide mixtures and sequences.
+The model uses a similar approach that has been used to compare insecticide mixtures and sequences [@Levick2017,@South2018].
 The implementation is simpler here because for rotations and sequences there is only a need to follow one insecticide at a time.
@@
 
-Sequences : one insecticide is used until resistance threshold frequency (0.5 in this case) is reached, then switch to new insecticide. 
-Rotations : use insecticide for set time, switch to other.
+1. Sequences : one insecticide is used until resistance threshold frequency (0.5 in this case) is reached, then switch to a new insecticide.
 
-In both cases continue until no more insecticides below the resistance threshold remain.
+2. Rotations : use insecticide for set time, switch to another.
+
+In both cases the choice of the new insecticide to use was made by going through the list of available insecticides and choosing the next one for which resistance was below the threshold.
+Thus there is the potential to switch back to insecticides that had been used earlier in a model run if resistance had not reached the threshold or if resistance had declined below the threshold during a period when not in use.
+
+In both cases a simulation run continued until no more insecticides remained below their resistance thresholds.
 
@@
 The model was run under four main scenarios :
+
 1. no resistance fitness costs or dispersal link to untreated areas
@@
 
+Example model runs are shown with no costs or dispersal (Fig 1), with just resistanace costs (Fig 2) and with just dispersal from untreated refugis (Fig 3).
+
 #### Results1. no resistance fitness costs or dispersal from untreated refugia
 
-When there are no resistance fitness costs or dispersal links to untreated refugia resistance frequencies do not decline when an insecticide is not in use.
+When there are no resistance fitness costs or dispersal links to untreated refugia resistance frequencies do not decline when an insecticide is not in use (Fig 1).
 Thus for sequences once resistance thresholds are reached the resistance frequencies remain at that level and the insecticide cannot be re-used.
@@
 
-When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold.
+When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold (Fig 4.1).
 
@@
 
-In contrast to the previous section 7% of model runs generated an advantage of greater than 20% for rotations over sequences.
+In contrast to the previous section 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).
 
 #### Results3. dispersal from untreated refugia added
+
+Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 4.3).
+
+#### Results4. costs and dispersal
+
+With both costs and dispersal the benefits of rotations are similar to when runs have costs alone (Fig 4.4).
+
+#### Runs where rotations have substantial advantage over sequences
+As shown in Fig 4.2 some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 5.
+
+A Partial Rank Correlation analysis identified that of the inputs, dominance of cost and dominance of selection had the greatest influence on the difference between rotations and sequences when resistance costs were included (Fig 6).
+A combination of high values of dominance of cost and low values for dominance of selection are required to give the greatest advantages to rotations, but do not guarantee this advantage (Fig 7).
 
@@
 1. rotations and sequences the same when no costs or dispersal
-2. with resistance costs, 67% of runs still the same for rotations and sequences, but 7% of runs have > 20% advantage for rotations
-3. with dispersal, ...
+2. with resistance costs, 67% of runs the same for rotations and sequences, but 6-7% of runs have > 20% advantage for rotations
+3. with dispersal from untreated refugia but no costs, small advantage of rotations but very few runs < 0.1% have >20% advantage.
 
-In very rare circumstances rotations can lead to the resistance frequencies for a group of insecticides being kept below resistance thresholds for a very long period of time (and substantially longer than for sequential use).
+In rare circumstances, when there are costs of resistance, rotations can lead to the resistance frequencies for a group of insecticides being kept below resistance thresholds for the length of the simulations where sequences run out of insecticides within around 100 generations.
 Is it possible that these rare circumstances where parameter values are 'just-right' could be implemented operationally ?
+The problem is that these rare circumstances rely on high costs of resistance, high dominance of these costs and low dominance of the selection for the resistance itself.
+These parameters are all not easy to measure and there isn't agreement on their likely values [@Kliot2012, @Bourguet2000].
+
+The importance of resistance costs for substantial evolutionary benefits of rotations over sequences points to the importance of the debate over the fitness costs of insecticide resistance e.g.
+[@Kliot2012].
+*extend this discussion of fitness costs*
 
@@
 In their effect on the evolution of insecticide resistance, rotations and sequences are most often the same.
-Infrequently, in less than 10% of model runs set between plausible limits, there are substantial benefits to the rotation strategy over a sequence.
-There was naver a substantial benefit of a sequence over a rotation.
+If costs of resistance are included, infrequently (less than 7% of model runs set between plausible limits) there are substantial benefits to the rotation strategy over a sequence.
+There was never a substantial benefit of a sequence over a rotation.
 Whether or not insecticide rotations are predicted to offer sizeable benefits over sequences is dependent on coarse and fine scale issues about the nature of insecticide resistance that are yet to be agreed upon.
-The model shows that resistance costs and/or dispersal from untreated refugia are necessary to provide an evolutionary advantage to rotations.
+The model shows that resistance costs are necessary to provide a substantial evolutionary advantage to rotations.
 However even when these requirements are satisfied particular combinations of inputs are required to generate the advantage of rotations.
@@
 ```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4.
-Effect of cost and dispersal on the difference between rotations and sequences"}
+Effect of cost and dispersal on the distribution of differences between rotations and sequences from model runs.
+The y axis is square root transformed to emphasise the infrequent runs where there is a significant benefit of rotations over sequences.
+4.1 when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results.
+4.2, 4.4 when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences.
+For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences.
+4.3 when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}
 
-  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=rot_minus_seq)) +
+  # remove runs where thresholds not reached for seq & rot
+  # with cost_mig ~ 65% of runs don't reach thresholds
+  # could get this func to output perecnt not reaching
+  rem_thresh_not_reached <- function(x) {x %>% filter(gens_seq != max_gen | gens_rot != max_gen)}
+
+  df_nocost_nomig <- rem_thresh_not_reached(df_nocost_nomig)
+  df_cost <- rem_thresh_not_reached(df_cost) #6.6
+  df_mig <- rem_thresh_not_reached(df_mig) #0.05
+  df_cost_mig <- rem_thresh_not_reached(df_cost_mig) #6.9
+
+  # add a column for percent difference between rot & seq
+  df_nocost_nomig <- mutate(df_nocost_nomig, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
+  df_cost <- mutate(df_cost, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
+  df_mig <- mutate(df_mig, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
+  df_cost_mig <- mutate(df_cost_mig, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
+
+  # calc % of runs where rotations > 20% better
+  pcent_runs_rot_20pc_better <- function(x){ x %>%
+                  filter(rot_minus_seq/gens_seq >= 0.2 ) %>%
+                  summarise(n=n()) %>% 
+                  #unlist() %>%
+                  mutate(pcent=n*100/nrow(x)) }
+
+  # pcent_runs_rot_20pc_better(df_nocost_nomig)
+  # pcent_runs_rot_20pc_better(df_cost)
+  # pcent_runs_rot_20pc_better(df_mig)
+  # pcent_runs_rot_20pc_better(df_cost_mig)
+
+  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=pcent_rot_adv)) +
 
@@
              xlim(-20,400) +
-             xlab('no cost or dispersal : rotation generations minus sequence') +
-             ylab('count (sqrt scale to show low values)') +
+             xlab('') +
+             ylab('count') +
+             ggtitle('1. no resistance costs or dispersal') +
              scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
@@
 
-  gg_cost <- ggplot( df_cost, aes(x=rot_minus_seq)) +
+  gg_cost <- ggplot( df_cost, aes(x=pcent_rot_adv)) +
 
@@
              xlim(-20,400) +
-             xlab('cost : rotation generations minus sequence') +
-             ylab('count (sqrt scale to show low values)') +
+             xlab('') +
+             ylab('count') +
+             ggtitle('2. with costs') +
              scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
@@
 
-  gg_mig <- ggplot( df_mig, aes(x=rot_minus_seq)) +
+  gg_mig <- ggplot( df_mig, aes(x=pcent_rot_adv)) +
 
@@
              xlim(-20,400) +
-             xlab('dispersal : rotation generations minus sequence') +
-             ylab('count (sqrt scale to show low values)') +
+             xlab('') +
+             ylab('count') +
+             ggtitle('3. with dispersal refugia') +
              scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
@@
 
-  grid.arrange(gg_nocost_nomig, gg_cost, gg_mig, ncol=1)
+  #gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_minus_seq)) + 
+  gg_cost_mig <- ggplot( df_cost_mig, aes(x=pcent_rot_adv)) +              
+             geom_histogram(binwidth=10) +
+             xlim(-20,400) +
+             xlab('percent benefit of rotation versus sequence') +
+             ylab('count') +
+             ggtitle('4. with costs and dispersal') +
+             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+             theme_minimal() +
+             theme(panel.grid.minor.y=element_blank())
+
+  grid.arrange(gg_nocost_nomig, gg_cost, gg_mig, gg_cost_mig, ncol=1)
+
+```
+
+```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 5.
+Example of a model run where there is a substantial benefit of rotations over a sequence.
+For the sequence resistance thresholds are reached for all insecticides in a little over 100 generations or 10 years.
+For the rotation resistance levels for all insecticides are maintained well below resistance thresholds for the length of the simulation (500 generations or 50 years)."}
+# ? are these run ids before or after the runs that don't reach thresholds have been removed ?
+# seems OK that they are after other runs removed
+#scenarios_to_plot <- c(624,2167,602,4028,5771)
+scenarios_to_plot <- c(2167)
+
+for(scenario in scenarios_to_plot)
+{
+  #cat(scenario,"\n")
+
+  linputs <- as.list( df_cost[scenario, 1:15] )
+
+  # run one scenario
+  dfres <- do.call(run_rot, linputs)
+
+  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+
+  #now run as sequence
+  linputs$rot_interval <- 0
+  dfres <- do.call(run_rot, linputs)
+
+  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+
+  grid.arrange(ggrot, ggseq, ncol=1)
+
+}
+
+```
+
+```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 6.
+Which model inputs have the greatest influence on the difference between rotations and sequences when costs alone are included ?
+Measured using Partial Rank Correlation Coefficients (PRCC).
+Dominance of cost and dominance of selection have the greatest influence."}
+
+library(sensitivity)
+
+# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
+# 
+# Arguments
+# X a data frame containing model input variables
+# y a vector containing the responses 
+# rank logical. If TRUE, the analysis is done on the ranks.
+# nboot the number of bootstrap replicates.
+# conf the confidence level of the bootstrap confidence intervals.
+
+#same as above
+in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')
+
+    x <- df_cost[,in_cols]
+    #y <- df_cost['rot_minus_seq']
+    y <- df_cost['pcent_rot_adv']
+
+    pcc_res <- pcc(x, y, rank=TRUE)
+
+    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
+    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
+    #plot(pcc_res)
+
+    #results are here I can probably rbind them together into a df that I can ggplot
+    #pcc_res$PRCC
+
+    to_plot <- pcc_res$PRCC
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    to_plot$inputs <- rownames(to_plot)
+
+    axis_text_x <- element_text(angle = 45,hjust = 1, vjust = 1)
+    height <- 0.9
+
+    gg <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
+             geom_point(shape=1, colour='red') +
+             theme(axis.text.x = axis_text_x) +
+             geom_hline(yintercept = 0, linetype=3) +
+             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
+             ylim(-1,1) +
+             #ggtitle(paste(strat_name[strategy_num])) +
+             xlab(NULL)
+
+    plot(gg)
+
+  #to put back in some ref gridlines that are removed by cowplot
+  #theme_set(theme_gray())
+  #theme_set(theme_bw())
+
+  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
+  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.6), labels='AUTO')
+
+```
+
+```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 7.
+Model runs with substantial benefits of rotations over sequences are associated with high dominance of cost and low dominance of selection.
+Dominace inputs are shown on the x and y axes.
+Small grey points indicate all model runs.
+Model runs in which rotations last greater than 20% longer than sequences are shown in colour with darker blue indicating those that have the greatest difference."}
+
+#'cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs'
+#'cost','dom_cos','dom_sel'
+
+y <- 'pcent_rot_adv'
+
+gg <- ggplot(df_cost, aes(x=dom_sel, y=dom_cos)) + 
+     geom_point(shape='.', size=1, alpha=0.2, show.legend=FALSE) +
+     theme_minimal() +
+     #geom_point(data=filter(df_cost, pcent_rot_adv >= 20 ), shape=1, size=1, col='red', alpha=0.8)
+     #shape=19 filled circle, 1=open circle
+     geom_point(data=filter(df_cost, pcent_rot_adv >= 20), aes(col=pcent_rot_adv), shape=19, size=1, alpha=0.9) +
+     viridis::scale_color_viridis(direction=-1) #discrete = TRUE)
+
+plot(gg)
 

### 2018-05-02 inst/documents/sensi_an_rotations1.Rmd
paper3 first draft

@@
 plot_scenarios  <- FALSE
-n_scenarios     <- 1000 # try 1000 runs with no cost
-#already_ran     <- FALSE
-already_ran     <- TRUE
+n_scenarios     <- 10000 # try 1000 runs with no cost
+already_ran     <- FALSE
+#already_ran     <- TRUE
 
@@
     ## variable inputs ##
-    #linputs$cost <-           runif(1, min=0,   max=0.1) 
-    # try 1000 runs with no cost
-    linputs$cost <-           0 
+    linputs$cost <-           runif(1, min=0,   max=0.1) 
+    #linputs$cost <-           0 
     linputs$expo_hi <-        runif(1, min=0.4, max=0.9)  #min=0.1, max=0.9

### 2018-05-08 inst/documents/paper3/pap3_rot_analyses.Rmd
paper3 writing

---
title: "pap3_rot_analyses : trial analyses for paper3"
output: pdf_document
#output: word_document
author: "Andy South"
date: "`r Sys.Date()`"
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(rotations)
library(ggplot2)
library(gridExtra)

outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"

  # load simulation outputs for some plots
  # beware that each object is called df_res_all
  # some plots are created by directly running the model
  load("df_res_all_rot_nocost_nomig1000.rda") #nocost nomig  
  df_nocost_nomig <- df_res_all
  load("df_res_all_rot10000.rda") #cost
  df_cost <- df_res_all
  load("df_res_all_rot_mig10000.rda") #mig 10k
  df_mig <- df_res_all
  load("df_res_all_rot_cost_mig10000.rda") #costmig 10k
  df_cost_mig <- df_res_all

```

```{r success, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}

#looking at classifying runs as to whether succeed_both (2), succeed_rot (1), fail_both (0). Then see what are the best predictors of these ?

# add a column for percent difference between rot & seq
df_cost <- mutate(df_cost, pcent_rot_adv = rot_minus_seq * 100/gens_seq)

df_cost <- mutate(df_cost, success_rot_seq = ifelse(gens_seq==max_gen & gens_rot==max_gen,'rot_seq',
                                               ifelse(gens_rot==max_gen,'rot',
                                               ifelse(gens_seq==max_gen,'seq', 'neither'        
                                                      ))))
# count(df_cost, success_rot_seq)
# 1         neither  5948
# 2             rot   209
# 3         rot_seq  3841
# 4             seq     2

# not very informative
# 2D plot dom_sel, dom_cos, facetted by whether neither succeed, rot only or both
# does show that rot_only and rot_seq more abundant at low dom_sel and high dom_cos  
gg <- ggplot(df_cost, aes(x=dom_sel, y=dom_cos, col=success_rot_seq)) + 
  geom_point(shape=1, size=1, alpha=0.5) +
  #geom_bin2d(bins=5) + #didn't look great
  facet_grid(~success_rot_seq) +
  viridis::scale_color_viridis(discrete = TRUE) #direction=-1)

plot(gg)

#effectiveness & exposure
gg <- ggplot(df_cost, aes(x=eff, y=expo_hi, col=success_rot_seq)) + 
  geom_point(shape=1, size=1, alpha=0.5) +
  #geom_bin2d(bins=5) + #didn't look great
  facet_grid(~success_rot_seq) +
  viridis::scale_color_viridis(discrete = TRUE) #direction=-1)

plot(gg)

```

```{r tree, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}
# have a quick go at a decision tree for chhosing between neither, rot or rot_seq

library(rpart)
library(rpart.plot)

#tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')

tree <- rpart::rpart("success_rot_seq ~ cost + expo_hi + male_expo_prop + eff + rr + dom_sel + dom_cos + start_freqs + n_insecticides + rot_interval", data = df_cost, method = 'class')

prp(tree, extra=1, varlen = 0, under=TRUE, 
    cex=0.6, round=0, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= " )

#initial tree results kind of make sense, but not very informative
#doesn't give rot_only as an output ? 
#cost is first branch
#odd that higher rr and expo_hi favour both working
#odd that dom_sel >0.15 favours both working ?

#try it with the 'successes' removed
df_cost2 <- filter(df_cost, success_rot_seq != 'rot_seq') #'neither'

tree <- rpart::rpart("success_rot_seq ~ cost + expo_hi + male_expo_prop + eff + rr + dom_sel + dom_cos + start_freqs + n_insecticides + rot_interval", data = df_cost2, method = 'class')

prp(tree, extra=1, varlen = 0, under=TRUE, 
    cex=0.6, round=0, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= " )

#dom_sel & dom_cos first branches but tree not very helpful

#with 'neither' removed lower cost & lower rr favour rot_only

```

```{r gen_trendlines, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}

# trying simple dot & trend plots for gens_rot & seq versus inputs
# work as expected when 'successes' not removed
# n_insecticides doesn't work because not enough points

inputs <- c('cost', 'expo_hi', 'male_expo_prop', 'eff', 'rr', 'dom_sel', 'dom_cos', 'start_freqs', 'n_insecticides', 'rot_interval')

for(input in inputs)
{
    y <- 'gens_rot'

    gg <- ggplot(df_cost, aes_string(x=input, y=y)) +

             geom_point(shape='.', size=1, alpha=0.2, show.legend=FALSE) + 
             geom_smooth(linetype='dashed',size=0.5)
    plot(gg)
}

# try it with the 'successes' removed
# doesn't look as good for cost
df_cost2 <- filter(df_cost, success_rot_seq != 'rot_seq')

for(input in inputs)
{
    y <- 'gens_rot'

    gg <- ggplot(df_cost2, aes_string(x=input, y=y)) +

             geom_point(shape='.', size=1, alpha=0.2, show.legend=FALSE) + 
             geom_smooth(linetype='dashed',size=0.5)
    plot(gg)
}

```

```{r threshold_violins, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}

# this could do it to show how the different results (neither, rot, or rot_seq)
# have different distributions of input values

inputs <- c('cost', 'expo_hi', 'male_expo_prop', 'eff', 'rr', 'dom_sel', 'dom_cos', 'start_freqs', 'n_insecticides', 'rot_interval')

for(input in inputs)
{

  print(ggplot(df_cost, aes_string(x='success_rot_seq',y=input, color='success_rot_seq')) + 
        #ylim(0,450) +
        #coord_cartesian( ylim=c(0, 350)) +
        geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = 1 ) +
        theme_minimal() +
        #this is an example of how you might facet
        #facet_grid(n_insecticides ~.) +  
        theme(panel.grid.minor.y=element_blank())) #,
        #           axis.text.x=element_blank())
        #geom_jitter(shape='.',height=0,width=0.2) )
}

```

```{r threshold_violins_facetted, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}

# I could maybe make the data longer and then facet by input
# think I would filter out, n_insecticides, start_freqs, male_expo_prop & rot_interval that don't show much rship
#TODO improve labels for these
inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos')
# something like this example from SO https://stackoverflow.com/a/48940750/1718356
#data_long <- tidyr::gather(data, key = type_col, value = categories, -col4)
#ggplot(data_long, aes(x = categories, fill = col4)) +
#  geom_bar() + 
#  facet_wrap(~ type_col, scales = "free_x")

#data_long <- tidyr::gather(data, key = type_col, value = categories, -col4)
#first just select the columns I want
df_cost_long <- select(df_cost,c(inputs,'success_rot_seq'))
#gather data to be longer
df_cost_long <- tidyr::gather(df_cost_long, key=input_name, value=input_value, -success_rot_seq)

#there were just an odd couple of results where seq better, remove them here or earlier
df_cost_long <- filter(df_cost_long, success_rot_seq != 'seq')

ggplot(df_cost_long, aes_string(x='success_rot_seq',y='input_value', color='success_rot_seq')) + 
        #ylim(0,450) +
        #coord_cartesian( ylim=c(0, 350)) +
        geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = 1 ) +
        theme_minimal() +
        #this is an example of how you might facet
        facet_wrap(~input_name, scales = "free_y") +  
        theme(panel.grid.minor.y=element_blank()) #,
        #           axis.text.x=element_blank())
        #geom_jitter(shape='.',height=0,width=0.2) )

```

### 2018-05-08 inst/documents/paper3/paper3_rotations.Rmd
paper3 writing

@@
                cache=FALSE,
-               fig.path='pap2figs/')
+               fig.path='pap3figs/')
 ```
@@
 
-A population genetic model was developed simulating changes over time in the frequencies of resistant and susceptible alleles in response to fitness differences.
-The model uses a similar approach that has been used to compare insecticide mixtures and sequences [@Levick2017,@South2018].
-The implementation is simpler here because for rotations and sequences there is only a need to follow one insecticide at a time.
+A population genetic model was developed tracking the change in resistance allele frequency as consequence of the genetic determination of the resistance phenotype, and the deployment policy for the insecticides.
+The model uses a similar approach to that previously used to compare insecticide mixtures and sequences [@Levick2017,@South2018].
+The implementation is simpler here because for rotations and sequences there is only a need to follow one insecticide at a time and linkage disequilibrium can be ignored.
 The simpler implementation allows the model to be run for an unlimited number of insecticides.
@@
 
-2. Rotations : use insecticide for set time, switch to another.
+2. Rotations : use insecticide for set time (i.e. number of insect generations), switch to another.
 
@@
 Thus there is the potential to switch back to insecticides that had been used earlier in a model run if resistance had not reached the threshold or if resistance had declined below the threshold during a period when not in use.
-
-In both cases a simulation run continued until no more insecticides remained below their resistance thresholds.
+This means that the “sequential” policy is effectively a “rotate when resistant” policy but we use the former term for consistency with previous work.
+*todo cite different trems from Rex consortium review* For both rotations and sequences the simulation continues until no more insecticides remain below their resistance thresholds.
 
@@
 
+**Table 1. Model inputs and the ranges used in simulations**
+
+Input  | Description | Min | Max
+---------------------- | ------------------------------------- | ----- | -----
+Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide | 0.5 | 1
+Exposure | proportion of females exposed to insecticide | 0.4 | 0.9
+Male exposure | proportion of males exposed to insecticide as proportion of females | 0 | 1
+Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide | 0.1 | 0.9
+Dominance of resistance | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide | 
+Frequency  | frequency of resistance alleles within the population | 0.005 | 0.1
+Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide | 0 | 0.1
+Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide | 0.1 | 0.9
+Insecticide number | number of independent insecticides available | 2 | 5  
+Rotation interval | number of generations to use insecticide before switching | 5 | 50
+Coverage | for scenarios with dispersal, the proportion of insects in the treated area rather than untreated refugia |  0.1 | 0.9
+Dispersal | proportion of population exchanged between treated and untreated areas per generation, 0=none, 1=random mixing |  0.1 | 0.9
+
 The model was run under four main scenarios :
 
-1. no resistance fitness costs or dispersal link to untreated areas
+1. baseline : no resistance fitness costs or dispersal to/from untreated refugia
 2. fitness costs of resistance added
@@
 
-To generate a decline in resistance over time when an insecticide is not being used either fitness costs or a dispersal link to an untreated area are required.
+For each scenario 10,000 runs were performed randomly selecting inputs according to the ranges specified in Table 1 using a uniform distribution in all cases.
 
@@
 
+*todo change this bit*
 It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative.
@@
 
-Example model runs are shown with no costs or dispersal (Fig 1), with just resistanace costs (Fig 2) and with just dispersal from untreated refugis (Fig 3).
+Example model runs are shown with no costs or dispersal (Fig 1), with just resistance costs (Fig 2) and with just dispersal from untreated refugia (Fig 3).
 
-#### Results1. no resistance fitness costs or dispersal from untreated refugia
+*todo describe these example runs first, before talking about overall results*
+
+#### Results1. form of resistance evolution with and without costs and dispersal.
 
@@
 For rotations resistance frequencies step upwards when insecticides are in use and remain at a plateau when they are not.
+If the rotation interval is short enough then the insecticide can be used a few times before the resistance thresholds are reached.
+
+### Results2 : comparing rotations and sequences across model runs.
+
+In all the example runs shown the resistance threshold (allele frequency of 0.5) was reached within the course 500 generations which was the maximum limit imposed in the model.
+Across all model runs there were many in which the thresholds were not reached in this time period.
+
+*but first need to explain that sequences pretty much never better*
+*so show the modified histograms first, i.e. don't filter out the threshold not reached runs*
+
+Thus to aid the comparison of rotations and sequences model runs can be broadly classified into three groups :
 
@@
 
-#### Results2. fitness costs of resistance added
-
+*todo : maybe I should just leave the 'succeeded' runs in here which would change these %s slightly*
 With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold.
-
-In contrast to the previous section 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).
-
-#### Results3. dispersal from untreated refugia added
+In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).
 
 Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 4.3).
-
-#### Results4. costs and dispersal
 
@@
 #### Runs where rotations have substantial advantage over sequences
+
+*Talk about classing runs into both 'succeeded', both 'failed' or just rotation succeeded.
+*show violin plots here ?*
+
 As shown in Fig 4.2 some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 5.
@@
 
+This can be understood intuitively.
+When there are no costs, the dynamics are the same, it is simply that sequential deployment proceeds in a few, larger steps while rotations proceeds in more numerous smaller steps; both reach the same point in around the same time (e.g.
+figs 1 and 3).
+However dominance effects on cost and resistance can act to keep frequencies low in a rotations policy.
+Recall that at low frequencies of resistance most resistance alleles are present as heterozygotes RS and only a vey small proportions heterozygotes.
+However costs are dominant and still act at low frequencies.
+This is illustrated on Figure Y.
+Relatively rapid rotations keep the allele frequencies at the LH side of Figure Y where costs during the periods when the insecticide is not deployed greatly outweigh the selection for IR during the periods when the insecticide is deployed.
+We believe the explanation is therefore that rotations keep allele frequencies in the parameter space to the left of Figure Y and hence acts to slow the evolution of IR compared to sequences which allow frequencies to rise to relatively high levels before the insecticide is replaced.
+(does this sound reasonable??).
+Note also that starting frequency, as expected, had an impact (Figure 7), lower starting frequencies slightly favouring rotations by allowing more time in the “low frequency” parameter space.
+
 In rare circumstances, when there are costs of resistance, rotations can lead to the resistance frequencies for a group of insecticides being kept below resistance thresholds for the length of the simulations where sequences run out of insecticides within around 100 generations.
@@
 
+Note however that this combination of recessive resistance and high dominance costs are exactly the features most desirable in any insecticide irrespective of how it is deployed.
+We discuss fitness below but note here that it is very difficult to maintain insecticide concentration at the level required to make resistance recessive.
+Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant; see Figure 1 of <cite Levick et al> and our more extensive discussion in <cite South et al WoS manuscript).
+
 The importance of resistance costs for substantial evolutionary benefits of rotations over sequences points to the importance of the debate over the fitness costs of insecticide resistance e.g.
@@
 
+We also found frequency of rotations had little effect.
+This is important in the control of vectors of human diseases such as malaria and dengue.
+These predominantly occur in resource-poor areas often with poor healthcare infrastructure.
+Experience suggests that rotation targets would often be missed (e.g rotations occur approximately every 15 generations rather than the planned “official” 10 generations) but our simulations suggest this would not fatally undermine a rotations policy.
+
 ### Caveats
 
-The model does not (although it could) include the potential effects of modifier genes ameliorating the costs of resistance alleles.
+The model does not include the potential effects of modifier genes ameliorating the costs of resistance alleles.
+To include this would require a more detailed model tracking linkage disequilibrium between each resistance locus and its modifier.
 It has been suggested that rotations could be favoured as a strategy because they keep resistance frequences lower and restrict selection pressure for modifier genes [@WHO2012].
@@
 
+In summary, we show that rotations are invariably better than sequential use, although the difference may often be small (figure 4); certainly rotations should not be seen as a panacea capable of removing the threat of resistance evolution.
+Rotations may have a large impact if fitness costs are prints and dominant (these are genetic/physiological factors outside our control)  and if resistance can be kept recessive.
+The latter is under our control to some extent (fig 1 of Levick et al ) but it is extremely difficult to deploy and maintain insecticide concentrations at levels that ensure recessively (cite WoS ms): as concentrations decline resistance become dominant and any befits of rotations largely disappear.
+
 ## TODO
 *try to come up with brief name for model output to make it easier to refer to it later*
+*add caveat about polygenic resistance*
+*The caption to each graph needs to have the list of parameter values used to generate them*
+*work on Ians fig Y to help explanation*
 
@@
 Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia.
-The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached.
-In this example the rotation and the sequence are stopped at the same time.
+The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached.
+In this example the rotation and the sequence reach the endpoint, resistance frequency > 0.5 to all insecicides, at the same time.
 For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again.
@@
 There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use.
-The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold.
 All other inputs are kept constant between the rotation and the sequence."}
@@
 ```
+
+\pagebreak
 
@@
 
+\pagebreak
+
 ```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4.
-Effect of cost and dispersal on the distribution of differences between rotations and sequences from model runs.
-The y axis is square root transformed to emphasise the infrequent runs where there is a significant benefit of rotations over sequences.
+Effect of cost and dispersal on the difference between rotations and sequences from model runs.
+The y axis shows the number of model runs for the different caegories of difference between rotations and sequences.
 4.1 when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results.
@@
 
-  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=pcent_rot_adv)) +
+  # log scale overemphasised low values
+  # sqrt scale looked good but Ian didn't like
 
-             geom_histogram(binwidth=10) +
-             xlim(-20,400) +
+  custom_bins <- c(-20,-1,1,20,100,450)
+  df_nocost_nomig$rot_adv_bins <- cut(df_nocost_nomig$pcent_rot_adv, custom_bins)
+  df_cost$rot_adv_bins <- cut(df_cost$pcent_rot_adv, custom_bins)
+  df_mig$rot_adv_bins <- cut(df_mig$pcent_rot_adv, custom_bins)
+  df_cost_mig$rot_adv_bins <- cut(df_cost_mig$pcent_rot_adv, custom_bins)
+
+  bin_labs <- levels(df_cost$rot_adv_bins)
+
+  #gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=pcent_rot_adv)) + 
+  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=rot_adv_bins)) +
+
+             geom_bar() +
+             #xlim(-20,400) +
+             xlim(bin_labs) +
              xlab('') +
@@
              ggtitle('1. no resistance costs or dispersal') +
-             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
              theme_minimal() +
-             theme(panel.grid.minor.y=element_blank())
+             theme(panel.grid.minor.y=element_blank(),
+                   axis.text.x=element_blank())
 
-  gg_cost <- ggplot( df_cost, aes(x=pcent_rot_adv)) +
+  gg_cost <- ggplot( df_cost, aes(x=rot_adv_bins)) +
 
-             geom_histogram(binwidth=10) +
-             xlim(-20,400) +
+             geom_bar() +
+             #geom_histogram(binwidth=10) +
+             #xlim(-20,400) +
              xlab('') +
@@
              ggtitle('2. with costs') +
-             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
+             #scale_x_continuous(breaks=c(-10,-1,0,1,20,100,400)) +
+             scale_x_discrete( limits=bin_labs,
+                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
              theme_minimal() +
-             theme(panel.grid.minor.y=element_blank())
+             theme(panel.grid.minor.y=element_blank(),
+                   axis.text.x=element_blank())
 
-  gg_mig <- ggplot( df_mig, aes(x=pcent_rot_adv)) +
+  gg_mig <- ggplot( df_mig, aes(x=rot_adv_bins)) +
 
-             geom_histogram(binwidth=10) +
-             xlim(-20,400) +
+             geom_bar() +
+             xlim(bin_labs) +
              xlab('') +
@@
              ggtitle('3. with dispersal refugia') +
-             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
              theme_minimal() + 
-             theme(panel.grid.minor.y=element_blank())
+             theme(panel.grid.minor.y=element_blank(),
+                   axis.text.x=element_blank())
 
   #gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_minus_seq)) + 
-  gg_cost_mig <- ggplot( df_cost_mig, aes(x=pcent_rot_adv)) +              
-             geom_histogram(binwidth=10) +
-             xlim(-20,400) +
-             xlab('percent benefit of rotation versus sequence') +
+  gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_adv_bins)) +
+
+             geom_bar() +
+             xlab('Percent benefit of rotation over sequence') +
              ylab('count') +
              ggtitle('4. with costs and dispersal') +
-             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
              theme_minimal() +
+             scale_x_discrete( limits=bin_labs,
+                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+       
              theme(panel.grid.minor.y=element_blank())
@@
 
+\pagebreak
+
 ```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 6.
-Which model inputs have the greatest influence on the difference between rotations and sequences when costs alone are included ?
+Which model inputs have the greatest influence on the difference between rotations and sequences ?
+In this case for scenario 2 where resistance costs are included but dispersal is not.
 Measured using Partial Rank Correlation Coefficients (PRCC).
-Dominance of cost and dominance of selection have the greatest influence."}
+Positive values indicate a positive influence on the benefit of rotations, only cost and dominance of cost in this case.
+Dominance of cost and dominance of selection have the greatest magnitude of influence."}
 
@@
 
+\pagebreak
+
 ```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 7.
-Model runs with substantial benefits of rotations over sequences are associated with high dominance of cost and low dominance of selection.
-Dominace inputs are shown on the x and y axes.
+Model runs in which rotations perform more than 20% better than sequences are associated with high dominance of cost and low dominance of selection.
+Dominance inputs are shown on the x and y axes.
 Small grey points indicate all model runs.
@@
 
+\pagebreak
+
 ```{r FigX, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure X.

### 2018-05-08 inst/documents/rotations_notes.Rmd
paper3 writing

@@
 
+*2/5/18*
+
+exchanging emails with Ian
+
+Thanks Ian, I'm getting closer to understanding. I'm just trying to understand why the lines cross, which suggests to me that resistance wouldn't be selected for at low frequencies ?
+
+This is a simple text description of how I'm starting to understand. Is it true ? (we can work on improving the wording later).
+
+Resistance increases when an insecticide is in use and, if there are resistance costs, decreases when it is not.
+
+If the difference between this increase and decrease over periods of use and non-use is positive then resistance will keep increasing.
+
+If the positive difference between the increase and decrease is lower at lower frequencies then there is an advantage to short rotation intervals that keep frequencies low.
+
+In this latter scenario sequential use leads to higher frequencies where increases during periods of use are more likely to outweigh decreases during periods of non-use.
+
+done ~ Can I change histograms of count of rot advantage to have different bins, we're only interested in whether difference is 0, moderate or substantial.
+
+Trying to recreate scenario 2167
+after removal of non threshold runs :
+
+  scenario <- 2167
+  linputs <- as.list( df_cost[scenario, 1:15] )
+  dfres <- do.call(run_rot, linputs)
+  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+  linputs$rot_interval <- 0
+  dfres <- do.call(run_rot, linputs)
+  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE) 
+  grid.arrange(ggrot, ggseq, ncol=1)
+
+linputs$dom_cos <- 0.8
+x11()
+
+  linputs$rot_interval <- 10
+  dfres <- do.call(run_rot, linputs)
+  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+  linputs$rot_interval <- 0
+  dfres <- do.call(run_rot, linputs)
+  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE) 
+  grid.arrange(ggrot, ggseq, ncol=1)
+
+# setting params for 2167
+linputs <- list()
+linputs$n_insecticides <- 3
+linputs$coverage <- 1
+linputs$migration <- 0
+linputs$cost <-     0.0582 
+linputs$max_gen <-  500
+linputs$expo_hi <-  0.64
+linputs$eff <-      0.96
+linputs$rr <-      0.78
+linputs$dom_sel <-  0.06
+linputs$dom_cos <-  0.7
+linputs$start_freqs <- 0.0731
+linputs$male_expo_prop <- 0.167
+linputs$plot <- FALSE
+
+  linputs$rot_interval <- 10
+  dfres <- do.call(run_rot, linputs)
+  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+  linputs$rot_interval <- 0
+  dfres <- do.call(run_rot, linputs)
+  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE) 
+  grid.arrange(ggrot, ggseq, ncol=1)
+
+# this shows a small tweek of the params leads to the sequence lasting forever too
+# and I don't know which input change causes the seq to last forever ?
+# Aha seems that I hadn't included rr=0.78, rr default is 0.5, 
+# with lower rr less selection for r and sequence does better 
+linputs <- list()
+linputs$n_insecticides <- 3
+linputs$coverage <- 1
+linputs$migration <- 0
+linputs$cost <-     0.058 
+linputs$max_gen <-  500
+linputs$expo_hi <-  0.64
+linputs$eff <-      0.96
+linputs$rr <-      0.5 #0.78
+linputs$dom_sel <-  0.06
+linputs$dom_cos <-  0.7
+linputs$male_expo_prop <- 0.167
+linputs$start_freqs <- 0.0731
+linputs$plot <- FALSE
+
+  linputs$rot_interval <- 10
+  dfres <- do.call(run_rot, linputs)
+  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+  linputs$rot_interval <- 0
+  dfres <- do.call(run_rot, linputs)
+  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE) 
+  grid.arrange(ggrot, ggseq, ncol=1)
+
+*3/5/18*
+Targets for today
+a) check how stable the rotations advantage is
+b) understand and display the mechanism
+
+But I went off that track because I got more to thinking we should be trying to identify circumstances where both strategies 'succeed'.
+However I suspect that 'success' here is due to low effectiveness and exposure which lead to slow evolution but possible cause low control too.
+
+Looking at trees to differentiate between success, failure and rot_only.
+Not very helpful.
+
+Looking at input effects on n generations, they go the way I would expect, but may have to be part of the story.
+
+The violin plots for each input var for rot_seq, rot, neither look pretty good.
+
+e.g. for cost they show that successful runs tend to have higher costs, those failing for both have lower costs. Those that work just for rotations tend to have lower costs.
+
+*8/5/18*
+
 # TODO
 
-~ run 10k runs of dispersal overnight
-~ try histograms on log scale ?
+Write out some text explaining the story & maybe the order of the plots. (keep num plots manageable).
+
+order could be
+1. example plots
+2. histograms 
+3. violin plots
+4.
+
+Not sure how best to look at effect of inputs on rot_seq, rot, neither. I could bin input ranges and then look at proportions of the 3 within them ?
+
+This would tell me how changing an input value changes the proportion of successes, failures and rot_only.
+Might be OK to display.
+Tricky to calculate, but again gets mired in the multivariate nature of the data & potential interactions.
+
+If I go back to using a continuous variable of generations until thresholds I lose out on the successful results.
+
+Maybe a more useful question, rather than when are rotations or sequence better is when are both rotations and sequences sustainable.
+i.e.
+under which combination of conditions do both strategies last 'forever' ?
+
+In that case the runs that I have been excluding become the useful ones.
+
+Because actually, finding the runs that are different for rotations and sequences after excluding those that 'succeed' for both is actually trying to find those that 'fail' for sequences.
+
+I think there could be a boundary area in between those that 'succeed' for both and those that 'fail' for both, where some 'succeed' only for rotations.
+
+I think takehome message maybe that rotations are much better than sequences only in rare circumstances.
+
+May also be something useful I can say about the number of insecticides too ? (Which is of interest to IVCC).
+
+Maybe reclass all 10k runs to succeed_both (2), succeed_rot (1), fail_both (0).
+Then see what are the best predictors of these ?
+Or do series of xy plots with points coloured, coloured points with all data will probably be a mess ...
+
+~ try changing scenario 2167 from df_cost
+~~ initial attempts in UI not working
+~~ could it be that it requires just right combination of params and that modifying them just slightly removes the advantage for rotations ?
+
+~ can I create some simple figs to illustrate how an increasing rate of resistance evolution with frequency can lead to rotations being favoured   
+~~ maybe just see how dom_cos changes shape of frequency plot
+~~ can I look at through UI ?
+
+~ save Ians comments on figY of
+
 ~ try decision tree to predict rot_minus_seq from inputs
-
-~ send MS to Ian before 11
 

### 2018-05-09 inst/documents/paper3/paper3_rotations.Rmd
pap3

@@
 
-*todo describe these example runs first, before talking about overall results*
-
 #### Results1. form of resistance evolution with and without costs and dispersal.
@@
 
+*todo continue describing the example runs first, before talking about overall results*
+
 ### Results2 : comparing rotations and sequences across model runs.
+
+*todo : change the %s because 'succeeded' runs now left in*
+When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold (Fig 4.1).
+With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold.
+In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).
+Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 4.3).
+With both costs and dispersal the benefits of rotations are similar to when runs have costs alone (Fig 4.4).
 
@@
 Thus to aid the comparison of rotations and sequences model runs can be broadly classified into three groups :
+1. both strategies succeeded : resistance thresholds for all insecticides not reached within 500 generations
+2. rotations only succeeded
+3. neither strategies succeeded
 
-When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold (Fig 4.1).
-
-*todo : maybe I should just leave the 'succeeded' runs in here which would change these %s slightly*
-With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold.
-In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).
-
-Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 4.3).
-
-With both costs and dispersal the benefits of rotations are similar to when runs have costs alone (Fig 4.4).
+*todo show a histogram plot of the counts of model runs in each category, across all startegies and scenarios*
+*decide whether to put before or after fig4*
 
@@
 We discuss fitness below but note here that it is very difficult to maintain insecticide concentration at the level required to make resistance recessive.
-Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant; see Figure 1 of <cite Levick et al> and our more extensive discussion in <cite South et al WoS manuscript).
+Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant; see Figure 1 of [@Levick2017] and our more extensive discussion in [cite South et al WoS manuscript].
 
@@
 ## TODO
+*add to methods description of how fitness determined in model, move & modify fig 1*
 *try to come up with brief name for model output to make it easier to refer to it later*
@@
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1.
-Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia.
-The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached.
-In this example the rotation and the sequence reach the endpoint, resistance frequency > 0.5 to all insecicides, at the same time.
+```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1. "}
+
+#temp moved 
+#Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia.
+The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence in which the insecticide is changed when a resistance threshold of 0.5 is reached.
+In this example the rotation and the sequence reach the endpoint, with all insecticides having a resistance frequency above 0.5, at the same time.
 For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again.
@@
 There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use.
-All other inputs are kept constant between the rotation and the sequence."}
+All other inputs are kept constant between the rotation and the sequence.
+Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1.
 
@@
 linputs$dom_cos <-  1
+linputs$rr <-       0.5
+linputs$male_expo_prop <-  0.5
 linputs$plot <-     FALSE
@@
 The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold.
-All other inputs are kept constant between the rotation and the sequence."}
+All other inputs are kept constant between the rotation and the sequence.
+Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0.05, dominance of cost=1."}
 
@@
 The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold.
-All other inputs are kept constant between the rotation and the sequence."}
+All other inputs are kept constant between the rotation and the sequence.
+Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0, coverage=0.5, dispersal=0.1."}
 
@@
 
-```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4.
+```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='center', fig.cap="Figure 4.
 Effect of cost and dispersal on the difference between rotations and sequences from model runs.
@@
   # with cost_mig ~ 65% of runs don't reach thresholds
-  # could get this func to output perecnt not reaching
-  rem_thresh_not_reached <- function(x) {x %>% filter(gens_seq != max_gen | gens_rot != max_gen)}
-
-  df_nocost_nomig <- rem_thresh_not_reached(df_nocost_nomig)
-  df_cost <- rem_thresh_not_reached(df_cost) #6.6
-  df_mig <- rem_thresh_not_reached(df_mig) #0.05
-  df_cost_mig <- rem_thresh_not_reached(df_cost_mig) #6.9
+  # could get this func to output percent not reaching
+  # now don't remove the runs where thresholds not reached
+  # rem_thresh_not_reached <- function(x) {x %>% filter(gens_seq != max_gen | gens_rot != max_gen)}
+  # 
+  # df_nocost_nomig <- rem_thresh_not_reached(df_nocost_nomig)
+  # df_cost <- rem_thresh_not_reached(df_cost) #6.6
+  # df_mig <- rem_thresh_not_reached(df_mig) #0.05
+  # df_cost_mig <- rem_thresh_not_reached(df_cost_mig) #6.9
 
@@
 
-  custom_bins <- c(-20,-1,1,20,100,450)
+  #custom_bins <- c(-20,-1,1,20,100,450)
+  custom_bins <- c(-20,-10,-1,1,10,20,100,450)
   df_nocost_nomig$rot_adv_bins <- cut(df_nocost_nomig$pcent_rot_adv, custom_bins)
@@
              scale_x_discrete( limits=bin_labs,
-                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
+                               #labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
+                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+    
              theme_minimal() +
@@
              scale_x_discrete( limits=bin_labs,
-                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+       
+                               #labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
+                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+     
              theme(panel.grid.minor.y=element_blank())
@@
 For the sequence resistance thresholds are reached for all insecticides in a little over 100 generations or 10 years.
-For the rotation resistance levels for all insecticides are maintained well below resistance thresholds for the length of the simulation (500 generations or 50 years)."}
+For the rotation resistance levels for all insecticides are maintained well below resistance thresholds for the length of the simulation (500 generations or 50 years).
+Effectiveness=0.57, exposure=0.88, male exposure proportion=0.72, resistance restoration=0.68, dominance of resistance=0.04, cost=0.07, dominance of cost=0.59."}
 # ? are these run ids before or after the runs that don't reach thresholds have been removed ?
@@
 #scenarios_to_plot <- c(624,2167,602,4028,5771)
-scenarios_to_plot <- c(2167)
+
+#now that non thershold runs are not removed :
+#which(df_cost$pcent_rot_adv > 200)
+#987 1023 1324 1365 1511 2051 2717 3187 3196 3582 3589 3768 4128 4219 4835 4880 6233 6538 6913 8666 9361
+scenarios_to_plot <- c(2717)
 

### 2018-05-09 inst/documents/rotations_notes.Rmd
pap3

@@
 *8/5/18*
+Getting violin plots working.
+
+*9/5/18*
+Leaving in all runs now
+
+Something weird has happened to fig1 in the paper3 ?? because of a > in fig legend
+
+Also because we are leaving all runs in, the indices of the runs with large rot advantage are now wrong ! I should have recoded a unique id for runs.
+
+ARg, now all plots have stopped appearing in doc and I don't know why ?????
+It was a < symbol in the MS text, could tell because the text stopped at that point.
 

### 2018-05-10 inst/documents/paper3/paper3_rotations.Rmd
pap3

@@
                cache=FALSE,
-               fig.path='pap3figs/')
+               fig.path='pap3figs/',
+               echo=FALSE, warning=FALSE, message=FALSE)
+
+#don't work for word
+#fig.show='hold', fig.align='center'
 ```
@@
 This means that the “sequential” policy is effectively a “rotate when resistant” policy but we use the former term for consistency with previous work.
-*todo cite different trems from Rex consortium review* For both rotations and sequences the simulation continues until no more insecticides remain below their resistance thresholds.
+*todo cite different terms from Rex consortium review* For both rotations and sequences the simulation continues until no more insecticides remain below their resistance thresholds.
 
@@
 
-*todo change this bit*
+*todo change this bit, we could actually use time-to-resistance for all strategies ?*
+*make it clearer exactly what we do*
 It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative.
@@
 
-Example model runs are shown with no costs or dispersal (Fig 1), with just resistance costs (Fig 2) and with just dispersal from untreated refugia (Fig 3).
-
 #### Results1. form of resistance evolution with and without costs and dispersal.
 
-When there are no resistance fitness costs or dispersal links to untreated refugia resistance frequencies do not decline when an insecticide is not in use (Fig 1).
+Example model runs are shown with no costs or dispersal (Fig 2), with just resistance costs (Fig 3) and with just dispersal from untreated refugia (Fig 4).
+
+When there are no resistance fitness costs or dispersal links to untreated refugia resistance frequencies do not decline when an insecticide is not in use (Fig 2).
 Thus for sequences once resistance thresholds are reached the resistance frequencies remain at that level and the insecticide cannot be re-used.
@@
 
-*todo : change the %s because 'succeeded' runs now left in*
-When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold (Fig 4.1).
-With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold.
-In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).
-Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 4.3).
-With both costs and dispersal the benefits of rotations are similar to when runs have costs alone (Fig 4.4).
+In all the example runs shown the resistance threshold (allele frequency of 0.5) was reached for all insecticides within the course 500 generations which was the maximum limit imposed in the model.
+There were many model runs [*quantify later when run finalised*] in which resistance frequencies for at least some insecticides were below thresholds at the end of 500 generations.
+Such a run could be considered a successful strategy given the conditions of the scenario because insecticides below the resistance threshold are still available to be used.
 
-In all the example runs shown the resistance threshold (allele frequency of 0.5) was reached within the course 500 generations which was the maximum limit imposed in the model.
-Across all model runs there were many in which the thresholds were not reached in this time period.
+Thus to aid the comparison of rotations and sequences model runs can be broadly classified into four groups :
+1. both strategies succeeded : for both rotations and sequences resistance thresholds for all insecticides not reached within 500 generations.
+2. neither strategy succeeded
+3. rotation only succeeded
+4. sequence only succeeded
 
-*but first need to explain that sequences pretty much never better*
-*so show the modified histograms first, i.e. don't filter out the threshold not reached runs*
+The number of model runs in each of these groups for the different scenarios are shown in Fig 5.
+In the base scenario with no costs or dispersal (Fig 5A) the majority of runs do not succeed, a small proportion succeed for both strategies and none succeed for either strategy alone.
+When costs of reistance are added (Fig 5B) more runs succeed for both strategies, up to about half of those that succeed for neither, and a small proportion (< 4%) succeed for rotations alone.
+Adding both costs and dispersal results in more runs succeeding for both than neither (Fig 5C).
 
-Thus to aid the comparison of rotations and sequences model runs can be broadly classified into three groups :
-1. both strategies succeeded : resistance thresholds for all insecticides not reached within 500 generations
-2. rotations only succeeded
-3. neither strategies succeeded
+To look at the potential benefit of rotations the number of generations below resistance thresholds was compared across strategies.
+This will be influenced by the previous analysis because only those runs that do not succeed for both strategies can generate a difference between them.
 
-*todo show a histogram plot of the counts of model runs in each category, across all startegies and scenarios*
-*decide whether to put before or after fig4*
+When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold (Fig 6A) (for comparison with other strategies the results are grouped into those between plus and minus 1%).
+With fitness costs included 80% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold.
+In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 6B).
+Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 6C).
+With both costs and dispersal the benefits of rotations are similar to when runs have costs alone (Fig 6D).
 
@@
 
-*Talk about classing runs into both 'succeeded', both 'failed' or just rotation succeeded.
 *show violin plots here ?*
@@
 
-Comparing time below resistance thresholds for rotations and sequences, the model shows :
+The model shows that rotations and sequences most often have very similar performance (within 1 %) in terms of time to reach resistance thresholds (Fig 6).
 
-1. rotations and sequences the same when no costs or dispersal
-2. with resistance costs, 67% of runs the same for rotations and sequences, but 6-7% of runs have > 20% advantage for rotations
+*todo add something about how costs and dispersal generally improve the performance of both strategies*
+*todo add about how rotations generally perform better relative to sequences under 'intermediate' scenarios.
+In 'better' scenarios both strategies do equally well and in 'poorer' scenarios they do equally badly.*
+
+1. rotations and sequences give the same results when there are no costs or dispersal.
+2. with resistance costs, 80% of runs are the same for rotations and sequences, but 6-7% of runs have > 20% advantage for rotations
 3. with dispersal from untreated refugia but no costs, small advantage of rotations but very few runs < 0.1% have >20% advantage.
@@
 This can be understood intuitively.
-When there are no costs, the dynamics are the same, it is simply that sequential deployment proceeds in a few, larger steps while rotations proceeds in more numerous smaller steps; both reach the same point in around the same time (e.g.
-figs 1 and 3).
+When there are no costs, the dynamics are the same, it is simply that sequential deployment proceeds in a few, larger steps while rotations proceeds in more numerous smaller steps; both reach the same point in around the same time.
 However dominance effects on cost and resistance can act to keep frequencies low in a rotations policy.
@@
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1. "}
+```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1.
+[todo need to modify caption from MJ paper, start by saying that only in left panel during application of the insecticide] The effect of model inputs on the fitness of genotypes for a single insecticide.
+Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
+Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
+For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of resistance determines how the fitness for SR lies between that of SS and RR.
+For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
+In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of resistance=0.7 which sets the fitness of the SR closer to RR than SS.
+Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}
 
-#temp moved 
-#Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia.
+library(resistance) 
+plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.7, dominance_cost=0.7, cost=0.3 )
+
+```
+
+\pagebreak
+
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2.
+Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia.
 The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence in which the insecticide is changed when a resistance threshold of 0.5 is reached.
@@
 All other inputs are kept constant between the rotation and the sequence.
-Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1.
+Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1."}
 
@@
 
-```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2.
+```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3.
 Comparing the increase in resistance frequencies over time for a rotation and a sequence when there are resistance fitness costs.
@@
 
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3.
+```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4.
 Comparing the increase in resistance frequencies over time for a rotation and a sequence with dispersal from untreated refugia.
@@
 
-```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='center', fig.cap="Figure 4.
+```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.align='center', fig.cap="Figure 5.
+Success of model runs under different scenarios and strategies.
+Model runs classed as a success if there are still insecticides for which resistance is below the threshold of 0.5 after 500 generations (circa 50 years for Anopheles)."}
+
+#remove seq maybe have a check that are very few
+# count(df_cost, success)
+# 1         neither  5948
+# 2             rot   209
+# 3         rot_seq  3841
+# 4             seq     2
+
+# if I put all scenarios in one df with a column for scenario I can do one ggplot and facet 
+# would work for the previous plot too and would simplify adding the new columns
+# first add the scenario column
+df_nocost_nomig$scenario <- 'no cost or dispersal'
+df_cost$scenario <- 'costs'
+df_mig$scenario <- 'dispersal'
+df_cost_mig$scenario <- 'costs and dispersal'
+
+df_all <- dplyr::bind_rows(df_nocost_nomig,
+                           df_cost,
+                           df_mig,
+                           df_cost_mig)
+#order for ggplot
+df_all$scenario <- factor(df_all$scenario, levels=c('no cost or dispersal','costs','dispersal','costs and dispersal'))
+
+# add a column for classification of whether, both, rotation, or neither succeeded
+df_all <- mutate(df_all, success = ifelse(gens_seq==max_gen & gens_rot==max_gen,'both',
+                                               ifelse(gens_rot==max_gen,'rotation only',
+                                               ifelse(gens_seq==max_gen,'sequence only', 'neither'  
+                                                      ))))
+#count(df_all, success)
+# 1          both 12413
+# 2       neither 18196
+# 3 rotation only   382
+# 4 sequence only     9
+
+#hack to add subplot labels to facets
+abcd <- data.frame(label = c("A","B","C","D"),
+                   scenario = levels(df_all$scenario))
+
+ggplot( df_all, aes(x=success)) + 
+        #geom_bar() +
+        stat_count(mapping = aes(y=..prop.., group=1)) + #puts proprtions in axes
+        scale_y_continuous(breaks=seq(0.2,0.8,0.2)) +
+        theme_minimal() +
+        ylab('proportion of model runs') +
+        facet_wrap(~scenario, ncol=1) + #, scales='free_y') + #can remove free_y later if I do 10k nocost runs    
+        geom_text(data=abcd,
+                  aes(x = -Inf, y = +Inf, label=label),
+                  hjust = 0,
+                  vjust = 0.9 )
+
+```
+
+\pagebreak
+
+```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.align='center', fig.cap="Figure 6.
 Effect of cost and dispersal on the difference between rotations and sequences from model runs.
 The y axis shows the number of model runs for the different caegories of difference between rotations and sequences.
-4.1 when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results.
-4.2, 4.4 when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences.
+A) when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results.
+B),D) when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences.
 For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences.
-4.3 when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}
+C) when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}
 
@@
   # add a column for percent difference between rot & seq
+  # LEFT IN FOR NOW BECAUSE used in later chunks
   df_nocost_nomig <- mutate(df_nocost_nomig, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
@@
 
-  # log scale overemphasised low values
-  # sqrt scale looked good but Ian didn't like
+  #try quicker way with df_all and facet
+  df_all <- mutate(df_all, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
+  custom_bins <- c(-20,-10,-1,1,10,20,100,450)
+  df_all$rot_adv_bins <- cut(df_all$pcent_rot_adv, custom_bins)  
+  bin_labs <- levels(df_all$rot_adv_bins)
 
-  #custom_bins <- c(-20,-1,1,20,100,450)
-  custom_bins <- c(-20,-10,-1,1,10,20,100,450)
-  df_nocost_nomig$rot_adv_bins <- cut(df_nocost_nomig$pcent_rot_adv, custom_bins)
-  df_cost$rot_adv_bins <- cut(df_cost$pcent_rot_adv, custom_bins)
-  df_mig$rot_adv_bins <- cut(df_mig$pcent_rot_adv, custom_bins)
-  df_cost_mig$rot_adv_bins <- cut(df_cost_mig$pcent_rot_adv, custom_bins)
+  #hack to add subplot labels to facets
+  abcd <- data.frame(label = c("A","B","C","D"),
+                   scenario = levels(df_all$scenario))
 
-  bin_labs <- levels(df_cost$rot_adv_bins)
+  ggplot( df_all, aes(x=rot_adv_bins)) +
 
-  #gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=pcent_rot_adv)) + 
-  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=rot_adv_bins)) +
+             #geom_bar() +
+             stat_count(mapping = aes(y=..prop.., group=1)) + #puts proprtions in axes
+             scale_y_continuous(breaks=seq(0,0.8,0.2)) +
+             xlab('Percent benefit of rotation over sequence') +
+             ylab('proportion of model runs') +
+             #ggtitle('') +
+             theme_minimal() +
+             facet_wrap(~scenario, ncol=1) +
+             scale_x_discrete( limits=bin_labs,
+                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+     
+             theme(panel.grid.minor.y=element_blank(),
+                   axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1)) +
+             geom_text(data=abcd,
+                  aes(x = -Inf, y = +Inf, label=label),
+                  hjust = 0,
+                  vjust = 0.9 )
 
-             geom_bar() +
-             #xlim(-20,400) +
-             xlim(bin_labs) +
-             xlab('') +
-             ylab('count') +
-             ggtitle('1. no resistance costs or dispersal') +
-             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
-             theme_minimal() +
-             theme(panel.grid.minor.y=element_blank(),
-                   axis.text.x=element_blank())
-
-  gg_cost <- ggplot( df_cost, aes(x=rot_adv_bins)) +
-
-             geom_bar() +
-             #geom_histogram(binwidth=10) +
-             #xlim(-20,400) +
-             xlab('') +
-             ylab('count') +
-             ggtitle('2. with costs') +
-             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
-             #scale_x_continuous(breaks=c(-10,-1,0,1,20,100,400)) +
-             scale_x_discrete( limits=bin_labs,
-                               #labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
-                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+    
-             theme_minimal() +
-             theme(panel.grid.minor.y=element_blank(),
-                   axis.text.x=element_blank())
-
-  gg_mig <- ggplot( df_mig, aes(x=rot_adv_bins)) +
-
-             geom_bar() +
-             xlim(bin_labs) +
-             xlab('') +
-             ylab('count') +
-             ggtitle('3. with dispersal refugia') +
-             theme_minimal() + 
-             theme(panel.grid.minor.y=element_blank(),
-                   axis.text.x=element_blank())
-
-  #gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_minus_seq)) + 
-  gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_adv_bins)) +
-
-             geom_bar() +
-             xlab('Percent benefit of rotation over sequence') +
-             ylab('count') +
-             ggtitle('4. with costs and dispersal') +
-             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
-             theme_minimal() +
-             scale_x_discrete( limits=bin_labs,
-                               #labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
-                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+     
-             theme(panel.grid.minor.y=element_blank())
-
-  grid.arrange(gg_nocost_nomig, gg_cost, gg_mig, gg_cost_mig, ncol=1)
+  # gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=rot_adv_bins)) + 
+  #                
+  #            geom_bar() +
+  #            #xlim(-20,400) +
+  #            xlim(bin_labs) +
+  #            xlab('') +
+  #            ylab('count') +
+  #            ggtitle('1. no resistance costs or dispersal') +
+  #            #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
+  #            theme_minimal() +
+  #            theme(panel.grid.minor.y=element_blank(),
+  #                  axis.text.x=element_blank())
+  # 
+  # 
+  # gg_cost <- ggplot( df_cost, aes(x=rot_adv_bins)) + 
+  #            
+  #            geom_bar() +
+  #            #geom_histogram(binwidth=10) +
+  #            #xlim(-20,400) +
+  #            xlab('') +
+  #            ylab('count') +
+  #            ggtitle('2. with costs') +
+  #            scale_x_discrete( limits=bin_labs))+    
+  #            theme_minimal() +
+  #            theme(panel.grid.minor.y=element_blank(),
+  #                  axis.text.x=element_blank())
+  # 
+  # gg_mig <- ggplot( df_mig, aes(x=rot_adv_bins)) + 
+  #            
+  #            geom_bar() +
+  #            xlim(bin_labs) +
+  #            xlab('') +
+  #            ylab('count') +
+  #            ggtitle('3. with dispersal refugia') +
+  #            theme_minimal() + 
+  #            theme(panel.grid.minor.y=element_blank(),
+  #                  axis.text.x=element_blank())
+  # 
+  # #gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_minus_seq)) + 
+  # gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_adv_bins)) +  
+  #   
+  #            geom_bar() +
+  #            xlab('Percent benefit of rotation over sequence') +
+  #            ylab('count') +
+  #            ggtitle('4. with costs and dispersal') +
+  #            #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
+  #            theme_minimal() +
+  #            scale_x_discrete( limits=bin_labs,
+  #                              labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+     
+  #            theme(panel.grid.minor.y=element_blank(),
+  #                  axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))  
+  # 
+  # grid.arrange(gg_nocost_nomig, gg_cost, gg_mig, gg_cost_mig, ncol=1)
 
@@
 
-```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 5.
+```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 7.
 Example of a model run where there is a substantial benefit of rotations over a sequence.
@@
 
-```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 6.
+```{r Fig8, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 8.
 Which model inputs have the greatest influence on the difference between rotations and sequences ?
@@
 
-```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 7.
+```{r Fig9, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 9.
 Model runs in which rotations perform more than 20% better than sequences are associated with high dominance of cost and low dominance of selection.
@@
 
-```{r FigX, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure X.
-[todo need to modify caption from MJ paper, start by saying that only in left panel during application of the insecticide] The effect of model inputs on the fitness of genotypes for a single insecticide.
-Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
-Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of resistance determines how the fitness for SR lies between that of SS and RR.
-For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of resistance=0.7 which sets the fitness of the SR closer to RR than SS.
-Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}
+```{r Fig10, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 10.
+Simplified indication of how fitness differences between selective advantages and costs are expected to change according to dominance and resistance allele frequencies.
+Top left : high dominance of cost, low dominance of selection, difference between cost and selection fitness components can be very low at low allele frequencies, difference increases as allele frequencies increase.
+Lower left: low dominance of cost, low dominance of selection : increased difference at the lowest frequencies.
+Top right: increased dominance of selection markedly increases benefits of resistance at low frequencies and thus the difference between costs and benefits even when dominance of cost is high.
+This shows why dominance of selection needs to be low and dominance of cost needs to be high, when dominance of selection is high the difference between costs and benefits is high even at low frequencies irrespective of the value of dominance of cost.
+When dominance of selection is low a high dominance of cost can lead to very small differences between costs and benefits at low frequencies.
+In these latter circumstances if rotations can keep frequencies low they can reduce the long term development of resistance.
+Plot calculation assumes Hardy-Weinberg equilibrium which will not be the case under selection but the general pattern will be similar.
+Fitness cost set to 0.05 and selective advantage of resistance to 0.4."}
 
-library(resistance) 
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.7, dominance_cost=0.7, cost=0.3 )
+# high dominance of cost, low dominance of selection
+# difference between cost and selection fitness components can be very low at low allele frequencies
+# difference increases as allele frequencies increase
+df1 <- freq_fitness_simple(cos_rr = 0.05, dom_cos = 0.9, sel_rr = 0.4, dom_sel = 0.1, plot=FALSE)
+# low dominance of cost, low dominance of selection
+# reduced difference at the lowest frequencies
+df2 <- freq_fitness_simple(cos_rr = 0.05, dom_cos = 0.1, sel_rr = 0.4, dom_sel = 0.1, plot=FALSE)
+# low dominance of cost, low dominance of selection
+# hmmm similar to previous - underwhelming may not want to include
+df3 <- freq_fitness_simple(cos_rr = 0.05, dom_cos = 0.1, sel_rr = 0.4, dom_sel = 0.9, plot=FALSE)
+# high, high
+# difference between the fitness components
+df4 <- freq_fitness_simple(cos_rr = 0.05, dom_cos = 0.9, sel_rr = 0.4, dom_sel = 0.9, plot=FALSE)
+
+#this shows why dominance of selection needs to be low and dominance of cost needs to be high
+#when dominance of selection is high the difference between costs and benefits is high even at low frequencies irrespective
+#of the value of dominance of cost
+#when dominance of selection is low a high dominance of cost can lead to very small differences between costs and benefits at low frequencies
+#thus if rotations can keep frequencies low they can reduce the long term development of resistance
+
+# bind the rows together
+dfgrid <- dplyr::bind_rows(df1,df2,df3,df4)
+
+#change levels of dom_cos to get higher value upper (facet_grid(as.table=FALSE) would do same)
+dfgrid$dom_cos <- factor(dfgrid$dom_cos)#, levels=c(0.9,0.1)) 
+dfgrid$dom_sel <- factor(dfgrid$dom_sel)#, levels=c(0.9,0.1))
+
+#hack to add subplot labels to facets
+abcd <- data.frame(label = c("A","B","C","D"),
+                   dom_cos = levels(factor(dfgrid$dom_cos)),
+                   dom_sel = levels(factor(dfgrid$dom_sel)))
+                   #fitness_component = NA)
+
+# plot with facets for dom_sel & dom_cos
+ggplot(dfgrid, aes(x=freq, y=fitness_value, col=fitness_component)) +
+  geom_line() +
+  theme_minimal() +
+  xlab('allele frequency') +
+  facet_grid(dom_cos~dom_sel, labeller = "label_both", as.table = FALSE) #+
+  # geom_text(data=abcd,
+  #   aes(x = -Inf, y = +Inf, label=label),
+  #   #inherit.aes=FALSE,
+  #   colour='black',
+  #   hjust = 0,
+  #   vjust = 0.9 )
+
+#my subplot labelling hack fails and I'm not sure why?
+#initially extra facets were added for dom_sel values of 1 & 2 that aren't in the data
+#now labels AC & BD appear on same plots
 
@@
 
-\pagebreak
-
 ## References

### 2018-05-10 inst/documents/paper3/paper3_rotations.tex
pap3

--- 2018-05-08 inst/documents/paper3/paper3_rotations.Rmd
@@
----
-title: "Insecticide rotations delay evolution of resistance in a minority of model runs when compared to sequential use."
-# output: 
-#   pdf_document:
-#     keep_tex: true
-#     fig_caption: true
-#     latex_engine: pdflatex
+\documentclass[11pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Insecticide rotations delay evolution of resistance in a minority of model runs when compared to sequential use.},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{longtable,booktabs}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-output: word_document
-#author: "Andy South and Ian Hastings"
-#date: "`r Sys.Date()`"
-bibliography: library.bib
-csl: biomed-central.csl
-fontsize: 11pt
-spacing: double
-# below worked in test_linenumbers.Rmd
-header-includes:
-   - \usepackage{setspace}
-   - \doublespacing
-   - \usepackage{lineno}
-   - \linenumbers
----
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+  \title{Insecticide rotations delay evolution of resistance in a minority of
+model runs when compared to sequential use.}
+  \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
+
+\usepackage{setspace}
+\doublespacing
+\usepackage{lineno}
+\linenumbers
+
+\begin{document}
+\maketitle
 
@@
 
-Department of Parasitology, Liverpool School of Tropical Medicine, Liverpool L3 5QA, UK.
+Department of Parasitology, Liverpool School of Tropical Medicine,
+Liverpool L3 5QA, UK.
 
-southandy@gmail.com, andy.south@lstmed.ac.uk
+\href{mailto:southandy@gmail.com}{\nolinkurl{southandy@gmail.com}},
+\href{mailto:andy.south@lstmed.ac.uk}{\nolinkurl{andy.south@lstmed.ac.uk}}
 
-ian.hastings@lstmed.ac.uk
+\href{mailto:ian.hastings@lstmed.ac.uk}{\nolinkurl{ian.hastings@lstmed.ac.uk}}
 
-\* Corresponding author
+* Corresponding author
 
-```{r, eval=TRUE, include=FALSE}
-#!!!! PAPER PROTOCOL
-#!!!! submission MS, knit to word_document 
-#!!!!  in WORD: delete figs, set double space, add line numbers 
-#!!!!  (spacing & line numbers work for pdf output but not for word)
-library(knitr)
-opts_chunk$set(dev="tiff",
-               dev.args=list(compression="lzw"),
-               dpi=300,
-               cache=FALSE,
-               fig.path='pap3figs/')
-```
+\subsection{Abstract}\label{abstract}
 
-```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
-  library(rotations)
-  library(ggplot2)
-  library(gridExtra)
+Insecticide resistance threatens the control of the vectors of dangerous
+diseases including malaria, dengue and zika. Recent increases in
+insecticide resistance in public health are an evolutionary process
+caused by sustained exposure of insect populations to a small number of
+available insecticides. Resistance is a particular problem in public
+health as compared to agriculture because insecticides need to be longer
+lasting to provide affordable protection and there are fewer available
+active ingredients. Efforts to limit the development of insecticide
+resistance are grouped under the term Insecticide Resistance Management
+(IRM). The main strategies advocated to reduce the development of
+resistance are 1) rotations, 2) sequences, 3) mixtures and 4) mosaics.
+Rotations regularly switch between the use of different insecticides
+with a short time step (one or a few years) irrespective of resistance
+levels. Sequences, in contrast, switch from one insecticide to another
+only when resistance levels have reached a critical, defined threshold
+usually over a longer timescale.
 
-  outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
+Rotations are often advocated as one of the best options for limiting
+the development of resistance.
 
-  # load simulation outputs for some plots
-  # beware that each object is called df_res_all
-  # some plots are created by directly running the model
-  load("df_res_all_rot_nocost_nomig1000.rda") #nocost nomig  
-  df_nocost_nomig <- df_res_all
-  load("df_res_all_rot10000.rda") #cost
-  df_cost <- df_res_all
-  load("df_res_all_rot_mig10000.rda") #mig 10k
-  df_mig <- df_res_all
-  load("df_res_all_rot_cost_mig10000.rda") #costmig 10k
-  df_cost_mig <- df_res_all
+Testing different IRM strategies in the field is difficult and there has
+been little recent modelling work on insecticide rotations.
 
-```
+In this paper a model is described allowing rotations and sequences to
+be compared in terms of their effect on the evolution of insecticide
+resistance. The model is used to develop a mechanistic understanding of
+when and why insecticide resistance is likely to evolve faster with
+rotations or sequences.
 
-## Abstract
+The results suggest that under more likely circumstances than not the
+evolution of insecticide resistance will reach resistance thresholds at
+very similar times for rotations and sequences.
 
-Insecticide resistance threatens the control of the vectors of dangerous diseases including malaria, dengue and zika.
-Recent increases in insecticide resistance in public health are an evolutionary process caused by sustained exposure of insect populations to a small number of available insecticides.
-Resistance is a particular problem in public health as compared to agriculture because insecticides need to be longer lasting to provide affordable protection and there are fewer available active ingredients.
-Efforts to limit the development of insecticide resistance are grouped under the term Insecticide Resistance Management (IRM).
-The main strategies advocated to reduce the development of resistance are 1) rotations, 2) sequences, 3) mixtures  and 4) mosaics.
-Rotations regularly switch between the use of different insecticides with a short time step (one or a few years) irrespective of resistance levels.
-Sequences, in contrast, switch from one insecticide to another only when resistance levels have reached a critical, defined threshold usually over a longer timescale.
+However under a less common, but still plausible, set of circumstances
+there are predicted to be large adavnatages to using a rotations as
+compared to a sequence.
 
-Rotations are often advocated as one of the best options for limiting the development of resistance.
-
-Testing different IRM strategies in the field is difficult and there has been little recent modelling work on insecticide rotations.
-
-In this paper a model is described allowing rotations and sequences to be compared in terms of their effect on the evolution of insecticide resistance.
-The model is used to develop a mechanistic understanding of when and why insecticide resistance is likely to evolve faster with rotations or sequences.
-
-The results suggest that under more likely circumstances than not the evolution of insecticide resistance will reach resistance thresholds at very similar times for rotations and sequences.
-
-However under a less common, but still plausible, set of circumstances there are predicted to be large adavnatages to using a rotations as compared to a sequence.
-
-The advantages to a rotation only occur when there are costs of resistance or dispersal from untreated areas, and there is high dominance of cost and low dominance of selection.
+The advantages to a rotation only occur when there are costs of
+resistance or dispersal from untreated areas, and there is high
+dominance of cost and low dominance of selection.
 
@@
 
-Other operational factors will favour rotations or sequences aside from the implications for the evolution of resistance.
-Developing an understanding of the evolutionary implications allows a more explicit consideration of these other factors on their own merits.
+Other operational factors will favour rotations or sequences aside from
+the implications for the evolution of resistance. Developing an
+understanding of the evolutionary implications allows a more explicit
+consideration of these other factors on their own merits.
 
@@
 
-## Introduction
-Despite much field and modelling work on the evolution of resistance under different insecticide or drug strategies [@Consortium2013] there is a small evidence base to support decisions about the use of insecticides in public health [@Sternberg2017].
+\subsection{Introduction}\label{introduction}
 
-Rotations are advocated as one of the most favoured approaches for Insecticide Resistance Management (IRM)[@WHO2012].
+Despite much field and modelling work on the evolution of resistance
+under different insecticide or drug strategies {[}1{]} there is a small
+evidence base to support decisions about the use of insecticides in
+public health {[}2{]}.
 
-In a rotation the insecticide used is changed on a set time interval irrespective of resistance levels.
-Sequential use, in contrast, denotes the practice of using an insecticide until resistance to it reaches a certain threshold and then switching to another.
-Different rotation time intervals can be considered, e.g.
-in vector control for Malaria, Indoor Residual Spraying may be repeated every year (around 10 anopheles mosquito generations) and Insecticide Treated Nets may be replaced every two years (around 20 generations).
-Multiples of these intervals could be used as the rotation interval.
-Multiple generational rotations like this contrast with the situation in agriculture where a rotation is sometimes more tightly defined as having an interval of a single generation [@Sudo2017].
+Rotations are advocated as one of the most favoured approaches for
+Insecticide Resistance Management (IRM){[}3{]}.
 
-Here a modern modelling approach is described to allow an assessment of the potential benefits, in terms of the evolution of resistance, of rotations over sequences.
+In a rotation the insecticide used is changed on a set time interval
+irrespective of resistance levels. Sequential use, in contrast, denotes
+the practice of using an insecticide until resistance to it reaches a
+certain threshold and then switching to another. Different rotation time
+intervals can be considered, e.g.~in vector control for Malaria, Indoor
+Residual Spraying may be repeated every year (around 10 anopheles
+mosquito generations) and Insecticide Treated Nets may be replaced every
+two years (around 20 generations). Multiples of these intervals could be
+used as the rotation interval. Multiple generational rotations like this
+contrast with the situation in agriculture where a rotation is sometimes
+more tightly defined as having an interval of a single generation
+{[}4{]}.
 
-## Methods
+Here a modern modelling approach is described to allow an assessment of
+the potential benefits, in terms of the evolution of resistance, of
+rotations over sequences.
 
-A population genetic model was developed tracking the change in resistance allele frequency as consequence of the genetic determination of the resistance phenotype, and the deployment policy for the insecticides.
-The model uses a similar approach to that previously used to compare insecticide mixtures and sequences [@Levick2017,@South2018].
-The implementation is simpler here because for rotations and sequences there is only a need to follow one insecticide at a time and linkage disequilibrium can be ignored.
-The simpler implementation allows the model to be run for an unlimited number of insecticides.
-The algebra behind the model is described in the supplementary information.
+\subsection{Methods}\label{methods}
 
-1. Sequences : one insecticide is used until resistance threshold frequency (0.5 in this case) is reached, then switch to a new insecticide.
+A population genetic model was developed tracking the change in
+resistance allele frequency as consequence of the genetic determination
+of the resistance phenotype, and the deployment policy for the
+insecticides. The model uses a similar approach to that previously used
+to compare insecticide mixtures and sequences {[}5{]}. The
+implementation is simpler here because for rotations and sequences there
+is only a need to follow one insecticide at a time and linkage
+disequilibrium can be ignored. The simpler implementation allows the
+model to be run for an unlimited number of insecticides. The algebra
+behind the model is described in the supplementary information.
 
-2. Rotations : use insecticide for set time (i.e. number of insect generations), switch to another.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Sequences : one insecticide is used until resistance threshold
+  frequency (0.5 in this case) is reached, then switch to a new
+  insecticide.
+\item
+  Rotations : use insecticide for set time (i.e.~number of insect
+  generations), switch to another.
+\end{enumerate}
 
-In both cases the choice of the new insecticide to use was made by going through the list of available insecticides and choosing the next one for which resistance was below the threshold.
-Thus there is the potential to switch back to insecticides that had been used earlier in a model run if resistance had not reached the threshold or if resistance had declined below the threshold during a period when not in use.
-This means that the “sequential” policy is effectively a “rotate when resistant” policy but we use the former term for consistency with previous work.
-*todo cite different trems from Rex consortium review* For both rotations and sequences the simulation continues until no more insecticides remain below their resistance thresholds.
+In both cases the choice of the new insecticide to use was made by going
+through the list of available insecticides and choosing the next one for
+which resistance was below the threshold. Thus there is the potential to
+switch back to insecticides that had been used earlier in a model run if
+resistance had not reached the threshold or if resistance had declined
+below the threshold during a period when not in use. This means that the
+``sequential'' policy is effectively a ``rotate when resistant'' policy
+but we use the former term for consistency with previous work.
+\emph{todo cite different terms from Rex consortium review} For both
+rotations and sequences the simulation continues until no more
+insecticides remain below their resistance thresholds.
 
-Within the model inputs effect fitness as shown in Figure 1, differences in fitness lead to the changes in allele frequencies over time.
+Within the model inputs effect fitness as shown in Figure 1, differences
+in fitness lead to the changes in allele frequencies over time.
 
-**Table 1. Model inputs and the ranges used in simulations**
+\textbf{Table 1. Model inputs and the ranges used in simulations}
 
-Input  | Description | Min | Max
----------------------- | ------------------------------------- | ----- | -----
-Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide | 0.5 | 1
-Exposure | proportion of females exposed to insecticide | 0.4 | 0.9
-Male exposure | proportion of males exposed to insecticide as proportion of females | 0 | 1
-Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide | 0.1 | 0.9
-Dominance of resistance | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide | 
-Frequency  | frequency of resistance alleles within the population | 0.005 | 0.1
-Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide | 0 | 0.1
-Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide | 0.1 | 0.9
-Insecticide number | number of independent insecticides available | 2 | 5  
-Rotation interval | number of generations to use insecticide before switching | 5 | 50
-Coverage | for scenarios with dispersal, the proportion of insects in the treated area rather than untreated refugia |  0.1 | 0.9
-Dispersal | proportion of population exchanged between treated and untreated areas per generation, 0=none, 1=random mixing |  0.1 | 0.9
+\begin{longtable}[]{@{}llll@{}}
+\toprule
+\begin{minipage}[b]{0.28\columnwidth}\raggedright\strut
+Input\strut
+\end{minipage} & \begin{minipage}[b]{0.46\columnwidth}\raggedright\strut
+Description\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
+Min\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
+Max\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Effectiveness\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+proportion of susceptible (SS) insects killed by exposure to
+insecticide\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.5\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+1\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+proportion of females exposed to insecticide\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.4\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.9\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Male exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+proportion of males exposed to insecticide as proportion of
+females\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+1\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Resistance restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+ability of resistance (RR) to restore fitness of insects exposed to
+insecticide\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.9\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Dominance of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+sets fitness of heterozygous (SR) insects between that of SS \& RR in
+presence of insecticide\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Frequency\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+frequency of resistance alleles within the population\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.005\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.1\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Cost of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+fitness of resistant (RR) insects in absence of insecticide\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.1\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Dominance of cost\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+sets fitness of heterozygous (SR) insects between that of SS \& RR in
+absence of insecticide\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.9\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Insecticide number\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+number of independent insecticides available\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+2\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+5\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Rotation interval\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+number of generations to use insecticide before switching\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+5\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+50\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Coverage\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+for scenarios with dispersal, the proportion of insects in the treated
+area rather than untreated refugia\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.9\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+Dispersal\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+proportion of population exchanged between treated and untreated areas
+per generation, 0=none, 1=random mixing\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+0.9\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 
-1. baseline : no resistance fitness costs or dispersal to/from untreated refugia
-2. fitness costs of resistance added
-3. dispersal link to untreated areas (no costs)
-4. costs and dispersal
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  baseline : no resistance fitness costs or dispersal to/from untreated
+  refugia
+\item
+  fitness costs of resistance added
+\item
+  dispersal link to untreated areas (no costs)
+\item
+  costs and dispersal
+\end{enumerate}
 
-For each scenario 10,000 runs were performed randomly selecting inputs according to the ranges specified in Table 1 using a uniform distribution in all cases.
+For each scenario 10,000 runs were performed randomly selecting inputs
+according to the ranges specified in Table 1 using a uniform
+distribution in all cases.
 
-#### Assessing relative performance of insecticide use strategies
+\paragraph{Assessing relative performance of insecticide use
+strategies}\label{assessing-relative-performance-of-insecticide-use-strategies}
 
-*todo change this bit*
-It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative.
-For this we summed the number of generations when an insecticide was deployed and the resistance allele frequency for that insecticide was below 0.5.
-This allowed comparison between strategies using a variable number of insecticides.
-Previously time-to-resistance, namely the number of generations it takes to reach a defined resistance threshold (usually a resistance allele frequency of 0.5), was used.
-Using time-to-resistance worked when considering just two insecticides, but when the number of insecticides is a variable it is not sufficient.
+\emph{todo change this bit} It is necessary to choose a measure to
+quantify the performance of an insecticide-use strategy and enable
+comparison with an alternative. For this we summed the number of
+generations when an insecticide was deployed and the resistance allele
+frequency for that insecticide was below 0.5. This allowed comparison
+between strategies using a variable number of insecticides. Previously
+time-to-resistance, namely the number of generations it takes to reach a
+defined resistance threshold (usually a resistance allele frequency of
+0.5), was used. Using time-to-resistance worked when considering just
+two insecticides, but when the number of insecticides is a variable it
+is not sufficient.
 
-## Results
+\subsection{Results}\label{results}
 
-Example model runs are shown with no costs or dispersal (Fig 1), with just resistance costs (Fig 2) and with just dispersal from untreated refugia (Fig 3).
+\paragraph{Results1. form of resistance evolution with and without costs
+and
+dispersal.}\label{results1.-form-of-resistance-evolution-with-and-without-costs-and-dispersal.}
 
-*todo describe these example runs first, before talking about overall results*
+Example model runs are shown with no costs or dispersal (Fig 2), with
+just resistance costs (Fig 3) and with just dispersal from untreated
+refugia (Fig 4).
 
-#### Results1. form of resistance evolution with and without costs and dispersal.
+When there are no resistance fitness costs or dispersal links to
+untreated refugia resistance frequencies do not decline when an
+insecticide is not in use (Fig 2). Thus for sequences once resistance
+thresholds are reached the resistance frequencies remain at that level
+and the insecticide cannot be re-used. For rotations resistance
+frequencies step upwards when insecticides are in use and remain at a
+plateau when they are not. If the rotation interval is short enough then
+the insecticide can be used a few times before the resistance thresholds
+are reached.
 
-When there are no resistance fitness costs or dispersal links to untreated refugia resistance frequencies do not decline when an insecticide is not in use (Fig 1).
-Thus for sequences once resistance thresholds are reached the resistance frequencies remain at that level and the insecticide cannot be re-used.
-For rotations resistance frequencies step upwards when insecticides are in use and remain at a plateau when they are not.
-If the rotation interval is short enough then the insecticide can be used a few times before the resistance thresholds are reached.
+\emph{todo continue describing the example runs first, before talking
+about overall results}
 
-### Results2 : comparing rotations and sequences across model runs.
+\subsubsection{Results2 : comparing rotations and sequences across model
+runs.}\label{results2-comparing-rotations-and-sequences-across-model-runs.}
 
-In all the example runs shown the resistance threshold (allele frequency of 0.5) was reached within the course 500 generations which was the maximum limit imposed in the model.
-Across all model runs there were many in which the thresholds were not reached in this time period.
+In all the example runs shown the resistance threshold (allele frequency
+of 0.5) was reached for all insecticides within the course 500
+generations which was the maximum limit imposed in the model. There were
+many model runs {[}\emph{quantify later when run finalised}{]} in which
+resistance frequencies for at least some insecticides were below
+thresholds at the end of 500 generations. Such a run could be considered
+a successful strategy given the conditions of the scenario because
+insecticides below the resistance threshold are still available to be
+used.
 
-*but first need to explain that sequences pretty much never better*
-*so show the modified histograms first, i.e. don't filter out the threshold not reached runs*
+Thus to aid the comparison of rotations and sequences model runs can be
+broadly classified into four groups : 1. both strategies succeeded : for
+both rotations and sequences resistance thresholds for all insecticides
+not reached within 500 generations. 2. neither strategy succeeded 3.
+rotation only succeeded 4. sequence only succeeded
 
-Thus to aid the comparison of rotations and sequences model runs can be broadly classified into three groups :
+The number of model runs in each of these groups for the different
+scenarios are shown in Fig 5.
 
-When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold (Fig 4.1).
+\emph{todo : change the \%s because `succeeded' runs now left in} When
+there were no costs or dispersal there was no difference between
+rotations and sequences in the number of generations below the
+resistance threshold (Fig 6A). With fitness costs included 67\% of model
+runs, similar to those in the previous section, gave no difference
+between rotations and sequences in the number of generations below the
+resistance threshold. In contrast, 7\% of model runs generated an
+advantage of greater than 20\% for rotations over sequences (Fig 6B).
+Dispersal from untreated refugia gives smaller benefits of rotations
+over sequences and in fewer runs (Fig 6C). With both costs and dispersal
+the benefits of rotations are similar to when runs have costs alone (Fig
+6D).
 
-*todo : maybe I should just leave the 'succeeded' runs in here which would change these %s slightly*
-With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold.
-In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).
+\paragraph{Runs where rotations have substantial advantage over
+sequences}\label{runs-where-rotations-have-substantial-advantage-over-sequences}
 
-Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 4.3).
+\emph{Talk about classing runs into both `succeeded', both `failed' or
+just rotation succeeded. }show violin plots here ?*
 
-With both costs and dispersal the benefits of rotations are similar to when runs have costs alone (Fig 4.4).
+As shown in Fig 4.2 some runs with costs of resistance can show
+substantial advantages for rotations over sequences. Examples of one
+such run is shown in Fig 5.
 
-#### Runs where rotations have substantial advantage over sequences
+A Partial Rank Correlation analysis identified that of the inputs,
+dominance of cost and dominance of selection had the greatest influence
+on the difference between rotations and sequences when resistance costs
+were included (Fig 6). A combination of high values of dominance of cost
+and low values for dominance of selection are required to give the
+greatest advantages to rotations, but do not guarantee this advantage
+(Fig 7).
 
-*Talk about classing runs into both 'succeeded', both 'failed' or just rotation succeeded.
-*show violin plots here ?*
+\subsection{Discussion}\label{discussion}
 
-As shown in Fig 4.2 some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 5.
+Comparing time below resistance thresholds for rotations and sequences,
+the model shows :
 
-A Partial Rank Correlation analysis identified that of the inputs, dominance of cost and dominance of selection had the greatest influence on the difference between rotations and sequences when resistance costs were included (Fig 6).
-A combination of high values of dominance of cost and low values for dominance of selection are required to give the greatest advantages to rotations, but do not guarantee this advantage (Fig 7).
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  rotations and sequences the same when no costs or dispersal
+\item
+  with resistance costs, 67\% of runs the same for rotations and
+  sequences, but 6-7\% of runs have \textgreater{} 20\% advantage for
+  rotations
+\item
+  with dispersal from untreated refugia but no costs, small advantage of
+  rotations but very few runs \textless{} 0.1\% have \textgreater{}20\%
+  advantage.
+\end{enumerate}
 
-## Discussion
+This can be understood intuitively. When there are no costs, the
+dynamics are the same, it is simply that sequential deployment proceeds
+in a few, larger steps while rotations proceeds in more numerous smaller
+steps; both reach the same point in around the same time (e.g.~figs 1
+and 3). However dominance effects on cost and resistance can act to keep
+frequencies low in a rotations policy. Recall that at low frequencies of
+resistance most resistance alleles are present as heterozygotes RS and
+only a vey small proportions heterozygotes. However costs are dominant
+and still act at low frequencies. This is illustrated on Figure Y.
+Relatively rapid rotations keep the allele frequencies at the LH side of
+Figure Y where costs during the periods when the insecticide is not
+deployed greatly outweigh the selection for IR during the periods when
+the insecticide is deployed. We believe the explanation is therefore
+that rotations keep allele frequencies in the parameter space to the
+left of Figure Y and hence acts to slow the evolution of IR compared to
+sequences which allow frequencies to rise to relatively high levels
+before the insecticide is replaced. (does this sound reasonable??). Note
+also that starting frequency, as expected, had an impact (Figure 7),
+lower starting frequencies slightly favouring rotations by allowing more
+time in the ``low frequency'' parameter space.
 
-Comparing time below resistance thresholds for rotations and sequences, the model shows :
+In rare circumstances, when there are costs of resistance, rotations can
+lead to the resistance frequencies for a group of insecticides being
+kept below resistance thresholds for the length of the simulations where
+sequences run out of insecticides within around 100 generations. Is it
+possible that these rare circumstances where parameter values are
+`just-right' could be implemented operationally ? The problem is that
+these rare circumstances rely on high costs of resistance, high
+dominance of these costs and low dominance of the selection for the
+resistance itself. These parameters are all not easy to measure and
+there isn't agreement on their likely values {[}7{]}.
 
-1. rotations and sequences the same when no costs or dispersal
-2. with resistance costs, 67% of runs the same for rotations and sequences, but 6-7% of runs have > 20% advantage for rotations
-3. with dispersal from untreated refugia but no costs, small advantage of rotations but very few runs < 0.1% have >20% advantage.
+Note however that this combination of recessive resistance and high
+dominance costs are exactly the features most desirable in any
+insecticide irrespective of how it is deployed. We discuss fitness below
+but note here that it is very difficult to maintain insecticide
+concentration at the level required to make resistance recessive.
+Concentrations decay post-application in most deployment strategies and
+this cause resistance to gradually change from being recessive to being
+dominant; see Figure 1 of {[}5{]} and our more extensive discussion in
+{[}cite South et al WoS manuscript{]}.
 
-This can be understood intuitively.
-When there are no costs, the dynamics are the same, it is simply that sequential deployment proceeds in a few, larger steps while rotations proceeds in more numerous smaller steps; both reach the same point in around the same time (e.g.
-figs 1 and 3).
-However dominance effects on cost and resistance can act to keep frequencies low in a rotations policy.
-Recall that at low frequencies of resistance most resistance alleles are present as heterozygotes RS and only a vey small proportions heterozygotes.
-However costs are dominant and still act at low frequencies.
-This is illustrated on Figure Y.
-Relatively rapid rotations keep the allele frequencies at the LH side of Figure Y where costs during the periods when the insecticide is not deployed greatly outweigh the selection for IR during the periods when the insecticide is deployed.
-We believe the explanation is therefore that rotations keep allele frequencies in the parameter space to the left of Figure Y and hence acts to slow the evolution of IR compared to sequences which allow frequencies to rise to relatively high levels before the insecticide is replaced.
-(does this sound reasonable??).
-Note also that starting frequency, as expected, had an impact (Figure 7), lower starting frequencies slightly favouring rotations by allowing more time in the “low frequency” parameter space.
+The importance of resistance costs for substantial evolutionary benefits
+of rotations over sequences points to the importance of the debate over
+the fitness costs of insecticide resistance e.g. {[}7{]}. \emph{extend
+this discussion of fitness costs}
 
-In rare circumstances, when there are costs of resistance, rotations can lead to the resistance frequencies for a group of insecticides being kept below resistance thresholds for the length of the simulations where sequences run out of insecticides within around 100 generations.
-Is it possible that these rare circumstances where parameter values are 'just-right' could be implemented operationally ?
-The problem is that these rare circumstances rely on high costs of resistance, high dominance of these costs and low dominance of the selection for the resistance itself.
-These parameters are all not easy to measure and there isn't agreement on their likely values [@Kliot2012, @Bourguet2000].
+We also found frequency of rotations had little effect. This is
+important in the control of vectors of human diseases such as malaria
+and dengue. These predominantly occur in resource-poor areas often with
+poor healthcare infrastructure. Experience suggests that rotation
+targets would often be missed (e.g rotations occur approximately every
+15 generations rather than the planned ``official'' 10 generations) but
+our simulations suggest this would not fatally undermine a rotations
+policy.
 
-Note however that this combination of recessive resistance and high dominance costs are exactly the features most desirable in any insecticide irrespective of how it is deployed.
-We discuss fitness below but note here that it is very difficult to maintain insecticide concentration at the level required to make resistance recessive.
-Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant; see Figure 1 of <cite Levick et al> and our more extensive discussion in <cite South et al WoS manuscript).
+\subsubsection{Caveats}\label{caveats}
 
-The importance of resistance costs for substantial evolutionary benefits of rotations over sequences points to the importance of the debate over the fitness costs of insecticide resistance e.g.
-[@Kliot2012].
-*extend this discussion of fitness costs*
+The model does not include the potential effects of modifier genes
+ameliorating the costs of resistance alleles. To include this would
+require a more detailed model tracking linkage disequilibrium between
+each resistance locus and its modifier. It has been suggested that
+rotations could be favoured as a strategy because they keep resistance
+frequences lower and restrict selection pressure for modifier genes
+{[}3{]}. However, there is little evidence to support the existence of
+such modifier genes in insectide resistance {[}{]}. Even with the
+existence of modifier genes it is unlikely that they would provide much
+of an advantage to rotations in the scenarios we describe. In the
+modelled scenarios the use of an insecticide is stopped when the
+resistance frequency for that insecticide reaches 0.5. Thus in all
+scenarios resistance frequencies do not remain at high levels for long.
+In this situation it seems unlikely that modifiers would be selected
+for.
 
-We also found frequency of rotations had little effect.
-This is important in the control of vectors of human diseases such as malaria and dengue.
-These predominantly occur in resource-poor areas often with poor healthcare infrastructure.
-Experience suggests that rotation targets would often be missed (e.g rotations occur approximately every 15 generations rather than the planned “official” 10 generations) but our simulations suggest this would not fatally undermine a rotations policy.
+\subsubsection{Conclusions}\label{conclusions}
 
-### Caveats
+In their effect on the evolution of insecticide resistance, rotations
+and sequences are most often the same. If costs of resistance are
+included, infrequently (less than 7\% of model runs set between
+plausible limits) there are substantial benefits to the rotation
+strategy over a sequence. There was never a substantial benefit of a
+sequence over a rotation. Whether or not insecticide rotations are
+predicted to offer sizeable benefits over sequences is dependent on
+coarse and fine scale issues about the nature of insecticide resistance
+that are yet to be agreed upon. The model shows that resistance costs
+are necessary to provide a substantial evolutionary advantage to
+rotations. However even when these requirements are satisfied particular
+combinations of inputs are required to generate the advantage of
+rotations.
 
-The model does not include the potential effects of modifier genes ameliorating the costs of resistance alleles.
-To include this would require a more detailed model tracking linkage disequilibrium between each resistance locus and its modifier.
-It has been suggested that rotations could be favoured as a strategy because they keep resistance frequences lower and restrict selection pressure for modifier genes [@WHO2012].
-However, there is little evidence to support the existence of such modifier genes in insectide resistance [].
-Even with the existence of modifier genes it is unlikely that they would provide much of an advantage to rotations in the scenarios we describe.
-In the modelled scenarios the use of an insecticide is stopped when the resistance frequency for that insecticide reaches 0.5.
-Thus in all scenarios resistance frequencies do not remain at high levels for long.
-In this situation it seems unlikely that modifiers would be selected for.
+The model only compares rotations and sequences in terms of their effect
+on the evolution of resistance. There are other operational reasons for
+why a sequence or a rotation may be favoured.
 
-### Conclusions
+In summary, we show that rotations are invariably better than sequential
+use, although the difference may often be small (figure 4); certainly
+rotations should not be seen as a panacea capable of removing the threat
+of resistance evolution. Rotations may have a large impact if fitness
+costs are prints and dominant (these are genetic/physiological factors
+outside our control) and if resistance can be kept recessive. The latter
+is under our control to some extent (fig 1 of Levick et al ) but it is
+extremely difficult to deploy and maintain insecticide concentrations at
+levels that ensure recessively (cite WoS ms): as concentrations decline
+resistance become dominant and any befits of rotations largely
+disappear.
 
-In their effect on the evolution of insecticide resistance, rotations and sequences are most often the same.
-If costs of resistance are included, infrequently (less than 7% of model runs set between plausible limits) there are substantial benefits to the rotation strategy over a sequence.
-There was never a substantial benefit of a sequence over a rotation.
-Whether or not insecticide rotations are predicted to offer sizeable benefits over sequences is dependent on coarse and fine scale issues about the nature of insecticide resistance that are yet to be agreed upon.
-The model shows that resistance costs are necessary to provide a substantial evolutionary advantage to rotations.
-However even when these requirements are satisfied particular combinations of inputs are required to generate the advantage of rotations.
+\subsection{TODO}\label{todo}
 
-The model only compares rotations and sequences in terms of their effect on the evolution of resistance. There are other operational reasons for why a sequence or a rotation may be favoured.
-
-In summary, we show that rotations are invariably better than sequential use, although the difference may often be small (figure 4); certainly rotations should not be seen as a panacea capable of removing the threat of resistance evolution.
-Rotations may have a large impact if fitness costs are prints and dominant (these are genetic/physiological factors outside our control)  and if resistance can be kept recessive.
-The latter is under our control to some extent (fig 1 of Levick et al ) but it is extremely difficult to deploy and maintain insecticide concentrations at levels that ensure recessively (cite WoS ms): as concentrations decline resistance become dominant and any befits of rotations largely disappear.
-
-## TODO
-*try to come up with brief name for model output to make it easier to refer to it later*
-*add caveat about polygenic resistance*
-*The caption to each graph needs to have the list of parameter values used to generate them*
-*work on Ians fig Y to help explanation*
+\emph{add to methods description of how fitness determined in model,
+move \& modify fig 1} \emph{try to come up with brief name for model
+output to make it easier to refer to it later} \emph{add caveat about
+polygenic resistance} \emph{The caption to each graph needs to have the
+list of parameter values used to generate them} \emph{work on Ians fig Y
+to help explanation}
 
@@
 
-## Figures
+\subsection{Figures}\label{figures}
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1.
+\begin{figure}
+
+\includegraphics{pap3figs/Fig1-1} \hfill{}
+
+\caption{Figure 1.
+[todo need to modify caption from MJ paper, start by saying that only in left panel during application of the insecticide] The effect of model inputs on the fitness of genotypes for a single insecticide.
+Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
+Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
+For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of resistance determines how the fitness for SR lies between that of SS and RR.
+For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
+In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of resistance=0.7 which sets the fitness of the SR closer to RR than SS.
+Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR.}\label{fig:Fig1}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}
+
+{\centering \includegraphics{pap3figs/Fig2-1}
+
+}
+
+\caption{Figure 2.
 Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia.
-The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached.
-In this example the rotation and the sequence reach the endpoint, resistance frequency > 0.5 to all insecicides, at the same time.
+The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence in which the insecticide is changed when a resistance threshold of 0.5 is reached.
+In this example the rotation and the sequence reach the endpoint, with all insecticides having a resistance frequency above 0.5, at the same time.
 For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again.
@@
 There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use.
-All other inputs are kept constant between the rotation and the sequence."}
-
-# revisit how best to do model runs from within the paper
-
-linputs <- list()
-
-linputs$n_insecticides <- 3
-linputs$coverage <- 1
-#linputs$migration <- 0.1
-linputs$cost <-     0 
-linputs$max_gen <-  200
-linputs$expo_hi <-  0.6
-linputs$eff <-      0.6
-linputs$dom_sel <-  1
-linputs$dom_cos <-  1
-linputs$plot <-     FALSE
-
-linputs$rot_interval=0
-dfseq <- do.call(run_rot, linputs)
-linputs$rot_interval=10
-dfrot <- do.call(run_rot, linputs)
-
-ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
-ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
-library(gridExtra)
-grid.arrange(ggrot, ggseq, ncol=1)
-
-```
+All other inputs are kept constant between the rotation and the sequence.
+Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1.}\label{fig:Fig2}
+\end{figure}
 
@@
 
-```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2.
+\begin{figure}
+
+{\centering \includegraphics{pap3figs/Fig3-1}
+
+}
+
+\caption{Figure 3.
 Comparing the increase in resistance frequencies over time for a rotation and a sequence when there are resistance fitness costs.
@@
 The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold.
-All other inputs are kept constant between the rotation and the sequence."}
-
-# revisit how best to do model runs from within the paper
-
-linputs <- list()
-
-linputs$n_insecticides <- 3
-linputs$coverage <- 1
-#linputs$migration <- 0.1
-linputs$cost <-     0.05 
-linputs$max_gen <-  200
-linputs$expo_hi <-  0.6
-linputs$eff <-      0.6
-linputs$dom_sel <-  1
-linputs$dom_cos <-  1
-linputs$plot <-     FALSE
-
-linputs$rot_interval=0
-dfseq <- do.call(run_rot, linputs)
-linputs$rot_interval=10
-dfrot <- do.call(run_rot, linputs)
-
-ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
-ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
-library(gridExtra)
-grid.arrange(ggrot, ggseq, ncol=1)
-
-```
+All other inputs are kept constant between the rotation and the sequence.
+Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0.05, dominance of cost=1.}\label{fig:Fig3}
+\end{figure}
 
@@
 
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3.
+\begin{figure}
+
+{\centering \includegraphics{pap3figs/Fig4-1}
+
+}
+
+\caption{Figure 4.
 Comparing the increase in resistance frequencies over time for a rotation and a sequence with dispersal from untreated refugia.
@@
 The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold.
-All other inputs are kept constant between the rotation and the sequence."}
-
-# revisit how best to do model runs from within the paper
-
-linputs <- list()
-
-linputs$n_insecticides <- 3
-linputs$coverage <- 0.5
-linputs$migration <- 0.1
-linputs$cost <-     0 
-linputs$max_gen <-  200
-linputs$expo_hi <-  0.6
-linputs$eff <-      0.6
-linputs$dom_sel <-  1
-linputs$dom_cos <-  1
-linputs$plot <-     FALSE
-
-linputs$rot_interval=0
-dfseq <- do.call(run_rot, linputs)
-linputs$rot_interval=10
-dfrot <- do.call(run_rot, linputs)
-
-ggrot <- rot_plot_resistance(dfrot, plot_refuge=TRUE, logy=TRUE, plot=FALSE)
-ggseq <- rot_plot_resistance(dfseq, plot_refuge=TRUE, logy=TRUE, plot=FALSE)
-library(gridExtra)
-grid.arrange(ggrot, ggseq, ncol=1)
-
-```
+All other inputs are kept constant between the rotation and the sequence.
+Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0, coverage=0.5, dispersal=0.1.}\label{fig:Fig4}
+\end{figure}
 
@@
 
-```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4.
-Effect of cost and dispersal on the difference between rotations and sequences from model runs.
-The y axis shows the number of model runs for the different caegories of difference between rotations and sequences.
-4.1 when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results.
-4.2, 4.4 when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences.
-For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences.
-4.3 when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}
+\begin{figure}
 
-  # remove runs where thresholds not reached for seq & rot
-  # with cost_mig ~ 65% of runs don't reach thresholds
-  # could get this func to output perecnt not reaching
-  rem_thresh_not_reached <- function(x) {x %>% filter(gens_seq != max_gen | gens_rot != max_gen)}
+{\centering \includegraphics{pap3figs/Fig5-1}
 
-  df_nocost_nomig <- rem_thresh_not_reached(df_nocost_nomig)
-  df_cost <- rem_thresh_not_reached(df_cost) #6.6
-  df_mig <- rem_thresh_not_reached(df_mig) #0.05
-  df_cost_mig <- rem_thresh_not_reached(df_cost_mig) #6.9
+}
 
-  # add a column for percent difference between rot & seq
-  df_nocost_nomig <- mutate(df_nocost_nomig, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
-  df_cost <- mutate(df_cost, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
-  df_mig <- mutate(df_mig, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
-  df_cost_mig <- mutate(df_cost_mig, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
-
-  # calc % of runs where rotations > 20% better
-  pcent_runs_rot_20pc_better <- function(x){ x %>%
-                  filter(rot_minus_seq/gens_seq >= 0.2 ) %>%
-                  summarise(n=n()) %>% 
-                  #unlist() %>%
-                  mutate(pcent=n*100/nrow(x)) }
-
-  # pcent_runs_rot_20pc_better(df_nocost_nomig)
-  # pcent_runs_rot_20pc_better(df_cost)
-  # pcent_runs_rot_20pc_better(df_mig)
-  # pcent_runs_rot_20pc_better(df_cost_mig)
-
-  # log scale overemphasised low values
-  # sqrt scale looked good but Ian didn't like
-
-  custom_bins <- c(-20,-1,1,20,100,450)
-  df_nocost_nomig$rot_adv_bins <- cut(df_nocost_nomig$pcent_rot_adv, custom_bins)
-  df_cost$rot_adv_bins <- cut(df_cost$pcent_rot_adv, custom_bins)
-  df_mig$rot_adv_bins <- cut(df_mig$pcent_rot_adv, custom_bins)
-  df_cost_mig$rot_adv_bins <- cut(df_cost_mig$pcent_rot_adv, custom_bins)
-
-  bin_labs <- levels(df_cost$rot_adv_bins)
-
-  #gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=pcent_rot_adv)) + 
-  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=rot_adv_bins)) +
-
-             geom_bar() +
-             #xlim(-20,400) +
-             xlim(bin_labs) +
-             xlab('') +
-             ylab('count') +
-             ggtitle('1. no resistance costs or dispersal') +
-             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
-             theme_minimal() +
-             theme(panel.grid.minor.y=element_blank(),
-                   axis.text.x=element_blank())
-
-  gg_cost <- ggplot( df_cost, aes(x=rot_adv_bins)) +
-
-             geom_bar() +
-             #geom_histogram(binwidth=10) +
-             #xlim(-20,400) +
-             xlab('') +
-             ylab('count') +
-             ggtitle('2. with costs') +
-             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
-             #scale_x_continuous(breaks=c(-10,-1,0,1,20,100,400)) +
-             scale_x_discrete( limits=bin_labs,
-                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
-             theme_minimal() +
-             theme(panel.grid.minor.y=element_blank(),
-                   axis.text.x=element_blank())
-
-  gg_mig <- ggplot( df_mig, aes(x=rot_adv_bins)) +
-
-             geom_bar() +
-             xlim(bin_labs) +
-             xlab('') +
-             ylab('count') +
-             ggtitle('3. with dispersal refugia') +
-             theme_minimal() + 
-             theme(panel.grid.minor.y=element_blank(),
-                   axis.text.x=element_blank())
-
-  #gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_minus_seq)) + 
-  gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_adv_bins)) +
-
-             geom_bar() +
-             xlab('Percent benefit of rotation over sequence') +
-             ylab('count') +
-             ggtitle('4. with costs and dispersal') +
-             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
-             theme_minimal() +
-             scale_x_discrete( limits=bin_labs,
-                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+       
-             theme(panel.grid.minor.y=element_blank())
-
-  grid.arrange(gg_nocost_nomig, gg_cost, gg_mig, gg_cost_mig, ncol=1)
-
-```
+\caption{Figure 5.
+Success of model runs under different scenarios and strategies.
+Model runs classed as a success if there are still insecticides for which resistance is below the threshold of 0.5 after 500 generations (circa 50 years for Anopheles).}\label{fig:Fig5}
+\end{figure}
 
@@
 
-```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 5.
-Example of a model run where there is a substantial benefit of rotations over a sequence.
-For the sequence resistance thresholds are reached for all insecticides in a little over 100 generations or 10 years.
-For the rotation resistance levels for all insecticides are maintained well below resistance thresholds for the length of the simulation (500 generations or 50 years)."}
-# ? are these run ids before or after the runs that don't reach thresholds have been removed ?
-# seems OK that they are after other runs removed
-#scenarios_to_plot <- c(624,2167,602,4028,5771)
-scenarios_to_plot <- c(2167)
+\textbackslash{}begin\{figure\}
 
-for(scenario in scenarios_to_plot)
-{
-  #cat(scenario,"\n")
+\{\centering \includegraphics{pap3figs/Fig6-1}
 
-  linputs <- as.list( df_cost[scenario, 1:15] )
+\}
 
-  # run one scenario
-  dfres <- do.call(run_rot, linputs)
+\textbackslash{}caption\{Figure 6. Effect of cost and dispersal on the
+difference between rotations and sequences from model runs. The y axis
+shows the number of model runs for the different caegories of difference
+between rotations and sequences. A) when there are no resistance costs
+or dispersal from untreated refugia, rotations and sequences produce the
+same results. B),D) when costs are added there are infrequent model runs
+where rotations last longer below resistance thresholds than sequences.
+For between 6 and 7\% of runs including costs, rotations last more than
+20\% longer than sequences. C) when dispersal from untreated refugia is
+included without costs there are even less frequent runs where rotations
+last longer than sequences (in just 4 out of 10000 runs rotations lasted
+greater than 20\% longer than sequences).\}\label{fig:Fig6}
+\textbackslash{}end\{figure\}
 
-  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+\pagebreak
 
-  #now run as sequence
-  linputs$rot_interval <- 0
-  dfres <- do.call(run_rot, linputs)
+\begin{figure}
 
-  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
-
-  grid.arrange(ggrot, ggseq, ncol=1)
+{\centering \includegraphics{pap3figs/Fig7-1}
 
@@
 
-```
+\caption{Figure 7.
+Example of a model run where there is a substantial benefit of rotations over a sequence.
+For the sequence resistance thresholds are reached for all insecticides in a little over 100 generations or 10 years.
+For the rotation resistance levels for all insecticides are maintained well below resistance thresholds for the length of the simulation (500 generations or 50 years).
+Effectiveness=0.57, exposure=0.88, male exposure proportion=0.72, resistance restoration=0.68, dominance of resistance=0.04, cost=0.07, dominance of cost=0.59.}\label{fig:Fig7}
+\end{figure}
 
@@
 
-```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 6.
+\begin{figure}
+
+{\centering \includegraphics{pap3figs/Fig8-1}
+
+}
+
+\caption{Figure 8.
 Which model inputs have the greatest influence on the difference between rotations and sequences ?
@@
 Positive values indicate a positive influence on the benefit of rotations, only cost and dominance of cost in this case.
-Dominance of cost and dominance of selection have the greatest magnitude of influence."}
-
-library(sensitivity)
-
-# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
-# 
-# Arguments
-# X a data frame containing model input variables
-# y a vector containing the responses 
-# rank logical. If TRUE, the analysis is done on the ranks.
-# nboot the number of bootstrap replicates.
-# conf the confidence level of the bootstrap confidence intervals.
-
-#same as above
-in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')
-
-    x <- df_cost[,in_cols]
-    #y <- df_cost['rot_minus_seq']
-    y <- df_cost['pcent_rot_adv']
-
-    pcc_res <- pcc(x, y, rank=TRUE)
-
-    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
-    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
-    #plot(pcc_res)
-
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
-
-    to_plot <- pcc_res$PRCC
-    #rename column 1 from 'original to PRCC
-    names(to_plot)[1] <- 'PRCC'
-    to_plot$inputs <- rownames(to_plot)
-
-    axis_text_x <- element_text(angle = 45,hjust = 1, vjust = 1)
-    height <- 0.9
-
-    gg <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
-             geom_point(shape=1, colour='red') +
-             theme(axis.text.x = axis_text_x) +
-             geom_hline(yintercept = 0, linetype=3) +
-             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
-             ylim(-1,1) +
-             #ggtitle(paste(strat_name[strategy_num])) +
-             xlab(NULL)
-
-    plot(gg)
-
-  #to put back in some ref gridlines that are removed by cowplot
-  #theme_set(theme_gray())
-  #theme_set(theme_bw())
-
-  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
-  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.6), labels='AUTO')
-
-```
+Dominance of cost and dominance of selection have the greatest magnitude of influence.}\label{fig:Fig8}
+\end{figure}
 
@@
 
-```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 7.
-Model runs in which rotations perform more than 20% better than sequences are associated with high dominance of cost and low dominance of selection.
-Dominance inputs are shown on the x and y axes.
-Small grey points indicate all model runs.
-Model runs in which rotations last greater than 20% longer than sequences are shown in colour with darker blue indicating those that have the greatest difference."}
+\textbackslash{}begin\{figure\}
 
-#'cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs'
-#'cost','dom_cos','dom_sel'
+\{\centering \includegraphics{pap3figs/Fig9-1}
 
-y <- 'pcent_rot_adv'
+\}
 
-gg <- ggplot(df_cost, aes(x=dom_sel, y=dom_cos)) + 
-     geom_point(shape='.', size=1, alpha=0.2, show.legend=FALSE) +
-     theme_minimal() +
-     #geom_point(data=filter(df_cost, pcent_rot_adv >= 20 ), shape=1, size=1, col='red', alpha=0.8)
-     #shape=19 filled circle, 1=open circle
-     geom_point(data=filter(df_cost, pcent_rot_adv >= 20), aes(col=pcent_rot_adv), shape=19, size=1, alpha=0.9) +
-     viridis::scale_color_viridis(direction=-1) #discrete = TRUE)
-
-plot(gg)
-
-```
+\textbackslash{}caption\{Figure 9. Model runs in which rotations perform
+more than 20\% better than sequences are associated with high dominance
+of cost and low dominance of selection. Dominance inputs are shown on
+the x and y axes. Small grey points indicate all model runs. Model runs
+in which rotations last greater than 20\% longer than sequences are
+shown in colour with darker blue indicating those that have the greatest
+difference.\}\label{fig:Fig9} \textbackslash{}end\{figure\}
 
@@
 
-```{r FigX, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure X.
-[todo need to modify caption from MJ paper, start by saying that only in left panel during application of the insecticide] The effect of model inputs on the fitness of genotypes for a single insecticide.
-Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
-Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of resistance determines how the fitness for SR lies between that of SS and RR.
-For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of resistance=0.7 which sets the fitness of the SR closer to RR than SS.
-Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}
+\subsection{References}\label{references}
 
-library(resistance) 
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.7, dominance_cost=0.7, cost=0.3 )
+\hypertarget{refs}{}
+\hypertarget{ref-Consortium2013}{}
+1. Consortium R. Heterogeneity of selection and the evolution of
+resistance. Trends in Ecology \& Evolution. 2013;28:110--8.
+doi:\href{https://doi.org/10.1016/j.tree.2012.09.001}{10.1016/j.tree.2012.09.001}.
 
-```
+\hypertarget{ref-Sternberg2017}{}
+2. Sternberg ED, Thomas MB. Insights from agriculture for the management
+of insecticide resistance in disease vectors. Evolutionary Applications.
+2017; February:1--11.
 
-\pagebreak
+\hypertarget{ref-WHO2012}{}
+3. WHO. Global plan for insecticide resistance management in malaria
+vectors (GPIRM). Geneva.; 2012.
 
-## References
-<!-- now auto created here-->
+\hypertarget{ref-Sudo2017}{}
+4. Sudo M, Takahashi D, Andow DA, Suzuki Y, Yamanaka T. Optimal
+management strategy of insecticide resistance under various insect life
+histories: Heterogeneous timing of selection and interpatch dispersal.
+2017. doi:\href{https://doi.org/10.1111/eva.12550}{10.1111/eva.12550}.
+
+\hypertarget{ref-Levick2017}{}
+5. Levick B, South A, Hastings IM. A Two-locus Model of The Evolution of
+Insecticide Resistance to Inform and Optimise Public Health Insecticide
+Deployment Strategies. PLOS Computational Biology. 2017;13:e1005327.
+
+\hypertarget{ref-South2018}{}
+6. South A, Hastings IM. Insecticide resistance evolution with mixtures
+and sequences: A model-based explanation. Malaria Journal.
+2018;17:1--20.
+doi:\href{https://doi.org/10.1186/s12936-018-2203-y}{10.1186/s12936-018-2203-y}.
+
+\hypertarget{ref-Kliot2012}{}
+7. Kliot A, Ghanim M. Fitness costs associated with insecticide
+resistance. Pest Management Science. 2012;68:1431--7.
+
+\hypertarget{ref-Bourguet2000}{}
+8. Bourguet D, Genissel A, Raymond M, Raymond AM. Insecticide Resistance
+and Dominance Levels. Journal of Economic Entomology. 2000;93:1588--95.
+doi:\href{https://doi.org/10.1603/0022-0493-93.6.1588}{10.1603/0022-0493-93.6.1588}.
+
+\end{document}

### 2018-05-10 inst/documents/rotations_notes.Rmd
pap3

@@
 
-*9/5/18*
-Leaving in all runs now
-
-Something weird has happened to fig1 in the paper3 ?? because of a > in fig legend
-
-Also because we are leaving all runs in, the indices of the runs with large rot advantage are now wrong ! I should have recoded a unique id for runs.
-
-ARg, now all plots have stopped appearing in doc and I don't know why ?????
-It was a < symbol in the MS text, could tell because the text stopped at that point.
-
-# TODO
-
-Write out some text explaining the story & maybe the order of the plots. (keep num plots manageable).
-
-order could be
-1. example plots
-2. histograms 
-3. violin plots
-4.
-
-Not sure how best to look at effect of inputs on rot_seq, rot, neither. I could bin input ranges and then look at proportions of the 3 within them ?
-
-This would tell me how changing an input value changes the proportion of successes, failures and rot_only.
-Might be OK to display.
-Tricky to calculate, but again gets mired in the multivariate nature of the data & potential interactions.
-
-If I go back to using a continuous variable of generations until thresholds I lose out on the successful results.
-
 Maybe a more useful question, rather than when are rotations or sequence better is when are both rotations and sequences sustainable.
@@
 
+*9/5/18*
+Leaving in all runs now
+
+Something weird has happened to fig1 in the paper3 ?? because of a > in fig legend
+
+Also because we are leaving all runs in, the indices of the runs with large rot advantage are now wrong ! I should have recoded a unique id for runs.
+
+ARg, now all plots have stopped appearing in doc and I don't know why ?????
+It was a < symbol in the MS text, could tell because the text stopped at that point.
+
+*10/5/18*
+
+Testing setting width of plots for publication at 300 dpi into word doc.
+
+Solution : just use fig.height & width (out.height etc not supported for word)
+This makes the tiffs correct. Word foten resizes them. To get right in word, right click set size & position, just click reset to go back to 100%.
+
+improved histogram figures
+
+# TODO
+
 I think takehome message maybe that rotations are much better than sequences only in rare circumstances.
@@
 May also be something useful I can say about the number of insecticides too ? (Which is of interest to IVCC).
-
-Maybe reclass all 10k runs to succeed_both (2), succeed_rot (1), fail_both (0).
-Then see what are the best predictors of these ?
-Or do series of xy plots with points coloured, coloured points with all data will probably be a mess ...
 

### 2018-05-10 inst/documents/test-plot-resize.Rmd
pap3

---
title: "test plot resize"
# output:
#   pdf_document:
#   keep_tex: true
# fig_caption: true
# latex_engine: pdflatex

output: word_document
---

Testing setting width of plots for publication at 300 dpi into word doc.

Solution : just use fig.height & width (out.height etc not supported for word)
This makes the tiffs correct. Word foten resizes them. To get right in word, right click set size & position, just click reset to go back to 100%.

Was finding that plots ignore my size settings both in the tiff files and in the word doc.

```{r, eval=TRUE, include=FALSE}
#!!!! PAPER PROTOCOL
#!!!! submission MS, knit to word_document 
#!!!!  in WORD: delete figs, set double space, add line numbers 
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='')
```

```{r test-resize1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=2, fig.show='hold', fig.align='center', fig.cap="test"}

plot(1:10)
```

```{r test-resize2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=2, fig.show='hold', fig.align='center', fig.cap="test"}

plot(1:10)
```

### 2018-05-10 R/Plot showing impact of freq.R

@@ -37,6 +37,98 @@ results.v[ii, 3]=selection
 matplot(x = results.v[,1], y = results.v[,2:3], type="l",lty=1,xlab='Frequency of R allele',ylab='value')
 
 
+# andy : how can I make Ians code more useable and improve the plot
+
+
+freq_fitness_simple <- function( cos_rr = 0.05, #fitness cost of the RR genotype in absence of insecticide
+                               dom_cos = 0.95, #dominance of cost
+
+                               #may later need to couch selective advantage in terms of rr_resoration, proportion treated in intervention site  etc
+                               sel_rr = 0.1, #selective advantage of RR genotype
+                               dom_sel = 0.05, # dominance of selective advantage
+                               plot = TRUE
+) {
+  
+num_freqs <- 100 
+df_ex <- data_frame(freq=rep(NA,num_freqs),
+                    cost=NA,
+                    selection=NA,
+                    dom_cos=dom_cos,
+                    dom_sel=dom_sel) #put inputs into the out df so they can be used
+
+for (ii in 1:100 ) {
+  
+  freq <- ii/100
+  df_ex$freq[ii] <- freq
+  
+  #now need weighted averages of fitness NB we assume Hardy-Weinburg which is only an approximation under selection
+  pop_size <- 1000# multiply by pop-size to make things explicit...not necessary as cancels out
+  # first term is number of alelles (multiply pop_size by 2 for diploidY; second term is Hardy-Weinberg proportions)
+  no_in_hetero <- (pop_size*2)*(2*freq*(1-freq)) 
+  # final multiplication by 2 because there are two R alleles in the homozygote
+  no_in_homo <- (pop_size*2)*(freq*freq)*2        
+  
+  prop_hetero <- no_in_hetero/(no_in_hetero+no_in_homo)   #the proportion of R allelels in heterozygotes
+  prop_homo <- 1-prop_hetero     #the proportion of R allelels in RR homozygotes
+  
+  fit_absence <- prop_hetero*(1-dom_cos*cos_rr) + prop_homo*(1-cos_rr)  
+  
+  df_ex$cost[ii] <- 1-fit_absence
+  
+  fit_presence = prop_hetero*(1+dom_sel*sel_rr) + prop_homo*(1+sel_rr) #todo why + here & - above ?
+  
+  df_ex$selection[ii]=fit_presence-1
+  
+}
+
+#gather data to be longer
[…]

### 2018-05-10 R/freq_fitness_simple.r

@@ -1,0 +1,76 @@
+#' freq_fitness_simple to show very simply how fitness is expected to vary by frequency 
+#' 
+#' based on ians code, external to main rotations model
+#' 
+#' @param dom_sel dominance of selection
+#' @param dom_cos dominance of cost
+#' @param sel_rr selective advantage of resistance 
+#' @param cos_rr fitness cost of RR in no insecticide
+#' 
+#' @param plot whether to plot results
+#' 
+#' @examples 
+#' df1 <- freq_fitness_simple(cos_rr = 0.05, dom_cos = 0.9, sel_rr = 0.4, dom_sel = 0.1)
+#' df2 <- freq_fitness_simple(cos_rr = 0.05, dom_cos = 0.1, sel_rr = 0.4, dom_sel = 0.1)
+#' df3 <- freq_fitness_simple(cos_rr = 0.05, dom_cos = 0.1, sel_rr = 0.4, dom_sel = 0.9)
+#' df4 <- freq_fitness_simple(cos_rr = 0.05, dom_cos = 0.9, sel_rr = 0.4, dom_sel = 0.9)
+#' 
+#' @import tidyverse 
+#' @return dataframe of results
+#' @export
+
+freq_fitness_simple <- function( cos_rr = 0.05, #fitness cost of the RR genotype in absence of insecticide
+                                 dom_cos = 0.95, #dominance of cost
+                                 
+                                 #may later need to couch selective advantage in terms of rr_resoration, proportion treated in intervention site  etc
+                                 sel_rr = 0.1, #selective advantage of RR genotype
+                                 dom_sel = 0.05, # dominance of selective advantage
+                                 plot = TRUE
+) {
+  
+  num_freqs <- 100 
+  df_ex <- data_frame(freq=rep(NA,num_freqs),
+                      cost=NA,
+                      selection=NA,
+                      dom_cos=dom_cos,
+                      dom_sel=dom_sel) #put inputs into the out df so they can be used
+  
+  for (ii in 1:100 ) {
+    
+    freq <- ii/100
+    df_ex$freq[ii] <- freq
+    
+    #now need weighted averages of fitness NB we assume Hardy-Weinburg which is only an approximation under selection
+    pop_size <- 1000# multiply by pop-size to make things explicit...not necessary as cancels out
[…]

### 2018-05-11 inst/documents/paper3/pap3_rot_analyses.Rmd
pap3

@@
 #TODO improve labels for these
-inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos')
+inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos','start_freqs','rot_interval')
 # something like this example from SO https://stackoverflow.com/a/48940750/1718356

### 2018-05-11 inst/documents/paper3/paper3_rotations.Rmd
pap3

@@
 
-Within the model inputs effect fitness as shown in Figure 1, differences in fitness lead to the changes in allele frequencies over time.
+*todo add to this description of how the model works and refer to SI where we can show the algebra*    
+Within the model, the inputs described in Table 1 determine the fitness in each generation as shown in Figure 1, differences in fitness lead to the changes in allele frequencies in the following generation.
 
@@
 
+*Ian : at this stage it would be good to revisit these model details before we continue further.*    
+*It is fairly easy for me to rerun all the analyses if we change our mind on these.*
+
+#### Optional model details    
+Two optional model details were included to aid giving a more meaningful comparison of strategies.
+Firstly allele frequencies in a run were not allowed to go below the starting frequencies randomly chosen for that run.
+This aims to represent a situation where in the absence of insecticide use allele frequencies are maintained in balance between mutation and selection.
+Without this option enabled it was found that
+
+Secoundly a constraint was added to stop rapid switching between insecticides every few generations which would be impractical in the field.
+Certain input combinations create this behaviour in the absence of a constraint.
+For example when two insecticides are close to the resistance frequency threshold of 0.5 it is possible for them to switch between each other every few generations as frequencies go above the threshold when in use and decline below it when not in use.
+*todo finish this part*
+
 The model was run under four main scenarios :
@@
 
-*todo change this bit, we could actually use time-to-resistance for all strategies ?*
-*make it clearer exactly what we do*
+*todo change this bit, think more about which outputs to use*
 It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative.
@@
 
-*show violin plots here ?*
+*refer to violin plots here ?*
 
@@
 ## TODO
-*add to methods description of how fitness determined in model, move & modify fig 1*
-*try to come up with brief name for model output to make it easier to refer to it later*
-*add caveat about polygenic resistance*
-*The caption to each graph needs to have the list of parameter values used to generate them*
-*work on Ians fig Y to help explanation*
+
+1. *add to methods description of how fitness determined in model, move & modify fig 1*
+1. *add caveat about polygenic resistance*
 
@@
 All other inputs are kept constant between the rotation and the sequence.
-Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1."}
+Effectiveness=0.7, exposure=0.7, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1."}
 
@@
 linputs$cost <-     0 
-linputs$max_gen <-  200
-linputs$expo_hi <-  0.6
-linputs$eff <-      0.6
+linputs$max_gen <-  150
+linputs$expo_hi <-  0.7
+linputs$eff <-      0.7
 linputs$dom_sel <-  1
@@
 
-ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
-ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='A. Rotation')
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='B. Sequence')
 library(gridExtra)
@@
 All other inputs are kept constant between the rotation and the sequence.
-Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0.05, dominance of cost=1."}
-
-# revisit how best to do model runs from within the paper
+Effectiveness=0.7, exposure=0.7, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0.05, dominance of cost=1."}
 
@@
 linputs$cost <-     0.05 
-linputs$max_gen <-  200
-linputs$expo_hi <-  0.6
-linputs$eff <-      0.6
+linputs$max_gen <-  150
+linputs$expo_hi <-  0.7
+linputs$eff <-      0.7
 linputs$dom_sel <-  1
@@
 All other inputs are kept constant between the rotation and the sequence.
-Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0, coverage=0.5, dispersal=0.1."}
+Effectiveness=0.7, exposure=0.7, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0, coverage=0.5, dispersal=0.1."}
 
@@
 
-ggrot <- rot_plot_resistance(dfrot, plot_refuge=TRUE, logy=TRUE, plot=FALSE)
-ggseq <- rot_plot_resistance(dfseq, plot_refuge=TRUE, logy=TRUE, plot=FALSE)
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='A. Rotation')
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='B. Sequence')
 library(gridExtra)
@@
 Success of model runs under different scenarios and strategies.
-Model runs classed as a success if there are still insecticides for which resistance is below the threshold of 0.5 after 500 generations (circa 50 years for Anopheles)."}
-
-#remove seq maybe have a check that are very few
-# count(df_cost, success)
-# 1         neither  5948
-# 2             rot   209
-# 3         rot_seq  3841
-# 4             seq     2
+Model runs classed as a success if there are still insecticides for which resistance is below the threshold of 0.5 after 500 generations (circa 50 years for Anopheles).
+With no costs or dispersal both strategies succeeded in around 5% of runs and neither did in 95%.
+When costs alone were included the proportion of runs in which both, neither and rotation succeeded was 38%, 60% and 2%.
+When dispersal alone was included both strategies succeeded in 20% or runs and neither did in 80%, in only 0.1% of runs did rotation alone succeed."}
 
@@
                                                       ))))
-#count(df_all, success)
-# 1          both 12413
-# 2       neither 18196
-# 3 rotation only   382
-# 4 sequence only     9
+#count(df_all, scenario, success)
+# 1 no cost or dispersal          both    47
+#  2 no cost or dispersal       neither   952
+#  3 no cost or dispersal sequence only     1
+#  4                costs          both  3841
+#  5                costs       neither  5948
+#  6                costs rotation only   209
+#  7                costs sequence only     2
+#  8            dispersal          both  1938
+#  9            dispersal       neither  8046
+# 10            dispersal rotation only    12
+# 11            dispersal sequence only     4
+# 12  costs and dispersal          both  6587
+# 13  costs and dispersal       neither  3250
+# 14  costs and dispersal rotation only   161
+# 15  costs and dispersal sequence only     2
 
@@
                   vjust = 0.9 )
+
+```
+
+\pagebreak
+
+```{r Fig5xtra, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.cap="Figure 5xtra.
+How model runs classed as success under the costs only scenario for the different strategies differ in the values of inputs.
+Violin plots showing how the density of input values changes according to the classification of success.
+Runs in which both strategies were successful were more likely to have higher costs and lower effectiveness, exposure and resistance restoration.
+Runs in which neither strategy was successful were the reverse.
+Runs for which rotations only were successful tended to have higher dominance of cost, lower dominance of resistance and lower rotation intervals.
+Remember that as shown in the previous figure the proportion of runs for which rotation alone succeeded was lower than for the other options."}
+
+# I could maybe make the data longer and then facet by input
+# think I would filter out, n_insecticides, start_freqs, male_expo_prop & rot_interval that don't show much rship
+#TODO improve labels for these
+inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos','start_freqs','rot_interval','male_expo_prop')
+# something like this example from SO https://stackoverflow.com/a/48940750/1718356
+#data_long <- tidyr::gather(data, key = type_col, value = categories, -col4)
+#ggplot(data_long, aes(x = categories, fill = col4)) +
+#  geom_bar() + 
+#  facet_wrap(~ type_col, scales = "free_x")
+
+#filter cost from all because the success calc was done on that
+df_cost_long <- filter(df_all,scenario=='costs')
+#first just select the columns I want
+df_cost_long <- select(df_cost_long,c(inputs,'success'))
+#gather data to be longer
+df_cost_long <- tidyr::gather(df_cost_long, key=input_name, value=input_value, -success)
+
+#there were just an odd couple of results where seq better, remove them here or earlier
+df_cost_long <- filter(df_cost_long, success != 'sequence only')
+
+ggplot(df_cost_long, aes_string(x='success',y='input_value', color='success')) + 
+        #ylim(0,450) +
+        #coord_cartesian( ylim=c(0, 350)) +
+        geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = 1 ) +
+        theme_minimal() +
+        #this is an example of how you might facet
+        facet_wrap(~input_name, scales = "free_y") +  
+        theme(panel.grid.minor.y=element_blank()) #,
+        #           axis.text.x=element_blank())
+        #geom_jitter(shape='.',height=0,width=0.2) )
+
+```
+
+\pagebreak
+
+```{r Fig5xtra_cost_dispersal, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.cap="Figure 5xtra cost and dispersal.
+Probably don't show in the paper.
+Similar to previous one.
+Just adds coverage and migration."}
+
+#TODO improve labels for these
+
+#coverage and migration added for costs and dispersal
+inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos','start_freqs','rot_interval','male_expo_prop','coverage','migration')
+
+#filter cost from all because the success calc was done on that
+df_cost_long <- filter(df_all,scenario=='costs and dispersal')
+#first just select the columns I want
+df_cost_long <- select(df_cost_long,c(inputs,'success'))
+#gather data to be longer
+df_cost_long <- tidyr::gather(df_cost_long, key=input_name, value=input_value, -success)
+
+#there were just an odd couple of results where seq better, remove them here or earlier
+df_cost_long <- filter(df_cost_long, success != 'sequence only')
+
+ggplot(df_cost_long, aes_string(x='success',y='input_value', color='success')) + 
+        #ylim(0,450) +
+        #coord_cartesian( ylim=c(0, 350)) +
+        geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = 1 ) +
+        theme_minimal() +
+        #this is an example of how you might facet
+        facet_wrap(~input_name, scales = "free_y") +  
+        theme(panel.grid.minor.y=element_blank()) #,
+        #           axis.text.x=element_blank())
+        #geom_jitter(shape='.',height=0,width=0.2) )
 
@@
   xlab('allele frequency') +
+  scale_x_continuous(breaks=seq(0,1,0.2)) +
   facet_grid(dom_cos~dom_sel, labeller = "label_both", as.table = FALSE) #+

### 2018-05-11 inst/documents/rotations_notes.Rmd
pap3

@@
 *10/5/18*
-
 Testing setting width of plots for publication at 300 dpi into word doc.
@@
 
+got Ians fig Y working in freq_fitness_simple() & added facetted grid plot to the paper.
+
+*11/5/18*
+- improved violin plots added to pap3
+- improving legends
+
+Finally grasping the nettle of insecticide switching.
+
+I want to stop switching to an insecticide that was used < X generations ago.
+
+To do that store an array which has a counter for each insecticide which records how long since it was last used. Then just add to that counter in each generation for each insecticide not in use.
+Don't allow an insecticide to be used that is below a user-defined level.
+
+in insecticide_check()
+BE CAREFUL!
+
+df_ins <- data.frame(last_used=rep(NA, n_insecticides))
+
 # TODO
+
+Before sending to Ian.
+
+~ add to methods about min rwr interval
+~ add about keeping resistance allele freq at starting level
+~ add to fig Y legend Ians caveats
 

### 2018-05-11 R/Plot showing impact of freq.R

@@ -1,142 +1,0 @@
-results.v<-matrix(1:300, ncol = 3)
-
-
-c<-0.05; #fitness cost of the RR genotype in absence of insecticide
-h_c<-0.95; #dominance of cost
-
-#may later need to couch selective advantage in terms of rr_resoration, proportion treated in intervention site  etc
-s<-0.1; #selective advantage of RR genotype
-h_s<-0.05; # dominance of selective advantage
-
-#s<-0.05; #selective advantage of RR genotype
-#h_s<-0.95; # dominance of selective advantage
-
-for (ii in 1:100 ) {
-
-freq=ii/100;
-
-#now need weighted averages of fitness NB we assume Hardy-Weinburg which is only an approximation under selection
-pop_size=1000# multiply by pop-size to make things explicit...not necessary as cancels out
-no_in_hetero= (pop_size*2)*(2*freq*(1-freq))  #first term is number of alelles (multiply pop_size by 2 for diploidY; second term is Hardy-Weinberg proportions)
-no_in_homo= (pop_size*2)*(freq*freq)*2        # final multiplication by 2 because there are two R alleles in the homozygote
-
-prop_hetero=no_in_hetero/(no_in_hetero+no_in_homo)   #the proportion of R allelels in heterozygotes
-prop_homo=1-prop_hetero     #the proportion of R allelels in RR homozygotes
-
-fit_absence=prop_hetero*(1-h_c*c) + prop_homo*(1-c)  
-cost=1-fit_absence;
-fitness = prop_hetero*(1+h_s*s) + prop_homo*(1+s)
-selection=fitness-1
-
-results.v[ii, 1]=freq
-results.v[ii, 2]=cost
-results.v[ii, 3]=selection
-
-}
-
-matplot(x = results.v[,1], y = results.v[,2:3], type="l",lty=1,xlab='Frequency of R allele',ylab='value')
-
-
-# andy : how can I make Ians code more useable and improve the plot
-
-
-freq_fitness_simple <- function( cos_rr = 0.05, #fitness cost of the RR genotype in absence of insecticide
-                               dom_cos = 0.95, #dominance of cost
-
-                               #may later need to couch selective advantage in terms of rr_resoration, proportion treated in intervention site  etc
-                               sel_rr = 0.1, #selective advantage of RR genotype
-                               dom_sel = 0.05, # dominance of selective advantage
-                               plot = TRUE
-) {
-  
-num_freqs <- 100 
[…]

### 2018-05-11 R/rot_plots.r

@@ -7,6 +7,7 @@
 #' @param add_gens_under50 whether to add a label of num generations under 50percent resistance
 #' @param df_resanother exploratory option to plot results of another scenario on the same graph
 #' @param lwd line thickness for resistance curves
+#' @param title optional title for plot, NULL for no title
 #' @param plot whether to plot results
 #'
 # check said that namespace dependencies not required
@@ -28,6 +29,7 @@ rot_plot_resistance <- function(df_res2,
                                 add_gens_under50 = TRUE,
                                 df_resanother = NULL,
                                 lwd = 1.5,
+                                title = NULL,
                                 plot = TRUE) {
   
   # column names of input dataframe
@@ -143,6 +145,8 @@ rot_plot_resistance <- function(df_res2,
     gg <- gg + geom_text(aes(x=Inf, y=y, label=gens_dep_under50), colour='black', show.legend=FALSE, hjust=1, vjust=0)
     
   }
+  
+  if (!is.null(title)) gg <- gg + ggtitle(title)
     
   if (plot) plot(gg)
   

### 2018-05-11 inst/documents/rotations_notes.Rmd
pap3

@@
 
-in insecticide_check()
+in insecticide_switch()
 BE CAREFUL!
 
-df_ins <- data.frame(last_used=rep(NA, n_insecticides))
+min_gens_switch_back minimum num gens before can switch back to an insecticide
+
+I set the default in run_rot to 10 generations.
+(I can leave min_rwr_interval in for now, when it is set to 1 it has no effect).
+
+Note that min_rwr_interval was assessed in insecticide_check() i.e. it stopped an insecticide from being stopped within a certain interval
+
+Which actually maybe wasn't so bad ?? Maybe I could stick with it ??
+i.e. that simulated not being able to withdraw an insecticide which has already been applied (which sounds pretty good now I think about it ...)
+
+Either of these could be OK. I had it set to 5 generations before which was half a year.
 
 # TODO
+
+Decide whether to use min_rwr_interval OR min_gens_switch_back
 

### 2018-05-18 inst/documents/paper3/pap3_rot_analyses.Rmd
paper3

@@
   df_cost_mig <- df_res_all
+  load("df_res_all_rot_nocost_nomig10000_mgs5.rda")
+  df_nocost_nomig_mgs5 <- df_res_all 
+  load("df_res_all_rot_cost10000_mgs5.rda")
+  df_cost_mgs5 <- df_res_all
 
@@
 ```
+
+\pagebreak
+Showing that using min_gens_switch_back=5 instead of min_rwr_interval=5 seems to make no difference in main results.
+
+```{r min_gens_switch_back on success, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}
+
+# if I put all scenarios in one df with a column for scenario I can do one ggplot and facet 
+# would work for the previous plot too and would simplify adding the new columns
+# first add the scenario column
+df_nocost_nomig$scenario <- 'no cost or dispersal'
+df_cost$scenario <- 'costs'
+df_mig$scenario <- 'dispersal'
+df_cost_mig$scenario <- 'costs and dispersal'
+
+df_nocost_nomig_mgs5$scenario <- 'no cost or dispersal mgs5'
+df_cost_mgs5$scenario <- 'costs mgs5'
+
+df_all <- dplyr::bind_rows(df_nocost_nomig,
+                           df_nocost_nomig_mgs5,
+                           df_cost,
+                           df_mig,
+                           df_cost_mig,
+                           df_cost_mgs5)
+#order for ggplot
+df_all$scenario <- factor(df_all$scenario, levels=c('no cost or dispersal','no cost or dispersal mgs5','costs','costs mgs5','dispersal','costs and dispersal'))
+
+# add a column for classification of whether, both, rotation, or neither succeeded
+df_all <- mutate(df_all, success = ifelse(gens_seq==max_gen & gens_rot==max_gen,'both',
+                                               ifelse(gens_rot==max_gen,'rotation only',
+                                               ifelse(gens_seq==max_gen,'sequence only', 'neither'  
+                                                      ))))
+#count(df_all, scenario, success)
+
+#hack to add subplot labels to facets
+abcd <- data.frame(label = c("A","B","C","D","E","F"),
+                   scenario = levels(df_all$scenario))
+
+ggplot( df_all, aes(x=success)) + 
+        #geom_bar() +
+        stat_count(mapping = aes(y=..prop.., group=1)) + #puts proprtions in axes
+        scale_y_continuous(breaks=seq(0.2,0.8,0.2)) +
+        theme_minimal() +
+        ylab('proportion of model runs') +
+        facet_wrap(~scenario, ncol=1) + #, scales='free_y') + #can remove free_y later if I do 10k nocost runs    
+        geom_text(data=abcd,
+                  aes(x = -Inf, y = +Inf, label=label),
+                  hjust = 0,
+                  vjust = 0.9 )
+
+```
+
+\pagebreak
+Using min_gens_switch_back=5 gives greater benefits to sequences when there are no costs or dispersal. I suspect this is an artefact. Try looking at some of the runs.
+```{r min_gens_switch_back on rotation benefit, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}
+
+  #try quicker way with df_all and facet
+  df_all <- mutate(df_all, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
+  custom_bins <- c(-20,-10,-1,1,10,20,100,450)
+  df_all$rot_adv_bins <- cut(df_all$pcent_rot_adv, custom_bins)  
+  bin_labs <- levels(df_all$rot_adv_bins)
+
+  #hack to add subplot labels to facets
+  abcd <- data.frame(label = c("A","B","C","D","E","F"),
+                   scenario = levels(df_all$scenario))
+
+  ggplot( df_all, aes(x=rot_adv_bins)) +
+
+             #geom_bar() +
+             stat_count(mapping = aes(y=..prop.., group=1)) + #puts proprtions in axes
+             scale_y_continuous(breaks=seq(0,0.8,0.2)) +
+             xlab('Percent benefit of rotation over sequence') +
+             ylab('proportion of model runs') +
+             #ggtitle('') +
+             theme_minimal() +
+             facet_wrap(~scenario, ncol=1) +
+             scale_x_discrete( limits=bin_labs,
+                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+     
+             theme(panel.grid.minor.y=element_blank(),
+                   axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1)) +
+             geom_text(data=abcd,
+                  aes(x = -Inf, y = +Inf, label=label),
+                  hjust = 0,
+                  vjust = 0.9 )
+
+```
+\pagebreak
+Using min_gens_switch_back=5 example of when sequence does better.
+```{r min_gens_switch_back sequence better example, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}
+
+df_seqbetter <- filter( df_all, pcent_rot_adv < -5 &
+                                scenario == 'no cost or dispersal mgs5' )
+
+#ooo scenario 5 has rot_adv of -97%
+#but remember it is not scenario5 in the original dataset !
+#Aha! all these big sequence advantages (well those >50%) 
+#were due to artfefact when rotation interval=5 and num insecticides=2 rotation couldn't return to first insecticide because it was only 5 generations ago !!
+
+df_nocost_nomig_mgs5 <- filter( df_all, scenario == 'no cost or dispersal mgs5' )
+
+#which(df_nocost_nomig_mgs5$pcent_rot_adv < -90)
+#217 1933 3435 4289 5649 7100 7925 7990
+
+scenarios_to_plot <- c(217)
+
+for(scenario in scenarios_to_plot)
+{
+  #cat(scenario,"\n")
+
+  linputs <- as.list( df_nocost_nomig_mgs5[scenario, 1:15] )
+
+  # run one scenario
+  dfres <- do.call(run_rot, linputs)
+
+  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+
+  #now run as sequence
+  linputs$rot_interval <- 0
+  dfres <- do.call(run_rot, linputs)
+
+  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+
+  grid.arrange(ggrot, ggseq, ncol=1)
+
+}
+
+```

### 2018-05-18 inst/documents/paper3/paper3_rotations.Rmd
paper3

@@
 *todo add to this description of how the model works and refer to SI where we can show the algebra*
+
 Within the model, the inputs described in Table 1 determine the fitness in each generation as shown in Figure 1, differences in fitness lead to the changes in allele frequencies in the following generation.
@@
 
-*Ian : at this stage it would be good to revisit these model details before we continue further.*    
+*Ian : at this stage it would be good to revisit the following model details.*    
 *It is fairly easy for me to rerun all the analyses if we change our mind on these.*
@@
 This aims to represent a situation where in the absence of insecticide use allele frequencies are maintained in balance between mutation and selection.
-Without this option enabled it was found that
+Without this option enabled, in runs including costs of resistance, resistance allele frequencies can decline below starting levels when an insecticide is not in use and this makes it more difficult to compare alternate strategies.
+*Ian, can you help out with a few sentences here ?*
 
@@
 For example when two insecticides are close to the resistance frequency threshold of 0.5 it is possible for them to switch between each other every few generations as frequencies go above the threshold when in use and decline below it when not in use.
-*todo finish this part*
+Two different ways of stopping this rapid switching behaviour were trialled.
+In one, a minimum-insecticide-use constraint stopped switching to another insecticide within a set number of generations (5 in this case) of starting using another.
+This could represent a field situation where once walls have been sprayed or nets distributed they cannot be unsprayed or recalled.
+In a second, a minimum-repeat-interval constraint stopped returning to the use of an insecticide within set number of generations (again we used 5 generations) since it was last used.
+*I'm looking into the implications of these different constraints now.*
 
@@
 
-*todo change this bit, think more about which outputs to use*
 It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative.
@@
 This allowed comparison between strategies using a variable number of insecticides.
-Previously time-to-resistance, namely the number of generations it takes to reach a defined resistance threshold (usually a resistance allele frequency of 0.5), was used.
-Using time-to-resistance worked when considering just two insecticides, but when the number of insecticides is a variable it is not sufficient.
+
+\pagebreak
 
@@
 
-*todo continue describing the example runs first, before talking about overall results*
+When resistance costs are included (Fig 3), resistance allele frequencies decline in periods when the corresponding insecticide is not in use.
+Thus for sequences the decline can allow insectides to be returned to use later on because resistance has declined below threshold levels in contrast with the situation in Figure 2.
 
-### Results2 : comparing rotations and sequences across model runs.
+Adding representation of an untreated refugia and dispersal between it and the treated area also allows resistance allele frequencies to decline when an insecticide is not in use (Fig 4).
+Resistance frequencies in the treated areas (red lines) increase when an insecticide is in use.
+Resistance frequencies in the untreated areas (blue lines) increase after those in the treated areas as a result of dispersal from the treated areas.
+When an insecticide stops being used the resistance frequency in the treated area declines due to dispersal from the untreated area.
+For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and consistent with the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again.
+For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines while not.
+
+#### Results2 : comparing rotations and sequences across model runs.
 
@@
 Thus to aid the comparison of rotations and sequences model runs can be broadly classified into four groups :
+
 1. both strategies succeeded : for both rotations and sequences resistance thresholds for all insecticides not reached within 500 generations.
@@
 When costs of reistance are added (Fig 5B) more runs succeed for both strategies, up to about half of those that succeed for neither, and a small proportion (< 4%) succeed for rotations alone.
-Adding both costs and dispersal results in more runs succeeding for both than neither (Fig 5C).
+Adding both costs and dispersal results in more runs succeeding for both than neither (Fig 5C) and the proportion of runs where rotations only succeed is similar to the scenario with costs alone.
+
+The runs in each of these three main classes (success in both, neither or rotations only) differ in their distribution of input values (Fig 5xtra).
+Runs in which both strategies were successful were more likely to have higher costs and lower effectiveness, exposure and resistance restoration.
+Runs in which neither strategy was successful were the reverse.
+Runs for which rotations only were successful tended to have higher dominance of cost, lower dominance of resistance and lower rotation intervals.
 
@@
 
-*refer to violin plots here ?*
+As shown in Fig 6B some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 7.
 
-As shown in Fig 4.2 some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 5.
+A Partial Rank Correlation analysis identified that of the inputs, dominance of cost and dominance of selection had the greatest influence on the difference between rotations and sequences when resistance costs were included (Fig 8).
+A combination of high values of dominance of cost and low values for dominance of selection are required to give the greatest advantages to rotations, but do not guarantee this advantage (Fig 9).
 
-A Partial Rank Correlation analysis identified that of the inputs, dominance of cost and dominance of selection had the greatest influence on the difference between rotations and sequences when resistance costs were included (Fig 6).
-A combination of high values of dominance of cost and low values for dominance of selection are required to give the greatest advantages to rotations, but do not guarantee this advantage (Fig 7).
+\pagebreak
 
@@
 This is illustrated on Figure Y.
-Relatively rapid rotations keep the allele frequencies at the LH side of Figure Y where costs during the periods when the insecticide is not deployed greatly outweigh the selection for IR during the periods when the insecticide is deployed.
-We believe the explanation is therefore that rotations keep allele frequencies in the parameter space to the left of Figure Y and hence acts to slow the evolution of IR compared to sequences which allow frequencies to rise to relatively high levels before the insecticide is replaced.
-(does this sound reasonable??).
-Note also that starting frequency, as expected, had an impact (Figure 7), lower starting frequencies slightly favouring rotations by allowing more time in the “low frequency” parameter space.
+Relatively rapid rotations keep the allele frequencies at the LH side of Figure 10 where costs during the periods when the insecticide is not deployed greatly outweigh the selection for IR during the periods when the insecticide is deployed.
+We believe the explanation is therefore that rotations keep allele frequencies in the parameter space to the left of Fig 10 and hence acts to slow the evolution of IR compared to sequences which allow frequencies to rise to relatively high levels before the insecticide is replaced.
+Note also that starting frequency, as expected, had an impact (Fig 8), lower starting frequencies slightly favouring rotations by allowing more time in the “low frequency” parameter space.
+[*Andy not sure that the effect of starting frequencies in the PRCC is notable, seems very close to the line of not having an effect either way*]
 
@@
 1. *add caveat about polygenic resistance*
+1. *talk about cross resistance*
 
@@
 
-# revisit how best to do model runs from within the paper
-
 linputs <- list()
@@
 
-ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='A. Rotation')
-ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='B. Sequence')
+ggrot <- rot_plot_resistance(dfrot, plot_refuge=TRUE, logy=TRUE, plot=FALSE, title='A. Rotation')
+ggseq <- rot_plot_resistance(dfseq, plot_refuge=TRUE, logy=TRUE, plot=FALSE, title='B. Sequence')
 library(gridExtra)
@@
 Effect of cost and dispersal on the difference between rotations and sequences from model runs.
-The y axis shows the number of model runs for the different caegories of difference between rotations and sequences.
+The y axis shows the number of model runs for the different categories of difference between rotations and sequences.
 A) when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results.
@@
 
-#now that non thershold runs are not removed :
+#now that non threshold runs are not removed :
 #which(df_cost$pcent_rot_adv > 200)
@@
              geom_hline(yintercept = 0, linetype=3) +
+             theme_minimal() +
              #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
@@
 In these latter circumstances if rotations can keep frequencies low they can reduce the long term development of resistance.
-Plot calculation assumes Hardy-Weinberg equilibrium which will not be the case under selection but the general pattern will be similar.
+Note that with higher costs or lower benefits of resistance there is the potential for the cost and benefit lines to cross and for there to be a region at low frequencies where resistance is not selected for.
+Plot calculation assumes Hardy-Weinberg equilibrium proportion of genotypes.
+The equilibriun will not be followed in this case because the resistance locus is under selection but it serves as an approximation to illustrate the general effect.
 Fitness cost set to 0.05 and selective advantage of resistance to 0.4."}

### 2018-05-18 inst/documents/rotations_notes.Rmd
paper3

@@
 
+done ~ add to methods about min rwr interval
+done ~ add about keeping resistance allele freq at starting level
+done ~ add to fig Y legend Ians caveats
+
+*14/5/2018*
+~ Paper3, decsribing results and adding to methods clarifiying model constraints.
+
+Before sending to Ian.
+~ resize no square figs, make sure one per page
+
+looking at how I made smaller reference bib file for uea work.
+
+Seems I can export a single non-synced bibtex file
+
+https://www.imperial.ac.uk/media/imperial-college/administration-and-support-services/library/public/LaTeX-and-Mendeley-sep-2017.pdf
+1. Exporting a BibTeX (.bib) file from your Mendeley library
+• Select the records for export
+• From top menus, go to File>Export…
+• In Popup box, choose a file name (no spaces); ensure file type is set as BibTeX
+• Save the .bib file in the same folder as the LaTeX document
+
+df_res_all_rot_nocost_nomig_10000_mgs5.rda
+
+mgs5 to indicate min_gens_switch_back=5
+
+Aha! all these big sequence advantages (well those >50%) 
+were due to artfefact when rotation interval=5 and num insecticides=2 rotation couldn't return to first insecticide because it was only 5 generations ago !!
+
+*17/5/2018*
+
+New papers to cite in rotations paper :
+
+Hackett (2016) 
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5026168/
+interesting model of resistance to transgenic crops with similar results on the importance of dominance of resistance and cost :
+"the most significant constraint upon model behaviour is the dominance of the resistant phenotype, with greater dominance decreasing Bt usage.
+For a given level of dominance of resistance, the type and dominance of any fitness costs may relax these constraints, allowing for larger areas of Bt crop [smaller refuges to control resistance evolution]"
+
+The R code is here and looks quite accessible :
+https://datadryad.org//resource/doi:10.5061/dryad.4s405
+
+Allen(2017) Reversing resistance: different routes and common themes across pathogens
+http://rspb.royalsocietypublishing.org/content/284/1863/20171619
+
+Mostly applied to microbes & fungi, some good general points, probably not useful to us.
+Does talk about different types of reversion and the importance of cost.
+
+My sentence from an example about negative cross resistance :
+Negative cross resistance meant that two fungicides were used successfully on Botrytis cinerea in European vineyards for 7 years ending in 1994 when a new mutation appeared that gave resistance to both.
+[87]
+
+Russell PE. 1995 Fungicide resistance—occurrence and management. J. Agric. Sci. 124, 317–323. (doi:10.1017/S0021859600073275)
+
+So negative cross resistance can work for a while but is susceptible to new mutations giving resistance to both ingredients.
+
+Alphey & Bonsal 2017 good accessible description of high dose-refuge strategy :
+Genetics‐based methods for agricultural insect pest management
+(but the paper itself is about sterile insect technique)
+
+For example, modelling the effects of larval competition and exploring late‐acting lethal phenotypes in mosquitoes predicted that this could be substantially more effective at population control than an early‐acting (e.g.
+embryonic) lethality or radiation‐induced sterility (Atkinson et al., 2007; Phuc et al., 2007; Alphey & Bonsall, 2014a).
+Indeed, if density‐dependent juvenile competition were over‐compensatory, genetic lethality that occurred at an earlier stage, thereby freeing survivors from regulation by intense competition, could push adult insect numbers higher than in the natural uncontrolled population.
+
+In terms of the genetics, consider a resistant allele r, which is initially rare.
+The dominant allele S is susceptible to Bt.
+If the high dose assumption is achieved, only some rr individuals can survive a full life cycle on transgenic plants and emerge as adults to mate.
+In the refuge, most emerging adults will be susceptible SS, especially if the r allele has fitness costs in the absence of the Bt toxin.
+If the refuge is located so that the two subpopulations are well‐mixed, most resistant rr survivors will mate with susceptible SS insects from the refuge.
+Their resulting Sr progeny cannot survive on the Bt crops and so will not pass on the resistant allele to future generations.
+Unfortunately, a few insect species, such as the economically important pests Helicoverpa zea (Boddie) and Helicoverpa armigera (Hübner) (both species, confusingly, known as both bollworm and corn earworm), have been identified in the past as ‘moderate dose pests’, where the toxins were unable to kill all heterozygotes (EPA, 1998; Tabashnik et al., 2008).
+Even where its main assumptions appear to hold, the high‐dose/refuge strategy is predicted only to delay resistance and, after two decades of commercially grown Bt cotton and Bt corn (maize), some field‐evolved resistance has now been observed and reported in a variety of insect species.
+In some cases, this has already led to reduced efficacy of crops, or even crop failures.
+A comprehensive review is provided by Tabashnik et al.
+(2013).
+
+In economic terms, the high‐dose/refuge strategy is an inter‐temporal trade‐off, sacrificing current value to retain value generated in future (Frisvold & Reeves, 2008).
+Refuge plants will be damaged, which reduces their yield if they are crops, or reduces the area allotted to crop production if alternative host plants are used.
+This damage or lost production is tolerated, in return for prolonging the efficacy of the protection afforded by the Bt crops.
+In principle, a conventional crop refuge could be sprayed with another pesticide (one with no cross‐resistance between its active ingredient and Bt), although doing so reduces its effectiveness as a refuge.
+
+Aha and a justification for using resistance threshold of 0.5 !
+Generally, a frequency of 0.5 is an indicator of a serious problem because resistance frequency increases rapidly once r becomes the more common allele (Carrière & Tabashnik, 2001).
+
+Carrière & Tabashnik, 2001. Modelling refuges in Bt corn
+Good paper to cite in rotations paper.
+They found some similar things to us, e.g. resistance not increasing when dominance of resistance low and dominance of cost high.
+
+*Carriere 2018* good one to include, recent data on fitness costs showing that there are highish fitness costs but that they are almost completely recessive.
+Genotype-specific fitness cost of resistance to Bt
+
+*18/5/2018*
+
+*Glunt 2018* Impact of temperature on insecticide toxicity.
+Relevant to window of selection. 
+In arabiensis & funestus with deltamethrin, generally mortalities higher when temperatures either higher or lower than standard.
+The difference between resistant and susceptible mortality is generally less, suggesting that the window of selction may be lower outside of standard lab conditions.
+Bendiocarb mortality increased with temperature and unsurprisingly was similar between strains susceptible and resistant to pyrethroids.
+
 # TODO
 
-Decide whether to use min_rwr_interval OR min_gens_switch_back
-
-Before sending to Ian.
-
-~ add to methods about min rwr interval
-~ add about keeping resistance allele freq at starting level
-~ add to fig Y legend Ians caveats
+*today*
+done ~ reread Ians text about dominance
+~ move rotations repo back to my github
+~ update R
+done ~ tweet to malaria journal
+~ think more about rotations datavis
+done ~ read the alphey oxitech modelling papers
+
+To talk with Ian about next week :
+~ min_rwr_interval vs min_gens_switch_back
+
+Look at Hackett R code. Could it be an even simpler model that we could use ?
+The R code is here and looks quite accessible :
+https://datadryad.org//resource/doi:10.5061/dryad.4s405
+
+Looking at moving rotations repo back to me. Seems transfer may be best way :
+https://help.github.com/articles/about-repository-transfers/
+Transfer is at the bottom of settings, options : from Ians account
+
+Look into whether to use min_rwr_interval OR min_gens_switch_back 
+I'm thinking that I'll stick with min_rwr_interval
+
+Revisit Ians paragraph about the rare conditions that give large benefit to rotations.
+
+"Note however that this combination of recessive resistance and high dominance costs are exactly the features most desirable in any insecticide irrespective of how it is deployed.
+...
+it is very difficult to maintain insecticide concentration at the level required to make resistance recessive.
+Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant.""
+me is there anything similar on the dominance of costs ?
 
@@
 May also be something useful I can say about the number of insecticides too ? (Which is of interest to IVCC).
+
+In rotations and WoS MS we should talk about high dose/refuge strategy for transgenic crops (specifically Bt corn).
+The high dose part is that the dose is sufficiently high to kill all heterozygotes i.e.
+making resistance recessive.
+But if this was used across a large area the evolution of resistance would be very rapid - because only homozygous resistant individuals would be left.
+The refuge part is to provide a source of susceptible immigrants which can generate heterozygous offspring that are vulnerable to the dose.
 
@@
 
-~ can I add figures from specific sensi_an runs ? or safer to run the model for set runs from the paper rmd.
-
-~ how to sort library.bib ? Have a much reduced copy in rotations folder now.
-
 think about killer plot to display difference between rot & seq generations
 
-be careful that min_rwr_interval doesn't come back to bite me. Ian wanted it different, i.e. not going back to an insecticide that has been used recently ...
-
-~ sensi_an_rotations1.Rmd
-
 ~ *decide on min_rwr_interval for sensi analysis
-~ decide on max_gen, what was it for paper1 ?
 ~ one generation per month, 12 generations per year, 500 gens is ~40 years
@@
 ~ maybe come back to i_in_use ribbons not meeting issue
-
 ~ rename rot_criterion to rot_threshold
-
 ~ add in some unit-tests ?
-
 ~ check how gens_under_thresh are counted in sensi_an_rotations1
@@
 
-Download papers on modelling oxitech
-Insecticide resistance
-https://doi.org/10.1093/jee/100.5.1642
-https://doi.org/10.1603/029.102.0233
-
 These are not about insecticide resistance (they are about resistance to the gene itself but may have useful pointers)
@@
 
-Type of fitness cost influences the rate of evolution of resistance to transgenic Bt crops
-Looks like could be very interesting :
-DOI: 10.1111/1365-2664.12680
-
 Combining the high-dose/refuge strategy and self-limiting transgenic
@@
 
-~ check again on current small differences between rotations & sequences
-
-~ maybe I can write a function to plot 2 scenarios on the same graph (in the same way that curtis fig2 plots mix & seq on same graph) ?
-~ I could have the insecticide use still along the top but with diff colours showing the two alternatives
-
-~ get sensi_an_rotations1.Rmd working for Ians Q1
-
 ~ can we get cost to decrease over time at high frequency as suggested in GPIRM ?
@@
 BEWARE active in active_or_refuge is potentially confusing, does not mean that insecticide is active/deployed, just means its is the active area (a different insecticide may be deployed)
-
-*Ian said first question is to assess whether there is any benefit to rotations when there are no costs and no emigration*
 
@@
 
-~ create a new version of run_rot2() allowing me to refactor the code and check that it still generates same results as run_rot()
-
 ~ maybe have a set_raf() function too
@@
 # useful
-~ one generation per month, 12 generations per year, 500 gens is ~40 years
+~ one generation per month, 12 (or 10) generations per year, 500 gens is ~40-50 years
 

### 2018-05-18 inst/documents/sensi_an_rotations1.Rmd
paper3 writing
paper3

@@
 #
-# 
-# Testing sensitvity analysis to answer question 1.
-Are sequences and rotations equivalent when coverage 100% and no fitness costs.
-On each page should see a run with 1 set of inputs first for sequence then for rotation.
 
@@
 plot_scenarios  <- FALSE
-n_scenarios     <- 10000 # try 1000 runs with no cost
+n_scenarios     <- 10000 
 already_ran     <- FALSE
@@
     ## constant inputs ##
-    #linputs$coverage <- 1
-    #linputs$migration <- 0
-    #linputs$cost <- 0 #variable cost below
+    linputs$coverage <- 1
+    linputs$migration <- 0
+    linputs$cost <- 0 #variable cost below
     linputs$max_gen <- max_gen
@@
     #beware min_rwr_interval, it's to stop repeated small rotations
-    linputs$min_rwr_interval = 5
+    linputs$min_rwr_interval = 1
+    linputs$min_gens_switch_back <- 5
     #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
@@
     ## variable inputs ##
-    linputs$cost <-           runif(1, min=0,   max=0.1) 
+    #linputs$cost <-           runif(1, min=0,   max=0.1) 
     #linputs$cost <-           0 
@@
     # dispersal (initially with cost set to 0)
-    linputs$coverage <-       runif(1, min=0.1, max=0.9)
-    linputs$migration <-      runif(1, min=0.1, max=0.9)
+    #linputs$coverage <-       runif(1, min=0.1, max=0.9)
+    #linputs$migration <-      runif(1, min=0.1, max=0.9)
 

### 2018-07-06 inst/documents/paper3/paper3_rotations.Rmd
for paper3

@@
   df_nocost_nomig <- df_res_all
-  load("df_res_all_rot10000.rda") #cost
+
+  #load("df_res_all_rot10000.rda") #cost
+  #df_cost <- df_res_all
+  #TEST looking at when freq limit removed :
+  load("df_res_all_rot_cost10000_no_freq_limit.rda")
   df_cost <- df_res_all
+
   load("df_res_all_rot_mig10000.rda") #mig 10k
@@
 Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide | 0.1 | 0.9
-Dominance of resistance | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide | 
+Dominance of resistance | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide | 0 | 1
 Frequency  | frequency of resistance alleles within the population | 0.005 | 0.1
 Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide | 0 | 0.1
-Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide | 0.1 | 0.9
+Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide | 0 | 1
 Insecticide number | number of independent insecticides available | 2 | 5  
@@
 *Ian : at this stage it would be good to revisit the following model details.*
-*It is fairly easy for me to rerun all the analyses if we change our mind on these.*
 
@@
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1.
+```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.cap="Figure 1.
 [todo need to modify caption from MJ paper, start by saying that only in left panel during application of the insecticide] The effect of model inputs on the fitness of genotypes for a single insecticide.
@@
 
-```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2.
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure 2.
 Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia.
@@
 linputs$plot <-     FALSE
+#5/7/18 try turning off
+linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
 
@@
 
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3.
+```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure 3.
 Comparing the increase in resistance frequencies over time for a rotation and a sequence when there are resistance fitness costs.
@@
 
-```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4.
+```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure 4.
 Comparing the increase in resistance frequencies over time for a rotation and a sequence with dispersal from untreated refugia.
@@
 
-```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.align='center', fig.cap="Figure 5.
+```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=9, fig.show='hold', fig.cap="Figure 5.
 Success of model runs under different scenarios and strategies.
@@
         ylab('proportion of model runs') +
-        facet_wrap(~scenario, ncol=1) + #, scales='free_y') + #can remove free_y later if I do 10k nocost runs    
+        facet_wrap(~scenario, nrow=1) + #, scales='free_y') + #can remove free_y later if I do 10k nocost runs    
+        theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))+
         geom_text(data=abcd,
@@
 
-```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.align='center', fig.cap="Figure 6.
+```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.cap="Figure 6.
 Effect of cost and dispersal on the difference between rotations and sequences from model runs.
@@
 
-```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 7.
+```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure 7.
 Example of a model run where there is a substantial benefit of rotations over a sequence.
@@
 For the rotation resistance levels for all insecticides are maintained well below resistance thresholds for the length of the simulation (500 generations or 50 years).
-Effectiveness=0.57, exposure=0.88, male exposure proportion=0.72, resistance restoration=0.68, dominance of resistance=0.04, cost=0.07, dominance of cost=0.59."}
+Effectiveness=0.57, exposure=0.88, male exposure proportion=0.72, resistance restoration=0.68, dominance of resistance=0.04, cost=0.07, dominance of cost=0.59, starting frequency=0.09."}
 # ? are these run ids before or after the runs that don't reach thresholds have been removed ?
@@
 
-```{r Fig8, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 8.
+```{r Fig8old, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure 8.
 Which model inputs have the greatest influence on the difference between rotations and sequences ?
@@
 Dominance of cost and dominance of selection have the greatest magnitude of influence."}
+
+#eval=FALSE because replaced by all scenarios in next chunk
 
@@
     #plot(pcc_res)
-
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
 
@@
 
+```
+
+\pagebreak
+
+```{r Fig8, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure 8.
+Which model inputs have the greatest influence on the difference between rotations and sequences ?
+For each scenario of resistance costs and dispersal.
+Measured using Partial Rank Correlation Coefficients (PRCC).
+Positive values indicate a positive influence on the benefit of rotations.
+In Resistance cost comes out as a negative impact on benefit of rotations.
+This could be because higher cost runs tend to be equal because they both sequences and rotations don't reach resistance thresholds.
+Maybe the PRCC is confusing ?
+Maybe Fig 5extra is clearer ?"}
+
+library(sensitivity)
+
+# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
+# 
+# Arguments
+# X a data frame containing model input variables
+# y a vector containing the responses 
+# rank logical. If TRUE, the analysis is done on the ranks.
+# nboot the number of bootstrap replicates.
+# conf the confidence level of the bootstrap confidence intervals.
+
+#in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs','coverage','migration')
+in_cols <- c('start_freqs','eff','expo_hi','male_expo_prop','rr','dom_sel','cost','dom_cos','coverage','migration')
+in_names <- c('start frequency','effectiveness','exposure','male exposure','resistance restoration','dominance of selection','cost','dominance of cost','coverage','migration')
+
+scenarios <- levels(df_all$scenario) #"no cost or dispersal" "costs" "dispersal" "costs and dispersal"
+#strat_name <- c("Sequential both","Adaptive both","Mix both")
+
+plotlist <- list(length(scenarios))
+
+for(scenario_num in 1:length(scenarios))
+{  
+  scen <- scenarios[scenario_num]
+
+  df_scenario <- filter(df_all, scenario==scen)
+
+    x <- df_scenario[,in_cols]
+    y <- df_scenario['pcent_rot_adv']
+
+    pcc_res <- pcc(x, y, rank=TRUE)
+
+    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
+    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
+    #plot(pcc_res)
+
+    to_plot <- pcc_res$PRCC
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    # sorts the ordering of x axis
+    to_plot$inputs <- factor(rownames(to_plot), levels=in_cols, labels=in_names)
+
+    #just labelling axis for final plot
+    if (scenario_num == length(scenarios))
+    {
+      axis_text_x <- element_text(angle = 45,hjust = 1, vjust = 1)
+      height <- 0.9
+    } else
+    {
+      axis_text_x <- element_blank() 
+      height = 0.5
+    }
+
+    library(cowplot) # to enable setting relative heights of plots
+
+    plotlist[[scenario_num]] <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
+             geom_point(colour='red') + #shape=1, 
+             theme(axis.text.x = axis_text_x) +
+             geom_hline(yintercept = 0, linetype=3) +
+             scale_y_continuous(breaks=c(-0.5,0,0.5)) +
+             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
+             #ylim(-1,1) +
+             ylim(-0.5, 0.5) +
+             #ggtitle(paste(strat_name[strategy_num])) +
+             ggtitle(scen) +      
+             xlab(NULL) 
+  }
+
   #to put back in some ref gridlines that are removed by cowplot
   #theme_set(theme_gray())
-  #theme_set(theme_bw())
+  theme_set(theme_bw())
 
   #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
-  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.6), labels='AUTO')
-
-```
-
-\pagebreak
-
-```{r Fig9, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 9.
+  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.6), labels='AUTO')
+
+```
+\pagebreak
+
+```{r FigX8, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure X8.
+PRCC of gens_rot modified from fig8 adding rot_interval and n_insecticides"}
+
+library(sensitivity)
+
+# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
+
+#in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs','coverage','migration')
+in_cols <- c('start_freqs','eff','expo_hi','male_expo_prop','rr','dom_sel','cost','dom_cos','coverage','migration','n_insecticides','rot_interval')
+in_names <- c('start frequency','effectiveness','exposure','male exposure','resistance restoration','dominance of selection','cost','dominance of cost','coverage','migration', 'num insecticides', 'rotation interval' )
+
+scenarios <- levels(df_all$scenario) #"no cost or dispersal" "costs" "dispersal" "costs and dispersal"
+#strat_name <- c("Sequential both","Adaptive both","Mix both")
+
+plotlist <- list(length(scenarios))
+
+for(scenario_num in 1:length(scenarios))
+{  
+  scen <- scenarios[scenario_num]
+
+  df_scenario <- filter(df_all, scenario==scen)
+
+    x <- df_scenario[,in_cols]
+    y <- df_scenario['gens_rot']
+
+    pcc_res <- pcc(x, y, rank=TRUE)
+
+    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
+    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
+    #plot(pcc_res)
+
+    to_plot <- pcc_res$PRCC
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    # sorts the ordering of x axis
+    to_plot$inputs <- factor(rownames(to_plot), levels=in_cols, labels=in_names)
+
+    #just labelling axis for final plot
+    if (scenario_num == length(scenarios))
+    {
+      axis_text_x <- element_text(angle = 45,hjust = 1, vjust = 1)
+      height <- 0.9
+    } else
+    {
+      axis_text_x <- element_blank() 
+      height = 0.5
+    }
+
+    library(cowplot) # to enable setting relative heights of plots
+
+    plotlist[[scenario_num]] <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
+             geom_point(colour='red') + #shape=1, 
+             theme(axis.text.x = axis_text_x) +
+             geom_hline(yintercept = 0, linetype=3) +
+             scale_y_continuous(breaks=c(-0.5,0,0.5)) +
+             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
+             #ylim(-1,1) +
+             ylim(-0.5, 0.5) +
+             #ggtitle(paste(strat_name[strategy_num])) +
+             ggtitle(scen) +      
+             xlab(NULL) 
+  }
+
+  #to put back in some ref gridlines that are removed by cowplot
+  #theme_set(theme_gray())
+  theme_set(theme_bw())
+
+  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
+  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.6), labels='AUTO')
+
+```
+
+\pagebreak
+
+```{r Fig9, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure 9.
 Model runs in which rotations perform more than 20% better than sequences are associated with high dominance of cost and low dominance of selection.
@@
      #shape=19 filled circle, 1=open circle
-     geom_point(data=filter(df_cost, pcent_rot_adv >= 20), aes(col=pcent_rot_adv), shape=19, size=1, alpha=0.9) +
+     geom_point(data=filter(df_cost, pcent_rot_adv >= 20), aes(col=pcent_rot_adv), shape=19, size=1.1, alpha=0.9) +
      viridis::scale_color_viridis(direction=-1) #discrete = TRUE)
@@
 
-```{r Fig10, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 10.
+```{r Fig10, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.cap="Figure 10.
 Simplified indication of how fitness differences between selective advantages and costs are expected to change according to dominance and resistance allele frequencies.

### 2018-07-06 inst/documents/paper3/recent_resist_literature.Rmd
for paper3

---
title: "Recent insecticide resistance literature for rotations and wos papers"

# output: pdf_document:

output: word_document
bibliography: library.bib
csl: biomed-central.csl

---

#### Gassmann 2009 fitness costs in Bt - Good
A review of 65 experimental studies of the costs of resistance to Bt toxins [@Gassmann2009] found fitness costs in around half. From 16 studies looking at fitness components they found that in 26% of comparisons fitness of rs was lower than that of ss indicating dominance of cost greater than zero. In four studies there was evidence that costs were dominant i.e. the fitness of the rs was equal to the rr. Based on these they suggest that for species with no data a recessive fitness cost of 25% be assumed for resistance to Bt.

Suppplemental table 4 shows that four studies and eight fitness components indicate dominant costs of resistance. Studies 6,8,13 and 16.

"Nonrecessive costs strongly favor decreased r allele frequency through selection against rs in refuges, despite selection on Bt crops favoring rr individuals. Accordingly, the dominance of fitness costs is critical, just as the dominance of resistance is critical."

#### Andow 2015 resistance to Bt in corn rootworm - Good
Good accessible description e.g. of high dose strategy.
[@Andow2015]

The different experiences with resistance to Bt maize in two different moth pests points to the importance of these population genetic parameters. Resistance has become a problem for the Western corn rootworm where resistance is not recessive, there are minimal costs of resistance and low dispersal. In contrast resistance has yet to become a problem for the European corn borer where resistance is recessive, there are higher costs of resistance and low dispersal. [@Andow2015]

(originally from tweet : https://twitter.com/alimanfoo/status/1014120657540575232)

High-dose in high-dose strategy means that all heterozgotes are killed and resistance is effectively recessive. But also defined operationally as 25 x the amount of toxin needed to kill 99.99% of a susceptible population.

#### Duke 2017 pesticide doses many implications
todo reread

#### Onstad 2001 rootworm model used to inform EPA  
Dominance of resistance was the most important parameter influencing evolution of resistance. If fully recessive (dominance 0) resistance not reached in 99 years, if fully dominant, resistance took just 2–5 years when refuges were 5% to 30% [@Onstad2001].

#### Carriere 2018
Measured the decline in frequencies of alleles conferring resistance to Bt toxins in pink bollworm raised in the absence of the toxin. By comparing observed and simulated changes in resistance they estimated that costs of resistance were between 14% and 36%. For one allele the results pointed to the costs being completely recessive and the other they suggested a small cost of 2% for the heterozygote suggesting a dominance of cost of 0.08.

The fitness costs of resistance to Bt toxins within the pink bollworm have been shown to be sufficient to reduce the frequency of resistance in the absence of the toxin but they are almost completely recessive. [@Carriere2018]

"When the frequency of resistance declines, selection against resistance alleles with recessive costs also declines because homozygous-resistant individuals become relatively rare. By contrast, selection against alleles with non-recessive costs is higher because heterozygotes remain relatively abundant. Accordingly, alleles inducing non-recessive costs should be eliminated at a faster rate than alleles inducing recessive costs." cites Falconer81 Intro to quantitative genetics.

"However, non-recessive costs of Bt resistance affecting fitness components seem relatively common. In 16 studies evaluating the dominance of costs, the percentage of non-recessive costs was 43% for development time, 21% for survival and 0% for growth rate.[@Gassmann2009]"

#### Tabashnik2005 Delayed resistance to Bt cotton
Monitoring of resistance in the pink bollworm to Bt toxins from transgenic plants showed no increase in resistance for 8 years despite widespread planting [@Tabashnik2005]. Combined experimental and simulation results suggest this could be due to untreated refuges, resistance being completely recessive, low resistance restoration, and resistance costs. A simulation showed that with 50% refuges and resistance restoration set to 0.6, fitness costs greater than 38% were required for resistance to decrease.

(Note that they assume that costs as well as resistance are recessive).

(We could try to produce something similar to Fig 2 in [@Tabashnik2005].)

#### Carriere 2001
Simple equation for whether resistance is expected to increase or decrease based on benefits and costs [@Carriere2001]. Be good for us to indicated how our model differs from this.

They found some similar things to us, e.g. resistance not increasing when dominance of resistance low and dominance of cost high.

#### Glunt 2018 Impact of temperature on insecticide toxicity.
Relevant to window of selection. 
In arabiensis & funestus with deltamethrin, generally mortalities higher when temperatures either higher or lower than standard. The difference between resistant and susceptible mortality is generally less, suggesting that the window of selction may be lower outside of standard lab conditions [@Glunt2018].

#### Hackett 2016 model of resistance to transgenic crops
Similar results to us on the importance of dominance of resistance and cost :
"the most significant constraint upon model behaviour is the dominance of the resistant phenotype, with greater dominance decreasing Bt usage. For a given level of dominance of resistance, the type and dominance of any fitness costs may relax these constraints, allowing for larger areas of Bt crop [smaller refuges to control resistance evolution]" [@Hackett2016]

#### Helps 2017 model of effect of high dose on resistance selection
*to do* look at this again. [@Helps2017] across a wide range of demographics etc.
"It is demonstrated that with high immigration resistance can be suppressed. This suppression how- ever, is rarely lost if the insecticide dose is reduced, and is absent altogether when individuals move from the treated population back into an untreated population. Reducing the dose of insecticide often resulted in slower development of resistance, except where the population combined a high influx of less resis- tant individuals into the treated population, a recessive resistance gene and a high efficacy, in which case reducing the dose of insecticide could result in faster selection for resistance.""

#### Immaraju 1990 Field Evaluation of Insecticide Rotation and Mixtures as Strategies for Citrus Thrips
[@Immaraju1990]

#### Kleinschmidt 2018 Implications of insecticide resistance for malaria vector control
WHO study
[@Kleinschmidt2018]

#### mbepera 2017 susceptibility higher in older anopheles
[@mbepera2017]

#### Curtis 1993 good quotes and references about rotations.
*todo, look at more*
It covers the oncheriasis control program which is taken by some to be evidence in favour of rotations.
Fig 3 & 4 is lab data comparing sequence & rotation, very similar to our plots, I should definitely cite it. [@Curtis1993]

Curtis cites how his 1987 work has already included effect of modifier genes, and shows it here in Fig 6.
"Supporters of prearranged rotations usually tacitly or explicitly assume that the general fitness due to resistance genes will improve if they are exposed to selection for many generations. Curtis (1987) simulated the behaviour of a modifer gene which could have this effect and found that it would only have an appreciable influence if one allowed resistance gene frequencies to become very high before switching insecticides.""

"One possible objection to the reactive system is that it requires an efficient and comprehensive system for testing for resistance to avoid ‘overshooting’ and
getting stuck as in the dotted line in Fig. 6. However, running a prearranged
rotation without a monitoring system would mean that one would have no way
of knowing if the system was still working."

## References
<!-- now auto created here-->

### 2018-07-06 inst/documents/rotations_notes.Rmd
for paper3

@@
 #rotations_notes.Rmd
-
-Just rough notes.
 
@@
 *3/4/18*
-
 ~ set up initial sensi analysis of 1000 runs
@@
 *4/4/2018*
-
 ooo with new ranges getting a few runs where there is a big advantage to rotation. Seems to be assoc with high dom_cos > 0.8
@@
 
-"One possible objection to the reactive system is that it requires an efficient and
-comprehensive system for testing for resistance to avoid ‘overshooting’ and
+"One possible objection to the reactive system is that it requires an efficient and comprehensive system for testing for resistance to avoid ‘overshooting’ and
 getting stuck as in the dotted line in Fig. 6. However, running a prearranged
@@
 I would like to get the Curtis 1987 paper
-Curtis CF. 1987. Genetic aspects of selection for resistance. In Ford MG, Holloman DW, Khambay BPS,
-Sawicki RM, eds. Combating resistance to xenobiotics. Chichester, U.K.: Ellis Horwood, 15e161.
+Curtis CF. 1987. Genetic aspects of selection for resistance. In Ford MG, Holloman DW, Khambay BPS, Sawicki RM, eds. Combating resistance to xenobiotics. Chichester, U.K.: Ellis Horwood, 15e161.
 
@@
 *27/4/2018*
-
 First draft of paper3
@@
 # 30/4/18
-
 100 runs with no cost all had identical times for rotations and sequences
@@
 *2/5/18*
-
 exchanging emails with Ian
@@
 Before sending to Ian.
-~ resize no square figs, make sure one per page
+~ resize non square figs, make sure one per page
 
@@
 
+*21/5/2018*
+Gould et al. Science paper.
+Wicked evolution: Can we address the sociobiological dilemma of pesticide resistance?
+
+If we are to address this recalcitrant issue of pesticide resistance, we must treat it as a “wicked
+problem,” in the sense that there are social, economic, and biological uncertainties and complexities
+interacting in ways that decrease incentives for actions aimed at mitigation. Here, we summarize
+the interacting factors and conclude with a call for government support of ambitious landscape-level experiments to assess which pesticide use strategies decrease resistance risks.
+
+In sexually reproducing weeds and insects, the rate of resistance evolution is strongly influenced by the relative fitness (dominant to recessive) of heterozygotes, and this sometimes depends on the dose of pesticide encountered in the field (Fig.
+3).
+Thus, it is difficult (and controversial) to determine whether resistance is expected to evolve more rapidly to higher or lower application concentrations of a pesticide [e.g., 16, 17].
+Even more complexity arises in attempts to predict resistance evolution when combinations of pesticides are applied (18, 19).
+Although the idea that such combinations will slow resistance evolution is theoretically controversial and lacks empirical support, mixtures are often recommended at the field level (15).
+
+Model-based analysis has shown that if disease vector resistance to pyrethroids
+becomes widespread, cases of malaria averted with ITNs could decline by 40% (25).
+Coupled with the estimate that bednets averted more than 65 million clinical malaria cases in sub- Saharan Africa in 2015 (7), and assuming that this figure provides a lower bound for potential cases averted in subsequent years, this would imply around 26 million additional clinical cases of malaria per year as a result of widespread vector resistance.
+Assuming an approximate lower bound cost of illness of at least $10 per malaria episode (26), insecticide resistance could conservatively cost sub-Saharan Africa at least $260 million per year.
+
+*24/5/18*
+replotted PRCC for all scenarios
+
+resistance cost comes out as a negative impact on benefit of rotations.
+This could be because higher cost runs tend to be equal because they both sequences and rotations don't reach resistance thresholds.
+Maybe the PRCC is confusing ?
+
+*25/5/18*
+While out for a run I thought that the constraint on allele frequencies not going below their starting values could have implications for difference between rotations & sequences.
+I did 10k cost simulations with the constraint removed to check.
+Either way I should include something about this in the discussion.
+
+*2/7/18* back to rotations after while on wos
+
+try to knit rotations paper get : error in do.call ... object 'run_rot' not found 
+Build & reload fixed it.
+
+Also the citations aren't being picked up now ? Because I messed up export of library.bib from Mendeley, exporting just a single selected reference.
+
+*4/7/2018* 
+looking at Hackett2016, downloaded R code which is <100 lines and the script works to produce the first figure.
+
+I was thinking that this could be a simpler model that we could also use with the same inputs that we use currently.
+This could possibly allow us to look at popn size, and to have a quick assessment of when resistance will increase and decrease.
+
+The growth of the pest population is described by a density-dependent selection model (Roughgarden 1971) which uses the two parameter density dependence function developed by Maynard Smith & Slatkin (1973).
+
+But maybe the way it works isn't appropriate for mosquitoes :
+
+1. Pests divided between refuge and toxic patch
+2. Pests within the Bt patch are exposed to toxins and mortality dependent on genotype.
+3. Pest mortality via intraspecific competition (incl. costs of resistance)
+4. Surviving larvae mature and mate at random to generate the next generation.
+
+In mosquitoes competition likely to occur at larval stage, maybe doesn't matter ?
+
 # TODO
 
-*today*
-done ~ reread Ians text about dominance
-~ move rotations repo back to my github
-~ update R
-done ~ tweet to malaria journal
-~ think more about rotations datavis
-done ~ read the alphey oxitech modelling papers
-
-To talk with Ian about next week :
-~ min_rwr_interval vs min_gens_switch_back
-
-Look at Hackett R code. Could it be an even simpler model that we could use ?
-The R code is here and looks quite accessible :
-https://datadryad.org//resource/doi:10.5061/dryad.4s405
+~ check on the runs with the freq limit removed :
+df_res_all_rot_cost10000_no_freq_limit.rda
+
+~ think about whether I can implement allele frequencies just being constrained to their starting values the first time that each insecticide is used.
+
+~ talk to Ian about Hackett2016 paper and whether we can use similarly simple model
+
+~ ask Ian to check through the rotations algebra when he writes up the SI.
+
+rotations paper :
+~ develop flow of story
+~ think about what questions managers will want answered
+
+A) when do both strategies work (include looking at insecticides and rotation level)
+ - maybe do PRCC
+B) if neither work, when are rotations better ?
+
+small changes :
+~ dominance of resistance
+~ how to term success. 'At least one insecticide still below resistance thresholds'
 
@@
 
-~ try decision tree to predict rot_minus_seq from inputs
-
 ~ modify sensi_an_rotations so that I have a reproduceable record of a) no cost or dispersal b)cost only c)dispersal only, d) cost & dispersal
@@
 think about killer plot to display difference between rot & seq generations
-
-~ *decide on min_rwr_interval for sensi analysis
-~ one generation per month, 12 generations per year, 500 gens is ~40 years
 
@@
 
-# good refs for paper
+# good refs for rotations paper
 resistance : malERA2017_updated-research-agenda-resistance-in-malaria.pdf
@@
 
+Slater2016. Use of an individual-based simulation model to explore and evaluate potential insecticide
+resistance management strategies
+
 # state of play

### 2018-07-06 inst/documents/sensi_an_rotations1.Rmd
for paper3

@@
 
-    # for Ians Q1 set coverage to 1 & cost to 0 
-    # April 2018, having go at including variable cost, may want to do some cost=0 runs later
-
     ## constant inputs ##
@@
     linputs$migration <- 0
-    linputs$cost <- 0 #variable cost below
+    #linputs$cost <- 0 #variable cost below
     linputs$max_gen <- max_gen
@@
     #beware min_rwr_interval, it's to stop repeated small rotations
-    linputs$min_rwr_interval = 1
-    linputs$min_gens_switch_back <- 5
+    linputs$min_rwr_interval <- 5
+    linputs$min_gens_switch_back <- 1
+    linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
     #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
@@
     ## variable inputs ##
-    #linputs$cost <-           runif(1, min=0,   max=0.1) 
+    linputs$cost <-           runif(1, min=0,   max=0.1) 
     #linputs$cost <-           0 

### 2018-07-06 R/run_rot.r

@@ -277,6 +277,8 @@ run_rot <- function(max_gen = 200,
     # to work with single start_freqs value too
     if (length(start_freqs) == 1) start_freqs <- rep(start_freqs, n_insecticides)
     
+    
+    #todo implement a no_r_below_mut option here
     if ( no_r_below_start )
     {
       for(insecticide in 1:n_insecticides)

### 2018-07-11 inst/documents/rotations_notes.Rmd
adding mutation-selection balance

@@
 
+*9/7/18* mutation-seln balance
+
+not quite working in run_rot() yet
+
+run_rot(no_r_below_start=FALSE,no_r_below_mut = TRUE)
+
+Ian said that mutation rate of 10e-9 is used in malaria studies. John Maynard Smith, Evolutionary Genetics p62 suggests mutation rate per generation per locus from drosophila of 10e-6.
+
+Ian suggested that he expected these frequencies to be so low that using them as a lower limit will not effect simulations.
+
+Checking ability to do runs where insecticides differ.
+Code at end of sensi_an_rotations1
+
+Currently getting these warnings, & seemingly more rows than scenarios in df_res_all ...
+1: In if (dom_cos == 0) { ... :
+  the condition has length > 1 and only the first element will be used
+
+fixed vectorising mutn_seln_bal.r
+
+Running for insecticides with different inputs now works :
+dfr <- run_rot(n_insecticides=3, eff=c(1,0.6,0.4), rot_interval=0)
+
+but I don't get one row per scenario in df_res_all from sensi_an_rotations
+
+for 10 runs I get 31 rows
+because I'm getting 1 row per insecticide in each run
+where does that happen ?
+seems it happens in this line :
+
+df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq))
+
+when linputs contains some elements that are vectors (repeated for each insecticide) the resulting dataframe has one row per insecticide.
+
+I would rather have one row per scenario and have all the vector elements saved to one record, perhaps as a comma delimited string ?
+
+seems that as.data.frame is the culprit. what about tidyverse option ? no didn't work
+
+this is the issue I have :
+as.data.frame(c(double=c(9,6), single=7))
+
+converts to this :
+double1                               9
+double2                               6
+single                                7
+
+when I want something like 
+double1                               9,6
+single                                7
+
 # TODO
+
+~ ian interested in starting frequencies at 10e-4, can do that for the runs with different values per insecticide.
+~ ian suggested talking to Simon if we need to run something for longer
+
+1. check out these runs that have the mut-sel balance option enabled
+df_res_all_rot_cost10000_mut_sel_bal.rda
+2. check ability to do runs where insecticides differ
+
+Currently getting these warnings, & seemingly more rows than scenarios in df_res_all ...
+1: In if (dom_cos == 0) { ... :
+  the condition has length > 1 and only the first element will be used
+
+~ think - are all rotation runs so far based on insecticides with equal effectiveness, resistance restoration, cost etc.
+?
+What might happen if we allow differences between insecticides ?
+Can we easily represent that ?
+I could possibly run one set of simulations where insecticides are allowed to be different.
+I think that could be set up by changing :
+linputs$expo_hi <-  runif(1, min=0.4, max=0.9)
+to:
+runif(n_insecticides, min=0.4, max=0.9)
+This might break stuff.
+Can just check that prop runs in which rotations are better doesn't increase.
+I suspect input params may be saved to outfiles as comma delimited strings ?
 

### 2018-07-11 inst/documents/sensi_an_rotations1.Rmd
adding mutation-selection balance

@@
 plot_scenarios  <- FALSE
-n_scenarios     <- 10000 
+n_scenarios     <- 10 #000 
 already_ran     <- FALSE
@@
     linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
+    linputs$no_r_below_mut <- TRUE #BEWARE trying out new mutation-selection balance option
     #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
@@
     # add scenario id column onto results
-    res$scenario <- i
+    # this isn't doing anything
+    # res$scenario <- i
     df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq))
@@
 ```
+
+# Have a go at running for insecticides with different parameters
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
+
+# set random seed so that results can be reproduced
+set.seed(2)
+max_gen         <- 500
+plot_scenarios  <- FALSE
+n_scenarios     <- 10 #000
+
+  df_res_all <- data.frame()
+
+  for(i in 1:n_scenarios)
+  {
+    #message should output to the R markdown console
+    if (i%%200 == 1) message("scenario ",i," of ",n_scenarios," ",Sys.time())
+
+    # inputs (could also be saved to a dataframe first & extracted one row at a time)
+    linputs <- list()
+
+    ## constant inputs ##
+    linputs$coverage <- 1
+    linputs$migration <- 0
+    #linputs$cost <- 0 #variable cost below
+    linputs$max_gen <- max_gen
+
+    #beware min_rwr_interval, it's to stop repeated small rotations
+    linputs$min_rwr_interval <- 5
+    linputs$min_gens_switch_back <- 1
+    linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
+    linputs$no_r_below_mut <- TRUE #BEWARE trying out new mutation-selection balance option
+    #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
+    linputs$plot <-     FALSE
+
+    ## variable inputs ##
+    linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
+    n_ins <- linputs$n_insecticides
+
+    ## new those below can be different for each insecticide
+    linputs$cost <-           runif(n_ins, min=0,   max=0.1) 
+    #linputs$cost <-           0 
+    linputs$expo_hi <-        runif(n_ins, min=0.4, max=0.9)  #min=0.1, max=0.9
+    linputs$male_expo_prop <- runif(1, min=0,   max=1)
+    linputs$eff <-            runif(n_ins, min=0.5, max=1) # min=0.3, max=1
+    linputs$rr  <-            runif(n_ins, min=0.1, max=0.9)
+    linputs$dom_sel <-        runif(n_ins, min=0,   max=1)
+    linputs$dom_cos <-        runif(n_ins, min=0,   max=1) 
+    linputs$start_freqs <-    runif(n_ins, min=0.005,   max=0.1) #note not log yet
+
+    # dispersal (initially with cost set to 0)
+    #linputs$coverage <-       runif(1, min=0.1, max=0.9)
+    #linputs$migration <-      runif(1, min=0.1, max=0.9)
+
+    # create null objects used later for plotting
+    ggrot <- ggseq <- NULL
+
+    # to compare sequence to rotation run all the scenarios above for both seq & rot 
+    for(rot_or_not in 0:1)
+    {  
+      #to get rot_int of 0 every other run
+      if (rot_or_not == 0) linputs$rot_interval <- 0
+      else   #random rot interval integer
+          linputs$rot_interval <-   sample( 5:50, 1 )  #beware set replace=T if more than 1
+
+      ##################
+      # run one scenario
+      dfres <- do.call(run_rot, linputs)
+
+      # summarise results per scenario
+      # note this is also currently calc at end of run_rot() here is probably better
+
+      res <- gens_under_thresh(dfres)
+
+      # to plot results for checking
+      if ( linputs$rot_interval == 0 )
+      {
+        gens_seq <- as.numeric(res)
+      } else
+      {
+        gens_rot <- as.numeric(res)
+      }
+
+      if (plot_scenarios)
+      {
+        # make plot
+        ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
+
+        # to plot results for checking
+        if ( linputs$rot_interval == 0 )
+        {
+          ggseq <- ggres
+          gens_seq <- as.numeric(res)
+        } else
+        {
+          ggrot <- ggres
+          gens_rot <- as.numeric(res)
+        }
+      }
+    } #end rot_or_not loop
+
+    # make an output row for each scenario with inputs, gens_rot, gens_seq
+    # need to make sure I retain the rot_interval for rot
+
+    # add scenario id column onto results
+    # this isn't doing anything
+    # res$scenario <- i
+    df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq))
+    df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
+
+    # rbind this scenario inputs & results onto previous
+    # not a good speed or memory way of doing ... but a start
+    df_res_all <- rbind(df_res_all, df_in_out)
+
+    if (plot_scenarios)
+    {  
+      # plot input vals
+      # extract out those between 0 & 1
+      dfi <- data.frame(t(as.data.frame(linputs)))
+      names(dfi) <- 'value'
+      dfi$inputs <- rownames(dfi)
+
+      #take out columns not between 0 & 1
+      dfi <- filter(dfi, !inputs %in% c('max_gen','rot_interval','n_insecticides','plot','min_rwr_interval'))
+      #also remove coverage & migration both set to 0 for now
+      dfi <- filter(dfi, !inputs %in% c('coverage','migration'))
+
+      ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
+                geom_point() +
+                theme_minimal() +
+                ylim(0,1) +
+                theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))
+
+      grid.arrange(ggrot, ggseq, ggins, ncol=1)
+
+      cat('scenario',i,'sum generations deployed insecticide under thresh rotation:', gens_rot,' sequence:',gens_seq,"\n\n")
+
+      #can I give the code to rerun this scenario
+      cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
+                  "cost=",signif(linputs$cost,2),", ",
+                  "start_freqs=",signif(linputs$start_freqs,2),", ", 
+                  "rot_interval=",linputs$rot_interval,", ",
+                  "eff=",signif(linputs$eff,2),",\n",
+                  "dom_sel=",signif(linputs$dom_sel,2),", ",
+                  "dom_cos=",signif(linputs$dom_cos,2),", ",
+                  "rr=",signif(linputs$rr,2),", ",
+                  "expo_hi=",signif(linputs$expo_hi,2),", ",
+                  "coverage=",signif(linputs$coverage,2),",\n",
+                  "migration=",signif(linputs$migration,2),", ",
+                  "max_gen=",linputs$max_gen,", ", 
+                  "min_rwr_interval=",linputs$min_rwr_interval,            
+                  ")\n") )
+
+    }
+
+    #I could save partial outputs every 200 scenarios
+    #just in case doesn't make to end
+    if (i%%1000 == 0)
+    {
+      #df_res_all$rot_minus_seq <- df_res_all$gens_rot - df_res_all$gens_seq
+
+      # save object containing inputs and results as rda for analysis
+      save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))    
+    }
+
+  } #end scenario loop
+
+```

### 2018-07-30 inst/documents/sensi_an_rotations1.Rmd
more efficient results storage and add scenario id column

@@
     # add scenario id column onto results
-    # this isn't doing anything
-    # res$scenario <- i
-    df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq))
+    df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, id=i))
     df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
@@
     # not a good speed or memory way of doing ... but a start
-    df_res_all <- rbind(df_res_all, df_in_out)
+    #df_res_all <- rbind(df_res_all, df_in_out)
+
+    #instead set up the dataframe to the required size in scenario1
+    if (i==1)
+    {
+      df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=n_scenarios))
+      names(df_res_all) <- names(df_in_out)
+    }
+
+    #putting results into one row of overall results dataframe
+    df_res_all[i,] <- df_in_out
 
@@
 # Have a go at running for insecticides with different parameters
-
+# starting to work ...
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
@@
 plot_scenarios  <- FALSE
-n_scenarios     <- 10 #000
+n_scenarios     <- 100 #1000 #000
 
-  df_res_all <- data.frame()
+  #df_res_all <- data.frame()
 
@@
     linputs$dom_cos <-        runif(n_ins, min=0,   max=1) 
-    linputs$start_freqs <-    runif(n_ins, min=0.005,   max=0.1) #note not log yet
+    #DIFFERENT
+    linputs$start_freqs <-    runif(n_ins, min=0.0001,   max=0.01) #note not log yet
 
@@
 
-    # add scenario id column onto results
-    # this isn't doing anything
-    # res$scenario <- i
-    df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq))
+    # to cope with multiple values per scenario convert them to strings
+    linputs2 <- lapply(linputs, function(x) ifelse(length(x)>1,toString(x),x))
+
+    #added scenario id column
+    df_in_out <- as.data.frame(c(linputs2, gens_rot=gens_rot, gens_seq=gens_seq, id=i), stringsAsFactors=FALSE)
+
     df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
@@
     } 
-
   } #end scenario loop
 
+#to allow plotting of chosen scenarios by re-running them 
+#initially didn't work for diff insecticides because the inputs are saved as strings 
+#now the strings get converted back to vectors below
+
+scenarios_to_plot <- c(13)
+
+for(scenario in scenarios_to_plot)
+{
+  #cat(scenario,"\n")
+  linputs <- as.list( df_res_all[scenario, 1:15] )
+
+  #converting multiple insecticide inputs back from strings
+  #have to ignore the ones that aren't strings
+  #linputs2 <- lapply(linputs, function(x) as.numeric(unlist(strsplit(as.character(x),split=", "))))
+
+  #note the space after the comma in split=", "
+  linputs <- lapply(linputs, function(x) if (!grepl(",",x)) x else
+                                          as.numeric(unlist(strsplit(as.character(x),split=", "))))
+
+  #TESTING changing min_rwr_interval from 5
+  linputs$min_rwr_interval <- 0
+
+  # run one scenario (rotation)
+  dfres <- do.call(run_rot, linputs)
+
+  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+
+  #now run as sequence
+  linputs$rot_interval <- 0
+  dfres <- do.call(run_rot, linputs)
+
+  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
+
+  grid.arrange(ggrot, ggseq, ncol=1)
+
+}
+
 ```

### 2018-07-30 inst/documents/paper3/recent_resist_literature.Rmd
try different insecticides

@@
 
-#### Helps 2017 model of effect of high dose on resistance selection
+#### Helps 2017 model of effect of high dose on resistance selection in agriculture
 *to do* look at this again. [@Helps2017] across a wide range of demographics etc.
+for wos
+"One management strategy that has long been advocated is the application of insecticides at the maximum permitted dose.
+This has been found, under some circumstances, to be able to prevent the resistance allele frequency from increasing.
+However this approach may, under different circumstances, lead to rapid selection for resistance to the insecticide."
+
 "It is demonstrated that with high immigration resistance can be suppressed.
-This suppression how- ever, is rarely lost if the insecticide dose is reduced, and is absent altogether when individuals move from the treated population back into an untreated population.
-Reducing the dose of insecticide often resulted in slower development of resistance, except where the population combined a high influx of less resis- tant individuals into the treated population, a recessive resistance gene and a high efficacy, in which case reducing the dose of insecticide could result in faster selection for resistance.""
+This suppression however, is rarely lost if the insecticide dose is reduced, and is absent altogether when individuals move from the treated population back into an untreated population.
+Reducing the dose of insecticide often resulted in slower development of resistance, except where the population combined a high influx of less resis- tant individuals into the treated population, a recessive resistance gene and a high efficacy, in which case reducing the dose of insecticide could result in faster selection for resistance."
+
+"Because insecticides can have adverse impacts on non-target organisms, regulators have to ensure that the maximum permitted dose on the product la- bel is also the minimum dose required to achieve effective control ( Anon, 2012 ).
+The process of determining that dose was described by Finney (1993) , as follows.
+In farm crops, there are many fac- tors that affect the level of control achieved with any given dose.
+Hence, if dose-response experiments are repeated across many sites and seasons, many different dose-response curves result.
+For practical purposes, the label recommended dose is usually set at the level which gives a high level of control in a high proportion of circumstances.
+Finney suggested 80–90% control in 80–90% of field experiments as a typical aim, although higher levels of control may be required for insect pests, for example, to prevent insect contam- ination in horticultural crops.
+There are two consequences of this process of determining the full label dose which are relevant to the analysis presented here.
+Firstly, the label dose will not be set at a level which gives 100% control, even of sensitive strains, since, as previously mentioned, this is not the aim of the maximum al- lowed dose specified on a product label.
+Hence, the modelling ap- proach presented assumes that it is not possible to drive resistant strains to extinction by insecticide treatment, within the range of doses permitted."
 
@@
 
+#### Venter2017 Benchmarking insecticide resistance intensity bioassays
+The operational effect of insecticides will depend on doses and the resistancve phenotypes of mosquitoes.
+Resistance intensity assays are a way of qauntifying this, here intensity assays are related to the resistance phenotypes of lab strains with characterised resistance mechanisms.
+probably Cite this in windows paper.
+
+#### Mbida2018 nets collected from the field against kdr resistant mosquitoes
+Shows that field collected nets kill all susceptible mosquitoes but low proportions of field collected mosquitoes.
+Genotyping the field collected mosquitoes exposed to the field collected nets (Table 4) showed that 7-35 % of the surviviors had the susceptible kdr allele and none of those killed had the resistance allele.
+This demonstrates that the nets are within the window of selection and will be selecting for kdr resistance.
+
+#### Denholm1992
+In an agricultural context points out that theory suggests that rotations and sequences should be equivalent.
+"Despite theoretical misgivings regarding the benefit of alternations, the prevalence of this tactic in existing strategies on cotton and elsewhere demonstrates its intuitive appeal and relative ease of implementation." [@Denholm1992]
+
 ## References

### 2018-07-30 inst/documents/rotations_notes.Rmd
try different insecticides

@@
 
+*13/7/18*
+
+ltst <- list(d=c(9,6),s=7)
+
+This nearly does what I want but loses names of columns
+paste0(ltst)
+[1] "c(9, 6)" "7"
+
+cool this does it :
+lapply(ltst,toString)
+$`d`
+[1] "9, 6"
+$s
+[1] "7"
+
+linputs2 <- lapply(linputs, toString)
+df_in_out <- as.data.frame(c(linputs2, gens_rot=gens_rot, gens_seq=gens_seq))
+
+Seems that with variable insecticides there are more runs where sequences are better (i.e. highly negative rot_minus_seq) ?
+
+See :
+df_res_all_rot1000-diffI-lowstart.rda
+
+*17/7/18*
+To be able to plot specific scenarios with different insecticides I need to get the strings that I saved above back into vectors.
+
+strsplit seems it should do the job :
+
+tst <- as.character(linputs$rr)
+unlist(strsplit(tst,split=", ")) #note the space after the comma
+
+*30/7/18*
+continuing getting it to cope with multiple insecticides in sensi_an_rotaions
+
+I'm trying to convert string representations of mutliple parameters (that were needed to save in an output file) back to vector versions so I can rerun function with them.
+
+  linputs2 <- lapply(linputs, function(x) ifelse(grepl(",",x),
+    as.numeric(unlist(strsplit(as.character(x),split=", "))),x))
+
+Currently I just get the first value
+> linputs$cost
+[1] "0.0209911216050386, 0.00808626485522836, 0.0159290118375793"
+> linputs2$cost
+[1] 0.02099112
+
+There was something a bit weird about the ifelse that meant just the first dimension was returned.
+This now works :
+
+  linputs <- lapply(linputs, function(x) if (!grepl(",",x)) x else
+                                          as.numeric(unlist(strsplit(as.character(x),split=", "))))
+
+Examples with different insecticides where sequences last much longer than rotations seem to be caused by one of the 3 insecticides lasting much longer in the sequence, in the rotation it gets to a stage where it is the only one remaining, and our changing algorithm doesn't allow switching back to it from itself.
+
 # TODO
 
+~ and run_id to outputs in sensi_an_rotations to make things easier in future, e.g. to identify runs with particular behaviour.
+
+~ check on 1000 runs with lower startfreqs and differing insecticides 
+~ is something weird going on favouring rotations more when insecticides differ ?
+~ yes
+
 ~ ian interested in starting frequencies at 10e-4, can do that for the runs with different values per insecticide.
+~ we won't run sensitivity analysis on those runs, but just to check that rotations and sequence runs still similar
 ~ ian suggested talking to Simon if we need to run something for longer
@@
 df_res_all_rot_cost10000_mut_sel_bal.rda
-2. check ability to do runs where insecticides differ
-
-Currently getting these warnings, & seemingly more rows than scenarios in df_res_all ...
-1: In if (dom_cos == 0) { ... :
-  the condition has length > 1 and only the first element will be used
+done 2. check ability to do runs where insecticides differ
 

### 2018-07-31 inst/documents/rotations_notes.Rmd
allow insecticides to continue to be used in a rotation if there are no other options remaining

@@
 
+~ added id to outputs in sensi_an_rotations to make things easier in future, e.g. to identify runs with particular behaviour.
+
+~ also made more efficient by declaring size of results df at start of loop
+
+The scenarios with the biggest differences seem to be where insecticide1 is the better one.
+Just want to check it isn't something else screwy ???
+
+*31/7/18*
+Getting this error pushing to github :
+0 [main] us 0 init_cheap: VirtualAlloc pointer is null, Win32 error 487
+AllocationBase 0x0, BaseAddress 0x68570000, RegionSize 0x340000, State 0x10000
+C:\Program Files (x86)\Git\bin\ssh.exe: *** Couldn't reserve space for cygwin's heap, Win32 error 0
+fatal: Could not read from remote repository.
+
+Please make sure you have the correct access rightsand the repository exists.
+
+updating RStudio didn't fix, had once before and rebooting fixed it then ...
+
+rebooting didn't fix, this post suggested updating git to v2 from https://gitforwindows.org/
+https://stackoverflow.com/a/31970708/1718356
+
+git --version 
+git version 1.9.5.msysgit.0
+
+now I have 2.18.0.windows1, the installer sorted uninstalling the old version, and I didn't have to change anything in RStudio.
+
 # TODO
 
-~ and run_id to outputs in sensi_an_rotations to make things easier in future, e.g. to identify runs with particular behaviour.
+~ try a solution to the diff insecticide problem. If no other insecticides are below resistance thresh and this one is then can keep using it in a rotation ...
 
@@
 
-Slater2016. Use of an individual-based simulation model to explore and evaluate potential insecticide
-resistance management strategies
+Slater2016. Use of an individual-based simulation model to explore and evaluate potential insecticide resistance management strategies
 

### 2018-07-31 inst/documents/sensi_an_rotations1.Rmd
allow insecticides to continue to be used in a rotation if there are no other options remaining

@@
 
+scenarios_to_plot <- c(21,23,32) #big difference ones all have insecticide1 as the good one
+scenarios_to_plot <- c(29,38,46)
 scenarios_to_plot <- c(13)
@@
   #TESTING changing min_rwr_interval from 5
-  linputs$min_rwr_interval <- 0
+  #if min_rwr_interval is set to 0 you get frequent switch backs to other insecticides 
+  #for 1 generation
+  #but when it's set to 5it pushes frequences above the resistance freq threshold
+  linputs$min_rwr_interval <- 5
+  #min_gens_switch_back has to be 0 to allow one insecticide to continue when all others dead
+  linputs$min_gens_switch_back = 0
 

### 2018-07-31 R/insecticide_check.r

@@ -32,8 +32,6 @@ insecticide_check <- function( RAF1gen,
   {  
     #BEWARE that switch criterion is female only
     
-    #TODO later alternative stop switching back to insecticide that has been used within 5 generations
-    
     # because have drop=false to preserve site dimension the gen dimension is preserved too
     # so just has a single item in the final dimension : RAF1gen[,,,1]
     
@@ -43,27 +41,51 @@ insecticide_check <- function( RAF1gen,
          #e.g. can stop simulation ending when resistance threshold has been reached
          gens_this_insecticide > min_rwr_interval )
     {
-      change_insecticide <- TRUE        
+      change_insecticide <- TRUE 
+      
+      #message(paste0("change from insecticide",current_insecticide, " freq=",RAF1gen[current_insecticide, 'f','intervention',1],"\n"))
     }
     #message(sprintf("confirm. RAF=%f, change_insecticide=%d \n", RAF1gen[current_insecticide, 'f','intervention'], change_insecticide))
   } 
 
   # even if in a rotation it my have assessed in previous loop that insecticide needs to be changed  
-  if (rot_interval != 0)
-  # if a rotation  
+  if (rot_interval != 0)   # if a rotation 
   {
-    if (gens_this_insecticide != rot_interval) 
-    {
-      # keeps the rotation going
-      # not needed here because done in run_rot()
-      gens_this_insecticide <- gens_this_insecticide+1  
-      # to make responsive rotation add a check of resistance frequency here
-      # and stop the rotation if threshold exceeded
+    # if (gens_this_insecticide != rot_interval) 
+    # {
+    #   # keeps the rotation going
+    #   # not needed here because done in run_rot()
+    #   gens_this_insecticide <- gens_this_insecticide+1  
+    #   # to make responsive rotation add a check of resistance frequency here
+    #   # and stop the rotation if threshold exceeded
+    # } else 
       
-    } else 
-    {
-      # time to rotate so need to identify the next insecticide in the rotation
-      change_insecticide <- TRUE 
+    #end of rotation reached
+    if (gens_this_insecticide >= rot_interval) 
+      {
+      # 31/7/18 I could add a check in here for case
+      # where only this insecticide remains below resistance threshold
+      # in which case don't want to change
+      # if I reset gens-this-insecticide then it will be used for another rot_interval
+      #I can use either of these to get the freqs for all other insecticides
+      # RAF[-current_insecticide, 'f','intervention',1]
[…]

### 2018-12-06 inst/documents/paper3/recent_resist_literature.Rmd
notes

@@
 ---
+
+#### Record of which publications show mortality vs concentration for RR SR SS
+[@Gould2018, @Bourguet2000, @Mallett1989]
+Dose response plots for RR,SR and SS showing change in effective resistance with increasing concentration are shown in a large number of publications but no one has turned them around the other way and related them to time.
 
@@
 
-#### Duke 2017 pesticide doses many implications
-todo reread
+#### Tabshnik2017 review of insecticide resistance to Bt crops, another good accessible summary
+In a review of insect resistance to engineered plants they found that 0 out of 17 studies where resistance has been detected met the high-dose standard, in contrast where susceptibility remains, 9 out of 13 cases did not meet the high-dose standards.
+[@Tabashnik2017].
+Points to low doses prompting evolution of resistance through reduced dominance.
+One example also suggests that having larger refuges can compensate for the high-dose standard not being met.
+
+Concludes that the main lesson overall is that abundant refuges delay resistance to transgeni crops.
+In a public health context this points to the importance of understanding the proportion of insect vector populations that are exposed to insecticide interventions.
+It also warns of a potential increase in the evolution of resistance as intervention coverage is increased for example as a part of malaria elimination efforts.
+
+#### Gressel 2017
+High doses likely to select for resistance conferred by single genes of major effect whereas low doses can allow the accumulation of many genes of minor effect. [@Gressel2017].
 
@@
 In arabiensis & funestus with deltamethrin, generally mortalities higher when temperatures either higher or lower than standard.
-The difference between resistant and susceptible mortality is generally less, suggesting that the window of selction may be lower outside of standard lab conditions [@Glunt2018].
+The difference between resistant and susceptible mortality is generally less, suggesting that the window of selection may be lower outside of standard lab conditions [@Glunt2018].
 
@@
 #### Venter2017 Benchmarking insecticide resistance intensity bioassays
-The operational effect of insecticides will depend on doses and the resistancve phenotypes of mosquitoes.
+The operational effect of insecticides will depend on doses and the resistance phenotypes of mosquitoes.
 Resistance intensity assays are a way of qauntifying this, here intensity assays are related to the resistance phenotypes of lab strains with characterised resistance mechanisms.
-probably Cite this in windows paper.
+probably cite this in windows paper.
+
+#### lopezmonroy2018
+Relates the frequency and intensity of pyr resistance in aedes to the frequency of kdr genes. Useful in translating between our modelling work and field measurements of resistance.
 
@@
 
+#### Matowo2015 rapid increase in Bendiocarb resistance following IRS, support for our modelling approach
+There was a sharp decrease in mortality in An.
+gambiae s.l.
+exposed to bendiocarb (carbamate) from 84% in November 2011 to 31% in December 2012 after two rounds of bendiocarb-based indoor residual spraying (IRS).
+Anopheles gambiae s.l.
+remained susceptible to pirimiphos-methyl
+(organophosphate). Bendiocarb-based IRS did not lead to the reversion of pyrethroid resistance.
+We could maybe model this. They have some data on various gene frequencies.
+
+#### Gullberg2011 Selection of Resistant Bacteria at Very Low Antibiotic Concentrations
+Shows windows of selection for antibiotic resistance in bacteria and how these extend to very low concentrations where the antibiotics are not effective. [@Gullberg2011]
+
 ## References

### 2018-12-06 inst/documents/rotations_notes.Rmd
notes

@@
 
+All rotation runs before were based on insecticides with equal effectiveness, resistance restoration, cost etc. These new runs allow differences between insecticides.
+
+One thing that they showed was that a sequence could do much better than a rotation due to the logic of when to change an insecticide.
+
+because just one insecticide remained below the resistance threshold and our logic didn't allow it to keep being used. These were generated when I allowed insecticides to differ.
+Below is the result for the same inputs of adding a condition to check all the other insecticides, and to allow us to keep using one if it is the last remaining.
+I have set the minimum rotation interval to 5 otherwise it keeps switching back to the other insecticides for a single generation.
+This gives more similar results for the rotation and sequence.
+
+*1/8/2018*
+Reading Curtis(1987) Genetic aspects of selection for resistance. Good short article.
+Summary :
+~ rotations and sequences very similar
+~ fitness modifiers unlikely to give benefit to rotations (unless resistance thresholds and costs very high)
+~ there maybe operational reasons for favouring sequences
+
+I should try to make sure our rotations paper is as easy to read as this.
+
+It is a good justification for not using modifiers so I don't think we need to implement modifiers but if we did, this is how he describes what he did - although all the details not there.
+"assumed a single modifier of each major resistance gene.
+The fitness disdavantage of each major gene was eliminated when the modifier was homozygous and halved when it was heterozygous.
+The modifier itself caused slight fitness disadvantage and was initially maintained in equilibrium by a blance of origin by mutation and elimination by selection."
+
+Other papers in the book:
+Denholm, Laboratory simulation of selection for resistance.
+Pyrethroid resistance in houseflies.
+Compares model results and lab experiments.
+Concludes that more field work needed before models can be operationally useful.
+
+Dennehy, Decision making for managing pest resistance to pesticides. Brief possibly useful stuff on decision making criteria in agriculture.
+
+Sawicki, chap9. Definition, detection and documentation of insecticide resistance.
+
+Tried a solution to the diff insecticide problem. If no other insecticides are below resistance thresh and this one is then can keep using it in a rotation ...
+
+*12/11/2018*
+Coming back to this after a while before presenting to IVCC.
+
+from UI :
+Warning: Error in raf_get: object 'RAF' not found
+but works for 2nd plot on right ?
+
+must be from run_rot()
+I don't understand why sceanrio A fails and B doesn't ?
+
+  178: raf_get [C:/rsprojects/rotations/R/insecticide_check.r#70]
+  177: insecticide_check [C:/rsprojects/rotations/R/insecticide_check.r#70]
+  176: run_rot [C:/rsprojects/rotations/R/run_rot.r#325]
+
+Also fails from the command line :
+
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 10 , eff = 0.8 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.01 , logy = TRUE , no_r_below_start = TRUE , min_rwr_interval = 5 )
+
+Error in raf_get(RAF, insecticide = -current_insecticide, sex = "f", gen = 1,  : 
+  object 'RAF' not found
+
+BUT now seems to be working from the command line ???
+and is working from shiny ... weird.
+
+this error then popped up again from shiny later on.
+
+*13/11/2018*
+preparing slides for IVCC talk next week
+
+~ add a slide about decision rules :
+because just one insecticide remained below the resistance threshold and our logic didn't allow it to keep being used. 
+Below is the result for the same inputs of adding a condition to check all the other insecticides, and to allow us to keep using one if it is the last remaining.
+I have set the minimum rotation interval to 5 otherwise it keeps switching back to the other insecticides for a single generation.
+This gives more similar results for the rotation and sequence.
+
+23/11/18
+#to output script as pdf for Ian to print
+library(knitr)
+stitch(script="R/run_rot.r")
+
+stitch(script="R/run_rot.r", system.file("misc", "knitr-template.Rhtml", package="knitr"))
+stitch(script="R/rot_migrate.r", system.file("misc", "knitr-template.Rhtml", package="knitr"))
+
+Then open html in a browser, print to pdf, set landscape and scale=150
+
+4/12/18
+Met with Jim he offered to look at parallelising.
+
+5/12/18
+making scaled down version of sensi analysis for Jim
+
+6/12/18
+Check why git changes are thinking that whole file has changed. How did I fix that for rnaturalearth ??
+
+GitHub suggests that you should make sure to only use \n as a newline character in git-handled repos. There's an option to auto-convert:
+
+git config --global core.autocrlf true #global didn't work
+git config core.autocrlf true #just this project
+
+Doing it for rotations project did fix the problem straight away. I didn't even have to restart RStudio.
+
 # TODO
 
-~ try a solution to the diff insecticide problem. If no other insecticides are below resistance thresh and this one is then can keep using it in a rotation ...
+For rotations send Jim a scaled down version of the rotations sensitivity analysis code. That just runs the sensitivity analysis and Jim will have a go at parallelising it.
+
+Still occasionally getting error from shinyUI resistrot :
+Warning: Error in raf_get: object 'RAF' not found
+
+It is only from scenarioA.
+Could it be something about sequence versus rotations ?
+Yes it doesn't happen when rot_interval=0 but did happen when rot_interval=10.
+But then when I start trying to look at it even from a new session it seems to dissappear.
+
+~ check on Rex consortium table of when rot better, what studies were they ? ag or pub health
 
@@
 ~ we won't run sensitivity analysis on those runs, but just to check that rotations and sequence runs still similar
-~ ian suggested talking to Simon if we need to run something for longer
 
@@
 df_res_all_rot_cost10000_mut_sel_bal.rda
-done 2. check ability to do runs where insecticides differ
-
-~ think - are all rotation runs so far based on insecticides with equal effectiveness, resistance restoration, cost etc.
-?
-What might happen if we allow differences between insecticides ?
-Can we easily represent that ?
-I could possibly run one set of simulations where insecticides are allowed to be different.
-I think that could be set up by changing :
-linputs$expo_hi <-  runif(1, min=0.4, max=0.9)
-to:
-runif(n_insecticides, min=0.4, max=0.9)
-This might break stuff.
-Can just check that prop runs in which rotations are better doesn't increase.
-I suspect input params may be saved to outfiles as comma delimited strings ?
 
@@
 ~~ Now I can possibly change RStudio to point to the repo on my Github (do when I have a bit more time)
-~~ update R
-
-These are not about insecticide resistance (they are about resistance to the gene itself but may have useful pointers)
-Alphey2011. Modeling resistance to genetic control of insects
-https://doi.org/10.1016/j.jtbi.2010.11.016
-
-Watkinson-Powell2017 Resistance to genetic insect control: Modelling the effects of space
-https://doi.org/10.1016/j.jtbi.2016.10.014
 
@@
 
-~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ?
+~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ? I don't think so because it requires linkage.
 

### 2018-12-06 inst/documents/sensi_an_rotations1.Rmd
comments

@@
     linputs$dom_cos <-        runif(1, min=0,   max=1) 
+    #TODO make this lognormal to get lower frequencies too
     linputs$start_freqs <-    runif(1, min=0.005,   max=0.1) #note not log yet   
@@
 
-    # rbind this scenario inputs & results onto previous
-    # not a good speed or memory way of doing ... but a start
+    # previously I used rbind this scenario inputs & results onto previous
+    # not a good speed or memory way of doing
     #df_res_all <- rbind(df_res_all, df_in_out)
 
-    #instead set up the dataframe to the required size in scenario1
+    # now replaced with this better way
+    # set a dataframe for the outputs to the required size in scenario1
     if (i==1)

### 2018-12-06 inst/documents/rotations_notes.Rmd
tweaks

@@
 
+fix :
+Warning: Error in raf_get: object 'RAF' not found
+run_rot() can create it but when I add breakpoints it disappears
+
+where is raf_get called ?
+insecticide_check()
+rot_migrate()
+
+I suspect problem is here in insecticide_check()    
+#end of rotation reached
+    if (gens_this_insecticide >= rot_interval) 
+      {
+      # 31/7/18 I could add a check in here for case
+      # where only this insecticide remains below resistance threshold
+      # in which case don't want to change
+      # if I reset gens-this-insecticide then it will be used for another rot_interval
+      #I can use either of these to get the freqs for all other insecticides
+      # RAF[-current_insecticide, 'f','intervention',1]
+
+e.g.
+raf_get(a)
+Error in raf_get(a) : object 'a' not found
+
 For rotations send Jim a scaled down version of the rotations sensitivity analysis code. That just runs the sensitivity analysis and Jim will have a go at parallelising it.

### 2019-02-06 inst/documents/rotations_notes.Rmd
document

@@
 
+For rotations sent Jim a scaled down version of the rotations sensitivity analysis code. That just runs the sensitivity analysis and Jim will have a go at parallelising it.
+
 6/12/18
@@
 
+6/2/19
+
+before meeting with Jim to look at rotations code
+
+previous error in raf_get() seems OK now
+
 # TODO
+
+# todo: we want to look at changing insecticides according to mortality rather than resistance frequency
+
+# Jim suggest
 
@@
 
-For rotations send Jim a scaled down version of the rotations sensitivity analysis code. That just runs the sensitivity analysis and Jim will have a go at parallelising it.
-
-Still occasionally getting error from shinyUI resistrot :
-Warning: Error in raf_get: object 'RAF' not found
-
 It is only from scenarioA.
@@
 ~ talk to Ian about Hackett2016 paper and whether we can use similarly simple model
-
-~ ask Ian to check through the rotations algebra when he writes up the SI.
 
@@
 
-Combining the high-dose/refuge strategy and self-limiting transgenic
-insects in resistance management- a test in experimental mesocosms 
-Zhou2017
-
 submit links to UIs and paper once it is published to :

### 2019-02-06 R/run_rot.r

@@ -296,7 +296,7 @@ run_rot <- function(max_gen = 200,
     if (length(start_freqs) == 1) start_freqs <- rep(start_freqs, n_insecticides)
     
     
-    #todo implement a no_r_below_mut option here
+    #options to stop resistance frequency going below start level or the mutation-selection balance
     if ( no_r_below_start | no_r_below_mut )
     {
       for(insecticide in 1:n_insecticides)

### 2019-02-13 inst/documents/rotations_notes.Rmd
comments
starting mortality calculations

@@
 
+13/2/2019
+
+looking at saving tested mortality levels for the different insecticides over time
+(as a prelude to using it as a criteria for switching insecticides)
+
+resistance frequency results are saved in df_results
+think I can create df_mort with the same structure and then put it through the same plotting function.
+
+but df_results summaries are only calculated at the end so df_mortali couldn't allow mortality to be used as a switching criteria.
+
+BUT also I need an equivalent of RAF[] (without the sex dimension, or maybe easier to leave it in)
+to enable switching
+
+  # RAF[, 'm', 'intervention', 1] = freqs
+  # dim1 :n_insecticides
+  # dim2 :sex: m,f
+  # dim3 :site: intervention,refuge
+  # dim4 : generations
+
+see run_rot for initial test mortality calculation
+
 # TODO
@@
 
-# Jim suggest
+# Jim suggests making flow chart
+# make branch of rotations repo and send instructions to Jim
+# get rotations repo back from Ian
 

### 2019-02-13 R/insecticide_check.r

@@ -37,10 +37,10 @@ insecticide_check <- function( RAF1gen,
     
     # TODO this is where to add mortality-based change criteria
     # how to calculate mortality from allele frequency and effectiveness etc. ?
-    # I think I want to calculate freq of RR,SR,SS (from HW I assume)
-    # then calculate expected proportion mortality for each (for SR will need dominance)
-    # freqRR = func(RAF)
-    # mortSS = effectiveness*
+    # Ian defines in the rotations SI
+    # calculate freq of RR,SR,SS (from HW)
+    # multiply each by fitness in a_fitloc[current_insecticide, 'SS', 'hi']
+    # calculated in fitness_single_locus()
     
     
     if ( RAF1gen[current_insecticide, 'f','intervention',1] > rot_criterion &

### 2019-02-13 R/run_rot.r

@@ -98,6 +98,11 @@ run_rot <- function(max_gen = 200,
                                          stringsAsFactors = FALSE,
                                          l_gene_plus_activity))  
   
+  # 13/2/2019 adding a df (maybe temporary) to record tested mortality of females
+  # as a prelude to using mortality as an alternative criteria to switch between insecticides
+  df_mortali <- df_results
+  
+  
   # 11/5/18 adding an df to store for each insecticide the last time it was used
   df_ins <- data.frame(last_used=rep(Inf, n_insecticides))
   # set value for this to 0 each time an insecticide is in use
@@ -275,6 +280,26 @@ run_rot <- function(max_gen = 200,
         #if (diagnostics) message(sprintf("generation %d: completed selection against locus %d in refugia\n", gen, insecticide))
       
       } # end if coverage < 1
+      
+      
+      # 13/2/2019 put something here to calculate mortality
+      # maybe can just pass RAF and fitness to a function ?
+      # can I do for all insecticides and sites at once ?
+      # this does for m & f, even though only probably need for f
+      # TODO make this dryer and more efficient
+      # freqs are calculated by HardyWeinberg
+      # currently in first example mortalities always below 0.785 ?
+      freqRR <- RAF[insecticide,,,gen-1]^2
+      mortRR <- freqRR * (1 - fitness[insecticide, 'RR', 'hi'])
+      freqSR <- 2 * (RAF[insecticide,,,gen-1] * 1-RAF[insecticide,,,gen-1])
+      mortSR <- freqSR * (1 - fitness[insecticide, 'RS', 'hi'])
+      freqSS <- (1-RAF[insecticide,,,gen-1])^2
+      mortSS <- freqSS * (1 - fitness[insecticide, 'SS', 'hi'])      
+      # only for f in the intervention site
+      mort <- mortRR['f','intervention'] + mortSR['f','intervention'] + mortSS['f','intervention']
+      #cat(mort," ")
+      # end temp new bit
+      
      } #end of cycling insecticides
     
     
@@ -388,6 +413,7 @@ run_rot <- function(max_gen = 200,
  
     # recording the insecticide that's going to be used in next timestep
     df_results$insecticide[gen] <- current_insecticide    
+    df_mortali$insecticide[gen] <- current_insecticide        
   
    } #### end of max_gen loop
 
@@ -410,7 +436,6 @@ run_rot <- function(max_gen = 200,
[…]

### 2019-02-20 inst/documents/rotations_notes.Rmd
in prog mortality calculations

@@
 
+Removed these comments from todo because think we have moved on :
+1. check out these runs that have the mut-sel balance option enabled
+df_res_all_rot_cost10000_mut_sel_bal.rda
+
+~ check on the runs with the freq limit removed :
+df_res_all_rot_cost10000_no_freq_limit.rda
+
+~ think about whether I can implement allele frequencies just being constrained to their starting values the first time that each insecticide is used.
+
+14/2/2019
+
+starting to get mortality recording working
+
+currently looking at the df_mortali array from within a stop put inside run_rot()
+
+with current starting resistance frequencies and effectiveness the mortalities start below the threshold where we would stop using the insecticide (90%)
+
+we can set starting frequencies lower and effectiveness higher e.g. 
+run_rot(eff=1, start_freqs=0.001)
+
+# moved rotations repo back from Ian
+https://help.github.com/articles/about-repository-transfers/
+Transfer is at the bottom of settings, options : from Ians account
+
+#20/2/2019
+
+watched Ellies talk
+~ she particularly talked about the benefit of combining complementary interventions in future (e.g. community larvel source management, ATSBs, etc.)
+~ How these combine will depend on which part of the transmission cycle they target, e.g. outdoor biters
+~ she showed quite nice cartoons of how untreated nets, treated nets and IRS can reduce tranmsission in different ways
+~ made me think there may be potential to develop a clearer visualisation of the transmission cycle and how interventions can work, similar to stuff I did with Gerry
+~ other interesting thing that Hilary asked about is how the different insecticides can modify both repellance and killing and how this may change due to resistance.
+Maybe similar to converstaions we had with Francis Rowland.
+~ but I don't really have the inclination to move in the modelling direction but may be options to keep some interests going parallel with it.
+
+~ more on how to record and display mortality
+
+currently resistance frequencies : 
+stored in an array (RAF), 
+saved to a dataframe (df_res) 
+converted to a diff format (df_res2) 
+plotted with rot_plot_resistance()
+
+for mortality :
+store in (a_mort)
+saved to df_mortali
+...
+
+Alternative Options
+A. convert df_mortali to df_mort2 so that can use rot_plot_resistance()
+
+B. save resistance and mortality in same structure, maybe able to add mortality on as a column to df_res2
+
+Tried the above, seems to work.
+
+This shows there is a very consistent relationship between mortality & resistance.
+plot(df_res2$mortality ~ df_res2$resistance)
+
+So maybe we don't even need to save mortality ???
+
+this showed not quite linear
+plot(lm(df_res2$mortality ~ df_res2$resistance))
+
+So can we just calculate a conversion from resistance to mortality that will allow us to estimate the resistance frequency that corresponds to 90% mortality ??
+
+I can probably just calculate the conversion from the code I've written recently ?
+
+Does this mean that the frequency differences of the different genotypes are not directly important ?
+
+see genotype_freq() and mort_from_resist()
+
+sent msg to Ian
+
 # TODO
 
+~ recording mortality - is now the time to try and record data in a 'tidier' way ?
+
+~ do I need to be able to plot mortality over time ?
+~~ on the same graph as resistance frequency or a different one ?
+
 # todo: we want to look at changing insecticides according to mortality rather than resistance frequency
@@
 # Jim suggests making flow chart
-# make branch of rotations repo and send instructions to Jim
-# get rotations repo back from Ian
 
@@
 
-1. check out these runs that have the mut-sel balance option enabled
-df_res_all_rot_cost10000_mut_sel_bal.rda
-
-~ check on the runs with the freq limit removed :
-df_res_all_rot_cost10000_no_freq_limit.rda
-
-~ think about whether I can implement allele frequencies just being constrained to their starting values the first time that each insecticide is used.
-
 ~ talk to Ian about Hackett2016 paper and whether we can use similarly simple model
@@
 ~ how to term success. 'At least one insecticide still below resistance thresholds'
-
-Looking at moving rotations repo back to me. Seems transfer may be best way :
-https://help.github.com/articles/about-repository-transfers/
-Transfer is at the bottom of settings, options : from Ians account
 

### 2019-02-20 R/genotype_freq.r

@@ -1,0 +1,28 @@
+#' calculate genotype frequencies from allele frequencies
+#' 
+#' PROVISIONAL
+
+#' @param resist_freq resistance allele frequency
+#' 
+#' @examples 
+#a_genfreq <- genotype_freq(0.001)
+
+#' @return fitness values
+#' @export
+
+genotype_freq <- function ( resist_freq = 0.01)
+{
+  
+  
+  a_genfreq<- array_named(genotype=c('SS','RS', 'RR'))
+
+  a_genfreq['SS'] <- (1 - resist_freq) ^ 2
+  
+  a_genfreq['RS'] <- 2 * resist_freq * (1 - resist_freq)
+  
+  a_genfreq['RR'] <- resist_freq ^ 2
+  
+  
+  return(a_genfreq)
+}
+

### 2019-02-20 R/mort_from_resist.r

@@ -1,0 +1,53 @@
+#' calculate mortality from resistance
+#' 
+#' PROVISIONAL
+
+#' @param eff effectiveness, for all insecticides or individually
+#' @param dom_sel dominance of selection, for all insecticides or individually
+#' @param dom_cos dominance of cost, for all insecticides or individually
+#' @param rr resistance restoration, for all insecticides or individually 
+#' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
+#' @param fitSS fitness of SS if no insecticide, for all insecticides or individually
+#' 
+#' 
+#' @examples 
+
+#' @return fitness values
+#' @export
+
+mort_from_resist <- function ( resist_freq = 0.01,
+                                   eff = 0.9,
+                                   dom_sel = 0.5,
+                                   dom_cos = 0.5,
+                                   rr = 0.5,
+                                   cost = 0,
+                                   fitSS = 1,
+                                   plot = FALSE)
+{
+  
+  a_genfreq <- genotype_freq(resist_freq)
+  #just to set up array
+  a_mort <- a_genfreq
+  
+  sel <- rr * eff #selection coeff is resistance restoration * effectiveness
+  
+  #TODO check this
+  #TODO have an intermediate step that returns the mortality of each genotype
+  a_mort['RR'] <- (eff-sel)
+  a_mort['RS'] <- (eff-(sel*dom_sel))
+  a_mort['SS'] <- eff
+  
+  # a_mort['RR'] <- a_genfreq['RR'] * (eff-sel)
+  # a_mort['RS'] <- a_genfreq['RS'] * (eff-(sel*dom_sel))
+  # a_mort['SS'] <- a_genfreq['SS'] * eff
+  
+  return(a_mort)
+
+  #TODO implement the below
+  #proportional mortality
+  #a_propmort <- a_mort * a_genfreq   
+  #mort <- sum(a_propmort)
+  
+  #return(mort)
+}
+

### 2019-02-21 R/mort_by_genotype.R
starting to get mortality in plots

@@ -1,0 +1,41 @@
+#' calculate mortality by genotype from resistance freq
+#' 
+#' PROVISIONAL
+
+#' @param eff effectiveness, for all insecticides or individually
+#' @param dom_sel dominance of selection, for all insecticides or individually
+#' @param dom_cos dominance of cost, for all insecticides or individually
+#' @param rr resistance restoration, for all insecticides or individually 
+#' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
+#' @param fitSS fitness of SS if no insecticide, for all insecticides or individually
+#' 
+#' 
+#' @examples 
+
+#' @return fitness values
+#' @export
+
+mort_by_genotype <- function ( resist_freq = 0.01,
+                               eff = 0.9,
+                               dom_sel = 0.5,
+                               dom_cos = 0.5,
+                               rr = 0.5,
+                               cost = 0,
+                               fitSS = 1,
+                               plot = FALSE)
+{
+  
+  #just to set up array
+  a_mort <- genotype_freq(0)
+  
+  sel <- rr * eff #selection coeff is resistance restoration * effectiveness
+  
+  a_mort['RR'] <- (eff-sel)
+  a_mort['RS'] <- (eff-(sel*dom_sel))
+  a_mort['SS'] <- eff
+  
+  
+  return(a_mort)
+  
+}
+

### 2019-02-26 inst/documents/rotations_notes.Rmd
starting to get mortality in plots
insecticide change using mortality now working - but may not for diff insecticides

@@
 
+# 21/2/2019
+
+Cool getting there with mortality checking :
+Best to set logy to F to be able to see 90% mort
+
+dfr <- run_rot(start_freqs = 0.0001, eff=1, logy=FALSE)
+
+# 26/2/2019
+
+starting to get mortality switch working.
+
+seems OK from messages but plots don't look right
+
+in the plots looks like still responding to frequency rather than mortality ??
+
+Something a bit weird going on ...
+Currently the threshold mortality is not reached in the plots, but seems that frequency is.
+
+Fixed!!! mostly a problem that eff,dom_sel,rr was not passed to new calc functions (and I didn't notice because they had defaults).
+
+dfr <- run_rot(start_freqs = 0.000000001, eff=1, logy=FALSE, mort_or_freq='mort', rot_interval=0, threshold=0.9)
+
 # TODO
+
+~ fix that mortality calc in plots and that for switching is currently repeated
+~ check it works with diff insecticides i.e. eff=c(0.7,0.8,0.9) I suspect not because mort_from_resist() or something else not properly vectorised.
+
+whatever I do I need to make mort/survival thing more transparent
+
+options
+~ pass as mortality and then convert to survival for calculation
+~ pass as survival convert to mortality for plot
+
+~ change origin of git repo in RStudio, probably need to do from git
 
@@
 ~~ on the same graph as resistance frequency or a different one ?
-
-# todo: we want to look at changing insecticides according to mortality rather than resistance frequency
 
@@
 ~ maybe come back to i_in_use ribbons not meeting issue
-~ rename rot_criterion to rot_threshold
 ~ add in some unit-tests ?

### 2019-02-26 inst/documents/rotations_story.Rmd
insecticide change using mortality now working - but may not for diff insecticides

@@
 
-dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 0, rot_criterion = 0.9, start_freqs = 0.001)
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 0, threshold = 0.9, start_freqs = 0.001)
 
@@
 
-dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 0, rot_criterion = 0.9, start_freqs = 0.001)
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 0, threshold = 0.9, start_freqs = 0.001)
 
@@
 
-dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 1, rot_criterion = 0.9, start_freqs = 0.001)
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=0, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 1, threshold = 0.9, start_freqs = 0.001)
 
@@
 
-dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 1, rot_criterion = 0.9, start_freqs = 0.001)
+dfr <- run_rot(n_insecticides=2, max_gen = 200, rot_interval=10, expo_hi=0.5, eff=0.8, cost=0.005, coverage=1, migration = 0, dom_cos = 1, threshold = 0.9, start_freqs = 0.001)
 

### 2019-02-26 R/genotype_freq.r

@@ -2,7 +2,7 @@
 #' 
 #' PROVISIONAL
 
-#' @param resist_freq resistance allele frequency
+#' @param rfreq resistance allele frequency
 #' 
 #' @examples 
 #a_genfreq <- genotype_freq(0.001)
@@ -10,17 +10,17 @@
 #' @return fitness values
 #' @export
 
-genotype_freq <- function ( resist_freq = 0.01)
+genotype_freq <- function ( rfreq )
 {
   
   
   a_genfreq<- array_named(genotype=c('SS','RS', 'RR'))
 
-  a_genfreq['SS'] <- (1 - resist_freq) ^ 2
+  a_genfreq['SS'] <- (1 - rfreq) ^ 2
   
-  a_genfreq['RS'] <- 2 * resist_freq * (1 - resist_freq)
+  a_genfreq['RS'] <- 2 * rfreq * (1 - rfreq)
   
-  a_genfreq['RR'] <- resist_freq ^ 2
+  a_genfreq['RR'] <- rfreq ^ 2
   
   
   return(a_genfreq)

### 2019-02-26 R/gens_under_thresh.r

@@ -4,7 +4,9 @@
 #' threshold while that insecticide is deployed.
 #'
 #' @param dfres dataframe of resistance results from run_rot()
-#' @param rot_criterion resistant allele frequency threshold
+#' @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, also precludes switch to an insecticide.
+#' @param mort_or_freq whether threshold for insecticide change is mortality 'mort' or resistance frequency 'freq'
+#' 
 # @import ggplot2
 #' @importFrom rlang .data
 #' @return integer num generations
@@ -16,23 +18,35 @@
 #' 
 
 gens_under_thresh <- function(dfres,
-                              rot_criterion = 0.5)
+                              threshold = 0.5,
+                              mort_or_freq = 'freq')
 {
 
 
 res <- dfres %>%
   # only assess in control areas not in refugia  
   dplyr::filter(.data$active_or_refuge=='active') %>%
-  group_by(.data$resist_gene) %>%
+  group_by(.data$resist_gene)
   # for all insecticides in all generations  
   # summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
   # summarise(mean_gens_under50 = mean(gens_under50)) %>%
   # just for deployed insecticides 
-  summarise(gens_dep_under50 = sum(.data$resistance < rot_criterion &
-                                     #finds insecticide in use = this one
-                                     .data$resist_gene==paste0('insecticide',.data$insecticide), na.rm=TRUE)) %>%
-  summarise(tot_gens_dep_under50 = sum(.data$gens_dep_under50)) %>%    
-  unlist()
+  ## TODO check change this to 1-.data$mortality for when freq_or_mort == 'mort 
+if ( mort_or_freq == 'freq' )
+  res <- summarise(res, gens_dep_under50 = sum(.data$resistance < threshold &
+                                   #finds insecticide in use = this one
+                                   .data$resist_gene==paste0('insecticide',.data$insecticide), na.rm=TRUE))
+
+else if ( mort_or_freq == 'mort' )
+{
+  res <- summarise(res, gens_dep_under50 = sum(.data$mortality < threshold &
+                                                 #finds insecticide in use = this one
+                                                 .data$resist_gene==paste0('insecticide',.data$insecticide), na.rm=TRUE))
+}
+        
+    
+summarise(res, tot_gens_dep_under50 = sum(.data$gens_dep_under50)) %>%    
+unlist()
 
 return(res)
 

### 2019-02-26 R/insecticide_check.r

@@ -5,10 +5,16 @@
 #' @param RAF1gen single generation array of resistance allele frequencies
 #' @param current_insecticide id number of current insecticide
 #' @param rot_interval frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
-#' @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
+#' @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, also precludes switch to an insecticide.
+#' @param mort_or_freq whether threshold for insecticide change is mortality 'mort' or resistance frequency 'freq'
 #' @param gens_this_insecticide num generations this insecticide has been used for
 #' @param min_rwr_interval minimum rotate-when-resistant interval to stop short switches, only used when rot_interval==0. set to 0 to have no effect.
-#' @param exit_rot whether to exit rotation interval if rot_criterion is reached
+#' @param exit_rot whether to exit rotation interval if threshold is reached
+# BEWARE these can be vectors and this func prob won't cope
+#' @param eff effectiveness propn. SS killed by insecticide, for all insecticides or individually
+#' @param dom_sel dominance of selection, for all insecticides or individually
+#' @param dom_cos dominance of cost, for all insecticides or individually
+#' @param rr resistance restoration, for all insecticides or individually 
 
 # @examples 
 #' 
@@ -19,12 +25,19 @@
 insecticide_check <- function( RAF1gen,
                                current_insecticide,
                                rot_interval, 
-                               rot_criterion,
+                               threshold,
+                               mort_or_freq,
                                gens_this_insecticide,
                                min_rwr_interval,
-                               exit_rot
+                               exit_rot,
+                               eff,
+                               dom_sel,
+                               rr
                               ) 
 {
+  #check that required value
+  stopifnot( mort_or_freq == 'mort' | mort_or_freq == 'freq')
+  
   change_insecticide <- FALSE
   
   # if rotate-when resistant OR exit_rot==TRUE so rotation can be exited
@@ -35,15 +48,31 @@ insecticide_check <- function( RAF1gen,
     # because have drop=false to preserve site dimension the gen dimension is preserved too
[…]

### 2019-02-26 R/insecticide_switch.r

@@ -5,15 +5,18 @@
 #' @param RAF array of resistance allele frequencies
 #' @param current_insecticide id number of current insecticide
 #' @param n_insecticides number of insecticides (and hence loci)
-#' @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
+#' @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, also precludes switch to an insecticide.
+#' @param mort_or_freq whether threshold for insecticide change is mortality 'mort' or resistance frequency 'freq'
 #' @param gen generation number
 #' @param min_gens_switch_back minimum num gens before can switch back to an insecticide
 #' @param df_ins number of generations since each insecticide used
 #' @param df_results results of sim so far
 #' @param diagnostics whether to output running info
 # @param rot_interval frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
-# @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
 # @param rot_count num generations this insecticide has been used for
+#' @param eff effectiveness, for all insecticides or individually
+#' @param dom_sel dominance of selection, for all insecticides or individually
+#' @param rr resistance restoration, for all insecticides or individually 
 #' 
 # @examples 
 #' 
@@ -24,12 +27,16 @@
 insecticide_switch <- function( RAF,
                                 current_insecticide,
                                 n_insecticides,
-                                rot_criterion,
+                                threshold,
+                                mort_or_freq,
                                 gen,
                                 min_gens_switch_back,
                                 df_ins,
                                 df_results,
-                                diagnostics)
+                                diagnostics,
+                                eff,
+                                dom_sel,
+                                rr )
 {
   next_insecticide_found <- FALSE
   candidate <- current_insecticide 
@@ -39,8 +46,23 @@ insecticide_switch <- function( RAF,
     #search through insecticides and go back to start if reach end
     candidate <- ifelse(candidate==n_insecticides, yes=1, no=candidate+1)
     
-    #11/5/18 adding condition of not going back to recently used insecticide
-    if (RAF[candidate, 'f','intervention', gen] < rot_criterion & 
+    # 26/2/19 allow mortality-based change criteria
+    
+    # set from resistance frequency first
+    check_value <- RAF[candidate, 'f','intervention',gen] 
+    
[…]

### 2019-02-26 R/mort_by_genotype.R

@@ -4,10 +4,10 @@
 
 #' @param eff effectiveness, for all insecticides or individually
 #' @param dom_sel dominance of selection, for all insecticides or individually
-#' @param dom_cos dominance of cost, for all insecticides or individually
+# @param dom_cos dominance of cost, for all insecticides or individually
 #' @param rr resistance restoration, for all insecticides or individually 
-#' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
-#' @param fitSS fitness of SS if no insecticide, for all insecticides or individually
+# @param cost fitness cost of RR in no insecticide, for all insecticides or individually
+# @param fitSS fitness of SS if no insecticide, for all insecticides or individually
 #' 
 #' 
 #' @examples 
@@ -15,14 +15,12 @@
 #' @return fitness values
 #' @export
 
-mort_by_genotype <- function ( resist_freq = 0.01,
-                               eff = 0.9,
-                               dom_sel = 0.5,
-                               dom_cos = 0.5,
-                               rr = 0.5,
-                               cost = 0,
-                               fitSS = 1,
-                               plot = FALSE)
+mort_by_genotype <- function ( eff,# = 0.9,
+                               dom_sel,# = 0.5,
+                               rr )# = 0.5)
+                               #cost = 0,
+                               #fitSS = 1,
+                               #plot = FALSE)
 {
   
   #just to set up array

### 2019-02-26 R/mort_from_resist.r

@@ -1,33 +1,31 @@
 #' calculate mortality from resistance
 #' 
-#' PROVISIONAL
-
+#' @param rfreq resistance allele frequency
 #' @param eff effectiveness, for all insecticides or individually
 #' @param dom_sel dominance of selection, for all insecticides or individually
-#' @param dom_cos dominance of cost, for all insecticides or individually
+# @param dom_cos dominance of cost, for all insecticides or individually
 #' @param rr resistance restoration, for all insecticides or individually 
-#' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
-#' @param fitSS fitness of SS if no insecticide, for all insecticides or individually
+# @param cost fitness cost of RR in no insecticide, for all insecticides or individually
+# @param fitSS fitness of SS if no insecticide, for all insecticides or individually
 #' 
 #' 
 #' @examples 
 
-#' @return total mortality
+#' @return total mortality single value
 #' @export
 
-mort_from_resist <- function ( resist_freq = 0.01,
-                                   eff = 0.9,
-                                   dom_sel = 0.5,
-                                   dom_cos = 0.5,
-                                   rr = 0.5,
-                                   cost = 0,
-                                   fitSS = 1)
+mort_from_resist <- function ( rfreq, # = 0.01,
+                                   eff,# = 0.9,
+                                   dom_sel,# = 0.5,
+                                   rr )# = 0.5,
+                                   #cost = 0,
+                                   #fitSS = 1)
 {
   
-  a_genfreq <- genotype_freq(resist_freq)
+  a_genfreq <- genotype_freq(rfreq)
   
   #formals() passes all args from this function
-  a_mort <- mort_by_genotype(formals())
+  a_mort <- mort_by_genotype(eff=eff, dom_sel=dom_sel, rr=rr)
   
   #proportional mortality
   a_propmort <- a_mort * a_genfreq   

### 2019-02-26 R/rot_plots.r

@@ -2,14 +2,15 @@
 #'
 #' @param df_res2 dataframe of resistance results from run_rot()
 #' @param plot_refuge whether to plot refuge as well as intervention
-#' @param rot_criterion resistant allele frequency threshold to add to plots, set to NULL to not add
+#' @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, also precludes switch to an insecticide.
+#' @param mort_or_freq whether threshold for insecticide change is mortality 'mort' or resistance frequency 'freq'
 #' @param logy whether to use log scale for y axis
 #' @param add_gens_under50 whether to add a label of num generations under 50percent resistance
 #' @param df_resanother exploratory option to plot results of another scenario on the same graph
 #' @param lwd line thickness for resistance curves
 #' @param title optional title for plot, NULL for no title
 #' @param plot whether to plot results
-#' @param plot_mortality whether to add mortality to plots
+#' @param plot_mort whether to add mortality to plots
 #'
 # check said that namespace dependencies not required
 # @import ggplot2
@@ -25,14 +26,15 @@
 #' 
 rot_plot_resistance <- function(df_res2,
                                 plot_refuge = TRUE,
-                                rot_criterion = 0.5,
+                                threshold = 0.5,
+                                mort_or_freq = 'mort',
                                 logy = TRUE,
                                 add_gens_under50 = TRUE,
                                 df_resanother = NULL,
                                 lwd = 1.5,
                                 title = NULL,
                                 plot = TRUE,
-                                plot_mortality = TRUE) {
+                                plot_mort = TRUE) {
   
   # column names of input dataframe
   # "generation"  "insecticide"     "resist_gene"  "active_or_refuge" "resistance"
@@ -67,13 +69,16 @@ rot_plot_resistance <- function(df_res2,
 
   # testing adding mortality to the plot
   # note mortality currently added in run_rot but actually we can calc from fixed conversion
-  if (plot_mortality) {
-    #gg <- ggplot( df_res2, aes_string(x='generation',y='mortality')) +
-    #  geom_line( alpha=0.5, lwd=lwd, colour='red3' )
+  if (plot_mort) {
+
+    # set y intercept for horiz line to threshold if mort, otherwise default to 0.9
[…]

### 2019-02-26 R/run_rot.r

@@ -5,15 +5,16 @@
 #' @param max_gen maximum number of mosquito generations to run the simulation
 #' @param n_insecticides number of insecticides (and hence loci)
 #' @param start_freqs starting frequencies of resistance either one per insecticide or same for all
-#' @param rot_interval frequency of rotation (in generations) NB if set to zero mean RwR i.e. rotate when resistant
-#' @param rot_criterion resistant allele frequency that triggers a RwR change or precludes a insecticide from being rotated in.
+#' @param rot_interval frequency of rotation (in generations) if set to zero means sequence, i.e. change when threshold reached
+#' @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, also precludes switch to an insecticide.
+#' @param mort_or_freq whether threshold for insecticide change is mortality 'mort' or resistance frequency 'freq'
 #' @param migration migration rate between treated & untreated areas 0-1. We assume that immigration=emigration.
 #' @param coverage proportion of mosquitoes that are covered by the intervention (and 1-C is the proportion of the population in the untreated refugia).
 #' @param start_insecticide which insecticide to start with
 #' @param expo_hi exposure to insecticide in hi niche, either single or vector of 1 per insecticide
 #' @param expo_lo exposure to insecticide in lo niche, either single or vector of 1 per insecticide
 #' @param male_expo_prop proportion tht males are exposed relative to f, default 1, likely to be <1 (could possibly be a vector per insecticide)
-#' @param eff effectiveness, for all insecticides or individually
+#' @param eff effectiveness propn. SS killed by insecticide, for all insecticides or individually
 #' @param dom_sel dominance of selection, for all insecticides or individually
 #' @param dom_cos dominance of cost, for all insecticides or individually
 #' @param rr resistance restoration, for all insecticides or individually 
@@ -22,7 +23,7 @@
 #' @param min_rwr_interval minimum rotate-when-resistant interval to stop short switches, only used when rot_interval==0. set to 0 to have no effect.
 #' @param no_r_below_start to stop resistance frequencies going below starting values TRUE or FALSE
 #' @param no_r_below_mut to stop resistance frequencies going below mutation-selection balance TRUE or FALSE
-#' @param exit_rot whether to exit rotation interval if rot_criterion is reached
+#' @param exit_rot whether to exit rotation interval if threshold is reached
 #' @param min_gens_switch_back minimum num gens before can switch back to an insecticide
 # @param df_ins number of generations since each insecticide used
 #' 
@@ -33,6 +34,7 @@
[…]

### 2019-03-07 inst/documents/rotations_notes.Rmd
visualising mortality threshold

@@
 
+1/3/19
+
+Had idea to try to plot regio of parameter space where mortality less than 90%.
+Not sure it is very useful, but did make a start.
+
+see plot_mort_thresh.r
+
+5/3/19
+~ added mort_or_freq switch to rotresist1/ui.r
+~ added threshold to UI
+
+7/3/19
+
+Using mortality switch for sequences, currently it switches back to other insecticides even when they are below threshold.
+
+Aha I think that is due to the min_rwr_interval. When it is set to 0 then the sim doesn't proceed.
+
+~ struggled to add legend for green mortality lines
+fixed now
+
 # TODO
+
+fix error
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 1e-04 , rot_interval = 10 , eff = 0.95 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.01 , logy = FALSE , no_r_below_start = FALSE , min_rwr_interval = 5 , threshold = 0.9 , mort_or_freq ='mort' )
+
+~ rot_interval > 0
+in insecticide_check()
+Error in genotype_freq(rfreq) : 
+  argument "rfreq" is missing, with no default
+
+~ publish to shinyapps and send link to Ian
+
+~ try to fix that min_rwr_interval seems to allow switching back to insecticides
+
+~ check that per-diem for Geneva made it to my bank
+
+dfr <- run_rot(start_freqs = 0.000000001, eff=1, logy=FALSE, mort_or_freq='mort', rot_interval=0, threshold=0.9)
 

### 2019-03-07 R/insecticide_check.r

@@ -111,16 +111,20 @@ insecticide_check <- function( RAF1gen,
       # then check is if value is < survival same as for frequency
       # ARG can't do this because mort_from_resist is not vectorised
       # sapply fixes it
-      if (mort_or_freq == 'mort') 
-        #other_ins_checks <- 1 - mort_from_resist(other_ins_checks)       
-        other_ins_checks <- 1 - sapply(other_ins_checks, mort_from_resist(eff=eff, dom_sel=dom_sel, rr=rr))
-      
-      # 31/7/18 new condition, only change if one to change to
-      if ( min(other_ins_checks) <= threshold )
+      #PROBLEM HERE giving arg 'rfreq' is missing error from UI
+      if (mort_or_freq == 'mort' & length(other_ins_checks)>0 ) 
       {
-        # time to rotate so need to identify the next insecticide in the rotation
-        change_insecticide <- TRUE         
+        #other_ins_checks <- 1 - mort_from_resist(other_ins_checks)       
+        other_ins_checks <- 1 - sapply(other_ins_checks, function(x) mort_from_resist(rfreq=x,eff=eff, dom_sel=dom_sel, rr=rr))
+        
+        # 31/7/18 new condition, only change if one to change to
+        if ( min(other_ins_checks) <= threshold )
+        {
+          # time to rotate so need to identify the next insecticide in the rotation
+          change_insecticide <- TRUE         
+        }        
       }
+
       
       # message(paste0("insecticide",current_insecticide, 
       #                " freq=",RAF1gen[current_insecticide, 'f','intervention',1],

### 2019-03-07 R/plot_mort_thresh.r

@@ -1,0 +1,51 @@
+#' plot_mort_thresh to try plotting the regions of param space where mortality > 90%
+
+
+plot_mort_thresh <- function( threshold = 0.9) {
+ 
+# create vectors of inputs  
+vrfreq <- 1/10^(1:10)
+veff <- seq(0.8,1,0.05)
+ 
+# dataframe with all combinations 
+df1 <- expand.grid(rfreq=vfreq,
+                   eff=veff)
+  
+# calc mortality
+# first go assume set dom and rr
+dom_sel <- 0.5
+rr <- 0.5
+
+#df1$mort <- mort_from_resist(df1$freq, eff=eff, dom_sel=dom_sel, rr=rr)
+#sapply(mort_from_resist(eff=veff,rfreq=vfreq,rr=1,dom_sel=1))
+
+#bodge start
+df1$mort <- NA
+for(i in 1:nrow(df1))
+{
+  df1$mort[i] <-mort_from_resist(rfreq=df1$rfreq[i], eff=df1$eff[i], dom_sel=dom_sel, rr=rr) 
+}
+
+# can change the scale so it just shows > 90% mort
+ggplot(df1, aes(eff, rfreq )) +
+  geom_tile(aes(fill = mort), color = "white") +
+  scale_fill_gradient(low = "white", high = "steelblue") +
+  scale_y_continuous(trans='log10') #, 
+                       #breaks=c(0.001,0.01,0.5,1),
+                       #labels=c('0.1%','1%','50%','100%'),
+  
+
+
+#df1$mort <- sapply(df1$freq, mort_from_resist(eff=eff, dom_sel=dom_sel, rr=rr))
+
+  # for(freq in vfreq)
+  # {
+  #   for(eff in veff)
+  #   {
+  #     cat("f:",freq, " eff:",eff,'\n')
+  #   }
+  # }
+  
+  
+   
+}

### 2019-03-07 R/run_rot.r

@@ -516,6 +516,7 @@ run_rot <- function(max_gen = 200,
   if (plot) rot_plot_resistance(df_res2, plot_refuge=plot_refuge, 
                                 logy=logy, add_gens_under50=add_gens_under50,
                                 threshold=threshold,
+                                mort_or_freq=mort_or_freq,
                                 plot_mort=plot_mort)
 
   

### 2019-03-07 inst/shiny/rotresist1/server.r

@@ -37,7 +37,10 @@ shinyServer(function(input, output) {
                 cost =              input$cost_A,
                 logy =              input$logy, #inefficient that it reruns when changing to log
                 no_r_below_start =  input$no_r_below_start,
-                min_rwr_interval =  input$min_rwr_interval )
+                min_rwr_interval =  input$min_rwr_interval,
+                threshold =         input$threshold,
+                mort_or_freq =      input$mort_or_freq
+                )
                 
         # a hack to output the inputs, so it can be run from the console
         cat("A:\n")
@@ -53,9 +56,11 @@ shinyServer(function(input, output) {
                       "coverage =",input$coverage_A,",",
                       "migration =",input$migration_A,",",
                       "cost =",input$cost_A,",",
-                      "logy =",input$logy,
-                      "no_r_below_start =",input$no_r_below_start,
-                      "min_rwr_interval =",input$min_rwr_interval,            
+                      "logy =",input$logy,",",
+                      "no_r_below_start =",input$no_r_below_start,",",
+                      "min_rwr_interval =",input$min_rwr_interval,",",
+                      "threshold =",input$threshold,",",
+                      "mort_or_freq ='",input$mort_or_freq,"'",
                       ")\n" )
         
                
@@ -85,7 +90,9 @@ shinyServer(function(input, output) {
                 cost =              input$cost_B,
                 logy =              input$logy, #inefficient that it reruns when changing to log
                 no_r_below_start =  input$no_r_below_start,
-                min_rwr_interval =  input$min_rwr_interval )
+                min_rwr_interval =  input$min_rwr_interval,
+                threshold =         input$threshold,
+                mort_or_freq =      input$mort_or_freq )
         
         
         # a hack to output the inputs, so it can be run from the console
@@ -102,9 +109,11 @@ shinyServer(function(input, output) {
             "coverage =",input$coverage_B,",",
             "migration =",input$migration_B,",",
             "cost =",input$cost_B,",",
-            "logy =",input$logy,
-            "no_r_below_start =",input$no_r_below_start,
-            "min_rwr_interval =",input$min_rwr_interval,
+            "logy =",input$logy,",",
+            "no_r_below_start =",input$no_r_below_start,",",
+            "min_rwr_interval =",input$min_rwr_interval,",",
+            "threshold =",input$threshold,",",
+            "mort_or_freq ='",input$mort_or_freq,"'",
             ")\n" )
         
       }) #end isolate  

### 2019-03-08 inst/documents/rotations_notes.Rmd
preparing shinyapps deployment with mortality switch

@@
 
+8/3/19
+eeek, found that there is a cran package called rotations, and this seems to be messing up deployment to shinyapps
+
 # TODO
 
-fix error
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 1e-04 , rot_interval = 10 , eff = 0.95 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.01 , logy = FALSE , no_r_below_start = FALSE , min_rwr_interval = 5 , threshold = 0.9 , mort_or_freq ='mort' )
-
-~ rot_interval > 0
-in insecticide_check()
-Error in genotype_freq(rfreq) : 
-  argument "rfreq" is missing, with no default
+why in this scenario does it rotate insecticides even though it starts below mortality threshold
+(to do with min_rwr threshold)
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 1e-04 , rot_interval = 0 , eff = 0.9 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.6 , migration = 0.01 , cost = 0.01 , logy = FALSE , no_r_below_start = FALSE , min_rwr_interval = 5 , threshold = 0.9 , mort_or_freq ='mort' )
+
+~ fix plot error when coverage==1
 

### 2019-03-08 inst/shiny/rotresist1/ui.r

@@ -180,7 +180,7 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
     column(1, h6("Advanced options :")),
     column(1, checkboxInput("no_r_below_start", "no_r_below_start", FALSE)), 
     column(1, NULL),
-    column(1, sliderInput("min_rwr_interval", "min rwr interval", val=5, min = 0, max = 20, step = 1, ticks=FALSE))
+    column(1, sliderInput("min_rwr_interval", "min rwr interval", val=0, min = 0, max = 20, step = 1, ticks=FALSE))
            
   ) #end fluid row
 

### 2019-03-12 inst/documents/rotations_notes.Rmd
notes

--- 2019-03-07 inst/documents/rotations_notes.Rmd
@@
 
+8/3/19
+eeek, found that there is a cran package called rotations, and this seems to be messing up deployment to shinyapps
+but was OK when I installed my rotations from github
+
+~ published to shinyapps and sent link to Ian
+
 # TODO
 
-fix error
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 1e-04 , rot_interval = 10 , eff = 0.95 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.01 , logy = FALSE , no_r_below_start = FALSE , min_rwr_interval = 5 , threshold = 0.9 , mort_or_freq ='mort' )
-
-~ rot_interval > 0
-in insecticide_check()
-Error in genotype_freq(rfreq) : 
-  argument "rfreq" is missing, with no default
-
-~ publish to shinyapps and send link to Ian
+~ change origin of git repo in RStudio, probably need to do from git
+
+~~ why in this scenario does it rotate insecticides even though it starts below mortality threshold
+(to do with min_rwr threshold)
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 1e-04 , rot_interval = 0 , eff = 0.9 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.6 , migration = 0.01 , cost = 0.01 , logy = FALSE , no_r_below_start = FALSE , min_rwr_interval = 5 , threshold = 0.9 , mort_or_freq ='mort' )
+
+~ fix plot error when coverage==1
 
@@
 ~ check it works with diff insecticides i.e. eff=c(0.7,0.8,0.9) I suspect not because mort_from_resist() or something else not properly vectorised.
+
+~ TIDY UP CODE AND IMPROVE DOCUMENTATION e.g. function args
 
@@
 
-~ change origin of git repo in RStudio, probably need to do from git
-
 ~ recording mortality - is now the time to try and record data in a 'tidier' way ?
-
-~ do I need to be able to plot mortality over time ?
-~~ on the same graph as resistance frequency or a different one ?
 
@@
 
-~ move github repo back to mine
-~~ From Github I merged changes from Ians with a pull request after comparing across the forks.
-~~ Now I can possibly change RStudio to point to the repo on my Github (do when I have a bit more time)
-
 submit links to UIs and paper once it is published to :

### 2019-03-12 inst/documents/rotations_notes.Rmd
notes

@@
 
+12/03/19
+~ change origin of git repo in RStudio, probably need to do from git
+remote: This repository moved. Please use the new location:        
+remote:   git@github.com:AndySouth/rotations.git    
+from the commandline :
+git remote rm origin
+git remote add origin git@github.com:AndySouth/rotations.git
+
 # TODO
-
-~ change origin of git repo in RStudio, probably need to do from git
 

### 2019-06-14 inst/documents/rotations_notes.Rmd
tidying

@@
 
-12/03/19
+#12/03/19
 ~ change origin of git repo in RStudio, probably need to do from git
@@
 
+#14/06/2019
+Testing before sending to Sam.
+Error in mortRR["f", "intervention"] : incorrect number of dimensions
+
+This is when coverage set to 1, because there is no intervention dimension in mortRR[] in run_tot()
+
+fixed
+
+then got this from the script because it is trying to cast a tibble with an insecticide column and gens to a number
+Error: (list) object cannot be coerced to type 'double'
+
+so I modified gens_under_thresh() to just return the numeric column
+
+BUT this doesn't solve and I don't know why this problem appeared when it wasn't there before ???
+
+gens_under_thresh() is giving one row per insecticide, is that what I'm expecting ??
+No it is supposed to be :
+'the sum across all insecticides for the number of generations below threshold while that insecticide is deployed'
+
+so why has it changed ?? BUG introduced by the mortality addition
+##14/6/19 corrected bug here by adding back res <-      
+res <- summarise(res, tot_gens_dep_under50 = sum(.data$gens_dep_under50)) %>%
+
 # TODO
+
+~ fix plot error when coverage==1
 
@@
 run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 1e-04 , rot_interval = 0 , eff = 0.9 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.6 , migration = 0.01 , cost = 0.01 , logy = FALSE , no_r_below_start = FALSE , min_rwr_interval = 5 , threshold = 0.9 , mort_or_freq ='mort' )
-
-~ fix plot error when coverage==1
 

### 2019-08-20 inst/documents/rotations_notes.Rmd
notes

@@
 
-# TODO
-
-~ fix plot error when coverage==1
+# 20/8/19
+checking this message below from my notes, problem is that simulation continues even when mortality starts below thresholds.
+Inputs are set for sequence, what it does is switch between the insecticides for the min_rwr_interval.
+If the min_rwr_interval is set greater the length of the use periods increases too.
+
+Probably I need to check that mortality not below threshold at start of the r_w_r interval ? In insecticide_check()
 
@@
 
-~ try to fix that min_rwr_interval seems to allow switching back to insecticides
-
-~ check that per-diem for Geneva made it to my bank
+# this is the part in insecticide_check() that causes problem
+    if ( check_value > threshold &
+         #20/8/19 also this allows an inseticide to start being used even when outside threshold
+         gens_this_insecticide > min_rwr_interval )
+    {
+      change_insecticide <- TRUE
+
+OR does it ? surely the problem is that insecticide_switch() is choosing an insecticide that is already outside threshold ?
+
+something weird going on in insecticide_switch() mortality threshold should be converted to survival for the comparison but doesn't seem to be :
+
+switch from insecticide 1 to 3; new mortality vs threshold = 0.100044 vs 0.900000
+
+AHA! NASTY BUG! there are multiple insecticides and it cycles through them because I do threshold <- 1-threshold that can give uncertain results.
+
+# TODO
+
+~ fix plot error when coverage==1
+
+STill if have min_rwr_interval > 0 an insecticide at very start of simulation outside the threshold will continue to be used.
+Mostly I think min_rwr_interval not a good idea, but does kind of represent what you might expect to happen in field when it is set to a year i.e.
+that once sprayed can't be unsprayed.
 
@@
 ~ pass as survival convert to mortality for plot
-
-~ recording mortality - is now the time to try and record data in a 'tidier' way ?
-
-# Jim suggests making flow chart
 
@@
 
-~ check on Rex consortium table of when rot better, what studies were they ? ag or pub health
-
 ~ check on 1000 runs with lower startfreqs and differing insecticides 
@@
 
-~ talk to Ian about Hackett2016 paper and whether we can use similarly simple model
-
 rotations paper :
@@
 
-Revisit Ians paragraph about the rare conditions that give large benefit to rotations.
-
-"Note however that this combination of recessive resistance and high dominance costs are exactly the features most desirable in any insecticide irrespective of how it is deployed.
-...
-it is very difficult to maintain insecticide concentration at the level required to make resistance recessive.
-Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant.""
-me is there anything similar on the dominance of costs ?
-
 I think takehome message maybe that rotations are much better than sequences only in rare circumstances.
@@
 
-In rotations and WoS MS we should talk about high dose/refuge strategy for transgenic crops (specifically Bt corn).
-The high dose part is that the dose is sufficiently high to kill all heterozygotes i.e.
-making resistance recessive.
-But if this was used across a large area the evolution of resistance would be very rapid - because only homozygous resistant individuals would be left.
-The refuge part is to provide a source of susceptible immigrants which can generate heterozygous offspring that are vulnerable to the dose.
-
-~ try changing scenario 2167 from df_cost
-~~ initial attempts in UI not working
-~~ could it be that it requires just right combination of params and that modifying them just slightly removes the advantage for rotations ?
-
 ~ can I create some simple figs to illustrate how an increasing rate of resistance evolution with frequency can lead to rotations being favoured
-~~ maybe just see how dom_cos changes shape of frequency plot
-~~ can I look at through UI ?
-
-~ save Ians comments on figY of
 
@@
 
-submit links to UIs and paper once it is published to :
-vectorlearningxchange.com
-
 ~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ? I don't think so because it requires linkage.
@@
 ~ one generation per month, 12 (or 10) generations per year, 500 gens is ~40-50 years
-
-# funding opportunities
-https://lstmed.sharepoint.com/research/RBPS/Pages/Funding%20Information/Funding-Opportunities.aspx     
-https://www.researchprofessional.com/funding/simple?q=software#1509451534514

### 2019-09-05 inst/documents/rotations_notes.Rmd
tidying

@@
 *13/7/18*
-
 ltst <- list(d=c(9,6),s=7)
@@
 Seems that with variable insecticides there are more runs where sequences are better (i.e. highly negative rot_minus_seq) ?
-
-See :
-df_res_all_rot1000-diffI-lowstart.rda
+See : df_res_all_rot1000-diffI-lowstart.rda
 
@@
 ~ added id to outputs in sensi_an_rotations to make things easier in future, e.g. to identify runs with particular behaviour.
-
 ~ also made more efficient by declaring size of results df at start of loop
@@
 
+# good refs for rotations paper
+resistance : malERA2017_updated-research-agenda-resistance-in-malaria.pdf
+
+"What are the benefits of insecticide rotations, mixtures, or spatial mosaics of different compounds?
+What is the impact of adding nonpyrethroid IRS where LLINs are already deployed at high coverage and quality?
+When should new insecticides be adopted?
+What is the ideal rotation period or mosaic configuration?
+How many insecticide classes are needed for effective rotation or mosaic strategies?
+Despite the absence of data to answer these questions, some countries have already developed operational frameworks for resistance management that could be adopted by other programmes.""
+
+The malERA Refresh Consultative Panel on Insecticide and Drug Resistance (2017) malERA: An updated research agenda for insecticide and drug resistance in malaria elimination and eradication.
+PLoS Med 14(11): e1002450.
+https://doi.org/10.1371/journal.pmed.1002450
+
+Slater2016. Use of an individual-based simulation model to explore and evaluate potential insecticide resistance management strategies
+
 *12/11/2018*
@@
 
-*13/11/2018*
+# 13/11/2018
 preparing slides for IVCC talk next week
@@
 
-23/11/18
-#to output script as pdf for Ian to print
+# 23/11/18
+to output script as pdf for Ian to print
+
 library(knitr)
@@
 
-4/12/18
-Met with Jim he offered to look at parallelising.
-
-5/12/18
-making scaled down version of sensi analysis for Jim
-
+# 5/12/18
 For rotations sent Jim a scaled down version of the rotations sensitivity analysis code. That just runs the sensitivity analysis and Jim will have a go at parallelising it.
 
-6/12/18
+# 6/12/18
 Check why git changes are thinking that whole file has changed. How did I fix that for rnaturalearth ??
@@
 
-6/2/19
-
+# 6/2/19
 before meeting with Jim to look at rotations code
-
 previous error in raf_get() seems OK now
 
-13/2/2019
-
+# 13/2/2019
 looking at saving tested mortality levels for the different insecticides over time
@@
 
-14/2/2019
-
+#14/2/2019
 starting to get mortality recording working
-
 currently looking at the df_mortali array from within a stop put inside run_rot()
@@
 #20/2/2019
-
 watched Ellies talk
@@
 # 21/2/2019
-
 Cool getting there with mortality checking :
@@
 # 26/2/2019
-
 starting to get mortality switch working.
@@
 
-1/3/19
-
+#1/3/19
 Had idea to try to plot regio of parameter space where mortality less than 90%.
@@
 
-5/3/19
+#5/3/19
 ~ added mort_or_freq switch to rotresist1/ui.r
@@
 7/3/19
-
 Using mortality switch for sequences, currently it switches back to other insecticides even when they are below threshold.
@@
 
+# 4/9/19
+However I change the run workflow I should try to keep it so that I can do the runs on multiple cores.
+
+# 5/9/19
+
+refactoring inputs
+Currently the list linputs which has one row per run is created in sensi_an_rotations1.Rmd by hardcoded min-max limits
+
+new inputs plan
+Have a rotations function that accepts a csv file or defaults and uses that to create the linputs list, also it can save the csv. Closer to what resistance code did.
+
+How to name the different files ? what are similar names in openMalaria ?
+
+in_expt
+in_runs
+
 # TODO
+
+~ investigate getting runs to name the output file with the date 
+~ and create a similarly named metadata file that records run input params 
+~ think about moving run control from a rmd to a .R script
+~ think about being able to create named run inputs files (csv ?) that store the list of inputs (can be the same one that is saved as run metadata)
+
+# RUNS FOR PAPER September 2019
+# A identical insecticides, seq vs rot, resistance frequency switch, A1 no costs no dispersal, A2 costs no dispersal, A3 dispersal no costs, A4 costs & dispersal. Sensitivity analysis of inputs.
+# B different insecticides, as above, no sensitivity analysis.
+# C identical insecticides, mortality switch, reduced input range so mortality starts above 90%
+# D different insecticides, mortality switch
+
+Runs are combinations of :
+Strategy (2)       : sequence, rotation
+Insecticides (2)   : same, different
+Switch (2)         : resistance frequency, mortality
+Scenarios (4)      : no costs no dispersal, costs no dispersal, dispersal no costs, costs & dispersal
+Replicates (10000) :
+
+So number of runs 2*2*2*4*10000 = 320000
 
@@
 
-fix :
+I think this is fixed now :
 Warning: Error in raf_get: object 'RAF' not found
@@
 
-~ check on 1000 runs with lower startfreqs and differing insecticides 
-~ is something weird going on favouring rotations more when insecticides differ ?
-~ yes
-
 ~ ian interested in starting frequencies at 10e-4, can do that for the runs with different values per insecticide.
@@
 rotations paper :
-~ develop flow of story
 ~ think about what questions managers will want answered
@@
 
-# good refs for rotations paper
-resistance : malERA2017_updated-research-agenda-resistance-in-malaria.pdf
-
-"What are the benefits of insecticide rotations, mixtures, or spatial mosaics of different compounds?
-What is the impact of adding nonpyrethroid IRS where LLINs are already deployed at high coverage and quality?
-When should new insecticides be adopted?
-What is the ideal rotation period or mosaic configuration?
-How many insecticide classes are needed for effective rotation or mosaic strategies?
-Despite the absence of data to answer these questions, some countries have already developed operational frameworks for resistance management that could be adopted by other programmes.""
-
-The malERA Refresh Consultative Panel on Insecticide and Drug Resistance (2017) malERA: An updated research agenda for insecticide and drug resistance in malaria elimination and eradication.
-PLoS Med 14(11): e1002450.
-https://doi.org/10.1371/journal.pmed.1002450
-
-Slater2016. Use of an individual-based simulation model to explore and evaluate potential insecticide resistance management strategies
-
 # state of play

### 2019-09-05 inst/documents/sensi_an_rotations1.Rmd
tidying

@@
     linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
-    linputs$no_r_below_mut <- TRUE #BEWARE trying out new mutation-selection balance option
+    linputs$no_r_below_mut <- FALSE #BEWARE trying out new mutation-selection balance option
     #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
@@
 
-    # rbind this scenario inputs & results onto previous
-    # not a good speed or memory way of doing ... but a start
-    #df_res_all <- rbind(df_res_all, df_in_out)
-
     #instead set up the dataframe to the required size in scenario1
@@
 
-    #I could save partial outputs every 200 scenarios
-    #just in case doesn't make to end
+    #save partial outputs just in case doesn't make to end
     if (i%%1000 == 0)
@@
   #have to ignore the ones that aren't strings
-  #linputs2 <- lapply(linputs, function(x) as.numeric(unlist(strsplit(as.character(x),split=", "))))
 
@@
   #TESTING changing min_rwr_interval from 5
-  #if min_rwr_interval is set to 0 you get frequent switch backs to other insecticides 
-  #for 1 generation
-  #but when it's set to 5it pushes frequences above the resistance freq threshold
+  #if min_rwr_interval is set to 0 you get frequent switch backs to other insecticides for 1 generation
+  #but when it's set to 5 it pushes frequences above the resistance freq threshold
   linputs$min_rwr_interval <- 5
@@
 
-  #now run as sequence
+  # now run as sequence
   linputs$rot_interval <- 0

### 2019-09-11 inst/documents/rotations_notes.Rmd
fix docs

@@
 
-in_expt
-in_runs
+# 6/9/19
+
+GOOD getting there, now a csv file saved in extdata that specifies the experiment
+
+linputs_multi <- set_run_inputs()
+# to get all inputs for a single scenario
+linputs <- purrr::map(linputs_multi, 2)
+# run one scenario
+dfres <- do.call(run_rot, linputs)
+
+# 10/9/19
+got different insecticides working
+added get_one_in() to get inputs for one scenario from multi-senario file
+
+# 11/09/19
+
+This is the new workflow.
+
+# read an experiment file specifying ranges
+inex <- read_in_expt() #specify filename
+# create an object containing inputs for all scenarios
+linmulti <- set_run_inputs(inex=inex)
+# get inputs for a single scenario
+linputs <- get_one_in(linmulti, scen_num=2)
+# run one scenario
+dfres <- do.call(run_rot, linputs)
 
 # TODO
+
+~ add option for different insecticides to the experiment file
+
+~ add the random seed to the experiment file (and decide where it will be set)
 
@@
 ~ think about moving run control from a rmd to a .R script
-~ think about being able to create named run inputs files (csv ?) that store the list of inputs (can be the same one that is saved as run metadata)
 

### 2019-09-11 R/mort_from_resist.r

@@ -10,7 +10,8 @@
 #' 
 #' 
 #' @examples 
-
+#' mort_from_resist( rfreq=0.01, eff=0.9, dom_sel=0.5, rr=0.5 )
+#' 
 #' @return total mortality single value
 #' @export
 

### 2019-09-11 R/set_run_inputs.r

@@ -6,10 +6,11 @@
 #' @param inex object containing the experiment input values and ranges
 #' 
 #' @examples 
-#' linmulti <- set_run_inputs()
-#' # to get all inputs for a single scenario
-#' #linputs <- purrr::map(linmulti, 2)
-#' # this also includes conversions of strings for multi insecticide runs
+#' # read an experiment file specifying ranges
+#' inex <- read_in_expt() #specify filename
+#' # create an object containing inputs for all scenarios
+#' linmulti <- set_run_inputs(inex=inex)
+#' # get inputs for a single scenario
 #' linputs <- get_one_in(linmulti, scen_num=2)
 #' # run one scenario
 #' dfres <- do.call(run_rot, linputs)

### 2019-09-13 inst/documents/rotation_runs2019.Rmd
getting there with running multiple scenarios

---
title: "rotation runs 2019"
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(rotations)
library(ggplot2)
library(gridExtra)

outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}

# read an experiment file specifying ranges
inex <- read_in_expt("in_rotations_paper_2019_inssame_freq.csv", nscenarios=1)

#inex <- read_in_expt("_in_expt_rotations_default_one_scenario.csv", nscenarios=1)

# create an object containing inputs for all scenarios
linmulti <- set_run_inputs(inex=inex)
linmulti <- set_run_inputs()

# get inputs for a single scenario
linputs <- get_one_in(linmulti, scen_num=2)

# run one scenario
dfres <- do.call(run_rot, linputs)

```

### 2019-09-13 inst/documents/rotations_notes.Rmd
getting there with running multiple scenarios

@@
 
-#14/06/2019
-Testing before sending to Sam.
-Error in mortRR["f", "intervention"] : incorrect number of dimensions
-
-This is when coverage set to 1, because there is no intervention dimension in mortRR[] in run_tot()
-
-fixed
-
-then got this from the script because it is trying to cast a tibble with an insecticide column and gens to a number
-Error: (list) object cannot be coerced to type 'double'
-
-so I modified gens_under_thresh() to just return the numeric column
-
-BUT this doesn't solve and I don't know why this problem appeared when it wasn't there before ???
-
-gens_under_thresh() is giving one row per insecticide, is that what I'm expecting ??
-No it is supposed to be :
-'the sum across all insecticides for the number of generations below threshold while that insecticide is deployed'
-
-so why has it changed ?? BUG introduced by the mortality addition
-##14/6/19 corrected bug here by adding back res <-      
-res <- summarise(res, tot_gens_dep_under50 = sum(.data$gens_dep_under50)) %>%
-
-# 20/8/19
-checking this message below from my notes, problem is that simulation continues even when mortality starts below thresholds.
-Inputs are set for sequence, what it does is switch between the insecticides for the min_rwr_interval.
-If the min_rwr_interval is set greater the length of the use periods increases too.
-
-Probably I need to check that mortality not below threshold at start of the r_w_r interval ? In insecticide_check()
-
-~~ why in this scenario does it rotate insecticides even though it starts below mortality threshold
-(to do with min_rwr threshold)
-run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 1e-04 , rot_interval = 0 , eff = 0.9 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.6 , migration = 0.01 , cost = 0.01 , logy = FALSE , no_r_below_start = FALSE , min_rwr_interval = 5 , threshold = 0.9 , mort_or_freq ='mort' )
-
-# this is the part in insecticide_check() that causes problem
-    if ( check_value > threshold &
-         #20/8/19 also this allows an inseticide to start being used even when outside threshold
-         gens_this_insecticide > min_rwr_interval )
-    {
-      change_insecticide <- TRUE
-
-OR does it ? surely the problem is that insecticide_switch() is choosing an insecticide that is already outside threshold ?
-
-something weird going on in insecticide_switch() mortality threshold should be converted to survival for the comparison but doesn't seem to be :
-
-switch from insecticide 1 to 3; new mortality vs threshold = 0.100044 vs 0.900000
-
-AHA! NASTY BUG! there are multiple insecticides and it cycles through them because I do threshold <- 1-threshold that can give uncertain results.
-
-# 4/9/19
-However I change the run workflow I should try to keep it so that I can do the runs on multiple cores.
-
-# 5/9/19
-
-refactoring inputs
-Currently the list linputs which has one row per run is created in sensi_an_rotations1.Rmd by hardcoded min-max limits
-
-new inputs plan
-Have a rotations function that accepts a csv file or defaults and uses that to create the linputs list, also it can save the csv. Closer to what resistance code did.
-
-How to name the different files ? what are similar names in openMalaria ?
-
-# 6/9/19
-
-GOOD getting there, now a csv file saved in extdata that specifies the experiment
-
-linputs_multi <- set_run_inputs()
-# to get all inputs for a single scenario
-linputs <- purrr::map(linputs_multi, 2)
-# run one scenario
-dfres <- do.call(run_rot, linputs)
-
-# 10/9/19
-got different insecticides working
-added get_one_in() to get inputs for one scenario from multi-senario file
-
-# 11/09/19
-
-This is the new workflow.
-
-# read an experiment file specifying ranges
-inex <- read_in_expt() #specify filename
-# create an object containing inputs for all scenarios
-linmulti <- set_run_inputs(inex=inex)
-# get inputs for a single scenario
-linputs <- get_one_in(linmulti, scen_num=2)
-# run one scenario
-dfres <- do.call(run_rot, linputs)
-
-# TODO
-
-~ add option for different insecticides to the experiment file
-
-~ add the random seed to the experiment file (and decide where it will be set)
-
-~ investigate getting runs to name the output file with the date 
-~ and create a similarly named metadata file that records run input params 
-~ think about moving run control from a rmd to a .R script
-
-# RUNS FOR PAPER September 2019
-# A identical insecticides, seq vs rot, resistance frequency switch, A1 no costs no dispersal, A2 costs no dispersal, A3 dispersal no costs, A4 costs & dispersal. Sensitivity analysis of inputs.
-# B different insecticides, as above, no sensitivity analysis.
-# C identical insecticides, mortality switch, reduced input range so mortality starts above 90%
-# D different insecticides, mortality switch
-
-Runs are combinations of :
-Strategy (2)       : sequence, rotation
-Insecticides (2)   : same, different
-Switch (2)         : resistance frequency, mortality
-Scenarios (4)      : no costs no dispersal, costs no dispersal, dispersal no costs, costs & dispersal
-Replicates (10000) :
-
-So number of runs 2*2*2*4*10000 = 320000
-
-~ fix plot error when coverage==1
-
-STill if have min_rwr_interval > 0 an insecticide at very start of simulation outside the threshold will continue to be used.
-Mostly I think min_rwr_interval not a good idea, but does kind of represent what you might expect to happen in field when it is set to a year i.e.
-that once sprayed can't be unsprayed.
-
-dfr <- run_rot(start_freqs = 0.000000001, eff=1, logy=FALSE, mort_or_freq='mort', rot_interval=0, threshold=0.9)
-
-~ fix that mortality calc in plots and that for switching is currently repeated
-~ check it works with diff insecticides i.e. eff=c(0.7,0.8,0.9) I suspect not because mort_from_resist() or something else not properly vectorised.
-
-~ TIDY UP CODE AND IMPROVE DOCUMENTATION e.g. function args
-
-whatever I do I need to make mort/survival thing more transparent
-
-options
-~ pass as mortality and then convert to survival for calculation
-~ pass as survival convert to mortality for plot
-
 I think this is fixed now :
@@
 
+#14/06/2019
+Testing before sending to Sam.
+Error in mortRR["f", "intervention"] : incorrect number of dimensions
+
+This is when coverage set to 1, because there is no intervention dimension in mortRR[] in run_tot()
+
+fixed
+
+then got this from the script because it is trying to cast a tibble with an insecticide column and gens to a number
+Error: (list) object cannot be coerced to type 'double'
+
+so I modified gens_under_thresh() to just return the numeric column
+
+BUT this doesn't solve and I don't know why this problem appeared when it wasn't there before ???
+
+gens_under_thresh() is giving one row per insecticide, is that what I'm expecting ??
+No it is supposed to be :
+'the sum across all insecticides for the number of generations below threshold while that insecticide is deployed'
+
+so why has it changed ?? BUG introduced by the mortality addition
+##14/6/19 corrected bug here by adding back res <-      
+res <- summarise(res, tot_gens_dep_under50 = sum(.data$gens_dep_under50)) %>%
+
+# 20/8/19
+checking this message below from my notes, problem is that simulation continues even when mortality starts below thresholds.
+Inputs are set for sequence, what it does is switch between the insecticides for the min_rwr_interval.
+If the min_rwr_interval is set greater the length of the use periods increases too.
+
+Probably I need to check that mortality not below threshold at start of the r_w_r interval ? In insecticide_check()
+
+~~ why in this scenario does it rotate insecticides even though it starts below mortality threshold
+(to do with min_rwr threshold)
+run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 1e-04 , rot_interval = 0 , eff = 0.9 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.6 , migration = 0.01 , cost = 0.01 , logy = FALSE , no_r_below_start = FALSE , min_rwr_interval = 5 , threshold = 0.9 , mort_or_freq ='mort' )
+
+# this is the part in insecticide_check() that causes problem
+    if ( check_value > threshold &
+         #20/8/19 also this allows an inseticide to start being used even when outside threshold
+         gens_this_insecticide > min_rwr_interval )
+    {
+      change_insecticide <- TRUE
+
+OR does it ? surely the problem is that insecticide_switch() is choosing an insecticide that is already outside threshold ?
+
+something weird going on in insecticide_switch() mortality threshold should be converted to survival for the comparison but doesn't seem to be :
+
+switch from insecticide 1 to 3; new mortality vs threshold = 0.100044 vs 0.900000
+
+AHA! NASTY BUG! there are multiple insecticides and it cycles through them because I do threshold <- 1-threshold that can give uncertain results.
+
+# 4/9/19
+However I change the run workflow I should try to keep it so that I can do the runs on multiple cores.
+
+# 5/9/19
+
+refactoring inputs
+Currently the list linputs which has one row per run is created in sensi_an_rotations1.Rmd by hardcoded min-max limits
+
+new inputs plan
+Have a rotations function that accepts a csv file or defaults and uses that to create the linputs list, also it can save the csv. Closer to what resistance code did.
+
+How to name the different files ? what are similar names in openMalaria ?
+
+# 6/9/19
+
+GOOD getting there, now a csv file saved in extdata that specifies the experiment
+
+linputs_multi <- set_run_inputs()
+# to get all inputs for a single scenario
+linputs <- purrr::map(linputs_multi, 2)
+# run one scenario
+dfres <- do.call(run_rot, linputs)
+
+# 10/9/19
+got different insecticides working
+added get_one_in() to get inputs for one scenario from multi-senario file
+
+# 11/09/19
+This is the new workflow.
+
+# read an experiment file specifying ranges
+inex <- read_in_expt() #specify filename
+# create an object containing inputs for all scenarios
+linmulti <- set_run_inputs(inex=inex)
+# get inputs for a single scenario
+linputs <- get_one_in(linmulti, scen_num=2)
+# run one scenario
+dfres <- do.call(run_rot, linputs)
+
+# 12/09/19
+sorting mort_or_freq and other inputs in set_run_inputs() and get_one_in()
+
+something I've done means that now the first insecticide is not changed in any scenarios...
+
+linputs <- get_one_in(linmulti, scen_num=1)
+linputs <- get_one_in(linmulti, scen_num=2)
+linputs <- get_one_in(linmulti, scen_num=3)
+
+linputs <- get_one_in(linmulti, scen_num=4)
+linputs <- get_one_in(linmulti, scen_num=5)
+linputs <- get_one_in(linmulti, scen_num=6)
+
+# run one scenario
+dfres <- do.call(run_rot, linputs)
+
+# 13/09/19
+
+inex <- read_in_expt("_in_expt_rotations_default_one_scenario.csv", nscenarios=1) 
+linmulti <- set_run_inputs(inex=inex)
+linputs <- get_one_in(linmulti, scen_num=1)
+dfres <- do.call(run_rot, linputs)
+
+moving to have male_expo_prop default as 0
+
+# TODO
+
+CHECK THE DIFFERENCE between linputs in run_rot() which is working compared to the input csv which mostly doesn't work.
+
+~ something is wrong both when mort_or_freq set to mort and freq
+
+~ when set to freq the first insecticide doesn't stop being used
+  when set to mort the first insecticide is just used very briefly
+
+~ add log-normal distribution for frequency
+~ how to deal with running seq vs rotations
+~ there is other stuff I still need in sensi_an_rotations1.Rmd - e.g. post-processing of outputs
+~ could I have multiple experiments as columns in the same input file ?
+
+~ add the random seed to the experiment file (and decide where it will be set)
+
+~ investigate getting runs to name the output file with the date 
+~ and create a similarly named metadata file that records run input params
+
+# RUNS FOR PAPER September 2019
+# A identical insecticides, seq vs rot, resistance frequency switch, A1 no costs no dispersal, A2 costs no dispersal, A3 dispersal no costs, A4 costs & dispersal. Sensitivity analysis of inputs.
+# B different insecticides, as above, no sensitivity analysis.
+# C identical insecticides, mortality switch, reduced input range so mortality starts above 90%
+# D different insecticides, mortality switch
+
+Runs are combinations of :
+Strategy (2)       : sequence, rotation
+Insecticides (2)   : same, different
+Switch (2)         : resistance frequency, mortality
+Scenarios (4)      : no costs no dispersal, costs no dispersal, dispersal no costs, costs & dispersal
+Replicates (10000) :
+
+So number of runs 2*2*2*4*10000 = 320000
+
+~ fix plot error when coverage==1
+
+STill if have min_rwr_interval > 0 an insecticide at very start of simulation outside the threshold will continue to be used.
+Mostly I think min_rwr_interval not a good idea, but does kind of represent what you might expect to happen in field when it is set to a year i.e.
+that once sprayed can't be unsprayed.
+
+dfr <- run_rot(start_freqs = 0.000000001, eff=1, logy=FALSE, mort_or_freq='mort', rot_interval=0, threshold=0.9)
+
+~ fix that mortality calc in plots and that for switching is currently repeated
+~ check it works with diff insecticides i.e. eff=c(0.7,0.8,0.9) I suspect not because mort_from_resist() or something else not properly vectorised.
+
+~ TIDY UP CODE AND IMPROVE DOCUMENTATION e.g. function args
+
+whatever I do I need to make mort/survival thing more transparent
+
+options
+~ pass as mortality and then convert to survival for calculation
+~ pass as survival convert to mortality for plot
+
 ~ ian interested in starting frequencies at 10e-4, can do that for the runs with different values per insecticide.

### 2019-09-13 R/read_in_expt.r

@@ -3,7 +3,7 @@
 #' have a default file, be able to save files to reproduce experiments
 #' 
 #' @param file name of the text file containing the experiment ranges
-#' @param n_scenarios number of runs, set here and not in input file
+#' @param nscenarios number of runs, set here and not in input file
 #' 
 # @examples 
 #' 
@@ -12,12 +12,15 @@
 
 
 read_in_expt <- function( file = NULL,
-                          n_scenarios = 10)
+                          nscenarios = 10)
 {
 
 
   # have a default file
   if (is.null(file)) file <- system.file("extdata", "_in_expt_rotations_default.csv", package = "rotations", mustWork = TRUE)
+  # or file in the extdata folder
+  else file <- system.file("extdata", file, package = "rotations", mustWork = TRUE)
+  # TODO allow file to be in other specified location too, i.e. accept a path
     
   # function to read input file with experiment specifications
   
@@ -37,15 +40,16 @@ read_in_expt <- function( file = NULL,
   tmp4 <- tmp3[2,]
   
   #convert variables to numeric
-  #TODO later I can have character variables too
-  tmp5 <- dplyr::mutate_all(tmp4, ~as.numeric(as.character(.))) 
+  #except named character variables
+  #tmp5 <- dplyr::mutate_all(tmp4, ~as.numeric(as.character(.))) 
+  tmp5 <- dplyr::mutate_at(tmp4,vars(-mort_or_freq), ~as.numeric(as.character(.)))
   
   #dplyr other options
   #df %>% mutate_if(is.factor, ~as.numeric(as.character(.)))
   # by specific columns:
   #df %>%   mutate_at(vars(x, y, z), ~as.numeric(as.character(.))) 
   
-  tmp5$n_scenarios <- n_scenarios
+  tmp5$nscenarios <- nscenarios
   
   tmp5
 

### 2019-09-13 R/run_rot.r

@@ -6,8 +6,10 @@
 #' @param n_insecticides number of insecticides (and hence loci)
 #' @param start_freqs starting frequencies of resistance either one per insecticide or same for all
 #' @param rot_interval frequency of rotation (in generations) if set to zero means sequence, i.e. change when threshold reached
-#' @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, also precludes switch to an insecticide.
+# @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, defaults mort:0.9 freq:0.5, also precludes switch to an insecticide.
 #' @param mort_or_freq whether threshold for insecticide change is mortality 'mort' or resistance frequency 'freq'
+#' @param mort_thresh mortality threshold for switching insecticides, default 0.9, only used if mort_or_freq is mort
+#' @param freq_thresh resistance frequency threshold for switching insecticides, default 0.5, only used if mort_or_freq is freq
 #' @param migration migration rate between treated & untreated areas 0-1. We assume that immigration=emigration.
 #' @param coverage proportion of mosquitoes that are covered by the intervention (and 1-C is the proportion of the population in the untreated refugia), if coverage set to 1 no refugia.
 #' @param start_insecticide which insecticide to start with
@@ -54,24 +56,26 @@
 run_rot <- function(max_gen = 200, 
                     n_insecticides = 4, 
                     start_freqs = 0.01,
-                    rot_interval = 10, 
-                    threshold = 0.5, 
+                    rot_interval = 0, 
+                    #threshold = 0.5, now set in first lines of function
                     mort_or_freq = 'mort', #'freq'
+                    mort_thresh = 0.9,
+                    freq_thresh = 0.5,
                     migration = 0.01, 
                     #migrate_intervention = 0.01, 
                     coverage = 0.8, 
                     start_insecticide = 1,
                     expo_hi = 0.8,
                     expo_lo = 0,                              
-                    male_expo_prop = 1,
-                    eff = 0.8, #c(0.5, 0.7, 0.9),
+                    male_expo_prop = 0,
+                    eff = 0.95, #c(0.5, 0.7, 0.9),
                     dom_sel = 0.5, #c(0.5, 0.5, 0.5),
                     dom_cos = 0.5, #c(0.5, 0.5, 0.5),
                     rr = 0.5, #c(0.5, 0.5, 0.5),
[…]

### 2019-09-13 R/set_run_inputs.r

@@ -25,24 +25,53 @@ set_run_inputs <- function( inex = NULL )
   # have a default file
   if (is.null(inex)) inex <- read_in_expt()
   
+  # set random seed so that runs can be reproduced
+  set.seed(inex$rand_seed)
   
   # currently a list that grows, slow, but not a large file  
   linmulti <- list()
   
   
   # take code from sensi_an_rotations1
-  for(i in 1:inex$n_scenarios)
+  for(i in 1:inex$nscenarios)
   {
-    if (i%%200 == 1) message("scenario ",i," of ",inex$n_scenarios," ",Sys.time())
+    if (i%%200 == 1) message("scenario ",i," of ",inex$nscenarios," ",Sys.time())
 
-    linmulti$max_gen[i] <- inex$max_gen
- 
-    # n_insecticides different because choosing random integer
-    linmulti$n_insecticides[i] <- sample( inex$n_insecticides_min:inex$n_insecticides_max, 1 )  #beware set replace=T if more than 1
     
+    # constant inputs
+    linmulti$max_gen[i] <- inex$max_gen
+    linmulti$min_rwr_interval[i] <- inex$min_rwr_interval
+    linmulti$min_gens_switch_back[i] <- inex$min_gens_switch_back
+    linmulti$no_r_below_start[i] <- inex$no_r_below_start
+    linmulti$no_r_below_mut[i] <- inex$no_r_below_mut
+    linmulti$exit_rot[i] <- inex$exit_rot
+    linmulti$mort_or_freq[i] <- inex$mort_or_freq
+    linmulti$threshold[i] <- inex$threshold 
+    linmulti$plot[i] <- inex$plot
+    
+ 
+    # random integers BEWARE that sample selects from 1:x when x is length 1
+    # causes problem when want n_insecticides or rot_interval to be fixed
+    # this resample function from the sample help file sorts it
+    # so e.g when min & max are 4 sample is always 4
+    resample <- function(x, ...) x[sample.int(length(x), ...)]
+    
+    linmulti$n_insecticides[i] <- resample( inex$n_insecticides_min:inex$n_insecticides_max, 1 )  #beware set replace=T if more than 1
+    # rotation interval, 0 for sequence
+    linmulti$rot_interval[i]  <-  resample( inex$rot_interval_min:inex$rot_interval_max, 1)    
        
     # select random uniform numbers between the ranges
     # this copes if min & max are the same
+    
[…]

### 2019-09-13 inst/documents/rotation_runs2019.Rmd
settin up rot vs seq runs

@@
 ---
-title: "rotation runs 2019"
+title: "rotation_runs2019"
 output: pdf_document
@@
 
+# options
+# i) run each scenario for rotation then sequence and calculate difference in the loop
+# ii) run each scenario for rotation then sequence and save gens_under_thresh for each in the output file
+
+nscenarios <- 10
+
 # read an experiment file specifying ranges
-inex <- read_in_expt("in_rotations_paper_2019_inssame_freq.csv", nscenarios=1)
-
-#inex <- read_in_expt("_in_expt_rotations_default_one_scenario.csv", nscenarios=1)
+inex <- read_in_expt("expt_rot_mort_dispersal.csv", nscenarios=nscenarios)
 
@@
 linmulti <- set_run_inputs(inex=inex)
-linmulti <- set_run_inputs()
 
-# get inputs for a single scenario
-linputs <- get_one_in(linmulti, scen_num=2)
+# run all scenarios
+for(i in 1:nscenarios)
+{
+  linputs <- get_one_in(linmulti, scen_num=i)
 
+  # to compare sequence to rotation run all the scenarios above for both seq & rot 
+  for(rot_or_not in 0:1)
+  {  
+    #to get rot_int of 0 in alternate runs - the other run uses the value from the input file
+    if (rot_or_not == 0) linputs$rot_interval <- 0
+
+    ##################
     # run one scenario
@@
 
+    # summarise results per scenario
+    # note this is also currently calc at end of run_rot() here is probably better
+    gens <- gens_under_thresh(dfres)
+
+    # to be able to save as for rot or seq
+    if ( linputs$rot_interval == 0 ) {gens_seq <- as.numeric(gens)}
+    else                             {gens_rot <- as.numeric(gens)}
+
+  } #end rot_or_not loop
+
+  # add scenario id column and difference onto results
+  # df_in_out is one row
+  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, id=i))
+  df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
+
+  # if the first scenario create df to store results of all scenarios
+  if (i==1)
+  {
+    df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=n_scenarios))
+    names(df_res_all) <- names(df_in_out)
+  }
+
+  #putting results into one row of overall results dataframe
+  df_res_all[i,] <- df_in_out
+
+} #end scenario loop
+
 ```

### 2019-09-13 inst/documents/rotations_notes.Rmd
settin up rot vs seq runs

@@
 
+    #now set to log-uniform 
+    minlog <- log(inex$start_freqs_min,10)
+    maxlog <- log(inex$start_freqs_max,10)
+
+    linmulti$start_freqs[i] <-  10^runif(1, min=minlog, max=maxlog)
+
+Use hist to display the distribution of values.    
+hist(linmulti$start_freqs, breaks=100)
+
 # TODO
-
-CHECK THE DIFFERENCE between linputs in run_rot() which is working compared to the input csv which mostly doesn't work.
-
-~ something is wrong both when mort_or_freq set to mort and freq
-
-~ when set to freq the first insecticide doesn't stop being used
-  when set to mort the first insecticide is just used very briefly
-
-~ add log-normal distribution for frequency
+see rotation_runs2019.Rmd & rotations-experiments-creating-201909.xlsx
+
 ~ how to deal with running seq vs rotations
+~ should rotation intervals just be set yearly i.e. 10,20,30 etc.
+
+GET FREQUENCY RUNS sorted first (easier)
+I could put the expt files all into one csv to start creating them, and divide into separate files later.
+
 ~ there is other stuff I still need in sensi_an_rotations1.Rmd - e.g. post-processing of outputs

### 2019-09-13 R/set_run_inputs.r

@@ -70,8 +70,15 @@ set_run_inputs <- function( inex = NULL )
     linmulti$coverage[i] <-     runif(1, min=inex$coverage_min,    max=inex$coverage_max)
     linmulti$migration[i] <-     runif(1, min=inex$migration_min,   max=inex$migration_max)
     
-    #TODO make this lognormal to get lower frequencies too
-    linmulti$start_freqs[i] <-    runif(1, min=inex$start_freqs_min,   max=inex$start_freqs_max) 
+    #log-uniform to get lower frequencies represented
+    #select ‘x’ from uniform -1 to -4 and set to 10^x to give equal weight to each log interval.
+    #log(0.01,10)
+    #from resistance
+    #P_1 <- 10^-(runif(1, min=1, max=4))
+    minlog <- log(inex$start_freqs_min,10)
+    maxlog <- log(inex$start_freqs_max,10)
+    
+    linmulti$start_freqs[i] <-  10^runif(1, min=minlog, max=maxlog) 
     
     # some can be different for each insecticide if insecticides_different is set to 1
     n_ins <- linmulti$n_insecticides[i]

### 2019-10-02 inst/documents/rotation_runs2019.Rmd
corrected bugs in rotation_runs2019.Rmd to do with saving rot_interval

@@
 
+  if (linputs$rot_interval == 0) stop("in rotation_runs2019 rot_interval in input file needs to be > 0 to allow it to be compared to a sequence")
+
+  #record rotation rot interval locally otherwise it get soverwritten by 0
+  rot_rot_interval <- linputs$rot_interval
+
   # to compare sequence to rotation run all the scenarios above for both seq & rot 
+  # need to do the rotation 2nd so that rot_interval from the input file can be saved with the outputs
   for(rot_or_not in 0:1)
@@
     if (rot_or_not == 0) linputs$rot_interval <- 0
+    else linputs$rot_interval <- rot_rot_interval
 
@@
   {
-    df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=n_scenarios))
+    df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=nscenarios))
     names(df_res_all) <- names(df_in_out)
@@
 
+  #can save partial outputs in case doesn't make to end
+  if (i==nscenarios | i%%1000 == 0)
+  {
+
+    # save object containing inputs and results as rda for analysis
+    # by default saved to inst/documents folder
+    save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))    
+  }
+
 } #end scenario loop

### 2019-10-02 R/get_one_in.r

@@ -33,4 +33,7 @@ get_one_in <- function( linmulti = NULL,
   linputs <- lapply(linputs, function(x) if (!grepl(",",x)) x else
     as.numeric(unlist(strsplit(as.character(x),split=", "))))
   
+  # to make sure it gets returned
+  linputs
+  
 }

### 2019-10-04 inst/documents/rotation_runs2019.Rmd
multi-experiment workflow now seems to be working

@@
 library(gridExtra)
+library(readxl) # for reading the excel file that allows me to keep formatting to see inputs clearly
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5.5, comment=NA, results='asis'}
 
-# options
-# i) run each scenario for rotation then sequence and calculate difference in the loop
-# ii) run each scenario for rotation then sequence and save gens_under_thresh for each in the output file
+nscenarios <- 100
+plotyn <- 1 #allow turning plots on and off from here
 
-nscenarios <- 10
+# still deciding how best to do this ...
+# infile_multi <- 'rotations-experiments-creating-201909.xlsx'
+# infile_multi <- system.file("extdata", infile_multi, package = "rotations", mustWork = TRUE)
+# inex_all <- read_excel(infile_multi)
 
-# read an experiment file specifying ranges
-inex <- read_in_expt("expt_rot_mort_dispersal.csv", nscenarios=nscenarios)
+# NOTE to add new files currently need to reinstall package because infiles are in extdata
+#infile <- "expt-A2-freq-insame-costs-nodisp.csv"
+# infile <- "expt-A4-freq-insame-costs-disp.csv"
+# # get the id for the expot to include in outfilenames, specifies experiment attributes
+# id_expt <- substr(infile,6,7)
+# # read an experiment file specifying ranges
+# inex <- read_in_expt(infile, nscenarios=nscenarios)
+
+# now specify an experiment from the multi-experiment file
+id_expt <- 'A2'
+#id_expt <- 'A4'
+
+inex <- read_in_expt(nscenarios = nscenarios, multi=TRUE, id_expt=id_expt)
+
+# text to go at start of document
+cat("Experiment :",id_expt," scenario number is pagenum -1 \n\n    ")
+
+cat("Experiment key :\n    ")
+cat("A: resistance frequency threshold, insecticides the same\n    ")
+cat("B: resistance frequency threshold, different insecticides\n    ")
+cat("C: mortality threshold, insecticides the same\n    ")
+cat("D: mortality threshold, different insecticides\n    ")
+
+cat("1: no costs, no dispersal\n    ")
+cat("2: costs, no dispersal\n    ")
+cat("3: no costs, dispersal\n    ")
+cat("4: costs, dispersal\n    ")
 
@@
 {
+  #message should output to the R markdown console
+  if (i%%200 == 1) message("scenario ",i," of ",nscenarios," ",Sys.time())
+
   linputs <- get_one_in(linmulti, scen_num=i)
@@
 
+  linputs$plot <- plotyn #allows setting whether to plot from script
+  linputs$add_gens_under50 <- TRUE #to add text to plots of gens under threshold
+
   #record rotation rot interval locally otherwise it get soverwritten by 0
   rot_rot_interval <- linputs$rot_interval
+
+  #to add label between plots in output doc
+  #this adds label underneath, not my intention but does the job
+  cat("\n\n\\pagebreak\n    ") #needs results='asis' in the chunk options
+  #cat("scenario",i,"\n     ") stopped plots from aligning
+  #instead just remember that scenario number is pagenum -1
 
@@
     else                             {gens_rot <- as.numeric(gens)}
+
+    #code to rerun approx this scenario
+    cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
+                #"mort_or_freq='",linputs$mort_or_freq,"', ", #quotes dont work, freq is now default
+                #'mort_or_freq="',linputs$mort_or_freq,'", ',
+                "cost=",signif(linputs$cost,2),", ",
+                "start_freqs=",signif(linputs$start_freqs,2),", ", 
+                "rot_interval=",linputs$rot_interval,", ",
+                "eff=",signif(linputs$eff,2),",\n",
+                "dom_sel=",signif(linputs$dom_sel,2),", ",
+                "dom_cos=",signif(linputs$dom_cos,2),", ",
+                "rr=",signif(linputs$rr,2),", ",
+                "expo_hi=",signif(linputs$expo_hi,2),", ",
+                "coverage=",signif(linputs$coverage,2),",\n",
+                "migration=",signif(linputs$migration,2),", ",
+                #"max_gen=",linputs$max_gen,", ", 
+                "min_rwr_interval=",linputs$min_rwr_interval,            
+                ")\n\n") )
 
@@
     # by default saved to inst/documents folder
-    save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))    
+    save(df_res_all, file=paste0(outFolder,format(Sys.time(), "%Y-%m-%d-%H-%M-"),'df_res_all_rot',id_expt,nscenarios,'.rda')) #paste0(outFolder,'*.rda')) 
+    # also save the experiment input file
+    save(inex, file=paste0(outFolder,format(Sys.time(), "%Y-%m-%d-%H-%M-"),'inex-',id_expt,'.rda'))   
   }

### 2019-10-04 inst/documents/rotations_notes.Rmd
corrected bugs in rotation_runs2019.Rmd to do with saving rot_interval
multi-experiment workflow now seems to be working

@@
 
+# 2/10/19
+corrected bugs in rotation_runs2019.Rmd to do with saving rot_interval
+
+# 3/10/19
+rotation_runs2019.Rmd - working for 1000 runs
+but progress messages not appearing in terminal
+If I knit I do get progress messages and also get a document with all plots but not sure how that will work if running for 10k
+
+~ improved plots so seq and rot on same page
+~ changed gens_under_thresh to annotation to make text better
+
+~ gens_under_thresh seemed to be counting 11 for a 10gen rotation ? Seems that a 10 gen rotation doesn't switch until generation 11 etc.
+~ AND seemed that rot_interval not passing to run_rot ?
+
+BOTH of these were because exit_rot default was true, now changed default to be false
+
+TWO SEEMING PROBLEMS
+prob1) sequence insecticide continues to be used after threshold I expect may be a problem with exit_rot=FALSE
+prob2) migration in the new runs seems to make active & refugia the same, maybe I've set limits too high ?
+I have default migration set to 0.01 and limits currently set to 0.1 to 0.9
+df_res2 <- run_rot(rot_interval = 0, mort_or_freq = 'freq', migration=0.5)
+
+Having difficulty replicating prob1
+
+df_res2 <- run_rot(rot_interval = 0, mort_or_freq = 'freq', migration=0.5)
+
+I set min_rwr_interval to 10 to stop rapid switches for sequence.
+
+run_rot(n_insecticides=2, mort_or_freq='freq', cost=0.013, start_freqs=0.077, rot_interval=10, eff=0.98,
+ dom_sel=0.55, dom_cos=0.55, rr=0.47, expo_hi=0.69, coverage=0.23, migration=0.86, max_gen=500,
+ min_rwr_interval=1)
+
+note there is a problem with the quotes when copying from the pdf
+
+diagnostics not helpful
+
+Aha, think its just to do with exit_rot = FALSE
+run_rot(n_insecticides=2, mort_or_freq='freq', rot_interval=10)
+
+# 4/10/19
+
+Fixed rotation insecticide continues to be used after threshold
+which was a bug in insecticide_check() introduced when mort thresholds added
+
+NEW SEEMING PROBLEM
+run_rot(n_insecticides=2, mort_or_freq='freq', rot_interval=10, exit_rot=FALSE, diagnostics = TRUE, min_gens_switch_back = 0)
+Aha I think problem that insecticide itself is not assessed so keeps being used even when abovbe thresh
+FIXED
+
+fixed gens_under_thresh count which had changed to be for all insecticides, not just one.
+
+successfully GOT runs2019 to access the xls file for experiments directly
+
 # TODO
@@
 
-~ how to deal with running seq vs rotations
-~ should rotation intervals just be set yearly i.e. 10,20,30 etc.
+migration limits ???
+I need to check on them more ! to make sure I know what they are doing.
+> RAF2 <- rot_migrate(RAF1gen, migration=1, coverage=0.1, verbose=TRUE)
+  m.intervention f.intervention m.refugia f.refugia
+1              1              1         0         0
+2              1              1         0         0
+3              1              1         0         0
+  m.intervention f.intervention m.refugia f.refugia
+1            0.1            0.1       0.1       0.1
+2            0.1            0.1       0.1       0.1
+3            0.1            0.1       0.1       0.1
+
+CHECK ALL INPUT PARAMS
+
+~ create some plot functions to work on the output files (df_res_all) e.g. plot_rotseq_diff()
+
+~ try to get run progress message into console or terminal
+~ improve readme.md with details of how to run simulations
+
+~ maybe just do annual rotation interval i.e. 10 ?
 

### 2019-10-04 R/read_in_expt.r

@@ -4,31 +4,60 @@
 #' 
 #' @param file name of the text file containing the experiment ranges
 #' @param nscenarios number of runs, set here and not in input file
+#' @param multi whether to read in a multi experiment excel file
+#' @param id_expt if a multi experiment file which id_expt to use (first two chars of the column name)
 #' 
-# @examples 
+#' @examples 
+#' #single experiment
+#' inex <- read_in_expt(nscenarios = 10, multi=FALSE)
+#' 
+#' #multi experiment
+#' #filename of multiexpt file currently hardcoded into function
+#' inex <- read_in_expt(nscenarios = 10, multi=TRUE, id_expt='A4')
 #' 
 #' @return  
 #' @export
 
 
 read_in_expt <- function( file = NULL,
-                          nscenarios = 10)
+                          nscenarios = 10,
+                          multi = FALSE,
+                          id_expt = 'A2')
 {
 
 
-  # have a default file
-  if (is.null(file)) file <- system.file("extdata", "_in_expt_rotations_default.csv", package = "rotations", mustWork = TRUE)
-  # or file in the extdata folder
-  else file <- system.file("extdata", file, package = "rotations", mustWork = TRUE)
-  # TODO allow file to be in other specified location too, i.e. accept a path
+  if (multi)
+  {
+    infile_multi <- 'rotations-experiments-creating-201909.xlsx'
     
-  # function to read input file with experiment specifications
+    infile_multi <- system.file("extdata", infile_multi, package = "rotations", mustWork = TRUE)
+    
+    #read excel file
+    inex_multi <- readxl::read_excel(infile_multi)
+    #get just names column and 1 column containing inputs for this experiment id
+    #specified by first two chars of the column name
+    inex_one <- dplyr::select(inex_multi,1, dplyr::starts_with(paste0(id_expt)))
+    
+    
+  } else
+  {
+    # single experiment file
+    # have a default file
+    if (is.null(file)) file <- system.file("extdata", "_in_expt_rotations_default.csv", package = "rotations", mustWork = TRUE)
+    # or file in the extdata folder
+    else file <- system.file("extdata", file, package = "rotations", mustWork = TRUE)
+    # TODO allow file to be in other specified location too, i.e. accept a path
+    
+    # function to read input file with experiment specifications
+    
+    #read scenario input (single column)
+    inex_one <- read.csv(file)    
+  }
   
[…]

### 2019-10-04 R/rot_migrate.r

@@ -6,16 +6,21 @@
 #' @param verbose whether to output before and after migration
 #' 
 #' @examples
-#' #frequencies the same for all insecticides
-#' RAF <- set_start_freqs(n=3, freqs=0.001)
-#' RAF[,,,1] # to view generation 1
+#' 
 #' #frequencies different for each insecticide
 #' RAF <- set_start_freqs(freqs=c(0.1,0.01,0.001))
-#' RAF[,,,1] # to view generation 1
-#' #testing mutliplying frequency arrays
-#' RAF <- set_start_freqs(max_gen = 1, freqs=c(0.1,0.01,0.001))
-#' RAF2 <- set_start_freqs(max_gen = 1, freqs=c(1,2,3))
-#' RAF*RAF2
+#' 
+#' simplest example, freqs 1 in intervention, 0 in refugia
+#' RAF <- set_start_freqs(n=3, freqs=1)
+#' RAF1gen <- RAF[,,,1] # to view generation 1
+#' #set freqs in refugia to 0
+#' RAF1gen[,,'refugia']  <- 0
+#' #testing migration
+#' RAF2 <- rot_migrate(RAF1gen, migration=0.5, coverage=0.5, verbose=TRUE)
+#' 
+#' #max migration leads to random mixing, irrespective of coverage, intervention and refugia end with same freq
+#' RAF2 <- rot_migrate(RAF1gen, migration=1, coverage=0.5, verbose=TRUE)
+#' RAF2 <- rot_migrate(RAF1gen, migration=1, coverage=0.2, verbose=TRUE)
 #' 
 #' 
 #' @return array of resistance freqs in 1 gen following migration

### 2019-10-04 R/run_rot.r

@@ -58,7 +58,7 @@ run_rot <- function(max_gen = 200,
                     start_freqs = 0.01,
                     rot_interval = 0, 
                     #threshold = 0.5, now set in first lines of function
-                    mort_or_freq = 'mort', #'freq'
+                    mort_or_freq = 'freq', #'mort', 
                     mort_thresh = 0.9,
                     freq_thresh = 0.5,
                     migration = 0.01, 
@@ -77,7 +77,7 @@ run_rot <- function(max_gen = 200,
                     min_rwr_interval = 10,
                     no_r_below_start = FALSE,
                     no_r_below_mut = FALSE,
-                    exit_rot = TRUE,
+                    exit_rot = FALSE,
                     min_gens_switch_back = 10,
                     
                     #inputs below not needed to run model itself

### 2019-10-04 R/set_run_inputs.r

@@ -35,8 +35,7 @@ set_run_inputs <- function( inex = NULL )
   # take code from sensi_an_rotations1
   for(i in 1:inex$nscenarios)
   {
-    if (i%%200 == 1) message("scenario ",i," of ",inex$nscenarios," ",Sys.time())
-
+    #if (i%%200 == 1) message("scenario ",i," of ",inex$nscenarios," ",Sys.time())
     
     # constant inputs
     linmulti$max_gen[i] <- inex$max_gen

### 2019-10-07 inst/documents/rotation_runs2019.Rmd
improving experiment outputs

@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5.5, comment=NA, results='asis'}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5.4, comment=NA, results='asis'}
 
@@
 # now specify an experiment from the multi-experiment file
-id_expt <- 'A2'
+#id_expt <- 'A2'
 #id_expt <- 'A4'
+id_expt <- 'B2'
+
+# to move the pdf of a preceeding run of this file to a saved location
+# RUN the line above and below after knitting the file
+# ASSUMING id_expt still set to previous run and still same day
+# BEWARE that id_expt will not be set correctly in home env if file was knitted
+#file.rename("rotation_runs2019.pdf",paste0(outFolder,format(Sys.time(), "%Y-%m-%d-"),id_expt,"-results-figs.pdf"))
 
@@
 cat("4: costs, dispersal\n    ")
+
+cat("\nNumbers on the right are the number of generations that this insecticide is in use and the corresponding resistance is below threshold.\n   ")
+
+cat("\nOn each page a Sequence is at the top and an annual Rotation at the bottom.")
 
@@
 
-    #code to rerun approx this scenario
+    #code to rerun approx this scenario, can be copied from pdf and pasted into R
+    #has to cope with converting vectors to strings
+    #spaces between args stop the pdf from messing with words
     cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
+                "rot_interval=",linputs$rot_interval,", ",
                 #"mort_or_freq='",linputs$mort_or_freq,"', ", #quotes dont work, freq is now default
                 #'mort_or_freq="',linputs$mort_or_freq,'", ',
-                "cost=",signif(linputs$cost,2),", ",
                 "start_freqs=",signif(linputs$start_freqs,2),", ", 
-                "rot_interval=",linputs$rot_interval,", ",
-                "eff=",signif(linputs$eff,2),",\n",
-                "dom_sel=",signif(linputs$dom_sel,2),", ",
-                "dom_cos=",signif(linputs$dom_cos,2),", ",
-                "rr=",signif(linputs$rr,2),", ",
+                "cost=c(",paste(signif(linputs$cost,2),collapse=','),"), ", 
+                "eff=c(",paste(signif(linputs$eff,2),collapse=','),"), ",               
+                "dom_sel=c(",paste(signif(linputs$dom_sel,2),collapse=','),"), ",
+                "dom_cos=c(",paste(signif(linputs$dom_cos,2),collapse=','),"), ",
+                "rr=c(",paste(signif(linputs$rr,2),collapse=','),"), ",
                 "expo_hi=",signif(linputs$expo_hi,2),", ",
-                "coverage=",signif(linputs$coverage,2),",\n",
+                "coverage=",signif(linputs$coverage,2),", ",
                 "migration=",signif(linputs$migration,2),", ",

### 2019-10-07 R/read_in_expt.r

@@ -28,7 +28,7 @@ read_in_expt <- function( file = NULL,
 
   if (multi)
   {
-    infile_multi <- 'rotations-experiments-creating-201909.xlsx'
+    infile_multi <- 'rotations-paper-inputs-201910.xlsx'
     
     infile_multi <- system.file("extdata", infile_multi, package = "rotations", mustWork = TRUE)
     

### 2019-10-09 inst/documents/rotations_notes.Rmd
improving experiment outputs
new simpler dispersal working

@@
 
+# 7/10/19
+got diff insecticide runs working
+Got mortality to appear properly in plots & key in non dispersal runs.
+
+# 8/10/19 
+added to readme
+
+# 9/10/19
+
+spent the day sweating over getting migration working.
+
+Now I'm pleased I think it is simpler and I know what its doing.
+Migration 1 is random mixing.
+
+see the excel file for testing : rotations-dispersal-demo-andy201910.xlsx
+
 # TODO
@@
 
+check on dispersal limits using new method. Seems that anything above 0.4 leads to near complete mixing
+run_rot(migration=0.4, coverage=0.5)
+#interesting how small levels of migration cause resistance freqs to decline even with no costs
+
+write word description of what dispersal does
+rename migration to dispersal ?
+
 migration limits ???
 I need to check on them more ! to make sure I know what they are doing.
+
+RAF <- set_start_freqs(n=3, freqs=1)
+> RAF1gen <- RAF[,,,1] # to view generation 1
+> #set freqs in refugia to 0
+> RAF1gen[,,'refugia']  <- 0
+> #testing migration
+> RAF2 <- rot_migrate(RAF1gen, migration=0.5, coverage=0.5, verbose=TRUE)
+  m.intervention f.intervention m.refugia f.refugia
+1              1              1         0         0
+2              1              1         0         0
+3              1              1         0         0
+  m.intervention f.intervention m.refugia f.refugia
+1           0.75           0.75      0.25      0.25
+2           0.75           0.75      0.25      0.25
+3           0.75           0.75      0.25      0.25
+
 > RAF2 <- rot_migrate(RAF1gen, migration=1, coverage=0.1, verbose=TRUE)
@@
 
+~ consider renaming package to avoid conflict with CRAN rotations package insectrot rotinsect
+
 CHECK ALL INPUT PARAMS

### 2019-10-09 R/migration_test.r

@@ -1,4 +1,4 @@
-#' migration_test to testimplementation of migration between an implementation area and refugia
+#' migration_test to test implementation of migration between treated area and refugia
 #'
 #' @param migration migration between 0-1
 #' @param coverage proportion of popn in treated area

### 2019-10-09 R/rot_migrate.r

@@ -36,32 +36,45 @@ rot_migrate <- function( RAF1gen,
   # which might be the case if insects killed in the treated area caused greater than 
   # random migration in from the untreated area
   
+  # OLD WAY
   # to allow migration to be on a scale from 0-1 (1-coverage is the max)
   # calculating this in rot_migrate function allows easier testing of function
-  migrate_intervention <- migration*(1-coverage)
-  migrate_refugia <- migrate_intervention*coverage/(1-coverage)  
+  # migrate_intervention <- migration*(1-coverage)
+  # migrate_refugia <- migrate_intervention*coverage/(1-coverage)  
+  # 
+  # # RAF[,, ensures calc is repeated for each insecticide and sex 
+  # # more concise version of original code
+  # 
+  # mig_intervention <- (1-migrate_intervention)*RAF1gen[,, 'intervention']+
+  #   migrate_intervention *RAF1gen[,, 'refugia']
+  # 
+  # mig_refugia      <- (1-migrate_refugia)*RAF1gen[,, 'refugia']+
+  #   migrate_refugia *RAF1gen[,, 'intervention']
+  # 
+  # RAF1gen[,, 'intervention'] <- mig_intervention
+  # RAF1gen[,, 'refugia'] <- mig_refugia
+  # 
   
-  # not needed because now calculated
-  # if(migrate_intervention>(1-coverage)){
-  # message(sprintf("warning from calibration: migration rate in/out of intervention exceed 1 minus coverage\n"))   
-  # }
+  # 9/10/19 changing dispersal to make simpler and clearer
+  leavers_intervention <- migration*(1-coverage)
+  leavers_refugia <- migration * coverage
+  stay_intervention <- 1-leavers_intervention
+  stay_refugia <- 1-leavers_refugia
   
-  #RAF <- array_named(insecticide=1:n_insecticides, sex=c('m','f'), site=c('intervention','refugia'), gen=1:max_gen)
+  # after dispersal
+  # f_intervention <- f_int*stay_intervention + f_ref*leave_intervention
+  # because those that have left intervention are replaced by dispersers from refuge
+  # BE CAREFUL - it feels wrong to multiply f_refugia by leavers_intervention but it is right
[…]

### 2019-10-10 inst/documents/rotation_runs2019.Rmd
dispersal modified and seemingly working

@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5.4, comment=NA, results='asis'}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5.2, comment=NA, results='asis'}
 
@@
 #id_expt <- 'A2'
+#id_expt <- 'A3'
 #id_expt <- 'A4'
-id_expt <- 'B2'
+#id_expt <- 'B2'
+#id_expt <- 'B3'
+id_expt <- 'B4'
 
@@
     #spaces between args stop the pdf from messing with words
-    cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
-                "rot_interval=",linputs$rot_interval,", ",
+    cat(paste0( "run_rot(rot_interval=",linputs$rot_interval,", ", 
+                "n_insecticides=",linputs$n_insecticides,", ",
                 #"mort_or_freq='",linputs$mort_or_freq,"', ", #quotes dont work, freq is now default
@@
                 "expo_hi=",signif(linputs$expo_hi,2),", ",
+                "male_expo_prop=",signif(linputs$male_expo_prop,2),", ",
                 "coverage=",signif(linputs$coverage,2),", ",

### 2019-10-10 inst/documents/rotations_notes.Rmd
dispersal modified and seemingly working

@@
 # 9/10/19
-
 spent the day sweating over getting migration working.
@@
 see the excel file for testing : rotations-dispersal-demo-andy201910.xlsx
+
+# 10/10/9
+migration effects on gens under thresh
+run_rot(coverage=0.5, cost=0, migration=0) #64
+run_rot(coverage=0.5, cost=0, migration=0.01) #72
+run_rot(coverage=0.5, cost=0, migration=0.1) #92
+run_rot(coverage=0.5, cost=0, migration=0.3) #104
+run_rot(coverage=0.5, cost=0, migration=0.6) #116
+run_rot(coverage=0.5, cost=0, migration=0.8) #120
+run_rot(coverage=0.5, cost=0, migration=1) #124
+
+migration limits ? 0.01 - 0.6
+
+But other A3 scenarios showe that nearly all runs with migration > 0.2 or 0.3 have treated area & refugia with very similar frequency curves.
+
+I noticed that male_expo_prop min was 1 in earlier runs, changed to 0
 
@@
 I need to check on them more ! to make sure I know what they are doing.
-
-RAF <- set_start_freqs(n=3, freqs=1)
-> RAF1gen <- RAF[,,,1] # to view generation 1
-> #set freqs in refugia to 0
-> RAF1gen[,,'refugia']  <- 0
-> #testing migration
-> RAF2 <- rot_migrate(RAF1gen, migration=0.5, coverage=0.5, verbose=TRUE)
-  m.intervention f.intervention m.refugia f.refugia
-1              1              1         0         0
-2              1              1         0         0
-3              1              1         0         0
-  m.intervention f.intervention m.refugia f.refugia
-1           0.75           0.75      0.25      0.25
-2           0.75           0.75      0.25      0.25
-3           0.75           0.75      0.25      0.25
-
-> RAF2 <- rot_migrate(RAF1gen, migration=1, coverage=0.1, verbose=TRUE)
-  m.intervention f.intervention m.refugia f.refugia
-1              1              1         0         0
-2              1              1         0         0
-3              1              1         0         0
-  m.intervention f.intervention m.refugia f.refugia
-1            0.1            0.1       0.1       0.1
-2            0.1            0.1       0.1       0.1
-3            0.1            0.1       0.1       0.1
 
@@
 
-~ maybe have a set_raf() function too
-
 ~ sort outputs which need to be time to resistance for planned rotations and RwR rotate when resistant. Sequence.

### 2019-10-15 inst/documents/rotation_runs2019.Rmd
runs now mostly working for expts C and D

@@
 #id_expt <- 'B3'
-id_expt <- 'B4'
+#id_expt <- 'B4'
+#id_expt <- 'C1'
+#id_expt <- 'C4'
+id_expt <- 'D4'
 

### 2019-10-15 inst/documents/rotations_notes.Rmd
runs now mostly working for expts C and D

@@
 
+# 14/10/19
+
+Quotes from IRM in Aedes paper :
+Dusfour2019
+Management of insecticide resistance in the major Aedes vectors of arboviruses: Advances and challenges.
+
+in blue could be disagreed with :
+
+Box 1. IRM strategies
+Rotation
+# This strategy employs temporal alternation of insecticides with different modes of action and is based on the assumption that resistance diminishes insect fitness in the absence of insecticide.
+
+[in our rotation paper we could address this misconception that fitness costs are required for rotations to be beneficial].
+
+# All comparisons with mosaic, rotation, or mixture based on theoretical models or empirical studies demonstrate that “reactive alternation” underperforms the other IRM approaches in slowing down resistance evolution because alternation is implemented too late, when resistance alleles are already at high frequency [60].
+
+In Singapore, permethrin was
+replaced by pirimiphos-methyl for thermal fogging after reduced efficacy was observed [61], yet after 9 years of interruption, high permethrin resistance persisted in field populations.
+
+Persistence of high levels of pyrethroid resistance was also observed 7 to 10 years after they were replaced by organophosphates in the state of Sao Paulo (Brazil) [62].
+As mentioned before, however, maintenance of resistance can also be due to the exposure of vector populations to household and agricultural insecticides or other xenobiotics (pollutants, other pesticides, etc.) [63–65].
+
+On the bright side, the Onchocerciasis Control Programme in West Africa is probably the most emblematic case of successful IRM in public health.
+After the first detections of temephos resistance in Simulium damnosum, a rotation scheme of insecticides with alternative modes of action was implemented for larval treatments.
+The rotation strategy drove the reversion of organophosphate resistance and preserved S.
+damnosum insecticide susceptibility for more than 20 years until the program was stopped [66].
+
+For malaria, a large-scale trial was conducted in Mexico to evaluate the impact of different IRM strategies on the resistance of Anopheles albimamus [67].
+In this study, a 3-compound annual rotation, a 2-compound mosaic, and a long-term treatment with a single insecticide were compared.
+An increase of resistance was observed in all villages, but it demonstrated that mosaic or rotation selected for low resistance levels and remained stable compared with villages with single treatment where resistance increased rapidly and to levels significantly greater [68].
+Control failures associated with pyrethroid resistance in malaria vectors have been observed for IRSs using pyrethroids.
+Switching to DDT against An.
+funestus populations without kdr mutation but specific metabolic mechanisms of pyrethroid resistance in South Africa [69] or to bendiocarb against An.
+gambiae in Equatorial Guinea restored the control of resistant populations [70].
+...
+The limited evidence-based information to support any IRM strategy in public health (e.g., mixture, rotation, noninsecticidal tools, etc.) complicates the prioritization of interventions.
+
+Despite that, we have to advocate for the introduction of IRM policies into vector control programs by favoring a proactive approach over a response mode. 
+#For example, it has been demonstrated that the preset rotation of unrelated insecticides performed better than changing insecticides over time in response to resistance [60, 66].
+
+#[next sentence is good for justifying future work]
+Until now, it has been near impossible to anticipate how resistance will impact the efficacy of vector control tools in the field and how a given IRM strategy will slow or reverse the evolution of resistance in targeted populations.
+A continuous effort connecting research institutes, control programs, industry, and WHO is required to conduct operational research that evaluates vector control strategies under field conditions and the genetics of resistance in broad terms (genotype–phenotype relationships, mosquito behavior, modeling).
+The outcomes of these studies will provide crucial information to optimize decision-making systems and the procedures to adjust IRM strategy according to field situations.
+
+#15/10/2019
+
+seeming bug with mortality switch that for rotations it stops the run when one insecticide is still below threshold.
+
+e.g D4 scenario1
+run_rot(rot_interval=10, n_insecticides=2, start_freqs=0.0018, cost=c(0.083,0.047), eff=c(0.95,0.96),
+ dom_sel=c(0.18,0.41), dom_cos=c(0.85,0.98), rr=c(0.29,0.71), expo_hi=0.69, male_expo_prop=0.17,
+ coverage=0.86, migration=0.28, min_rwr_interval=10, mort_or_freq = 'mort')
+
+but compare with freq that does work :
+
+run_rot(rot_interval=10, n_insecticides=2, start_freqs=0.0018, cost=c(0.083,0.047), eff=c(0.95,0.96),
+dom_sel=c(0.18,0.41), dom_cos=c(0.85,0.98), rr=c(0.29,0.71), expo_hi=0.69, male_expo_prop=0.17,
+coverage=0.86, migration=0.28, min_rwr_interval=10)
+
+in insecticide_check()
+
 # TODO
 see rotation_runs2019.Rmd & rotations-experiments-creating-201909.xlsx
+
+see why insecticide_check() is failing for mort
+
+~ work out what we want Sam to do
+~~ do 16 sets of 10,000 runs
+~~ what analyses on outputs
 

### 2019-10-15 R/rot_plots.r

@@ -42,6 +42,13 @@ rot_plot_resistance <- function(df_res2,
   # filter out refuge if not wanted (if coverage=1 no refuge anyway)
   #if (!plot_refuge) df_res2 <- filter(df_res2, active_or_refuge != 'refuge')
   if (!plot_refuge) df_res2 <- df_res2[df_res2$active_or_refuge=='active',]
+
+
+  # set plot_mort true for mort    
+  if ( is.null(plot_mort) ) plot_mort <- ifelse(mort_or_freq=='mort',TRUE,FALSE)
+  # set logy false for mort    
+  if ( is.null(logy) ) logy <- ifelse(mort_or_freq=='mort',FALSE,TRUE)  
+  
   
   # to allow plotting of insecticide in use
   # add column which has a value if insecticide in use & NA if not
@@ -73,6 +80,7 @@ rot_plot_resistance <- function(df_res2,
   }
 
   # testing adding mortality to the plot
+  
   # note mortality currently added in run_rot but actually we can calc from fixed conversion
   if (plot_mort) {
 

### 2019-10-16 inst/documents/rotation_runs2019.Rmd
correcting couple bugs insecticide switching

@@
 ---
+
+An rmarkdown document that can be modified and knit to run scenarios within a specified experiment.
+Output files are saved to the specified outFolder.
 
@@
 nscenarios <- 100
-plotyn <- 1 #allow turning plots on and off from here
-
-# still deciding how best to do this ...
-# infile_multi <- 'rotations-experiments-creating-201909.xlsx'
-# infile_multi <- system.file("extdata", infile_multi, package = "rotations", mustWork = TRUE)
-# inex_all <- read_excel(infile_multi)
-
-# NOTE to add new files currently need to reinstall package because infiles are in extdata
-#infile <- "expt-A2-freq-insame-costs-nodisp.csv"
-# infile <- "expt-A4-freq-insame-costs-disp.csv"
-# # get the id for the expot to include in outfilenames, specifies experiment attributes
-# id_expt <- substr(infile,6,7)
-# # read an experiment file specifying ranges
-# inex <- read_in_expt(infile, nscenarios=nscenarios)
+plotyn <- 1 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
 
@@
 
+# old ways of doing it
+# infile_multi <- 'rotations-experiments-creating-201909.xlsx'
+# infile_multi <- system.file("extdata", infile_multi, package = "rotations", mustWork = TRUE)
+# inex_all <- read_excel(infile_multi)
+
+# older way with individual experiment files
+# to add new files currently need to reinstall package because infiles are in extdata
+# infile <- "expt-A2-freq-insame-costs-nodisp.csv"
+# infile <- "expt-A4-freq-insame-costs-disp.csv"
+# # get the id for the expot to include in outfilenames, specifies experiment attributes
+# id_expt <- substr(infile,6,7)
+# # read an experiment file specifying ranges
+# inex <- read_in_expt(infile, nscenarios=nscenarios)
+
 inex <- read_in_expt(nscenarios = nscenarios, multi=TRUE, id_expt=id_expt)
@@
 # text to go at start of document
+cat("\nThere are 16 potential experiments :\n\n    ")
+
 cat("Experiment :",id_expt," scenario number is pagenum -1 \n\n    ")

### 2019-10-16 R/insecticide_check.r

@@ -41,6 +41,13 @@ insecticide_check <- function( RAF1gen,
   #check that required value
   stopifnot( mort_or_freq == 'mort' | mort_or_freq == 'freq')
   
+  # convert threshold to survival if mortality option is selected
+  # BEWARE mort-survival conversion
+  if (mort_or_freq == 'mort') 
+  {
+    threshold <- 1-threshold
+  }
+  
   change_insecticide <- FALSE
   
   # if rotate-when-resistant OR exit_rot==TRUE so rotation can be exited
@@ -59,13 +66,10 @@ insecticide_check <- function( RAF1gen,
       
     check_value <- freq 
     
-    # convert threshold to survival if mortality option is selected
-    # BEWARE mort-survival conversion
     # then check is if value is > survival same as for frequency
     if (mort_or_freq == 'mort') 
     {
       check_value <- surv  
-      threshold <- 1-threshold
     }
 
     
@@ -88,8 +92,6 @@ insecticide_check <- function( RAF1gen,
       if (diagnostics)
          message(paste0("change from insecticide",current_insecticide, " check=",check_value,"\n"))
     }
-    
-
     
     
   } 
@@ -115,14 +117,17 @@ insecticide_check <- function( RAF1gen,
       # sapply allows mort_from_resist to be applied to all
       if (mort_or_freq == 'mort' & length(other_ins_checks)>0 ) 
       {
-        #other_ins_checks <- 1 - mort_from_resist(other_ins_checks)       
+        #16/10/19 bug that this_ins_check wasn't converted to mortality
+        #also bug that mortality hadn't been converted to survival, now is above
+        this_ins_check <- 1 - mort_from_resist(this_ins_check, eff=eff, dom_sel=dom_sel, rr=rr)       
         other_ins_checks <- 1 - sapply(other_ins_checks, function(x) mort_from_resist(rfreq=x,eff=eff, dom_sel=dom_sel, rr=rr))
       }
       
       #BEWARE 4/10/2019 used to be BUG here, this bracket was after next loop leading to no change for 'freq'
       
-      # 31/7/18 only change if one to change to
-      # 4/10/19 in rotation have to force stop if this insecticide above thresh, even if no other to change to 
+      # only change if one to change to
+      # in rotation force stop if this insecticide above thresh, even if no other to change to 
+      # also in rotation stick with this insecticide if it's the only one below threshold
       if ( this_ins_check > threshold | min(other_ins_checks) <= threshold)
       {
[…]

### 2019-10-16 R/mort_from_resist.r

@@ -15,7 +15,7 @@
 #' @return total mortality single value
 #' @export
 
-mort_from_resist <- function ( rfreq, # = 0.01,
+mort_from_resist <- function ( rfreq, # = 0.01, # no defaults to avoid code bugs with it being called with wrong values
                                    eff,# = 0.9,
                                    dom_sel,# = 0.5,
                                    rr )# = 0.5,

### 2019-11-06 inst/documents/sensi_an_rotations201911.Rmd
starting plots for paper

--- 2018-07-31 inst/documents/sensi_an_rotations1.Rmd
@@
 ---
-title: "sensi_an_rotations1. Start of sensitivity analysis for rotations."
+title: "sensi_an_rotations201911. To analyse rotation runs for paper"
 output: pdf_document
 #output: word_document
-author: "Andy South and Ian Hastings"
+author: "Andy South, Sam Jones and Ian Hastings"
 date: "`r Sys.Date()`"
 ---
+
+Code to plot and analyse rotations runs for paper.
+
+Runs themselves are done from rotation_runs2019.Rmd.
+
+16 expts of 10,000 runs each comparing rotations and sequences.
+
+Code below will need to be adapted to work on all 16 experiments, so far it just works on one.
 
@@
 
-  outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
-```
-
-10,000 runs comparing rotations and sequences.
-
-```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
-#eval=FALSE just to temporarily hide this text
-
-# Starting question : is there any benefit to rotations when there are no costs and no migration to an untreated area ?
-# 
-# Test 1. No refugia (i.e. Coverage=100%) or fitness costs. Would expect all policies to be equivalent.
-# 
-# Test 2. Dominance of resistance and fitness costs both set to 0.5.
-# I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles.
-I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same.
-I’m probably wrong but it’s a nice test…..
-#
-
-#\pagebreak
+  # where the rotation runs saved
+  infolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
 ```
@@
 
-#start small & build it up
-#maybe create an equiv. of resistance::sensiAnPaperPart.r later
+# load the data, start with just one experiment
+# dataframes will need to be renamed because they will all be called df_res_all to start
+infile <- "df_res_all_rot_mig1000.rda"
+infile <- paste0(infolder, infile)
 
-# example for 1 run
-# linputs <- list(max_gen = 50, n_insecticides = 2)
-# do.call(run_rot, linputs)
-# dfinputs <- as.data.frame(linputs)
-
-# So one way of doing is to create an linputs list for each run & pass to run_rot()
-# so run_rot() only ever runs for 1 scenario
-# in contrast resistance::runModel2() accepted an input matrix & would run for multiple scenarios
-# not sure which is better
-
-# set random seed so that results can be reproduced
-#set.seed(1)
-set.seed(2)
-max_gen         <- 500
-plot_scenarios  <- FALSE
-n_scenarios     <- 10 #000 
-already_ran     <- FALSE
-#already_ran     <- TRUE
-
-# if already done runs, just load data for plots
-if (already_ran)
-{
-  #C:/rsprojects/rotations/inst/documents/
-
-  #load("df_res_all_rot10000.rda")
-  load("df_res_all_rot_mig1000.rda")
-
-}else
-{
-  df_res_all <- data.frame()
-
-  for(i in 1:n_scenarios)
-  {
-    #message should output to the R markdown console
-    if (i%%200 == 1) message("scenario ",i," of ",n_scenarios," ",Sys.time())
-
-    # inputs (could also be saved to a dataframe first & extracted one row at a time)
-    linputs <- list()
-
-    ## constant inputs ##
-    linputs$coverage <- 1
-    linputs$migration <- 0
-    #linputs$cost <- 0 #variable cost below
-    linputs$max_gen <- max_gen
-
-    #beware min_rwr_interval, it's to stop repeated small rotations
-    linputs$min_rwr_interval <- 5
-    linputs$min_gens_switch_back <- 1
-    linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
-    linputs$no_r_below_mut <- TRUE #BEWARE trying out new mutation-selection balance option
-    #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
-    linputs$plot <-     FALSE
-
-    ## variable inputs ##
-    linputs$cost <-           runif(1, min=0,   max=0.1) 
-    #linputs$cost <-           0 
-    linputs$expo_hi <-        runif(1, min=0.4, max=0.9)  #min=0.1, max=0.9
-    linputs$male_expo_prop <- runif(1, min=0,   max=1)
-    linputs$eff <-            runif(1, min=0.5, max=1) # min=0.3, max=1
-    linputs$rr  <-            runif(1, min=0.1, max=0.9)
-    linputs$dom_sel <-        runif(1, min=0,   max=1)
-    linputs$dom_cos <-        runif(1, min=0,   max=1) 
-    linputs$start_freqs <-    runif(1, min=0.005,   max=0.1) #note not log yet   
-    linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
-    # dispersal (initially with cost set to 0)
-    #linputs$coverage <-       runif(1, min=0.1, max=0.9)
-    #linputs$migration <-      runif(1, min=0.1, max=0.9)
-
-    # create null objects used later for plotting
-    ggrot <- ggseq <- NULL
-
-    # to compare sequence to rotation run all the scenarios above for both seq & rot 
-    for(rot_or_not in 0:1)
-    {  
-      #to get rot_int of 0 every other run
-      if (rot_or_not == 0) linputs$rot_interval <- 0
-      else   #random rot interval integer
-          linputs$rot_interval <-   sample( 5:50, 1 )  #beware set replace=T if more than 1
-
-      ##################
-      # run one scenario
-      dfres <- do.call(run_rot, linputs)
-
-      # summarise results per scenario
-      # note this is also currently calc at end of run_rot() here is probably better
-
-      res <- gens_under_thresh(dfres)
-
-      # to plot results for checking
-      if ( linputs$rot_interval == 0 )
-      {
-        gens_seq <- as.numeric(res)
-      } else
-      {
-        gens_rot <- as.numeric(res)
-      }
-
-      if (plot_scenarios)
-      {
-        # make plot
-        ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
-
-        # to plot results for checking
-        if ( linputs$rot_interval == 0 )
-        {
-          ggseq <- ggres
-          gens_seq <- as.numeric(res)
-        } else
-        {
-          ggrot <- ggres
-          gens_rot <- as.numeric(res)
-        }
-      }
-
-    # TODO I may want to concat all dfres into one object (adda column for scenario num) to save for later analysis
-    # e.g. if I wanted to calculate average resistance frequencies over time in addition to gens below threshold
-
-    } #end rot_or_not loop
-
-    # make an output row for each scenario with inputs, gens_rot, gens_seq
-    # need to make sure I retain the rot_interval for rot
-
-    # add scenario id column onto results
-    df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, id=i))
-    df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
-
-    # rbind this scenario inputs & results onto previous
-    # not a good speed or memory way of doing ... but a start
-    #df_res_all <- rbind(df_res_all, df_in_out)
-
-    #instead set up the dataframe to the required size in scenario1
-    if (i==1)
-    {
-      df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=n_scenarios))
-      names(df_res_all) <- names(df_in_out)
-    }
-
-    #putting results into one row of overall results dataframe
-    df_res_all[i,] <- df_in_out
-
-    if (plot_scenarios)
-    {  
-      # plot input vals
-      # extract out those between 0 & 1
-      dfi <- data.frame(t(as.data.frame(linputs)))
-      names(dfi) <- 'value'
-      dfi$inputs <- rownames(dfi)
-
-      #take out columns not between 0 & 1
-      dfi <- filter(dfi, !inputs %in% c('max_gen','rot_interval','n_insecticides','plot','min_rwr_interval'))
-      #also remove coverage & migration both set to 0 for now
-      dfi <- filter(dfi, !inputs %in% c('coverage','migration'))
-
-      ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
-                geom_point() +
-                theme_minimal() +
-                ylim(0,1) +
-                theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))
-
-      grid.arrange(ggrot, ggseq, ggins, ncol=1)
-
-      cat('scenario',i,'sum generations deployed insecticide under thresh rotation:', gens_rot,' sequence:',gens_seq,"\n\n")
-
-      #can I give the code to rerun this scenario
-      cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
-                  "cost=",signif(linputs$cost,2),", ",
-                  "start_freqs=",signif(linputs$start_freqs,2),", ", 
-                  "rot_interval=",linputs$rot_interval,", ",
-                  "eff=",signif(linputs$eff,2),",\n",
-                  "dom_sel=",signif(linputs$dom_sel,2),", ",
-                  "dom_cos=",signif(linputs$dom_cos,2),", ",
-                  "rr=",signif(linputs$rr,2),", ",
-                  "expo_hi=",signif(linputs$expo_hi,2),", ",
-                  "coverage=",signif(linputs$coverage,2),",\n",
-                  "migration=",signif(linputs$migration,2),", ",
-                  "max_gen=",linputs$max_gen,", ", 
-                  "min_rwr_interval=",linputs$min_rwr_interval,            
-                  ")\n") )
-
-    }
-
-    #I could save partial outputs every 200 scenarios
-    #just in case doesn't make to end
-    if (i%%1000 == 0)
-    {
-      #df_res_all$rot_minus_seq <- df_res_all$gens_rot - df_res_all$gens_seq
-
-      # save object containing inputs and results as rda for analysis
-      save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))    
-    }
-
-  } #end scenario loop
-
-}
+load(infile)
 
@@
 
-## mostly (~67%) no difference but infrequent where rotations better
-## ~ 7% of runs, rotation at least 20% better than sequence
-
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
@@
 #eval false because smaller runs don't have these
+#TODO change which runs to show as examples
 
@@
 
-#How does the difference between rotations and sequences respond to model inputs ?
+#Plots of how the difference between rotations and sequences respond to model inputs.
+
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
@@
 # go through selected inputs and plot them on x with rot_minus_seq on y 
-# initially with 200 reps doesn't show anything obvious, try it with 2000 overnight runs
+# TODO check whether this covers all the inputs we want to look at
 in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')
@@
 
-# PRCC
+# PRCC to show the sensitivity of the difference between rotation and sequence to inputs
+# this can only be used for the experiments where all insecticides are the same
+# (because in the ones where insecticides differ there are multiple values for each input)
 
@@
 #same as above
+#TODO check that this covers all the inputs we want to look at
 in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')
@@
 
-# Runs where rotations are >20% better (in red) are associated with higher dominance of cost and lower dominance of selection
-
-# but that does not guarantee, must be interaction with other inputs too.
+# Runs where rotations are >20% better (in red) are associated with higher dominance of cost and lower dominance of selection, but doesn't guarantee, must be interaction with other inputs too.
+# this was for earlier runs when there was some strange model behaviour, may not be the case now
+# and may not be very meaningful anyway
 
@@
 
-# looking at modelling it
+# quick, not very successful look at modelling it
 # library(mgcv)
@@
 # summary(mod_gam3)
-
-#actually decision trees may be useful
 
@@
 ```
-
-# Have a go at running for insecticides with different parameters
-# starting to work ...
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
-
-# set random seed so that results can be reproduced
-set.seed(2)
-max_gen         <- 500
-plot_scenarios  <- FALSE
-n_scenarios     <- 100 #1000 #000
-
-  #df_res_all <- data.frame()
-
-  for(i in 1:n_scenarios)
-  {
-    #message should output to the R markdown console
-    if (i%%200 == 1) message("scenario ",i," of ",n_scenarios," ",Sys.time())
-
-    # inputs (could also be saved to a dataframe first & extracted one row at a time)
-    linputs <- list()
-
-    ## constant inputs ##
-    linputs$coverage <- 1
-    linputs$migration <- 0
-    #linputs$cost <- 0 #variable cost below
-    linputs$max_gen <- max_gen
-
-    #beware min_rwr_interval, it's to stop repeated small rotations
-    linputs$min_rwr_interval <- 5
-    linputs$min_gens_switch_back <- 1
-    linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
-    linputs$no_r_below_mut <- TRUE #BEWARE trying out new mutation-selection balance option
-    #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
-    linputs$plot <-     FALSE
-
-    ## variable inputs ##
-    linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
-    n_ins <- linputs$n_insecticides
-
-    ## new those below can be different for each insecticide
-    linputs$cost <-           runif(n_ins, min=0,   max=0.1) 
-    #linputs$cost <-           0 
-    linputs$expo_hi <-        runif(n_ins, min=0.4, max=0.9)  #min=0.1, max=0.9
-    linputs$male_expo_prop <- runif(1, min=0,   max=1)
-    linputs$eff <-            runif(n_ins, min=0.5, max=1) # min=0.3, max=1
-    linputs$rr  <-            runif(n_ins, min=0.1, max=0.9)
-    linputs$dom_sel <-        runif(n_ins, min=0,   max=1)
-    linputs$dom_cos <-        runif(n_ins, min=0,   max=1) 
-    #DIFFERENT
-    linputs$start_freqs <-    runif(n_ins, min=0.0001,   max=0.01) #note not log yet
-
-    # dispersal (initially with cost set to 0)
-    #linputs$coverage <-       runif(1, min=0.1, max=0.9)
-    #linputs$migration <-      runif(1, min=0.1, max=0.9)
-
-    # create null objects used later for plotting
-    ggrot <- ggseq <- NULL
-
-    # to compare sequence to rotation run all the scenarios above for both seq & rot 
-    for(rot_or_not in 0:1)
-    {  
-      #to get rot_int of 0 every other run
-      if (rot_or_not == 0) linputs$rot_interval <- 0
-      else   #random rot interval integer
-          linputs$rot_interval <-   sample( 5:50, 1 )  #beware set replace=T if more than 1
-
-      ##################
-      # run one scenario
-      dfres <- do.call(run_rot, linputs)
-
-      # summarise results per scenario
-      # note this is also currently calc at end of run_rot() here is probably better
-
-      res <- gens_under_thresh(dfres)
-
-      # to plot results for checking
-      if ( linputs$rot_interval == 0 )
-      {
-        gens_seq <- as.numeric(res)
-      } else
-      {
-        gens_rot <- as.numeric(res)
-      }
-
-      if (plot_scenarios)
-      {
-        # make plot
-        ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
-
-        # to plot results for checking
-        if ( linputs$rot_interval == 0 )
-        {
-          ggseq <- ggres
-          gens_seq <- as.numeric(res)
-        } else
-        {
-          ggrot <- ggres
-          gens_rot <- as.numeric(res)
-        }
-      }
-    } #end rot_or_not loop
-
-    # make an output row for each scenario with inputs, gens_rot, gens_seq
-    # need to make sure I retain the rot_interval for rot
-
-    # to cope with multiple values per scenario convert them to strings
-    linputs2 <- lapply(linputs, function(x) ifelse(length(x)>1,toString(x),x))
-
-    #added scenario id column
-    df_in_out <- as.data.frame(c(linputs2, gens_rot=gens_rot, gens_seq=gens_seq, id=i), stringsAsFactors=FALSE)
-
-    df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
-
-    # rbind this scenario inputs & results onto previous
-    # not a good speed or memory way of doing ... but a start
-    #df_res_all <- rbind(df_res_all, df_in_out)
-
-    #instead set up the dataframe to the required size in scenario1
-    if (i==1)
-    {
-      df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=n_scenarios))
-      names(df_res_all) <- names(df_in_out)
-    }
-
-    #putting results into one row of overall results dataframe
-    df_res_all[i,] <- df_in_out
-
-    if (plot_scenarios)
-    {  
-      # plot input vals
-      # extract out those between 0 & 1
-      dfi <- data.frame(t(as.data.frame(linputs)))
-      names(dfi) <- 'value'
-      dfi$inputs <- rownames(dfi)
-
-      #take out columns not between 0 & 1
-      dfi <- filter(dfi, !inputs %in% c('max_gen','rot_interval','n_insecticides','plot','min_rwr_interval'))
-      #also remove coverage & migration both set to 0 for now
-      dfi <- filter(dfi, !inputs %in% c('coverage','migration'))
-
-      ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
-                geom_point() +
-                theme_minimal() +
-                ylim(0,1) +
-                theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))
-
-      grid.arrange(ggrot, ggseq, ggins, ncol=1)
-
-      cat('scenario',i,'sum generations deployed insecticide under thresh rotation:', gens_rot,' sequence:',gens_seq,"\n\n")
-
-      #can I give the code to rerun this scenario
-      cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
-                  "cost=",signif(linputs$cost,2),", ",
-                  "start_freqs=",signif(linputs$start_freqs,2),", ", 
-                  "rot_interval=",linputs$rot_interval,", ",
-                  "eff=",signif(linputs$eff,2),",\n",
-                  "dom_sel=",signif(linputs$dom_sel,2),", ",
-                  "dom_cos=",signif(linputs$dom_cos,2),", ",
-                  "rr=",signif(linputs$rr,2),", ",
-                  "expo_hi=",signif(linputs$expo_hi,2),", ",
-                  "coverage=",signif(linputs$coverage,2),",\n",
-                  "migration=",signif(linputs$migration,2),", ",
-                  "max_gen=",linputs$max_gen,", ", 
-                  "min_rwr_interval=",linputs$min_rwr_interval,            
-                  ")\n") )
-
-    }
-
-    #I could save partial outputs every 200 scenarios
-    #just in case doesn't make to end
-    if (i%%1000 == 0)
-    {
-      #df_res_all$rot_minus_seq <- df_res_all$gens_rot - df_res_all$gens_seq
-
-      # save object containing inputs and results as rda for analysis
-      save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))    
-    } 
-  } #end scenario loop
-
-#to allow plotting of chosen scenarios by re-running them 
-#initially didn't work for diff insecticides because the inputs are saved as strings 
-#now the strings get converted back to vectors below
-
-scenarios_to_plot <- c(21,23,32) #big difference ones all have insecticide1 as the good one
-scenarios_to_plot <- c(29,38,46)
-scenarios_to_plot <- c(13)
-
-for(scenario in scenarios_to_plot)
-{
-  #cat(scenario,"\n")
-  linputs <- as.list( df_res_all[scenario, 1:15] )
-
-  #converting multiple insecticide inputs back from strings
-  #have to ignore the ones that aren't strings
-  #linputs2 <- lapply(linputs, function(x) as.numeric(unlist(strsplit(as.character(x),split=", "))))
-
-  #note the space after the comma in split=", "
-  linputs <- lapply(linputs, function(x) if (!grepl(",",x)) x else
-                                          as.numeric(unlist(strsplit(as.character(x),split=", "))))
-
-  #TESTING changing min_rwr_interval from 5
-  #if min_rwr_interval is set to 0 you get frequent switch backs to other insecticides 
-  #for 1 generation
-  #but when it's set to 5it pushes frequences above the resistance freq threshold
-  linputs$min_rwr_interval <- 5
-  #min_gens_switch_back has to be 0 to allow one insecticide to continue when all others dead
-  linputs$min_gens_switch_back = 0
-
-  # run one scenario (rotation)
-  dfres <- do.call(run_rot, linputs)
-
-  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
-
-  #now run as sequence
-  linputs$rot_interval <- 0
-  dfres <- do.call(run_rot, linputs)
-
-  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
-
-  grid.arrange(ggrot, ggseq, ncol=1)
-
-}
-
-```

### 2019-11-06 inst/documents/sensi_an_rotations201911.Rmd
add new output files

@@
   library(gridExtra)
+  library(tidyverse)
 
   # where the rotation runs saved
-  infolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
+  #infolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
+  infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\SJ_Runs_Nov\\"
+
 ```
@@
 # dataframes will need to be renamed because they will all be called df_res_all to start
-infile <- "df_res_all_rot_mig1000.rda"
+
+infile <- "2019-11-05-10-42-df_res_all_rotA210000.rda"
+
 infile <- paste0(infolder, infile)

### 2019-11-11 inst/documents/sensi_an_rotations201911.Rmd
plotting 16 expts

@@
 
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
+
+# sam is running all 16 scenarios with a single outfile
+# but in the meantime can I read the rda's in individually ?
+# e.g. need to be able to find files named like this
+# 2019-11-05-10-42-df_res_all_rotA210000.rda
+# how to get all combinations of A-D and 1-4
+id_columns <- expand.grid(c("A","B","C","D"),c(1:4))
+expt_ids <- paste0(id_columns[,1], id_columns[,2])
+#i think that if I add the 10000 in I can find the unique files
+expt_ids <- paste0(expt_ids,10000)
+
+#list all files
+#list.files(infolder)
+
+#list files matching a pattern, now get just on matching each
+#list.files(infolder,"A110000")
+
+df_res_16 <- data.frame(stringsAsFactors=FALSE)
+
+# read in all 16 files, rbind and add ID columns
+for(i in 1:length(expt_ids))
+{
+
+  # gets loaded as df_res_all because that is what it was saved as
+  load(paste0(infolder, list.files(infolder, expt_ids[i])))
+  # add experiment ids, first as separate columns to make facetting easier maybe
+  df_res_all$idAD <- substr(expt_ids[i],1,1)
+  df_res_all$id14 <- substr(expt_ids[i],2,2)
+
+  df_res_16 <- rbind(df_res_16, df_res_all)
+
+}
+
+save(df_res_16, file=paste0(infolder,'df_res_16.rda'))
+
+```
+
 # proportion of scenarios not reaching resistance thresholds
@@
 
-    print( ggplot(df_res_all, aes(x=rot_minus_seq)) +
+    # print( ggplot(df_res_all, aes(x=rot_minus_seq)) + 
+    #          
+    #          geom_histogram(binwidth=10) +
+    #          xlab('rot_minus_seq : rotation generations minus sequence') +
+    #          ylab('count (sqrt scale to show low values)') +
+    #          scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+    #          theme_minimal() +
+    #          theme(panel.grid.minor.y=element_blank()))
 
-             geom_histogram(binwidth=10) +
-             xlab('rot_minus_seq : rotation generations minus sequence') +
-             ylab('count (sqrt scale to show low values)') +
-             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
+# no colour ??
+# ggplot(df_res_all, aes(x=rot_minus_seq, colour=rot_minus_seq)) + 
+#      geom_histogram(binwidth=10) + 
+#      scale_color_brewer(palette="Dark2")
+
+#GOOD setting colours so can see easier when rot or seq better
+
+widthbin <- 10 #either 10 or 20 looks fairly similar
+
+# df_res_all$best_strategy <- ifelse(df_res_all$rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_all$rot_minus_seq > (widthbin/2), "rotation", "no_difference"))
+# 
+# ggplot(df_res_all, aes(x=rot_minus_seq, fill=best_strategy)) + 
+#      geom_histogram(binwidth=widthbin, colour='white') + 
+#      theme_minimal() +
+#      theme(panel.grid.minor.y=element_blank())
+
+df_res_16$best_strategy <- ifelse(df_res_16$rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_16$rot_minus_seq > (widthbin/2), "rotation", "no_difference"))
+
+# plot all 160k runs together
+# ggplot(df_res_16, aes(x=rot_minus_seq, fill=best_strategy)) + 
+#      geom_histogram(binwidth=widthbin, colour='white') + 
+#      theme_minimal() +
+#      theme(panel.grid.minor.y=element_blank())
+
+#facet by A-D and 1-4
+ggplot(df_res_16, aes(x=rot_minus_seq, fill=best_strategy)) + 
+     geom_histogram(binwidth=widthbin, colour='white') + 
      theme_minimal() +
-             theme(panel.grid.minor.y=element_blank()))
+     xlim(-100,100) +
+     theme(panel.grid.minor.y=element_blank()) +
+     facet_grid(rows=vars(id14), cols=vars(idAD))
+
+# lengthening facet labels
+
+idAD_lookup <- c(
+  A = "A freq i-same",
+  B = "B freq i-diff",
+  C = "C mort i-same",
+  D = "B mort i-diff"
+)
+
+id14_lookup <- c(
+  "1" = "1 nocost nodisp",
+  "2" = "2 costs nodisp",
+  "3" = "3 nocost disp",
+  "4" = "4 costs disp"
+)
+
+#colrs <- c("no_difference"=rgb(210,210,210,max=255),"rotation"=rgb(113,166,89,max=255),"sequence"=rgb(137,117,202,max=255))
+colrs <- c("no_difference"=rgb(255,255,255,max=255),"rotation"=rgb(2,98,215,max=255),"sequence"=rgb(227,0,169,max=255))
+
+#facet by A-D and 1-4
+ggplot(df_res_16, aes(x=rot_minus_seq, fill=best_strategy)) + 
+     geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     xlim(-100,100) +
+     xlab("advantage of rotations over sequence in generations") +
+     theme(panel.grid.minor.y=element_blank()) +
+     facet_grid(rows=vars(id14), cols=vars(idAD), labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+# p2 + facet_grid(vore ~ conservation, labeller = labeller(
+#   .default = capitalize,
+#   conservation = conservation_status
+# ))
+
+#df_res_16 %>% group_by(id12) %>% tally()
 

### 2019-11-11 inst/documents/sensi_an_rotations201911.Rmd
more plots

@@
   infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\SJ_Runs_Nov\\"
-
-```
-
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
-#results='hide',
-#comment=NA removes ## before text output
-
-# load the data, start with just one experiment
-# dataframes will need to be renamed because they will all be called df_res_all to start
-
-infile <- "2019-11-05-10-42-df_res_all_rotA210000.rda"
-
-infile <- paste0(infolder, infile)
-
-load(infile)
 
@@
 
+# best strategy calculation
+# no_difference set to +/- half of widthbin set above
 df_res_16$best_strategy <- ifelse(df_res_16$rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_16$rot_minus_seq > (widthbin/2), "rotation", "no_difference"))
+
+df_res_16$rms_percent <- 100 * df_res_16$rot_minus_seq / df_res_16$gens_seq
 
@@
 # lengthening facet labels
-
 idAD_lookup <- c(
@@
 
-#facet by A-D and 1-4
+#rot_minus_seq, facet by A-D and 1-4, better facet labels
 ggplot(df_res_16, aes(x=rot_minus_seq, fill=best_strategy)) + 
@@
      theme(panel.grid.minor.y=element_blank()) +
-     facet_grid(rows=vars(id14), cols=vars(idAD), labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
 
-# p2 + facet_grid(vore ~ conservation, labeller = labeller(
-#   .default = capitalize,
-#   conservation = conservation_status
-# ))
+#tally
+#df_res_16 %>% group_by(id12) %>% tally()
 
-#df_res_16 %>% group_by(id12) %>% tally()
+#rms (rot_minus_seq)_pcent, facet by A-D and 1-4, better facet labels
+ggplot(df_res_16, aes(x=rms_pcent, fill=best_strategy)) + 
+     geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     xlim(-100,100) +
+     xlab("percentage advantage of rotations over sequence in generations") +
+     theme(panel.grid.minor.y=element_blank()) +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+#just as a test
+#can I plot gens_seq and gens_rot in same facet
+#but I need to make data longer to do that, currently in same columns
+
+#oooo this is interesting
+#it shows what gens_rot the best strategies occur at
+ggplot(df_res_16, aes(x=gens_rot, fill=best_strategy)) + 
+     geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     #xlim(-100,100) +
+     xlab("generations to reach thresholds for ROTATION") +
+     theme(panel.grid.minor.y=element_blank()) +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+#gens_seq the best strategies occur at
+ggplot(df_res_16, aes(x=gens_seq, fill=best_strategy)) + 
+     geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     #xlim(-100,100) +
+     xlab("generations to reach thresholds for SEQUENCE") +
+     theme(panel.grid.minor.y=element_blank()) +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
 

### 2019-11-13 inst/documents/rotations_notes.Rmd
correcting couple bugs insecticide switching
starting plots for paper
investigating strange model behaviour

@@
 
+I think I fixed this issue ...?
+
+# 6/11/2019
+
+starting to get some of the analysis code ready for Sam.
+
+taking code from sensi_an_rotations1.Rmd
+
+# 13/11/2019
+
+Why does this sequence switch back when there is no cost or dispersal ?
+
+run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0021, cost=c(0), eff=c(0.98), dom_sel=c(0.51),dom_cos=c(0.63), rr=c(0.8), expo_hi=0.47, male_expo_prop=0.01, coverage=1, migration=0,min_rwr_interval=10, diagnostics = TRUE)
+
+Using a mortality switch it now doesn't switch back :
+run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0021, cost=c(0), eff=c(0.98), dom_sel=c(0.51),dom_cos=c(0.63), rr=c(0.8), expo_hi=0.47, male_expo_prop=0.01, coverage=1, migration=0, min_rwr_interval=10, mort_or_freq = 'mort')
+
+Not due to dom_cos >0 even though cost=0.
+
+insecticide1 check=0.488229891221731 freq=0.488229891221731 surv=0.406690062481814 gens_this_insecticide=39
+generation 41
+
+insecticide1 check=0.520515732451062 freq=0.520515732451062 surv=0.431997734603674 gens_this_insecticide=40
+change from insecticide1 check=0.520515732451062
+
+generation 121, switch from insecticide 3 to 1; new frequency vs threshold = 0.488603 and 0.500000
+
+Seems like freq has declined from 0.5205 to 0.4886 while insecticide not in use despite no costs or dispersal ???
+
+ooo weird seems to go down the generation after use stopped and then stays at that constant level:
+
+gen39 i1 freq=0.455590826478509
+gen40 i1 freq=0.488229891221731
+gen41 i1 freq=0.520515732451062
+gen42 i1 freq=0.488603067950344
+gen43 i1 freq=0.488603067950344
+gen44 i1 freq=0.488603067950344
+
 # TODO
@@
 
-see why insecticide_check() is failing for mort
-
-~ work out what we want Sam to do
-~~ do 16 sets of 10,000 runs
-~~ what analyses on outputs
+1. find & fix problem with C runs see above
+2. implement annual restrictions for sequences
+
+# I think we allow sequence to exit at anytime ? This gives a kind of advantage to sequences, rotations tend to end up with more generations above threshold.
+
+Questions to answer re 16k runs :
+
+1. why with mortality switch are there more runs favouring sequence ?
 

### 2019-11-13 inst/documents/sensi_an_rotations201911.Rmd
investigating strange model behaviour

@@
 
+# calc percentage difference between rotation and sequence
+df_res_16$rms_percent <- 100 * df_res_16$rot_minus_seq / df_res_16$gens_seq
+
 # best strategy calculation
@@
 
-df_res_16$rms_percent <- 100 * df_res_16$rot_minus_seq / df_res_16$gens_seq
+# OR use a % threshold for strategies to be the same
+# still uses widthbin which now represents a pcent
+# no_difference set to +/- half of widthbin% set above
+df_res_16$best_strategy <- ifelse(df_res_16$rms_percent <= -(widthbin/2),"sequence", ifelse(df_res_16$rms_percent > (widthbin/2), "rotation", "no_difference"))
 
@@
 #rms (rot_minus_seq)_pcent, facet by A-D and 1-4, better facet labels
-ggplot(df_res_16, aes(x=rms_pcent, fill=best_strategy)) + 
+#TODO make the best_strategy threshold basedon percent too
+ggplot(df_res_16, aes(x=rms_percent, fill=best_strategy)) + 
      geom_histogram(binwidth=widthbin, colour='darkgrey') + 

### 2019-11-19 inst/documents/rotation_runs2019.Rmd
added change_interval so all changes can only hppen e.g. annually, plus changes now assessed against mean of M&F

@@
 # now specify an experiment from the multi-experiment file
+#id_expt <- 'A1'
 #id_expt <- 'A2'
@@
 #id_expt <- 'C4'
-id_expt <- 'D4'
+#id_expt <- 'D4'
 
@@
                 #"max_gen=",linputs$max_gen,", ", 
-                "min_rwr_interval=",linputs$min_rwr_interval,            
+                "change_interval=",linputs$change_interval,
+                #"min_rwr_interval=",linputs$min_rwr_interval,  
                 ")\n\n") )

### 2019-11-19 inst/documents/rotations_notes.Rmd
added change_interval so all changes can only hppen e.g. annually, plus changes now assessed against mean of M&F

@@
 
-run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0021, cost=c(0), eff=c(0.98), dom_sel=c(0.51),dom_cos=c(0.63), rr=c(0.8), expo_hi=0.47, male_expo_prop=0.01, coverage=1, migration=0,min_rwr_interval=10, diagnostics = TRUE)
+run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0021, cost=c(0), eff=c(0.98), dom_sel=c(0.51),dom_cos=c(0.63), rr=c(0.8), expo_hi=0.47, male_expo_prop=0.01, coverage=1, migration=0,min_rwr_interval=10, diagnostics = FALSE)
 
@@
 
+Ian suggested this could be because we assess frequency after selection and before random mating.
+So what could be happening is the sequence is stopped because f is above threshold, then random mating happens to take it below threshold.
+
+In which case we could just make sure that insecticide_check is called after mating.
+But was tricky to do, partly a result of us assessing just Females for insecticide switch. Now the switch is checked based on mean of M&F.
+
+New setting to make it easier to bug-check
+
+run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.1, max_gen = 50, cost=c(0), eff=c(0.98), dom_sel=c(0.51),dom_cos=c(0.63), rr=c(0.8), expo_hi=0.47, male_expo_prop=0.01, coverage=1, migration=0,min_rwr_interval=10, diagnostics = FALSE)
+
+*storedgen16 freqs i1:0.5172  i2:0.1
+gen17 freqs i1:0.4852  i2:0.1292
+gen17 i1 freq=0.485228904909616
+
+# 19/11/2019
+
+confirm that insecticide check based on mean of M&F does what I'd expect
+
+trying 100 runs of A1
+
+NO sequence still not doing what I' expect, i.e. for A1 it's going back to insecticide 1 :
+Ah probably just because I haven't built and saved ...
+
+run_rot(rot_interval=0, n_insecticides=2, start_freqs=0.077, cost=c(0), eff=c(0.98), dom_sel=c(0.83),
+dom_cos=c(0.47), rr=c(0.2), expo_hi=0.69, male_expo_prop=0.17, coverage=1, migration=0,
+min_rwr_interval=10)
+
+Currently I have min_gens_switch_back and min_rwr_interval set to 10.
+Probably I should set them to 0.
+
+Instead have an input specifiying interval on which insecticides can be changed - applying both to sequence and rotation.
+
+change_interval = 10
+
+if (gens_this_insecticide %% change_interval == 0)
+
 # TODO
@@
 1. find & fix problem with C runs see above
-2. implement annual restrictions for sequences
+2. add change_interval as an input
+3. should I set min_gens_switch_back and min_rwr_interval to 0, they should have no effect
 
 # I think we allow sequence to exit at anytime ? This gives a kind of advantage to sequences, rotations tend to end up with more generations above threshold.
+# so now we want to only allow sequences to exit at annual time points (if an option is enabled)
 

### 2019-11-19 R/run_rot.r

@@ -23,6 +23,7 @@
 #' @param cost fitness cost of RR in no insecticide, for all insecticides or individually
 #' @param fitSS fitness of SS if no insecticide, usually assumed to be 1, for all insecticides or individually
 #' @param min_rwr_interval minimum rotate-when-resistant interval to stop short switches, only used when rot_interval==0. set to 0 to have no effect.
+#' @param change_interval interval that insecticide can be changed on for both rotation and sequence, default 10 gens 1 year
 #' @param no_r_below_start to stop resistance frequencies going below starting values TRUE or FALSE, beware can cause odd results
 #' @param no_r_below_mut to stop resistance frequencies going below mutation-selection balance TRUE or FALSE
 #' @param exit_rot whether to exit rotation interval if threshold is reached
@@ -74,11 +75,12 @@ run_rot <- function(max_gen = 200,
                     rr = 0.5, #c(0.5, 0.5, 0.5),
                     cost = 0.1, #c(0,0,0),
                     fitSS = 1,
-                    min_rwr_interval = 10,
+                    min_rwr_interval = 0,
+                    change_interval = 10,
                     no_r_below_start = FALSE,
                     no_r_below_mut = FALSE,
                     exit_rot = FALSE,
-                    min_gens_switch_back = 10,
+                    min_gens_switch_back = 0,
                     
                     #inputs below not needed to run model itself
                     plot = TRUE,
@@ -186,10 +188,15 @@ run_rot <- function(max_gen = 200,
   for(gen in 2:max_gen)
     { 
     
-    #TEMP check frequency of one insecticide
-    freq1 <- round(RAF[1, 'f','intervention',gen-1],3)
-    freq2 <- round(RAF[2, 'f','intervention',gen-1],3)    
-    message(paste0("*storedgen",gen-1," freqs i1:", freq1,"  i2:", freq2))
+    
+    #TEMP check frequency of two insecticides
+    # freq1f <- round(RAF[1, 'f','intervention',gen-1],3)
+    # freq2f <- round(RAF[2, 'f','intervention',gen-1],3) 
+    # freq1m <- round(RAF[1, 'm','intervention',gen-1],3)
+    # freq2m <- round(RAF[2, 'm','intervention',gen-1],3)     
+    # message(paste0("*storedgen",gen-1," freqs i1 f,m:", freq1f,",",freq1m,
+    #                "  i2 f,m:", freq2f,",",freq2m))
+    
[…]

### 2019-11-19 R/set_run_inputs.r

@@ -41,6 +41,7 @@ set_run_inputs <- function( inex = NULL )
     linmulti$max_gen[i] <- inex$max_gen
     linmulti$min_rwr_interval[i] <- inex$min_rwr_interval
     linmulti$min_gens_switch_back[i] <- inex$min_gens_switch_back
+    linmulti$change_interval[i] <- inex$change_interval    
     linmulti$no_r_below_start[i] <- inex$no_r_below_start
     linmulti$no_r_below_mut[i] <- inex$no_r_below_mut
     linmulti$exit_rot[i] <- inex$exit_rot

### 2019-11-20 inst/documents/rotation_runs2019.Rmd
delete old data and figs

@@
 #id_expt <- 'B3'
-#id_expt <- 'B4'
+id_expt <- 'B4'
 #id_expt <- 'C1'

### 2019-11-22 inst/documents/sensi_an_rotations201911.Rmd
starting new plots for paper

@@
 
-# proportion of scenarios not reaching resistance thresholds
+# difference between rotations and sequences, all runs
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-# check proportion of scenarios not reaching thresholds by max_runs for both seq & rot
-
-    nruns_fail <- df_res_all %>% 
-                  filter(gens_seq == max_gen & gens_rot == max_gen) %>%
-                  summarise(n=n()) %>% 
-                  unlist()
-
-    cat('proportion runs not reaching thresholds : ', nruns_fail/nrow(df_res_all) )
-
-    # remove runs where thresholds not reached for seq & rot 
-    df_res_all <- df_res_all %>% 
-                  filter(gens_seq != max_gen | gens_rot != max_gen)
-
-    nruns_same <- df_res_all %>% 
-                  filter(gens_seq == gens_rot) %>%
-                  summarise(n=n()) %>% 
-                  unlist()
-
-    cat('proportion runs (that did reach threshold) where rotations and sequence the same  : ', nruns_same/nrow(df_res_all) )
-
-    nruns_rot_20pc_better <- df_res_all %>% 
-                  filter(rot_minus_seq/gens_seq >= 0.2 ) %>%
-                  summarise(n=n()) %>% 
-                  unlist()
-
-    cat('proportion runs (that did reach threshold) where rotations >20 % better than sequence  : ', nruns_rot_20pc_better/nrow(df_res_all) )
-
-```
-
-# difference between rotations and sequences across all runs
-
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-
-    # print( ggplot(df_res_all, aes(x=rot_minus_seq)) + 
-    #          
-    #          geom_histogram(binwidth=10) +
-    #          xlab('rot_minus_seq : rotation generations minus sequence') +
-    #          ylab('count (sqrt scale to show low values)') +
-    #          scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
-    #          theme_minimal() +
-    #          theme(panel.grid.minor.y=element_blank()))
-
-# no colour ??
-# ggplot(df_res_all, aes(x=rot_minus_seq, colour=rot_minus_seq)) + 
-#      geom_histogram(binwidth=10) + 
-#      scale_color_brewer(palette="Dark2")
+```{r fig-si-1-abs-diff-rot-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-# plot all 160k runs together
-# ggplot(df_res_16, aes(x=rot_minus_seq, fill=best_strategy)) + 
-#      geom_histogram(binwidth=widthbin, colour='white') + 
-#      theme_minimal() +
-#      theme(panel.grid.minor.y=element_blank())
+# lengthening facet labels
+# idAD_lookup <- c(
+#   A = "A freq i-same",
+#   B = "B freq i-diff",
+#   C = "C mort i-same",
+#   D = "B mort i-diff"
+# )
+# id14_lookup <- c(
+#   "1" = "1 nocost nodisp",
+#   "2" = "2 costs nodisp",
+#   "3" = "3 nocost disp",
+#   "4" = "4 costs disp"
+# )
 
-#facet by A-D and 1-4
-ggplot(df_res_16, aes(x=rot_minus_seq, fill=best_strategy)) + 
-     geom_histogram(binwidth=widthbin, colour='white') + 
-     theme_minimal() +
-     xlim(-100,100) +
-     theme(panel.grid.minor.y=element_blank()) +
-     facet_grid(rows=vars(id14), cols=vars(idAD))
-
-# lengthening facet labels
 idAD_lookup <- c(
-  A = "A freq i-same",
-  B = "B freq i-diff",
-  C = "C mort i-same",
-  D = "B mort i-diff"
+  A = "A frequency switch\n insecticides same",
+  B = "B frequency switch\n insecticides different",
+  C = "C mortality switch\n insecticides same",
+  D = "D mortality switch\n insecticides different"
 )
@@
 id14_lookup <- c(
-  "1" = "1 nocost nodisp",
-  "2" = "2 costs nodisp",
-  "3" = "3 nocost disp",
-  "4" = "4 costs disp"
+  "1" = "1 no fit. costs\n no dispersal",
+  "2" = "2 fit. costs\n no dispersal",
+  "3" = "3 no fit. costs\n dispersal",
+  "4" = "4 fit. costs and\n dispersal"
 )
@@
 
+```
+
+# pcent difference between rotations and sequences, all runs
+
+```{r fig1-pcent-diff-rot-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
 #rms (rot_minus_seq)_pcent, facet by A-D and 1-4, better facet labels
-#TODO make the best_strategy threshold basedon percent too
 ggplot(df_res_16, aes(x=rms_percent, fill=best_strategy)) + 
@@
      theme_minimal() +
-     xlim(-100,100) +
+     #xlim(-100,100) +
+     scale_x_continuous(limits=c(-80,80), breaks = c(-60,-40,-20,0,20,40,60)) +
      xlab("percentage advantage of rotations over sequence in generations") +
-     theme(panel.grid.minor.y=element_blank()) +
+     theme(panel.grid.minor=element_blank()) +
      facet_grid(rows=vars(id14), cols=vars(idAD),
@@
 #but I need to make data longer to do that, currently in same columns
+
+```
+
+# at each generations below threshold for rotations, which strategy was best
+
+```{r fig2-which-best-each-gens-rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
                 labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+```
+
+# at each generations below threshold for rotations, which strategy was best
+
+```{r fig3-which-best-each-gens-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 \pagebreak
+
+# proportion of scenarios not reaching resistance thresholds
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+# check proportion of scenarios not reaching thresholds by max_runs for both seq & rot
+
+    nruns_fail <- df_res_16 %>% 
+                  filter(gens_seq == max_gen & gens_rot == max_gen) %>%
+                  summarise(n=n()) %>% 
+                  unlist()
+
+    cat('proportion runs not reaching thresholds : ', nruns_fail/nrow(df_res_16) )
+
+    # remove runs where thresholds not reached for seq & rot 
+    df_res_16 <- df_res_16 %>% 
+                  filter(gens_seq != max_gen | gens_rot != max_gen)
+
+    nruns_same <- df_res_16 %>% 
+                  filter(gens_seq == gens_rot) %>%
+                  summarise(n=n()) %>% 
+                  unlist()
+
+    cat('proportion runs (that did reach threshold) where rotations and sequence the same  : ', nruns_same/nrow(df_res_16) )
+
+    nruns_rot_20pc_better <- df_res_16 %>% 
+                  filter(rot_minus_seq/gens_seq >= 0.2 ) %>%
+                  summarise(n=n()) %>% 
+                  unlist()
+
+    cat('proportion runs (that did reach threshold) where rotations >20 % better than sequence  : ', nruns_rot_20pc_better/nrow(df_res_16) )
+
+```
 
@@
     #plotlist[[plotnum]] <- 
-    print( ggplot(df_res_all, aes_string(x=i, y=y)) +
+    print( ggplot(df_res_16, aes_string(x=i, y=y)) +
 
@@
 
-    x <- df_res_all[,in_cols]
-    y <- df_res_all['rot_minus_seq']
+    x <- df_res_16[,in_cols]
+    y <- df_res_16['rot_minus_seq']
 
@@
 
-ggplot(df_res_all, aes(x=dom_sel, y=dom_cos)) + 
+ggplot(df_res_16, aes(x=dom_sel, y=dom_cos)) + 
      geom_point(shape=3, size=1, alpha=0.2, show.legend=FALSE) +
      theme_minimal() +
-     geom_point(data=filter(df_res_all, rot_minus_seq/gens_seq >= 0.3 ), shape=1, size=1, col='red', alpha=0.8)
-
-# quick, not very successful look at modelling it
-# library(mgcv)
-# mod_lm2 <- gam(rot_minus_seq ~ cost + dom_cos + dom_sel, data=df_res_all)
-# summary(mod_lm2)  
-# 
-# mod_gam2 <- gam(rot_minus_seq ~ s(cost) + s(dom_cos) + s(dom_sel) + s(rr) + s(expo_hi) + s(male_expo_prop) + s(eff) + s(start_freqs), data=df_res_all)
-# 
-# summary(mod_gam2)  
-# #Deviance explained = 19.5%
-# 
-# mod_gam3 <- gam(rot_minus_seq ~ te(cost,dom_cos,dom_sel), data=df_res_all)
-# summary(mod_gam3)
+     geom_point(data=filter(df_res_16, rot_minus_seq/gens_seq >= 0.3 ), shape=1, size=1, col='red', alpha=0.8)
 
@@
 
-    gg <- ggplot( df_res_all, aes_string(x='rot_interval',
+    gg <- ggplot( df_res_16, aes_string(x='rot_interval',
                                          y='tot_gens_dep_under50',
@@
 
-    gg <- ggplot( df_res_all, aes_string(x='eff',
+    gg <- ggplot( df_res_16, aes_string(x='eff',
                                          y='tot_gens_dep_under50') ) +
@@
 
-    gg <- ggplot( df_res_all, aes_string(x='expo_hi',
+    gg <- ggplot( df_res_16, aes_string(x='expo_hi',
                                          y='tot_gens_dep_under50') ) +
@@
 
-    gg <- ggplot( df_res_all, aes_string(x='expo_hi*eff*male_expo_prop',
+    gg <- ggplot( df_res_16, aes_string(x='expo_hi*eff*male_expo_prop',
                                          y='tot_gens_dep_under50') ) +

### 2019-11-27 inst/documents/rotations_notes.Rmd
rename

@@
 
+# 25/11/19
+
+paper from Justin :
+
+Busi 2019 Rotations and mixtures of soil-applied herbicides delay resistance
+
+From the abstract :
+Simulations indicate that without rotation it takes twice as long to select for resistance to a particular soil-applied herbicide – trifluralin – than to any other herbicide option considered.
+Relative to trifluralin-only use, simple herbicide rotation patterns have no effect in delaying resistance, whereas more complex rotation patterns can delay resistance two- or three-fold.
+Herbicide mixtures further delay resistance up to six-fold in comparison to single use or simple herbicide rotations.
+
+The model is able to represent the evolution of resistance conferred by either single or multiple genes (monogenic or polygenic resistance) using explicit individual-based genetics
+
+The PERTH model, parametrized with data obtained from a decade of experimental evolution studies,14,30–32,35,42 predicts
+that herbicide mixtures are the most robust strategy to
+minimize the risk of resistance and cross-resistance evolution to
+soil-applied herbicides.
+The relative benefits of rotations and particularly mixtures in delaying resistance were found to be consistent across a number of genetic and fitness-penalty scenarios and 100-fold difference in initial allele frequencies.
+
+Unsurprisingly, the worst ‘strategies’ were those with no herbicide rotation (1111 best, 2222, 3333).
+
+Table 2. The rank column indicates the relative efficacy
+of the strategy in delaying resistance under GS1, which is not affected by allele frequency GS3 versus GS1.
+
+Pesticide mixtures at high dosages of each pesticide and/or combinations appears to be the most effective measure at delaying
+resistance evolution because they help ensure that pests are
+treated with two different pesticides. Similarly, weeds resistant to one herbicide mode of action are likely to be killed by a another herbicide acting at a different site of action.
+
+Me :
+
+Hi Justin,
+
+Thanks for sending this on, apologies it has taken me a whole month to get around to reading it in detail.
+I've just submitted th final proofs for the windows of selection paper and have finalised the simulations for our rotations paper.
+
+It was encouraging to see some plots that looked similar to ours.
+
+However I do find this paper somewhat confusing. The title 'Rotations and mixtures of soil-applied herbicides delay resistance' suggests a benefit to rotations.
+
+Then in the abstract it says : "Relative to trifluralin-only use, simple herbicide rotation patterns have no effect in delaying resistance, whereas more complex rotation patterns can delay resistance two- or three-fold."
+
+I wasn't able to establish what the mechanistic difference was between the simple and complex rotations.
+
+They do seem to show results and justification for mixtures being beneficial.
+
+I think, looking at Table 2, that they are comparing rotations to constant use of a single herbicide.
+I don't think that they consider sequential stratgies (i.e.
+switching insecticides once resistance thresholds are reached).
+
+In our work, that we are close to submitting. We find very little difference between sequential and rotation strategies.
+
+Andy
+
+So for one 'good' herbicide, simple rotations provide no advantage.
+Complex rotations apparently can provide an advantage, but I wasn't able to establish why from their results.
+They results do seem to provide clear support and justification for the benefit of mixtures.
+
 # TODO
@@
 3. should I set min_gens_switch_back and min_rwr_interval to 0, they should have no effect
-
-# I think we allow sequence to exit at anytime ? This gives a kind of advantage to sequences, rotations tend to end up with more generations above threshold.
-# so now we want to only allow sequences to exit at annual time points (if an option is enabled)
 

### 2019-11-27 inst/documents/paper_figs_rotations201911.Rmd
working on paper figs

--- 2019-11-22 inst/documents/sensi_an_rotations201911.Rmd
@@
 ---
-title: "sensi_an_rotations201911. To analyse rotation runs for paper"
-output: pdf_document
-#output: word_document
+title: "paper_figs_rotations201911 : create paper figures from rotation runs"
+#output: pdf_document
+output: word_document
 author: "Andy South, Sam Jones and Ian Hastings"
@@
 
-Code below will need to be adapted to work on all 16 experiments, so far it just works on one.
+```{r, eval=TRUE, include=FALSE}
+#set to eval=TRUE to make figs for submission
+
+library(knitr)
+
+# options to create final publication quality figures
+# word doc summarises & figs get stored
+opts_chunk$set(dev="tiff",
+               dev.args=list(compression="lzw"),
+               dpi=300,
+               cache=FALSE,
+               fig.path='rotfigs/',
+               fig.width=8.5,
+               fig.height=8.5)
+```
 
@@
   library(tidyverse)
+  library(patchwork) #for arranging multiple plots on page
 
@@
 save(df_res_16, file=paste0(infolder,'df_res_16.rda'))
+
+```
+
+# example scenarios
+# A1 frequency switch insecticides same, no costs no disp
+
+```{r fig1-examp-run-A1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# TODO put into a reusable function to plot seq vs rot
+# TODO probably allow run_rot to return the ggplot object
+
+dfseq <- run_rot(rot_interval=0, n_insecticides=4, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+
+dfrot <- run_rot(rot_interval=10, n_insecticides=4, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE )
+ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE )
+
+ggseq <- ggseq + theme(legend.position="none")
+ggrot <- ggrot + theme(legend.position="bottom",
+                       axis.title.y = element_blank())
+
+#using patchwork
+#plot seq and rot side-by-side
+#guides=collect to put legend between plots, doesn't seem to work
+(ggseq + ggrot) + plot_layout(guides = 'collect')
+
+```
+
+# A2 frequency switch insecticides same, costs no disp
+
+```{r fig2-examp-run-A2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+#TODO replace later, quick pasted test to see effect of adding cost
+
+dfseq <- run_rot(rot_interval=0, n_insecticides=4, start_freqs=0.0026, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+
+dfrot <- run_rot(rot_interval=10, n_insecticides=4, start_freqs=0.0026, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE )
+ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE )
+
+ggseq <- ggseq + theme(legend.position="none")
+ggrot <- ggrot + theme(legend.position="bottom",
+                       axis.title.y = element_blank())
+
+#using patchwork
+#plot seq and rot side-by-side
+#guides=collect to put legend between plots, doesn't seem to work
+(ggseq + ggrot) + plot_layout(guides = 'collect')
+
+```
+
+# A3 frequency switch insecticides same, no costs, disp
+# in this example rotations have a reasonable advantage but I'm not sure why !!
+
+```{r fig3-examp-run-A2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+#TODO replace later, quick pasted test to see effect of adding dispersal
+
+dfseq <- run_rot(rot_interval=0, n_insecticides=4, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+
+dfrot <- run_rot(rot_interval=10, n_insecticides=4, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE )
+ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE )
+
+ggseq <- ggseq + theme(legend.position="none")
+ggrot <- ggrot + theme(legend.position="bottom",
+                       axis.title.y = element_blank())
+
+#using patchwork
+#plot seq and rot side-by-side
+#guides=collect to put legend between plots, doesn't seem to work
+(ggseq + ggrot) + plot_layout(guides = 'collect')
+
+```
+
+# B1 frequency switch insecticides different, no costs no disp
+
+```{r fig4-examp-run-A1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
      xlab("advantage of rotations over sequence in generations") +
-     theme(panel.grid.minor.y=element_blank()) +
+     ylab("count of randomly sampled runs") +  
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
      facet_grid(rows=vars(id14), cols=vars(idAD),
@@
      xlab("percentage advantage of rotations over sequence in generations") +
-     theme(panel.grid.minor=element_blank()) +
+     ylab("count of randomly sampled runs") +    
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
      facet_grid(rows=vars(id14), cols=vars(idAD),
@@
      xlab("generations to reach thresholds for ROTATION") +
-     theme(panel.grid.minor.y=element_blank()) +
+     ylab("count of randomly sampled runs") +    
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
      facet_grid(rows=vars(id14), cols=vars(idAD),
@@
      xlab("generations to reach thresholds for SEQUENCE") +
-     theme(panel.grid.minor.y=element_blank()) +
+     ylab("count of randomly sampled runs") +    
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
      facet_grid(rows=vars(id14), cols=vars(idAD),
@@
 
-#Plots of how the difference between rotations and sequences respond to model inputs.
+```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
+# doesn't show much
+# Plots of how the difference between rotations and sequences respond to model inputs.
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r figx-PRCC, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
+```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
+
 # Runs where rotations are >20% better (in red) are associated with higher dominance of cost and lower dominance of selection, but doesn't guarantee, must be interaction with other inputs too.
@@
 # and may not be very meaningful anyway
-
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
 

### 2019-11-27 inst/documents/rotations_notes.Rmd
working on paper figs

@@
 
+FIXED and new 16k simulations run.
+
 # 25/11/19
@@
 
-Me :
+Me : (not sent)
 
@@
 
+# 27/11/2019
+
+paper_figs_rotations201911.Rmd
+
+choose example runs for early figs
+
+ARg, this may not matter,  but I had wondered if I could directly compare runs with and without costs.
+BUT it seems the way that random samples are drawn that th values for other input change.
+
+e.g. on p6, A1 no cost eff=0.79
+A2 with cost cost=0.79, eff=0.99
+
+run_rot(rot_interval=0, n_insecticides=5, start_freqs=0.092, cost=c(0), eff=c(0.79), dom_sel=c(0.16),
+dom_cos=c(0.94), rr=c(0.19), expo_hi=0.73, male_expo_prop=0.15, coverage=1, migration=0,
+min_rwr_interval=10)
+
+run_rot(rot_interval=0, n_insecticides=5, start_freqs=0.077, cost=c(0.079), eff=c(0.99), dom_sel=c(0.5),
+dom_cos=c(0.81), rr=c(0.38), expo_hi=0.46, male_expo_prop=0.16, coverage=1, migration=0,
+change_interval=10)
+
+This may not matter for the main runs. but may mean that I want to choose example runs differently.
+
+Happens in set_run_inputs()
+
+    # select random numbers for each insecticide
+    cost <-     runif(n_rands, min=inex$cost_min,        max=inex$cost_max) 
+    eff  <-     runif(n_rands, min=inex$eff_min,         max=inex$eff_max)
+
+Interestingly, (and possibly worryingly) seems that the random deviate isn't used if min and max are the same.
+
 # TODO
+
+paper_figs_rotations201911.Rmd
 see rotation_runs2019.Rmd & rotations-experiments-creating-201909.xlsx
 
-1. find & fix problem with C runs see above
-2. add change_interval as an input
-3. should I set min_gens_switch_back and min_rwr_interval to 0, they should have no effect
+a. create a function to compare seq and rot plots
+1. check random number issues
+2. look at paper3 folder
 

### 2019-12-02 inst/documents/paper_figs_rotations201911.Rmd
figs

@@
 
-dfseq <- run_rot(rot_interval=0, n_insecticides=4, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
 min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-dfrot <- run_rot(rot_interval=10, n_insecticides=4, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
 min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
@@
 
-ggseq <- ggseq + theme(legend.position="none")
-ggrot <- ggrot + theme(legend.position="bottom",
-                       axis.title.y = element_blank())
+ggseq <- ggseq + theme(axis.title.x = element_blank()) +
+                 ggtitle("A) sequence")
+ggrot <- ggrot + theme(legend.position="none")+
+                 ggtitle("B) rotation")
 
@@
 #guides=collect to put legend between plots, doesn't seem to work
-(ggseq + ggrot) + plot_layout(guides = 'collect')
+(ggseq / ggrot) + plot_layout(guides = 'collect')
 
@@
 
-dfseq <- run_rot(rot_interval=0, n_insecticides=4, start_freqs=0.0026, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0026, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
 min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-dfrot <- run_rot(rot_interval=10, n_insecticides=4, start_freqs=0.0026, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.0026, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
 min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
@@
 
-```{r fig3-examp-run-A2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig3-examp-run-A3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-dfseq <- run_rot(rot_interval=0, n_insecticides=4, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-dfrot <- run_rot(rot_interval=10, n_insecticides=4, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
@@
 
-# B1 frequency switch insecticides different, no costs no disp
+# B2 frequency switch insecticides different, costs
 
-```{r fig4-examp-run-A1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig4-examp-run-B2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+#in sequence first insecticide keeps being used for 500 gen
+#in rotation switch to other insecticides that reach thereshold means that generations are lost over threshold
+
+dfseq <- run_rot(n_insecticides=3, start_freqs=0.093, cost=c(0.07,0.067,0.032), eff=c(0.8,0.79,0.71),
+dom_sel=c(0.23,0.68,0.38), dom_cos=c(0.89,0.65,0.55), rr=c(0.12,0.75,0.33), expo_hi=0.73, male_expo_prop=0.25, coverage=1, migration=0, min_rwr_interval=10)
 
@@
 
-```{r fig1-pcent-diff-rot-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r figx1-pcent-diff-rot-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-```{r fig2-which-best-each-gens-rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r figx2-which-best-each-gens-rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-```{r fig3-which-best-each-gens-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r figx3-which-best-each-gens-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-# Examples of runs where rotations > 20% better than sequences
-
-```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=3.5}
-#eval false because smaller runs don't have these
-#TODO change which runs to show as examples
-
-scenarios_to_plot <- c(624,2167,602,4028,5771)
-
-for(scenario in scenarios_to_plot)
-{
-  cat(scenario,"\n")
-
-  linputs <- as.list( df_res_all[scenario, 1:15] )
-
-  # run one scenario
-  dfres <- do.call(run_rot, linputs)
-
-  rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=TRUE)
-
-  #now run as sequence
-  linputs$rot_interval <- 0
-  dfres <- do.call(run_rot, linputs)
-
-  rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=TRUE)
-
-  dfi <- data.frame(t(as.data.frame(linputs)))
-  names(dfi) <- 'value'
-  dfi$inputs <- rownames(dfi)
-
-  #take out columns not between 0 & 1
-  dfi <- filter(dfi, !inputs %in% c('max_gen','rot_interval','n_insecticides','plot','min_rwr_interval'))
-  #also remove coverage & migration both set to 0 for now
-  dfi <- filter(dfi, !inputs %in% c('coverage','migration'))
-
-  ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
-            geom_point() +
-            theme_minimal() +
-            ylim(0,1) +
-            theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))
-  plot(ggins)
-
-}
-
-```
-
 ```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
@@
 
-```{r figx-PRCC, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r figxx-PRCC, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 

### 2019-12-02 inst/documents/paper_figs_rotations201911.Rmd
figs

@@
 save(df_res_16, file=paste0(infolder,'df_res_16.rda'))
+
+# filter just the 4 A expts
+df_res_A <- dplyr::filter(df_res_16, idAD=="A")
 
@@
 
-# proportion of scenarios not reaching resistance thresholds
+# count proportion of scenarios not reaching resistance thresholds
 
@@
 # (because in the ones where insecticides differ there are multiple values for each input)
+# start by running on just A expts
 
@@
 
-    x <- df_res_16[,in_cols]
-    y <- df_res_16['rot_minus_seq']
+    x <- df_res_A[,in_cols]
+    y <- df_res_A['rot_minus_seq']
 
@@
              #ggtitle(paste(strat_name[strategy_num])) +
+             ggtitle("Effect of inputs on benefit of rotations, all A expts.") +
              xlab(NULL)

### 2019-12-10 inst/documents/rotation_runs2019.Rmd
cumulativ mortality recording, correcting gens_under_thresh for mortality

@@
 #id_expt <- 'B3'
-id_expt <- 'B4'
-#id_expt <- 'C1'
+#id_expt <- 'B4'
+id_expt <- 'C1'
 #id_expt <- 'C4'
@@
     # summarise results per scenario
-    # note this is also currently calc at end of run_rot() here is probably better
-    gens <- gens_under_thresh(dfres)
+    # this calcs for whole run, at end of run_rot() it calcs per insecticide
+    if ( linputs$mort_or_freq == 'freq' ) threshold <- linputs$freq_thresh
+    else if ( linputs$mort_or_freq == 'mort' ) threshold <- linputs$mort_thresh
+
+    #gens <- gens_under_thresh(dfres=dfres, threshold=threshold, mort_or_freq=linputs$mort_or_freq)
+
+    # 10/12/19 change to recording end generations and mean mortality
+    gens <- calc_max_gens(dfres=dfres)
+    cum_mort <- cumulat_mortality(dfres=dfres)
 
     # to be able to save as for rot or seq
-    if ( linputs$rot_interval == 0 ) {gens_seq <- as.numeric(gens)}
-    else                             {gens_rot <- as.numeric(gens)}
+    if ( linputs$rot_interval == 0 ) {gens_seq <- as.numeric(gens)
+                                      mort_seq <- cum_mort/gens_seq}
+    else                             {gens_rot <- as.numeric(gens)
+                                      mort_rot <- cum_mort/gens_rot}
 
@@
   # df_in_out is one row
-  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, id=i))
-  df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
+  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, mort_rot=mort_rot, mort_seq=mort_seq, id=i))
+
+  df_in_out$gens_rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
+
+  df_in_out$mort_rot_minus_seq <- df_in_out$mort_rot - df_in_out$mort_seq
 

### 2019-12-10 inst/documents/rotations_notes.Rmd
cumulativ mortality recording, correcting gens_under_thresh for mortality

@@
 
+# 9/12/19
+
+Thinking that I can output mean % mortality over all generations and that will give us a better comparison between sequences and rotations (alongside going back to recording total generations untill all insecticides exhausted).
+
+# 10/12/19
+
+corrected recordning of gens_under_thresh for mortality even though probably not going to use anymore
+
+~ set recording of 
+i  total generations (maybe from run_rot)
+ii cumulative or mean mortality (maybe post-processed)
+
+I think better to use mean mortality because it is more obviously meaningfull, and a potential way of comaparing runs even id they have different lengths or number of insecticides.
+
 # TODO

### 2019-12-10 R/cumulat_mortality.r

@@ -1,0 +1,33 @@
+#' cumulat_mortality to calculate cumulative mortality across all generations for a simulation
+#'  just for the currently active insecticide (and not in the refuge) 
+#'
+#' @param dfres dataframe of resistance results from run_rot()
+#' 
+#' @importFrom rlang .data
+#' @return integer num generations
+#' @export
+#'
+#' @examples
+#' df_res2 <- run_rot(plot=FALSE)
+#' cumulat_mortality(df_res2)
+#' 
+
+cumulat_mortality <- function(dfres)
+{
+
+  #just for the currently active insecticide (and not in the refuge) 
+  
+  res <- dfres %>%
+    # only assess in control areas not in refugia  
+    # just for deployed insecticides 
+    dplyr::filter(.data$active_or_refuge=='active' &
+                  .data$resist_gene==paste0('insecticide',.data$insecticide)) %>%
+    #group_by(.data$resist_gene) %>%
+    summarise(cum_mort = sum(.data$mortality, na.rm=TRUE)) 
+    #%>%
+    #unlist()
+  
+  return(res$cum_mort) 
+  
+
+}

### 2019-12-10 R/run_rot.r

@@ -540,23 +540,40 @@ run_rot <- function(max_gen = 200,
   # to get active & refuge into the same subplot
   df_res2 <- tidyr::separate(df_res2, region, into=c("resist_gene","active_or_refuge"))
   
-  # calculate number generations under 50% resistance to be used in plotting 
-  # probably should be somewhere else ! ? just for active area
-  # also calculated in sensi_an_rotations1.Rmd
-  # this does give the answer, but only 1 per insecticide
+  # calculate number generations under threshold to be used in plotting 
+  # just for active area
+  # BEWARE different for frequency and mortality and probably not calculated correctly here
+  
+  # calculated elsewhere for all insecticides
+  # this gives 1 per insecticide
   # NSE problem fixed by dplyr::filter 
-  df_res2 <- df_res2 %>%
+  df_res3 <- df_res2 %>%
     dplyr::filter(.data$active_or_refuge=='active') %>%
-    group_by(.data$resist_gene) %>%
-    # for all insecticides in all generations  
-    # summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
-    # just for deployed insecticides 
+    group_by(.data$resist_gene)
+  
+  # just for deployed insecticides 
+
+  if ( mort_or_freq == 'freq' )
+  {
+    df_res3 <- df_res3 %>%
     summarise(gens_dep_under50 = sum(.data$resistance < threshold &
                                        #finds insecticide in use = this one
-                                       .data$resist_gene==paste0('insecticide',.data$insecticide), na.rm=TRUE)) %>%
-    #summarise(tot_dep_gens_under50 = sum(gens_dep_under50)) %>%    
-    dplyr::ungroup() %>%
+                                       .data$resist_gene==paste0('insecticide',.data$insecticide), na.rm=TRUE))     
+  }
+  else if ( mort_or_freq == 'mort' )
+  {
+    df_res3 <- df_res3 %>%
+      summarise(gens_dep_under50 = sum(.data$mortality > threshold &
+                                         #finds insecticide in use = this one
+                                         .data$resist_gene==paste0('insecticide',.data$insecticide), na.rm=TRUE))     
+  }
+  
+
+  df_res2 <- dplyr::ungroup(df_res3) %>%
     dplyr::left_join(df_res2, by='resist_gene')
+  
+  #add on max generations that this run reached (just a single value repeated for all rows)
[…]

### 2020-01-06 inst/documents/rotation_runs2019.Rmd
sorted 3 outputs : gens_rot_minus_seq    #total generations gens_u_rot_minus_seq  #gens under threshold mort_rot_minus_seq    #average mortality

@@
 
-nscenarios <- 100
+nscenarios <- 20
 plotyn <- 1 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
@@
 #id_expt <- 'B4'
-id_expt <- 'C1'
+#id_expt <- 'C1'
 #id_expt <- 'C4'
-#id_expt <- 'D4'
+id_expt <- 'D4'
 
@@
 
-    #gens <- gens_under_thresh(dfres=dfres, threshold=threshold, mort_or_freq=linputs$mort_or_freq)
+    gens_u <- gens_under_thresh(dfres=dfres, threshold=threshold, mort_or_freq=linputs$mort_or_freq)
 
@@
     if ( linputs$rot_interval == 0 ) {gens_seq <- as.numeric(gens)
+                                      gens_u_seq <- as.numeric(gens_u)
                                       mort_seq <- cum_mort/gens_seq}
     else                             {gens_rot <- as.numeric(gens)
+                                      gens_u_rot <- as.numeric(gens_u)
                                       mort_rot <- cum_mort/gens_rot}
@@
   # df_in_out is one row
-  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, mort_rot=mort_rot, mort_seq=mort_seq, id=i))
+  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, 
+                                        gens_u_rot=gens_u_rot, gens_u_seq=gens_u_seq,
+                                        mort_rot=mort_rot, mort_seq=mort_seq, id=i))
 
   df_in_out$gens_rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
-
+  df_in_out$gens_u_rot_minus_seq <- df_in_out$gens_u_rot - df_in_out$gens_u_seq  
   df_in_out$mort_rot_minus_seq <- df_in_out$mort_rot - df_in_out$mort_seq

### 2020-01-06 inst/documents/rotations_notes.Rmd
sorted 3 outputs : gens_rot_minus_seq    #total generations gens_u_rot_minus_seq  #gens under threshold mort_rot_minus_seq    #average mortality

@@
 
-I think better to use mean mortality because it is more obviously meaningfull, and a potential way of comaparing runs even id they have different lengths or number of insecticides.
+I think better to use mean mortality because it is more obviously meaningfull, and a potential way of comparing runs even if they have different lengths or number of insecticides.
+
+# 2020-01-06 NEW YEAR
+
+sorted 3 outputs :
+gens_rot_minus_seq    #total generations
+gens_u_rot_minus_seq  #gens under threshold
+mort_rot_minus_seq    #average mortality
 
@@
 see rotation_runs2019.Rmd & rotations-experiments-creating-201909.xlsx
+
+# in rotation_runs2019.Rmd uncomment the line below and re-add calc gens-under-thresh
+# gens <- gens_under_thresh(dfres=dfres, threshold=threshold, mort_or_freq=linputs$mort_or_freq)
+
+*check on new outputs : tot-gens, mean-mort
+
+Try some rotation_runs2019.Rmd with new outputs.
+Get Sam to run new ones.
 

### 2020-01-06 inst/documents/rotation_runs2019.Rmd
gens under thresh now working after correction to mort_thresh and freq_thresh

@@
 
-nscenarios <- 20
+nscenarios <- 200
 plotyn <- 1 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
@@
 #id_expt <- 'A4'
-#id_expt <- 'B2'
+id_expt <- 'B2'
 #id_expt <- 'B3'

### 2020-01-06 inst/documents/rotations_notes.Rmd
gens under thresh now working after correction to mort_thresh and freq_thresh

@@
 
+problem that gens_u_rot and gens_u_seq both currently 0
+
+seems to be because linputs$freq_thresh & mort thresh are both NULL
+
 # TODO
@@
 
-# in rotation_runs2019.Rmd uncomment the line below and re-add calc gens-under-thresh
-# gens <- gens_under_thresh(dfres=dfres, threshold=threshold, mort_or_freq=linputs$mort_or_freq)
-
-*check on new outputs : tot-gens, mean-mort
+#check : gens_u_rot_minus_seq
+
+currently 0 for all 200 runs of b2
 

### 2020-01-06 R/set_run_inputs.r

@@ -46,7 +46,8 @@ set_run_inputs <- function( inex = NULL )
     linmulti$no_r_below_mut[i] <- inex$no_r_below_mut
     linmulti$exit_rot[i] <- inex$exit_rot
     linmulti$mort_or_freq[i] <- inex$mort_or_freq
-    linmulti$threshold[i] <- inex$threshold 
+    linmulti$mort_thresh[i] <- inex$mort_thresh 
+    linmulti$freq_thresh[i] <- inex$freq_thresh 
     linmulti$plot[i] <- inex$plot
     
  

### 2020-01-06 inst/documents/rotation_runs2019.Rmd
testing expts

--- 2020-01-06 inst/documents/rotation_runs2019.Rmd
@@
                 "migration=",signif(linputs$migration,2),", ",
-                #"max_gen=",linputs$max_gen,", ", 
+                "max_gen=",linputs$max_gen,", ", 
                 "change_interval=",linputs$change_interval,

### 2020-01-06 inst/documents/rotations_notes.Rmd
testing expts

@@
 
+This B2 scenario 33 p34 is interesting, big benefit of rotations. One insecticide stays low in rotation seemingly due to use of others.
+
+run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.013, cost=c(0.061,0.024,0.047), eff=c(0.97,0.88,0.72),
+dom_sel=c(0.49,0.15,0.92), dom_cos=c(0.72,0.8,0.37), rr=c(0.25,0.51,0.42), expo_hi=0.52, male_expo_prop=0.92,
+coverage=1, migration=0, max_gen=500, change_interval=10)
+
 # TODO

### 2020-01-08 inst/documents/rotations_runs2020_sam.rmd
edit sams file for 160k runs to output tot_gens and mean mortality

--- 2020-01-06 inst/documents/rotation_runs2019.Rmd
@@
 ---
-title: "rotation_runs2019"
-output: pdf_document
-#output: word_document
-author: "Andy South and Ian Hastings"
+title: "rotation_runs2020_sam"
+author: "Andy South, Sam Jones and Ian Hastings"
 date: "`r Sys.Date()`"
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
 ---
 
-An rmarkdown document that can be modified and knit to run scenarios within a specified experiment.
+An rmarkdown document to run 160k runs, 10k from each of 16 experiments.
+Sam Jones runs on fast machine overnight.
 Output files are saved to the specified outFolder.
+
+rotation_runs2019.Rmd is similar but allows experiments to be run individually and to generate a pdf of results for an experiment per run.
 
@@
 
-outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
+#Sam: Andy, it appears that TinyTex is also required 
+library(tinytex)
+outFolder <- "I:/Rotations/Results_Jan/"
 ```
@@
 
-nscenarios <- 200
-plotyn <- 1 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
+Stime = date ()  
+cat("startTime", Stime, "\n")
 
-# now specify an experiment from the multi-experiment file
-#id_expt <- 'A1'
-#id_expt <- 'A2'
-#id_expt <- 'A3'
-#id_expt <- 'A4'
-#id_expt <- 'B2'
-#id_expt <- 'B3'
-#id_expt <- 'B4'
-#id_expt <- 'C1'
-#id_expt <- 'C4'
-id_expt <- 'D4'
+plotyn <- 0 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
 
-# to move the pdf of a preceeding run of this file to a saved location
-# RUN the line above and below after knitting the file
-# ASSUMING id_expt still set to previous run and still same day
-# BEWARE that id_expt will not be set correctly in home env if file was knitted
-#file.rename("rotation_runs2019.pdf",paste0(outFolder,format(Sys.time(), "%Y-%m-%d-"),id_expt,"-results-figs.pdf"))
+runList<-c('A1','A2','A3','A4','B1','B2','B3','B4','C1','C2','C3','C4','D1','D2','D3','D4')
+#runList<-'A1'
 
-# old ways of doing it
-# infile_multi <- 'rotations-experiments-creating-201909.xlsx'
-# infile_multi <- system.file("extdata", infile_multi, package = "rotations", mustWork = TRUE)
-# inex_all <- read_excel(infile_multi)
+nscenarios <- 10
 
-# older way with individual experiment files
-# to add new files currently need to reinstall package because infiles are in extdata
-# infile <- "expt-A2-freq-insame-costs-nodisp.csv"
-# infile <- "expt-A4-freq-insame-costs-disp.csv"
-# # get the id for the expot to include in outfilenames, specifies experiment attributes
-# id_expt <- substr(infile,6,7)
+for(r in runList){
+  id_expt <- r
+
 # # read an experiment file specifying ranges
 # inex <- read_in_expt(infile, nscenarios=nscenarios) 
-
 inex <- read_in_expt(nscenarios = nscenarios, multi=TRUE, id_expt=id_expt)
-
-# text to go at start of document
-cat("\nThere are 16 potential experiments :\n\n    ")
-
-cat("Experiment :",id_expt," scenario number is pagenum -1 \n\n    ")
-
-cat("Experiment key :\n    ")
-cat("A: resistance frequency threshold, insecticides the same\n    ")
-cat("B: resistance frequency threshold, different insecticides\n    ")
-cat("C: mortality threshold, insecticides the same\n    ")
-cat("D: mortality threshold, different insecticides\n    ")
-
-cat("1: no costs, no dispersal\n    ")
-cat("2: costs, no dispersal\n    ")
-cat("3: no costs, dispersal\n    ")
-cat("4: costs, dispersal\n    ")
-
-cat("\nNumbers on the right are the number of generations that this insecticide is in use and the corresponding resistance is below threshold.\n   ")
-
-cat("\nOn each page a Sequence is at the top and an annual Rotation at the bottom.")
 
@@
 linmulti <- set_run_inputs(inex=inex)
-
 # run all scenarios
@@
 
-  #record rotation rot interval locally otherwise it get soverwritten by 0
+  #record rotation rot interval locally otherwise it gets overwritten by 0
   rot_rot_interval <- linputs$rot_interval
-
-  #to add label between plots in output doc
-  #this adds label underneath, not my intention but does the job
-  cat("\n\n\\pagebreak\n    ") #needs results='asis' in the chunk options
-  #cat("scenario",i,"\n     ") stopped plots from aligning
-  #instead just remember that scenario number is pagenum -1
 
@@
   #can save partial outputs in case doesn't make to end
-  if (i==nscenarios | i%%1000 == 0)
-  {
+  #if (i==nscenarios | i%%1000 == 0)
+  #{
 
@@
     # by default saved to inst/documents folder
-    save(df_res_all, file=paste0(outFolder,format(Sys.time(), "%Y-%m-%d-%H-%M-"),'df_res_all_rot',id_expt,nscenarios,'.rda')) #paste0(outFolder,'*.rda')) 
-    # also save the experiment input file
-    save(inex, file=paste0(outFolder,format(Sys.time(), "%Y-%m-%d-%H-%M-"),'inex-',id_expt,'.rda'))   
-  }
+
+  #}
 
@@
 
+save(df_res_all, file=paste0(outFolder,'df_res_all_rot',id_expt,nscenarios,'.rda')) #paste0(outFolder,'*.rda')) 
+    # also save the experiment input file
+    save(inex, file=paste0(outFolder,format(Sys.time(), "%Y-%m-%d-%H-%M-"),'inex-',id_expt,'.rda'))
+
+Etime = date ()	
+cat("endTime", Etime, "\n")
+
+cat(Stime,Etime)
+}
 ```

### 2020-01-10 inst/documents/rotation_runs2019.Rmd
edit runs files

@@
 # BEWARE that id_expt will not be set correctly in home env if file was knitted
-#file.rename("rotation_runs2019.pdf",paste0(outFolder,format(Sys.time(), "%Y-%m-%d-"),id_expt,"-results-figs.pdf"))
+file.rename("rotation_runs2019.pdf",paste0(outFolder,format(Sys.time(), "%Y-%m-%d-"),id_expt,"-results-figs.pdf"))
 

### 2020-01-10 inst/documents/rotation_runs2020_sam.rmd
edit runs files

--- 2020-01-08 inst/documents/rotations_runs2020_sam.rmd
@@
 ---
-title: "rotation_runs2020_sam"
+title: "rotation_runs2020_sam v1"
 author: "Andy South, Sam Jones and Ian Hastings"
@@
 outFolder <- "I:/Rotations/Results_Jan/"
+#outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
 ```

### 2020-01-20 inst/documents/paper_figs_rotations201911.Rmd
working on paper figures

@@
   #infolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
-  infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\SJ_Runs_Nov\\"
+  #infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\SJ_Runs_Nov\\"
+  infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\Results_Jan\\"
 
@@
 
-# df_res_all$best_strategy <- ifelse(df_res_all$rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_all$rot_minus_seq > (widthbin/2), "rotation", "no_difference"))
+# df_res_all$best_strategy <- ifelse(df_res_all$gens_rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_all$gens_rot_minus_seq > (widthbin/2), "rotation", "no_difference"))
 # 
-# ggplot(df_res_all, aes(x=rot_minus_seq, fill=best_strategy)) + 
+# ggplot(df_res_all, aes(x=gens_rot_minus_seq, fill=best_strategy)) + 
 #      geom_histogram(binwidth=widthbin, colour='white') + 
@@
 # calc percentage difference between rotation and sequence
-df_res_16$rms_percent <- 100 * df_res_16$rot_minus_seq / df_res_16$gens_seq
+df_res_16$rms_percent <- 100 * df_res_16$gens_rot_minus_seq / df_res_16$gens_seq
 
@@
 # no_difference set to +/- half of widthbin set above
-df_res_16$best_strategy <- ifelse(df_res_16$rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_16$rot_minus_seq > (widthbin/2), "rotation", "no_difference"))
+df_res_16$best_strategy <- ifelse(df_res_16$gens_rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_16$gens_rot_minus_seq > (widthbin/2), "rotation", "no_difference"))
 
@@
 
-#rot_minus_seq, facet by A-D and 1-4, better facet labels
-ggplot(df_res_16, aes(x=rot_minus_seq, fill=best_strategy)) + 
+#gens_rot_minus_seq, facet by A-D and 1-4, better facet labels
+ggplot(df_res_16, aes(x=gens_rot_minus_seq, fill=best_strategy)) + 
      geom_histogram(binwidth=widthbin, colour='darkgrey') + 
@@
 
-#rms (rot_minus_seq)_pcent, facet by A-D and 1-4, better facet labels
+#rms (gens_rot_minus_seq)_pcent, facet by A-D and 1-4, better facet labels
 ggplot(df_res_16, aes(x=rms_percent, fill=best_strategy)) + 
@@
     nruns_rot_20pc_better <- df_res_16 %>% 
-                  filter(rot_minus_seq/gens_seq >= 0.2 ) %>%
+                  filter(gens_rot_minus_seq/gens_seq >= 0.2 ) %>%
                   summarise(n=n()) %>% 
@@
 
-# go through selected inputs and plot them on x with rot_minus_seq on y 
+# go through selected inputs and plot them on x with gens_rot_minus_seq on y 
 # TODO check whether this covers all the inputs we want to look at
@@
 
-y <- 'rot_minus_seq'
+y <- 'gens_rot_minus_seq'
 
@@
     x <- df_res_A[,in_cols]
-    y <- df_res_A['rot_minus_seq']
+    y <- df_res_A['gens_rot_minus_seq']
 
@@
      theme_minimal() +
-     geom_point(data=filter(df_res_16, rot_minus_seq/gens_seq >= 0.3 ), shape=1, size=1, col='red', alpha=0.8)
+     geom_point(data=filter(df_res_16, gens_rot_minus_seq/gens_seq >= 0.3 ), shape=1, size=1, col='red', alpha=0.8)
 
@@
 ```
+
+# new - mean mortality over whole sim
+# slightly surprising that not much variation
+# this one colours by best strategy in generation terms
+# B1 suggests that runs where rotation is better may have
+# lower mortality under rotation
+# try to find a different way of showing that 
+# (e.g. just plot gens_diff against mortality diff)
+
+```{r figxmort-rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# ** mean mortality for rotations
+
+# have a go at classifying gens_rot so I can use it to colour the mean mort bars
+# cut(gens_rot, 5) this does that
+
+ggplot(df_res_16, aes(x=mort_rot, fill = cut(gens_rot, 5))) + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
+     geom_histogram(colour='darkgrey') + 
+     #geom_histogram(binwidth=0.02, colour='darkgrey') + 
+     #be interesting to color by generations to see if strategies that last longer have lower mortality ?
+     #scale_fill_manual(values=colrs) +
+     #scale_fill_continuous(aes(gens_rot)) +
+     scale_fill_brewer(palette = 'greens') +
+     theme_minimal() +
+     #xlim(-100,100) +
+     xlab("mean mortality over all generations for rotation") +
+     ylab("count of randomly sampled runs") +  
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+```
+
+```{r figxgens_rot_v_mort_rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# ** mean mortality for rotations
+
+ggplot(df_res_16, aes(x=gens_rot, fill = cut(mort_rot, 5))) + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
+     geom_histogram(colour='darkgrey') + 
+     #geom_histogram(binwidth=0.02, colour='darkgrey') + 
+     #be interesting to color by generations to see if strategies that last longer have lower mortality ?
+     #scale_fill_manual(values=colrs) +
+     #scale_fill_continuous(aes(gens_rot)) +
+     scale_fill_brewer(palette = 'greens') +
+     theme_minimal() +
+     #xlim(-100,100) +
+     #xlab("mean mortality over all generations for rotation") +
+     #ylab("count of randomly sampled runs") +  
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+```
+
+```{r figxmort_rot_v_start_freqs, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# i expect that mean mort may be largely influenced by startfreqs, but remember when diff. there are multiple
+
+# ** mean mortality for rotations
+
+ggplot(df_res_16, aes(x=start_freqs, y=mort_rot)) + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
+     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     geom_hline(yintercept = 0, linetype="dashed", color = 'darkgrey') +
+     geom_vline(xintercept = 0, linetype="dashed", color = 'darkgrey') +
+     geom_point(shape='.') +
+     geom_smooth() +
+     #geom_abline() +
+     # scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     # #xlim(-100,100) +
+     # xlab("advantage of rotations over sequence in mean mortality") +
+     # ylab("count of randomly sampled runs") +  
+     theme(panel.grid.minor=element_blank()) +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+```
+
+```{r figxmort-which-best-meanmort, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# ** using best strategy from generations calc above
+# ** so trying to see if being better in generations terms
+# ** has consequences for mean mortality
+
+# **hmmm doesn't seem to work, all x differences 0
+# because widthbin set to 10 above
+
+#gens_rot_minus_seq, facet by A-D and 1-4, better facet labels
+ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
+     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     geom_histogram(binwidth=0.02, colour='darkgrey') + 
+     scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     #xlim(-100,100) +
+     xlab("advantage of rotations over sequence in mean mortality") +
+     ylab("count of randomly sampled runs") +  
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+#tally
+#df_res_16 %>% group_by(id12) %>% tally()
+
+```
+
+#try to establish whether
+#runs that have higher generations for rotations
+#also have lower mean mortality
+
+```{r figxmort-v-gens, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+#gens_rot_minus_seq, facet by A-D and 1-4, better facet labels
+ggplot(df_res_16, aes(x=gens_rot_minus_seq, y=mort_rot_minus_seq)) + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
+     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     geom_hline(yintercept = 0, linetype="dashed", color = 'darkgrey') +
+     geom_vline(xintercept = 0, linetype="dashed", color = 'darkgrey') +
+     geom_point(shape='.') +
+     geom_smooth(method='lm') +
+     #geom_abline() +
+     # scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     # #xlim(-100,100) +
+     # xlab("advantage of rotations over sequence in mean mortality") +
+     # ylab("count of randomly sampled runs") +  
+     theme(panel.grid.minor=element_blank()) +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+```

### 2020-01-20 inst/documents/rotation_runs2019.Rmd
working on paper figures

@@
 #id_expt <- 'A4'
-#id_expt <- 'B2'
+#id_expt <- 'B1'
+id_expt <- 'B2'
 #id_expt <- 'B3'

### 2020-01-20 inst/documents/rotations_notes.Rmd
edit runs files
working on paper figures

@@
 
+saved run file for sam to rotations_runs2020_sam.rmd
+
+# 2020-01-10
+
+sam did new runs
+
+in paper_figs_rotations201911.rmd
+~ replace rot_minus_seq with gens_rot_minus_seq
+
+# 2020-01-13
+looking at mean mortality from new runs
+
+# 2020-01-17
+
+A2 runs
+run 94 (p95) rotation better
+p109 sequence better
+
+looking for the B1 runs where rotations are a bit better than sequences
+
 # TODO
@@
 
-#check : gens_u_rot_minus_seq
-
-currently 0 for all 200 runs of b2
+** convert single run plots to output total generations rather than gens-under-thresh **
 
 Try some rotation_runs2019.Rmd with new outputs.
-Get Sam to run new ones.
 
@@
 
-~ consider renaming package to avoid conflict with CRAN rotations package insectrot rotinsect
+~ consider renaming package to avoid conflict with CRAN rotations package insectrot rotinsect rotseq rotateinsect
 
@@
 ~ improve readme.md with details of how to run simulations
-
-~ maybe just do annual rotation interval i.e. 10 ?
-
-GET FREQUENCY RUNS sorted first (easier)
-I could put the expt files all into one csv to start creating them, and divide into separate files later.
-
-~ there is other stuff I still need in sensi_an_rotations1.Rmd - e.g. post-processing of outputs
-~ could I have multiple experiments as columns in the same input file ?
-
-~ add the random seed to the experiment file (and decide where it will be set)
-
-~ investigate getting runs to name the output file with the date 
-~ and create a similarly named metadata file that records run input params
 
@@
 
-~ fix plot error when coverage==1
-
 STill if have min_rwr_interval > 0 an insecticide at very start of simulation outside the threshold will continue to be used.
@@
 ~ add in some unit-tests ?
-~ check how gens_under_thresh are counted in sensi_an_rotations1
-    summarise(gens_dep_under50 = sum(resistance < 0.5 &
-                                     #finds insecticide in use = this one
-                                     resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
-                                 summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
-                                 unlist()
-
-~ how to deal with 'overshooting' leading to more generations under 50 when rotation intervals are shorter ? 
-dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
-dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
-rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)
 
@@
 ~ can we get cost to decrease over time at high frequency as suggested in GPIRM ?
-
-gens_dep_under50 is ugly and I need to rewrite in sensi_an_rotations !
 
@@
 
-~ sort outputs which need to be time to resistance for planned rotations and RwR rotate when resistant. Sequence.
-
-~ time under resitance threshold, e.g. 50%
-
-Test 1. No refugia (i.e. Coverage=100%) or fitness costs.
-Would expect all policies to be equivalent.
-
-Test 2. Dominance of resistance and fitness costs both set to 0.5.
-I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles.
-I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same.
-I’m probably wrong but it’s a nice test…..
-
 ~ Ian & I talked about using fitness to determine changes.

### 2020-02-12 inst/documents/paper_figs_rotations201911.Rmd
paper fig edits

@@
 
+#fig1 before this is model description
+
 # example scenarios
@@
 
-```{r fig1-examp-run-A1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig2-examp-run-A1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.001, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
 min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.001, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
 min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE )
-ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE )
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE )
+ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE  )
 
@@
 
-```{r fig2-examp-run-A2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig3-examp-run-A2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0026, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.001, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
 min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.0026, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.001, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
 min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE )
-ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE )
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE  )
+ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE  )
+
+ggseq <- ggseq + theme(axis.title.x = element_blank()) +
+                 ggtitle("A) sequence")
+ggrot <- ggrot + theme(legend.position="none")+
+                 ggtitle("B) rotation")
+
+#using patchwork
+#plot seq and rot side-by-side
+#guides=collect to put legend between plots, doesn't seem to work
+(ggseq / ggrot) + plot_layout(guides = 'collect')
+
+```
+
+# C2 mortality switch insecticides same, no costs, disp
+
+```{r fig4-examp-run-C2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# mortality switch
+
+dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.001, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'mort')
+
+dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.001, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
+min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'mort')
+
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=TRUE, plot_refuge = FALSE, logy =FALSE  )
+ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=TRUE, plot_refuge = FALSE, logy =FALSE  )
 
@@
 # A3 frequency switch insecticides same, no costs, disp
-# in this example rotations have a reasonable advantage but I'm not sure why !!
 
-```{r fig3-examp-run-A3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig5-examp-run-A3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.001, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.0026, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
+dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.001, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')
 
-ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE )
-ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE )
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE, logy =FALSE )
+ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE, logy =FALSE )
 
@@
 
-```{r fig4-examp-run-B2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig6-examp-run-B2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
 #in sequence first insecticide keeps being used for 500 gen
-#in rotation switch to other insecticides that reach thereshold means that generations are lost over threshold
+#in rotation switch to other insecticides that reach threshold means that generations are lost over threshold
 
-dfseq <- run_rot(n_insecticides=3, start_freqs=0.093, cost=c(0.07,0.067,0.032), eff=c(0.8,0.79,0.71),
-dom_sel=c(0.23,0.68,0.38), dom_cos=c(0.89,0.65,0.55), rr=c(0.12,0.75,0.33), expo_hi=0.73, male_expo_prop=0.25, coverage=1, migration=0, min_rwr_interval=10)
+dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0059, cost=c(0.099,0.088,0.062), eff=c(1,0.96,0.78), dom_sel=c(0.63,0.95,0.42), dom_cos=c(0.6,0.49,0.14), rr=c(0.36,0.11,0.86), expo_hi=0.65, male_expo_prop=0.087, coverage=1, migration=0, plot=FALSE, max_gen=500, change_interval=10, mort_or_freq = 'freq')
+
+dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.0059, cost=c(0.099,0.088,0.062), eff=c(1,0.96,0.78), dom_sel=c(0.63,0.95,0.42), dom_cos=c(0.6,0.49,0.14), rr=c(0.36,0.11,0.86), expo_hi=0.65, male_expo_prop=0.087, coverage=1, migration=0, plot=FALSE, max_gen=500, change_interval=10, mort_or_freq = 'freq')
+
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE )
+ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE )
+
+ggseq <- ggseq + theme(axis.title.x = element_blank()) +
+                 ggtitle("A) sequence")
+ggrot <- ggrot + theme(legend.position="none")+
+                 ggtitle("B) rotation")
+
+#using patchwork
+#plot seq and rot side-by-side
+#guides=collect to put legend between plots, doesn't seem to work
+(ggseq / ggrot) + plot_layout(guides = 'collect')
 

### 2020-02-12 inst/documents/rotation_runs2019.Rmd
paper fig edits

--- 2020-01-10 inst/documents/rotation_runs2019.Rmd
@@
 #id_expt <- 'A4'
+#id_expt <- 'B1'
 #id_expt <- 'B2'
 #id_expt <- 'B3'
-#id_expt <- 'B4'
+id_expt <- 'B4'
 #id_expt <- 'C1'
 #id_expt <- 'C4'
-id_expt <- 'D4'
+#id_expt <- 'D4'
 

### 2020-02-12 R/rot_plots.r

@@ -4,6 +4,8 @@
 #' @param plot_refuge whether to plot refuge as well as intervention
 #' @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, also precludes switch to an insecticide.
 #' @param mort_or_freq whether threshold for insecticide change is mortality 'mort' or resistance frequency 'freq'
+#' @param mort_thresh mortality threshold for switching insecticides, default 0.9, only used if mort_or_freq is mort
+#' @param freq_thresh resistance frequency threshold for switching insecticides, default 0.5, only used if mort_or_freq is freq
 #' @param logy whether to use log scale for y axis
 #' @param add_gens_under50 whether to add a label of num generations under 50percent resistance
 #' @param df_resanother exploratory option to plot results of another scenario on the same graph
@@ -26,10 +28,11 @@
 #' 
 rot_plot_resistance <- function(df_res2,
                                 plot_refuge = TRUE,
-                                threshold = 0.5,
+                                mort_thresh = 0.9,
+                                freq_thresh = 0.5,
                                 mort_or_freq = 'mort',
                                 logy = TRUE,
-                                add_gens_under50 = TRUE,
+                                add_gens_under50 = FALSE,
                                 df_resanother = NULL,
                                 lwd = 1.5,
                                 title = 'auto', #NULL
@@ -88,13 +91,13 @@ rot_plot_resistance <- function(df_res2,
     # TODO change yintercept bit only plot one thresh
     if (mort_or_freq=='mort') 
     {
-      yintercept <- threshold 
+      yintercept <- mort_thresh 
       thresh_name <- "threshold mortality"
       thresh_col <- "green"
     }
     else 
     {
-      yintercept <- 0.9
+      yintercept <- freq_thresh
       thresh_name <- "threshold resistance"
       thresh_col <- "red"      
     }
@@ -152,10 +155,10 @@ rot_plot_resistance <- function(df_res2,
     # set y intercept for horiz line to threshold if freq, otherwise default to 0.5
     # stopped plotting this when thresh is mortality
     # BEWARE value of plot_mort too
-    if (! is.null(threshold) & mort_or_freq=='freq')
+    if (! is.null(freq_thresh) & mort_or_freq=='freq')
     {
-      if (mort_or_freq=='freq') yintercept <- threshold
-      else yintercept <- 0.5
+      if (mort_or_freq=='freq') yintercept <- mort_thresh
[…]

### 2020-02-25 inst/documents/paper_figs_rotations201911.Rmd
finish past threshold change, bughunting and roxygen bumped. passes check

@@
 
-widthbin <- 10 #either 10 or 20 looks fairly similar
+widthbin <- 10 #either 10 or 20 looks fairly similar, 5 gives gaps because of course runs are constrained to 1 year
 
@@
 df_res_16$rms_percent <- 100 * df_res_16$gens_rot_minus_seq / df_res_16$gens_seq
+
+# 2020-02-25 checking B1 scenarios that are giving weird results
+df_res_B1 <- dplyr::filter(df_res_16, idAD=='B' & id14 == 1)
+
+df_res_B1 %>% group_by(gens_rot_minus_seq) %>% tally()
+# A tibble: 7 x 2
+#   gens_rot_minus_seq     n
+#                <dbl> <int>
+# 1                -10     1
+# 2                  0  6454
+# 3                 10  1653
+# 4                 20  1207
+# 5                 30   514
+# 6                 40   150
+# 7                 50    21
+# so this seems to suggest that rotations are indeed better for B1
+
+#possibilities :
+#A something different in the runs that Sam did and the 200 example ones done for the document
+#B a problem with the document
+
+#but the B1 document looks as expected ???
+
+# TODO 2020-02-25 check whats going on here
 

### 2020-02-25 inst/documents/rotation_runs2019.Rmd
finish past threshold change, bughunting and roxygen bumped. passes check

--- 2020-01-10 inst/documents/rotation_runs2019.Rmd
@@
 
-nscenarios <- 200
+nscenarios <- 10
 plotyn <- 1 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
@@
 #id_expt <- 'A4'
+id_expt <- 'B1'
 #id_expt <- 'B2'
@@
 
-  #record rotation rot interval locally otherwise it get soverwritten by 0
+  #record rotation rot interval locally otherwise it gets overwritten by 0
   rot_rot_interval <- linputs$rot_interval

### 2020-02-25 inst/documents/rotations_notes.Rmd
finish past threshold change, bughunting and roxygen bumped. passes check

@@
 
+# 2020-02-25
+
+checking on B1 runs following feedback from Sam.
+
+B1 scenarios, (no costs, no dispersal, insecticides different).
+
+In earlier runs there was no advantage to rotation when there is no cost and dispersal. In the latest runs there is.
+
+I think this is something to do with time spent above the resistance threshold.
+You may remember that insecticides can only be changed on a yearly timestep.
+This can result in resistance going above threshold, when that happens it can result in simulations lasting longer.
+
+It seems that something is happening with this that gives the advantage to rotations. This creates a slightly strange 'dog-tooth' pattern seen in fig x3 and pasted below.
+
+But sam checked that in the example 200 runs there were just a few that had minor benefits for sequences (not rotations)
+
+I could try doing 1000 runs to check ?
+or do an equiv of fig x3 for the 200 runs.
+
+its x2 which has the dogtooth
+OR first check how the fig x2 is produced to check for possible glitches
+
+fig x1 shows that of the runs that are different they mostly show benefit of rotation.
+
+fig-si-1-abs-diff-rot-seq shows NO B1 runs with benefit of sequence.
+
+# 2020-02-25 checking B1 scenarios that are giving weird results
+df_res_B1 <- dplyr::filter(df_res_16, idAD=='B' & id14 == 1)
+
+df_res_B1 %>% group_by(gens_rot_minus_seq) %>% tally()
+# A tibble: 7 x 2
+#   gens_rot_minus_seq     n
+#                <dbl> <int>
+# 1                -10     1
+# 2                  0  6454
+# 3                 10  1653
+# 4                 20  1207
+# 5                 30   514
+# 6                 40   150
+# 7                 50    21
+# so this seems to suggest that rotations are indeed better for B1
+
+#possibilities :
+#A something different in the runs that Sam did and the 200 example ones done for the document
+#B a problem with the document
+
+#but the B1 document looks as expected ???
+
+#should run 1,2,3 etc. from the doc be expected to be the same as the ones in df_res_B1 ?
+#remember pdfs with figs are in C:\Dropbox\resistanceResults\rotations
+
+#YES for run 1 the inputs look about the same
+#EXCEPT that e.g. eff is recorded in df_res_B1 as just a single value (0.98) whereas there are 2 insecticides should be c(0.98,0.74)
+
+#also starngely the pdf seems still to be showing gens in-use under thresh (maybe I didn't get around to changing that)
+
+a) install-restart and try redoing B1 pdf for first 10 runs
+1) check on changing gens-under-thresh output for pdf
+2) check on saving multiple values when insecticides are diff
+
+run1 
+in pdf  rot:40, seq:40
+in runs rot:50, seq:40
+
+I wonder if something went wrong with Sams runs that just running 1 insecticide (but then would expect them to be just like A1 runs ??)
+
+& maybe Sams reporting on the pdf was based on the gens-under-thresh numbers which are now superceded (my mistake for not updating package before re-running pdf)
+
+Can I run the rot for run1 from the pdf, does it generate 40 as in the pdf or 50 as in the runs ?
+eek maybe something hadn't updated before Sam did the runs ?
+
+run_rot(rot_interval=10, n_insecticides=2, start_freqs=0.077, cost=c(0,0), eff=c(0.98,0.74), dom_sel=c(0.55,0.55),
+dom_cos=c(0.24,0.76), rr=c(0.77,0.47), expo_hi=0.69, male_expo_prop=0.17, coverage=1, migration=0,
+max_gen=500, change_interval=10)
+
+this gives 40 gens, same as the pdf.
+
+Now can I rerun the first 10 runs in the same way that Sam did ?
+To do that can just click on the saved rda file.
+It does give the same results as the pdf.
+
+I think I may need to ask Sam to re-run. Making sure that we've updated everything.
+
+One remaining query, why are the inputs for multiple insecticides not saved into the df_res_all object (by both mine & Sams runs) ?
+
+it must happen here :
+  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, 
+                                        gens_u_rot=gens_u_rot, gens_u_seq=gens_u_seq,
+                                        mort_rot=mort_rot, mort_seq=mort_seq, id=i))
+
+linputs has the multiple entries for eff, df_in_out has lost them
+i thought I had done something to allow for that in past ?
+
 # TODO
+
+2020-02-25 : probably get Sam to rerun the 16k runs
 

### 2020-02-25 R/rot_plots.r

@@ -2,7 +2,6 @@
 #'
 #' @param df_res2 dataframe of resistance results from run_rot()
 #' @param plot_refuge whether to plot refuge as well as intervention
-#' @param threshold trigger for change of insecticide, either resistance frequency or mortality dependent on mort_or_freq, also precludes switch to an insecticide.
 #' @param mort_or_freq whether threshold for insecticide change is mortality 'mort' or resistance frequency 'freq'
 #' @param mort_thresh mortality threshold for switching insecticides, default 0.9, only used if mort_or_freq is mort
 #' @param freq_thresh resistance frequency threshold for switching insecticides, default 0.5, only used if mort_or_freq is freq

### 2020-02-25 R/run_rot.r

@@ -581,7 +581,8 @@ run_rot <- function(max_gen = 200,
   # do the plots
   if (plot) rot_plot_resistance(df_res2, plot_refuge=plot_refuge, 
                                 logy=logy, add_gens_under50=add_gens_under50,
-                                threshold=threshold,
+                                mort_thresh = mort_thresh,
+                                freq_thresh = freq_thresh,
                                 mort_or_freq=mort_or_freq,
                                 plot_mort=plot_mort)
 

### 2020-02-26 inst/documents/rotation_runs2019.Rmd
corrected bug in recording inputs for different insecticdes

@@
 
-nscenarios <- 10
+nscenarios <- 200
 plotyn <- 1 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
@@
   linputs <- get_one_in(linmulti, scen_num=i)
+  # creating 2nd version of inputs that retains multi-insectcide strings so that can be saved in an output file as a single row
+  linputs_tosave <- get_one_in(linmulti, scen_num=i, convert_diff_insecticide_strings = FALSE)
 
@@
   # df_in_out is one row
-  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, 
+
+  #2020-02-25 beware previously when insecticides are different df_in_out ended up with multiple rows with results repeated
+  #which got unintentionally semi-corrected below
+  #df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, 
+  #now using linputs_tosave which contains unchanged strings, and had to add stringsAsFactors = FALSE
+  df_in_out <- as.data.frame(c(linputs_tosave, gens_rot=gens_rot, gens_seq=gens_seq,                                
                                         gens_u_rot=gens_u_rot, gens_u_seq=gens_u_seq,
-                                        mort_rot=mort_rot, mort_seq=mort_seq, id=i))
+                                        mort_rot=mort_rot, mort_seq=mort_seq, id=i), stringsAsFactors = FALSE)
 
@@
 
+  #2020-02-25 beware (see above) previoulsy when insecticides are different this saved just first row of df_in_out
+  #result is that correct results are saved but with just the inputs for the first insecticide
+  #this shouldn't effect the output results
+  #now corrected above by keeping the values for multiple different insecticides as a string
   #putting results into one row of overall results dataframe

### 2020-02-26 inst/documents/rotation_runs2020_sam.rmd
corrected bug in recording inputs for different insecticdes

@@
   linputs <- get_one_in(linmulti, scen_num=i)
+
+  #2020-02-25 creating 2nd version of inputs that retains multi-insectcide strings so that can be saved in an output file as a single row
+  linputs_tosave <- get_one_in(linmulti, scen_num=i, convert_diff_insecticide_strings = FALSE)
 

### 2020-02-26 inst/documents/rotations_notes.Rmd
corrected bug in recording inputs for different insecticdes

@@
 
-it must happen here :
+it must happen here in rotation_runs2019 :
   df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, 
@@
 
+(to be clear I don't think the above is source of an error, it just might influence ability to analyse runs with diff insecticides later)
+
+BUT
+e.g. when there are 4 insecticides df_in_out ends up with 4 rows ... & the outputs repeated
+THEN this bit (i think) just saves one row so the outputs will be OK
+
+df_res_all[i,] <- df_in_out
+
+It's a bit of a mess though.
+
+  #2020-02-25 BEWARE (see above) when insecticides are different this saves just first row of df_in_out
+  #result is that correct results are saved but with just the inputs for the first insecticide
+  #this shouldn't effect the output results, but I should try to correct
+  #putting results into one row of overall results dataframe
+  df_res_all[i,] <- df_in_out
+
+# this demonstrates problem
+tst <- as.data.frame(c(linputs))
+
+How did I solve it somewhere else ?
+
+solution may be to use linmulti which contains the strings with multiple insecticide values in.
+e.g. these are the effectivenesses for the first run
+linmulti$eff[1]
+[1] "0.98304248759523, 0.738747693016194"
+
+get_one_in() converts linmulti to linputs which can be used to run a simulation, but maybe I just need to take a single row from linmulti to save into the output file.
+
+Maybe add an argument to get_one_in to be able to keep the strings ?
+
+tst <- as.data.frame(c(linputs_tosave))
+
+but still remaining problem that df_res_all can't store a string so it seems to store 1 instead, e.g. for eff
+
 # TODO
@@
 2020-02-25 : probably get Sam to rerun the 16k runs
+
+first try to fix the issues above, or check they don't cause problems
+
+Try to fix so that this saves multiple insecticide inputs, e.g. as c(x,y,z)
+  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, 
+                                        gens_u_rot=gens_u_rot, gens_u_seq=gens_u_seq,
+                                        mort_rot=mort_rot, mort_seq=mort_seq, id=i))
 

### 2020-03-23 inst/documents/paper_figs_rotations201911.Rmd
correct figure error

@@
   #infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\SJ_Runs_Nov\\"
-  infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\Results_Jan\\"
+  infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\2020-02-results-sam\\"
 
@@
     # remove runs where thresholds not reached for seq & rot 
-    df_res_16 <- df_res_16 %>% 
+    df_res_16_threshreached <- df_res_16 %>% 
                   filter(gens_seq != max_gen | gens_rot != max_gen)
 
-    nruns_same <- df_res_16 %>% 
+    nruns_same <- df_res_16_threshreached %>% 
                   filter(gens_seq == gens_rot) %>%
@@
 
-    nruns_rot_20pc_better <- df_res_16 %>% 
+    nruns_rot_20pc_better <- df_res_16_threshreached %>% 
                   filter(gens_rot_minus_seq/gens_seq >= 0.2 ) %>%
@@
 
-```{r figxmort-which-best-meanmort, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r figxmort-rot-minus-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
-# ** using best strategy from generations calc above
-# ** so trying to see if being better in generations terms
-# ** has consequences for mean mortality
+# ** mean mortality difference between rot & seq
+# why do these seem not to sum to totals ?
 
-# **hmmm doesn't seem to work, all x differences 0
-# because widthbin set to 10 above
-
-#gens_rot_minus_seq, facet by A-D and 1-4, better facet labels
-ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
-#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
-     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
+ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
      geom_histogram(binwidth=0.02, colour='darkgrey') + 
-     scale_fill_manual(values=colrs) +
+     #scale_fill_manual(values=colrs) +
      theme_minimal() +
@@
 
-#tally
-#df_res_16 %>% group_by(id12) %>% tally()
+```
+
+```{r figxmort-which-best-meanmort, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# ** using best strategy from generations calc above
+# ** so trying to see if being better in generations terms
+# ** has consequences for mean mortality
+
+ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
+     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     geom_histogram(binwidth=0.02, colour='darkgrey') + 
+     scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     xlim(-0.2,0.2) +
+     xlab("advantage of rotations over sequence in mean mortality") +
+     ylab("count of randomly sampled runs") +  
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
 

### 2020-03-23 inst/documents/rotation_runs2019.Rmd
correct figure error

@@
 
-nscenarios <- 200
-plotyn <- 1 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
+nscenarios <- 10000 #200
+plotyn <- 0 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs
 
@@
 # BEWARE that id_expt will not be set correctly in home env if file was knitted
-file.rename("rotation_runs2019.pdf",paste0(outFolder,format(Sys.time(), "%Y-%m-%d-"),id_expt,"-results-figs.pdf"))
+#file.rename("rotation_runs2019.pdf",paste0(outFolder,format(Sys.time(), "%Y-%m-%d-"),id_expt,"-results-figs.pdf"))
 
@@
   #this adds label underneath, not my intention but does the job
-  cat("\n\n\\pagebreak\n    ") #needs results='asis' in the chunk options
+  if (plotyn) cat("\n\n\\pagebreak\n    ") #needs results='asis' in the chunk options
   #cat("scenario",i,"\n     ") stopped plots from aligning
@@
     #spaces between args stop the pdf from messing with words
-    cat(paste0( "run_rot(rot_interval=",linputs$rot_interval,", ", 
+    if (plotyn) cat(paste0( "run_rot(rot_interval=",linputs$rot_interval,", ", 
                 "n_insecticides=",linputs$n_insecticides,", ",

### 2020-12-14 inst/documents/combined_SI.rmd
figure tweaks

---
title: "Supplementary Information: PRCC and Classification trees"
author: "Andy South, Sam Jones and Ian Hastings"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
  df_print: paged
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
#library(rotations)
library(ggplot2)
library(gridExtra)
library(tidyr)
library(dplyr)
library(knitr)
library(kableExtra)
# where the rotation runs saved
#andy
#infolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\2020-02-results-sam\\"
#sam
#infolder <- "I:/Rotations/Results_Jan/"

```

The results shown on Figures 7 and 8 of the main text were subjected to PRCC and decision tree analysis. To recap, the analysis was as follows i.e.

Column A: each insecticide had identical properties; switching criterion was resistance allele frequency exceeding 50%
Column B: each insecticide had different properties; switching criterion was resistance allele frequency exceeding 50%
 Column C: each insecticide had identical properties; switching criterion was mortality <90%
Column D: each insecticide had different properties; switching criterion was mortality <90%

Row 1: no costs of resistance or dispersal into/out of an untreated refugia\
Row 2: costs are present; untreated areas absent\
Row 3: costs absent; refugia present of an untreated refugia\
Row 4: costs and refugia and both present\

\pagebreak

PRCC

The results are given in the graphs at the end of this SI. The wealth of data they contain meant the most convenient method of evaluating the results is to print out each page, physically arrange  them by column/row and search for consistent patterns within rows or within columns. It is also possible to compare consistency between evaluation criteria in each row/column by comparing evaluation based on operational lifespan of the insecticide armoury (red crosses) with evaluation based on mean mortality during this lifespan (blue crosses); note that lifespan data are absent in plots of A1 and C1 as each policy had identical results in these circumstances.

There may be some slight patterns (e.g. “coverage” across row 3, “effectiveness” in column C) but these may well be coincidental and there seem no clear, consistent patterns that would identify circumstances under which one policy is consistently favoured over the other.

The results given in D3 and D4 are the most important as they reflect the most realistic deployment scenarios i.e. insecticides differ in their properties, switching is based on mortality), but PRCC values are small and inconsistent both within the plots (i.e. evaluation based on operation lifespan or on mean mortality) or between analyses of D3 and D4.

Finally, note the ‘p’ value may be very low (the sample size is very large) so the effects are ‘real’ but small in magnitude and inconsistent across scenarios

\pagebreak

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
#results='hide',
#comment=NA removes ## before text output
# load the data, start with just one experiment
# dataframes will need to be renamed because they will all be called df_res_all to start

experiment_vector<-c('A1','A2','A3','A4','B1','B2','B3','B4','C1','C2','C3','C4','D1','D2','D3','D4')
#experiment_vector<-c('A1','A2','A3','A4')
experiment_files<-grep("df_res_all",list.files(infolder),value=TRUE)

#infile <- "2019-11-05-10-42-df_res_all_rotA210000.rda"
#infile <- paste0(infolder, infile)
#load(infile)

experiment_files<-paste0(infolder,experiment_files)
```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

library(ppcor)
#library(ggpubr)
# for gens

index<-0

for(e in experiment_files){
  index<-index+1

  load(e)

  # all  
  PRCC_a <- c('n_insecticides','eff','rr','dom_sel','expo_hi','male_expo_prop','start_freqs')
  # cost present
  PRCC_b <- c('cost','dom_cos')
  # refugia
  PRCC_c <- c('coverage','migration')

  names_a <- c('Number of insecticides','Effectiveness','Resistance restoration','Dominance of   resistance','Exposure', 'Male exposure','Starting frequency')
  names_b <-c('Cost of resistance', 'Dominance of cost')
  names_c <-c('Coverage','Dispersal')

  if(index==1 | index==5 | index==9 | index==13){
    in_cols<-PRCC_a
    all_names<-names_a
  }
  if(index==2 | index==6 | index==10 | index==14){
    in_cols<-c(PRCC_a,PRCC_b)
    all_names <- c(names_a,names_b)
  }
  if(index==3 | index==7 | index==11 | index==15){
    in_cols<-c(PRCC_a,PRCC_c)
    all_names <- c (names_a,names_c)
  }
  if(index==4 | index==8 | index==12 | index==16){
    in_cols<-c(PRCC_a,PRCC_b,PRCC_c)
    all_names <-c (names_a,names_b,names_c)
  }

  # filter out runs where both reach 500? 
  df_res_all <- mutate(df_res_all, pcent_rot_adv = gens_rot_minus_seq * 100/gens_seq)

  df_res_all <- mutate(df_res_all, rot_seq_adv_new_params = ifelse(gens_seq==max_gen & gens_rot==max_gen,'Neither',
                                                                   ifelse(pcent_rot_adv>0,'Rotation',
                                                                          ifelse(pcent_rot_adv<0,'Sequence', 'Equivalent'        
                                                                          ))))

  df_cost<-filter(df_res_all,rot_seq_adv_new_params != 'Neither')

  # Add ratios to df_cost
  # Ratio is seq / rot (i.e., above 1 = rot better, below 1 = seq better)

  df_cost <- rowwise(df_cost)
  df_cost<-mutate ( df_cost, gens_ratio =ifelse (

    rot_seq_adv_new_params == 'Equivalent' , 1,

    # max / min

    gens_rot / gens_seq

  ))

  df_cost<-mutate( df_cost, mort_ratio =

                      mort_rot / mort_seq

  )

  df_cost<- mutate (df_cost, adj_gen_ratio=ifelse(abs(gens_rot-gens_seq)<11,1,gens_ratio))

  df_cost<-data.frame(df_cost)

  # Incredibly convulted method required to seperate the character/ listed variables... 
  # Which are cost, eff, rr, dom_sel, dom_cos. 
  if(index== 5 | index== 6 | index== 7 | index== 8 | index== 13 | index== 14 | index== 15 | index== 16){
    # First convert to list... 
    adj_cost1 <- as.list(strsplit(df_cost$cost,","))
    # Then for each element of list, force to numeric and calculate mean...
    # can't combine these for some reason,crashes.. 
    adj_cost2 <- lapply(adj_cost1, as.numeric)
    adj_cost3 <- lapply(adj_cost2, mean)
    df_cost$cost <- as.numeric(adj_cost3)

    adj_eff1 <- as.list(strsplit(df_cost$eff,","))
    adj_eff2 <- lapply(adj_eff1, as.numeric)
    adj_eff3 <- lapply(adj_eff2, mean)
    df_cost$eff <- as.numeric(adj_eff3)

    adj_rr1 <- as.list(strsplit(df_cost$rr,","))
    adj_rr2 <- lapply(adj_rr1, as.numeric)
    adj_rr3 <- lapply(adj_rr2, mean)
    df_cost$rr <- as.numeric(adj_rr3)

    adj_dom_sel1 <- as.list(strsplit(df_cost$dom_sel,","))
    adj_dom_sel2 <- lapply(adj_dom_sel1, as.numeric)
    adj_dom_sel3 <- lapply(adj_dom_sel2, mean)
    df_cost$dom_sel <- as.numeric(adj_dom_sel3)

    adj_dom_cos1 <- as.list(strsplit(df_cost$dom_cos,","))
    adj_dom_cos2 <- lapply(adj_dom_cos1, as.numeric)
    adj_dom_cos3 <- lapply(adj_dom_cos2, mean)
    df_cost$dom_cos <- as.numeric(adj_dom_cos3)
  }

  x <- df_cost[,in_cols]

  y_gen_2 <- df_cost[c(in_cols,'adj_gen_ratio')]
  y_mort_2<- df_cost[c(in_cols,'mort_ratio')]

  gen_prcc<-pcor(y_gen_2,method="spearman")
  mort_prcc<-pcor(y_mort_2,method="spearman")

  gen_plot2 <- gen_prcc$estimate
  mort_plot2 <- mort_prcc$estimate

  gen_p <- gen_prcc$p.value
  mort_p <- mort_prcc$p.value

  gen_p_format <- gen_p[1:(nrow(gen_p)-1),ncol(gen_p)]
  mort_p_format <- mort_p[1:(nrow(gen_p)-1),ncol(mort_p)]

  gen_estimate <- gen_plot2[1:(nrow(gen_plot2)-1),ncol(gen_plot2)]
  mort_estimate <- mort_plot2[1:(nrow(gen_plot2)-1),ncol(mort_plot2)]

  # Estimates are now all correct. Need to get them into a single data-frame (both_plot with columns inputs, PRCC-gen and PRCC_mort as below).

  # in_cols should automatically scale, use that to create df.. 
  both_plot<-data.frame(in_cols)
  both_plot$PRCC_gen<-gen_estimate
  both_plot$PRCC_mort<-mort_estimate
  names(both_plot)[1] <-"inputs"
  both_plot[1] <- all_names

  gg <- ggplot() +

    geom_point(aes(x=inputs,y=PRCC_gen,colour="Generations"),data=both_plot) +
    geom_point(aes(x=inputs,y=PRCC_mort,colour="Mortality"),data=both_plot)+

    scale_x_discrete()+
    theme_bw()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5),legend.title=element_blank()) +
    geom_hline(yintercept = 0, linetype=3) +
    geom_hline(yintercept = 0.3, linetype=3) +
    geom_hline(yintercept = -0.3, linetype=3) +
    ylim(-1,1)+ 
    scale_y_continuous(name="PRCC estimate",breaks=c(-1,-0.5,-0.3,0,0.3,0.5,1))+
    xlab("Parameters") +
    coord_cartesian(ylim=c(-1,1))+
    ggtitle(paste("Scenario",experiment_vector[index]))

  plot(gg)

  names(gen_p_format)<-all_names
  names(mort_p_format)<-all_names

  print("P values for generations")
  print(gen_p_format)

  print("P values for mortality")
  print(mort_p_format)

} # end e loop

```

\pagebreak

Classification trees

Classification trees were run on the same datasets, using the R package Rpart. When evaluated on generations, those runs with a difference of 10 generations or less were regarded as “equivalent”. A significant subset of runs evaluated on mortality were also equivalent as they had the same overall mortality under both policies (as might be expected if both policies reached the same endpoint but by different routes). The trees are interpreted as follows:
•	The numbers under each box give the numbers of runs in the box which were equivalent, favoured rotations, favours sequences, in that order.
•	The percentage given below each box is the percentage of runs that fall into that box

As for PRCC, the only reasonable way to look for consistency seems to require printing the graphs, spreading them on the floor in columns/rows as in Figs 7&8 in main ms and look for consistent patterns across row and down columns. There was a tendency for low exposure to favour rotations in Column A when evaluated on mean moratility but in general there was no consistent patterns.

As noted above, scenario A3 and A4 reflect the most realistic deployment scenarios but there was little consistence y between their trees.

Even when classification did occur they classes did not consistent identify parameter spaces in which one policy was clearly favoured. For example, Scenario D4 evaluated on mortality was relatively simple and identified parameter spaces in which sequences, rotations or equivalence was most likely. However, the numbers beneath each box show that although one policy or equivalence tended to be favoured, there were a large number of runs in the classification box which were contrary to the favoured policy/equivalence. For example, rotations was the favoured policy in the bottom right box of this figure with 134 run favouring rotations but the classification box still contained 68 runs where the policies were equivalent and 91 runs where sequences were favoured.

\pagebreak

```{r tree, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, comment=NA}

library(rpart)
library(rpart.plot)

index<-0

for(e in experiment_files){

  index<-index+1

  #print(paste("Scenario", experiment_vector[index]))

  load(e)

  # all  
  PRCC_a <- c('n_insecticides','eff','rr','dom_sel','expo_hi','male_expo_prop','st art_freqs')
  # cost present
  PRCC_b <- c('cost','dom_cos')
  # refugia
  PRCC_c <- c('coverage','migration')

  names_a <- c('Number of insecticides','Effectiveness','Resistance     restoration','Dominance of   resistance','Exposure', 'Male exposure','Starting frequency')
  names_b <-c('Cost of resistance', 'Dominance of cost')
  names_c <-c('Coverage','Dispersal')

  if(index==1 | index==5 | index==9 | index==13){
    in_cols<-PRCC_a
    all_names<-names_a
  }
  if(index==2 | index==6 | index==10 | index==14){
    in_cols<-c(PRCC_a,PRCC_b)
    all_names <- c(names_a,names_b)
  }
  if(index==3 | index==7 | index==11 | index==15){
    in_cols<-c(PRCC_a,PRCC_c)
    all_names <- c (names_a,names_c)
  }
  if(index==4 | index==8 | index==12 | index==16){
    in_cols<-c(PRCC_a,PRCC_b,PRCC_c)
    all_names <-c (names_a,names_b,names_c)
  }

  # filter out runs where both reach 500? 
  df_res_all <- mutate(df_res_all, pcent_rot_adv = gens_rot_minus_seq * 100/gens_seq)

  df_res_all <- mutate(df_res_all, rot_seq_adv_new_params = ifelse(gens_seq==max_gen & gens_rot==max_gen,'Neither',
                                                                   ifelse(pcent_rot_adv>0,'Rotation',
                                                                          ifelse(pcent_rot_adv<0,'Sequence', 'Equivalent'        
                                                                          ))))

  df_cost<-filter(df_res_all,rot_seq_adv_new_params != 'Neither')

  # Add ratios to df_cost
  # Ratio is seq / rot (i.e., above 1 = rot better, below 1 = seq better)
  # Apply would be better but changes the d.f into a tibble

  df_cost <- rowwise(df_cost)
  df_cost<-mutate ( df_cost, gens_ratio =ifelse (

    rot_seq_adv_new_params == 'Equivalent' , 1,

    # max / min

   gens_rot / gens_seq

  ))

  df_cost<-mutate ( df_cost, mort_ratio =

                      mort_rot / mort_seq

  )

  df_cost<- mutate (df_cost, adj_gen_ratio=ifelse(abs(gens_rot-gens_seq)<11,1,gens_ratio))

  df_cost<-data.frame(df_cost)

  # New mutated params to give gens advantage as discrete variable... 
  df_cost<- mutate (df_cost, final_gens_ratio=ifelse(
    gens_ratio == 1, 'Equivalent', ifelse(gens_ratio>1, 'Rotation','Sequence')))

  # And for mortality.. 
  df_cost<- mutate (df_cost, final_mort_ratio=ifelse(
    mort_ratio == 1, 'Equivalent', ifelse(mort_ratio>1, 'Rotation','Sequence')))

  # Incredibly convulted method required to seperate the character/ listed variables... 
  # Which are cost, eff, rr, dom_sel, dom_cos. 
  if(index== 5 | index== 6 | index== 7 | index== 8 | index== 13 | index== 14 | index== 15 | index== 16){
  # First convert to list... 
  adj_cost1 <- as.list(strsplit(df_cost$cost,","))
  # Then for each element of list, force to numeric and calculate mean...
  # can't combine these for some reason,crashes.. 
  adj_cost2 <- lapply(adj_cost1, as.numeric)
  adj_cost3 <- lapply(adj_cost2, mean)
  df_cost$cost <- as.numeric(adj_cost3)

  adj_eff1 <- as.list(strsplit(df_cost$eff,","))
  adj_eff2 <- lapply(adj_eff1, as.numeric)
  adj_eff3 <- lapply(adj_eff2, mean)
  df_cost$eff <- as.numeric(adj_eff3)

  adj_rr1 <- as.list(strsplit(df_cost$rr,","))
  adj_rr2 <- lapply(adj_rr1, as.numeric)
  adj_rr3 <- lapply(adj_rr2, mean)
  df_cost$rr <- as.numeric(adj_rr3)

  adj_dom_sel1 <- as.list(strsplit(df_cost$dom_sel,","))
  adj_dom_sel2 <- lapply(adj_dom_sel1, as.numeric)
  adj_dom_sel3 <- lapply(adj_dom_sel2, mean)
  df_cost$dom_sel <- as.numeric(adj_dom_sel3)

  adj_dom_cos1 <- as.list(strsplit(df_cost$dom_cos,","))
  adj_dom_cos2 <- lapply(adj_dom_cos1, as.numeric)
  adj_dom_cos3 <- lapply(adj_dom_cos2, mean)
  df_cost$dom_cos <- as.numeric(adj_dom_cos3)
  }

# this seems to be best way of doing it

if(min(df_cost[,"adj_gen_ratio"]==1) && max(df_cost[,"adj_gen_ratio"]==1)){
print(paste("Scenario", experiment_vector[index],"Generations tree cannot run"))
}

else{
  # Names a
   if(index==1 | index==5 | index==9 | index==13){  
    tree <- rpart::rpart("final_gens_ratio ~ n_insecticides + eff + rr +dom_sel + expo_hi + male_expo_prop + start_freqs", data = df_cost, method = 'class')

prp(tree, extra=101, type=3, varlen = 0, under=TRUE, 
    cex=0.6, round=1, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= " ,main=paste("Scenario",experiment_vector[index],"generations decision tree"))

   }
  # names a, b
  if(index==2 | index==6 | index==10 | index==14){
       tree <- rpart::rpart("final_gens_ratio ~ n_insecticides + eff + rr +dom_sel + expo_hi + male_expo_prop + start_freqs + cost + dom_cos", data = df_cost, method = 'class')

prp(tree, extra=101, type=3, varlen = 0, under=TRUE, 
    cex=0.6, round=1, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= " ,main=paste("Scenario",experiment_vector[index],"generations decision tree"))

  }

  # names a, c
  if(index==3 | index==7 | index==11 | index==15){
    tree <- rpart::rpart("final_gens_ratio ~ n_insecticides + eff + rr +dom_sel + expo_hi + male_expo_prop + start_freqs + coverage + migration", data = df_cost, method = 'class')

 prp(tree, extra=101, type=3, varlen = 0, under=TRUE, 
    cex=0.6, round=1, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= ",main=paste("Scenario",experiment_vector[index],"generations decision tree") )

  }
  # names abc
  if(index==4 | index==8 | index==12 | index==16){
    tree <- rpart::rpart("final_gens_ratio ~ n_insecticides + eff + rr +dom_sel + expo_hi + male_expo_prop + start_freqs + cost + dom_cos + coverage + migration", data = df_cost, method = 'class')

prp(tree, extra=101, type=3, varlen = 0, under=TRUE, 
    cex=0.6, round=1, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= ",main=paste("Scenario",experiment_vector[index],"generations decision tree") )

  }

}

   if(index==1 | index==5 | index==9 | index==13){  
    tree <- rpart::rpart("final_mort_ratio ~ n_insecticides + eff + rr +dom_sel + expo_hi + male_expo_prop +     start_freqs", data = df_cost, method = 'class')

    prp(tree, extra=101, type=3, varlen = 0, under=TRUE, 
    cex=0.6, round=1, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= " ,main=paste("Scenario",experiment_vector[index],"mortality decision tree"))

   }

  # names a, b
  if(index==2 | index==6 | index==10 | index==14){
       tree <- rpart::rpart("final_mort_ratio ~ n_insecticides + eff + rr +dom_sel + expo_hi + male_expo_prop + start_freqs + cost + dom_cos", data = df_cost, method = 'class')

    prp(tree, extra=101, type=3, varlen = 0, under=TRUE, 
    cex=0.6, round=1, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= " ,main=paste("Scenario",experiment_vector[index],"mortality decision tree"))

  }

  # names a, c
  if(index==3 | index==7 | index==11 | index==15){
    tree <- rpart::rpart("final_mort_ratio ~ n_insecticides + eff + rr +dom_sel + expo_hi + male_expo_prop + start_freqs + coverage + migration", data = df_cost, method = 'class')

    prp(tree, extra=101, type=3, varlen = 0, under=TRUE, 
    cex=0.6, round=1, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= " ,main=paste("Scenario",experiment_vector[index],"mortality decision tree"))

  }

  # names abc
  if(index==4 | index==8 | index==12 | index==16){
    tree <- rpart::rpart("final_mort_ratio ~ n_insecticides + eff + rr +dom_sel + expo_hi + male_expo_prop + start_freqs + cost + dom_cos + coverage + migration", data = df_cost, method = 'class')

    prp(tree, extra=101, type=3, varlen = 0, under=TRUE, 
    cex=0.6, round=1, split.font=1, ycompress = TRUE,
    eq="\n= ", lt="\n< ", ge="\n>= " ,main=paste("Scenario",experiment_vector[index],"mortality decision tree"))

  }

}
```

\pagebreak

### 2020-12-14 inst/documents/combined_SI.tex
figure tweaks

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplementary Information: PRCC and Classification trees},
  pdfauthor={Andy South, Sam Jones and Ian Hastings},
  pdfborder={0 0 0},
  breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-2}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Supplementary Information: PRCC and Classification trees}
\author{Andy South, Sam Jones and Ian Hastings}
\date{2020-05-19}

\begin{document}
\maketitle

The results shown on Figures 7 and 8 of the main text were subjected to
PRCC and decision tree analysis. To recap, the analysis was as follows
i.e.

Column A: each insecticide had identical properties; switching criterion
was resistance allele frequency exceeding 50\% Column B: each
insecticide had different properties; switching criterion was resistance
allele frequency exceeding 50\% Column C: each insecticide had identical
properties; switching criterion was mortality \textless{}90\% Column D:
each insecticide had different properties; switching criterion was
mortality \textless{}90\%

Row 1: no costs of resistance or dispersal into/out of an untreated
refugia\\
Row 2: costs are present; untreated areas absent\\
Row 3: costs absent; refugia present of an untreated refugia\\
Row 4: costs and refugia and both present\\

\pagebreak

PRCC

The results are given in the graphs at the end of this SI. The wealth of
data they contain meant the most convenient method of evaluating the
results is to print out each page, physically arrange them by column/row
and search for consistent patterns within rows or within columns. It is
also possible to compare consistency between evaluation criteria in each
row/column by comparing evaluation based on operational lifespan of the
insecticide armoury (red crosses) with evaluation based on mean
mortality during this lifespan (blue crosses); note that lifespan data
are absent in plots of A1 and C1 as each policy had identical results in
these circumstances.

There may be some slight patterns (e.g.~``coverage'' across row 3,
``effectiveness'' in column C) but these may well be coincidental and
there seem no clear, consistent patterns that would identify
circumstances under which one policy is consistently favoured over the
other.

The results given in D3 and D4 are the most important as they reflect
the most realistic deployment scenarios i.e.~insecticides differ in
their properties, switching is based on mortality), but PRCC values are
small and inconsistent both within the plots (i.e.~evaluation based on
operation lifespan or on mean mortality) or between analyses of D3 and
D4.

Finally, note the `p' value may be very low (the sample size is very
large) so the effects are `real' but small in magnitude and inconsistent
across scenarios

\pagebreak

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-1.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##                       NaN                       NaN                       NaN 
## Dominance of   resistance                  Exposure             Male exposure 
##                       NaN                       NaN                       NaN 
##        Starting frequency 
##                       NaN 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.279497e-05              4.132992e-09             2.839076e-271 
## Dominance of   resistance                  Exposure             Male exposure 
##             5.796504e-134              1.597436e-10              0.000000e+00 
##        Starting frequency 
##              1.962586e-28
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-2.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.431292e-10              6.866538e-02              7.066883e-02 
## Dominance of   resistance                  Exposure             Male exposure 
##             3.410289e-101              2.987057e-03              5.003460e-05 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              2.009100e-04              4.302566e-01             2.149941e-139 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##             6.256138e-103              6.819722e-58              8.500737e-01 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.204792e-72             1.070692e-190              1.688966e-70 
##        Starting frequency        Cost of resistance         Dominance of cost 
##             2.058805e-124             6.201080e-282              8.977118e-09
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-3.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.040190e-07              8.847019e-15              1.544334e-37 
## Dominance of   resistance                  Exposure             Male exposure 
##             1.590978e-248              3.102177e-38              5.230439e-10 
##        Starting frequency                  Coverage                 Dispersal 
##              2.436760e-98             3.112878e-287              2.048830e-04 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              1.736910e-28              2.438733e-12             4.530652e-206 
## Dominance of   resistance                  Exposure             Male exposure 
##              2.569953e-12              2.352183e-40              1.960913e-21 
##        Starting frequency                  Coverage                 Dispersal 
##              1.982413e-36              0.000000e+00              1.339645e-39
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-4.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              5.964994e-04              9.664349e-04              9.390170e-06 
## Dominance of   resistance                  Exposure             Male exposure 
##             1.232597e-103              1.023258e-13              8.119258e-01 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.217907e-04              9.860850e-01              2.460829e-40 
##                  Coverage                 Dispersal 
##              5.473628e-25              2.552781e-01 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              3.420232e-47              1.749062e-30              1.830962e-38 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.030718e-35              2.399366e-60              5.004423e-23 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              7.936141e-59              4.081867e-44              9.055371e-11 
##                  Coverage                 Dispersal 
##              1.993505e-78              2.970499e-01
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-5.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##                       NaN                       NaN                       NaN 
## Dominance of   resistance                  Exposure             Male exposure 
##                       NaN                       NaN                       NaN 
##        Starting frequency 
##                       NaN 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.527251e-51              8.903486e-01             2.042072e-319 
## Dominance of   resistance                  Exposure             Male exposure 
##             6.806321e-253              5.600780e-26              0.000000e+00 
##        Starting frequency 
##              4.181653e-01
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-6.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##             3.445174e-107              2.530469e-04              5.221642e-06 
## Dominance of   resistance                  Exposure             Male exposure 
##              3.489091e-51              2.074492e-14              3.151796e-05 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.292085e-03              2.661099e-52              2.604509e-16 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              7.385269e-43              6.990974e-01              2.016125e-13 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.221363e-02              3.963614e-04              9.717417e-01 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.914308e-04              3.310871e-13              3.074197e-06
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-7.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.621072e-44              1.190614e-03              3.178174e-11 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.321768e-59              4.243877e-35              5.891333e-06 
##        Starting frequency                  Coverage                 Dispersal 
##              1.081656e-78             7.213789e-200              1.513218e-64 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.922073e-25              2.194519e-01             4.420165e-108 
## Dominance of   resistance                  Exposure             Male exposure 
##              4.123880e-05              8.941555e-01              1.221618e-06 
##        Starting frequency                  Coverage                 Dispersal 
##              9.166678e-07             6.332364e-237              1.705509e-02
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-8.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              3.091741e-52              9.523047e-03              5.946935e-11 
## Dominance of   resistance                  Exposure             Male exposure 
##              9.285600e-44              2.692993e-09              7.693886e-02 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              3.210231e-04              1.814614e-25              4.636453e-06 
##                  Coverage                 Dispersal 
##              1.513121e-15              4.841748e-02 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.574257e-16              1.705323e-01              3.938262e-17 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.015848e-02              6.186243e-01              4.336015e-02 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              3.418948e-04              1.734081e-05              1.140939e-07 
##                  Coverage                 Dispersal 
##              2.918941e-03              1.555878e-02
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-9.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##                       NaN                       NaN                       NaN 
## Dominance of   resistance                  Exposure             Male exposure 
##                       NaN                       NaN                       NaN 
##        Starting frequency 
##                       NaN 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              5.536947e-04              0.000000e+00              0.000000e+00 
## Dominance of   resistance                  Exposure             Male exposure 
##              0.000000e+00             2.891056e-217             9.843181e-174 
##        Starting frequency 
##             1.976263e-323
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-10.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              6.613208e-02              9.126458e-01              2.107584e-03 
## Dominance of   resistance                  Exposure             Male exposure 
##              9.689983e-50              2.363406e-03              1.312496e-03 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              2.947348e-03              7.551858e-01              1.364583e-63 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              1.901525e-24             3.162762e-177             1.467414e-309 
## Dominance of   resistance                  Exposure             Male exposure 
##              0.000000e+00             9.265694e-195              1.352644e-33 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              0.000000e+00              1.442660e-47              1.341460e-20
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-11.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              1.324957e-06              3.097366e-17              5.149826e-91 
## Dominance of   resistance                  Exposure             Male exposure 
##             6.974852e-181              2.268646e-56              1.139877e-05 
##        Starting frequency                  Coverage                 Dispersal 
##             9.857767e-129             1.725897e-225              2.580508e-05 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.962626e-02             2.666263e-161             2.505642e-144 
## Dominance of   resistance                  Exposure             Male exposure 
##             4.098326e-179              9.857408e-70              9.155748e-18 
##        Starting frequency                  Coverage                 Dispersal 
##             1.562930e-131              5.436548e-24              7.645491e-02
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-12.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.121770e-02              4.043657e-05              2.535163e-26 
## Dominance of   resistance                  Exposure             Male exposure 
##             7.777280e-145              5.690995e-33              3.447412e-03 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              7.900528e-53              3.439374e-02              1.318700e-25 
##                  Coverage                 Dispersal 
##              2.240292e-66              1.772746e-02 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.831730e-05             2.018771e-175             1.446793e-136 
## Dominance of   resistance                  Exposure             Male exposure 
##             2.082715e-232             5.502352e-109              3.536926e-11 
##        Starting frequency        Cost of resistance         Dominance of cost 
##             4.439919e-225              3.034425e-04              1.179055e-07 
##                  Coverage                 Dispersal 
##              1.230631e-03              4.419448e-01
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-13.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##                       NaN                       NaN                       NaN 
## Dominance of   resistance                  Exposure             Male exposure 
##                       NaN                       NaN                       NaN 
##        Starting frequency 
##                       NaN 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              6.242765e-52             1.633581e-122             1.296183e-124 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.561774e-79             7.049997e-117             9.553510e-153 
##        Starting frequency 
##             1.035562e-320
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-14.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              8.200503e-55              1.443596e-06              1.441594e-10 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.201341e-51              2.943605e-09              1.116616e-02 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              8.701257e-09              5.228213e-14              1.053013e-05 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              1.325812e-44              7.853500e-01              2.859244e-09 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.031419e-05              2.370490e-08              8.829040e-01 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.159202e-15              3.963494e-11              3.563846e-01
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-15.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              7.434045e-29              2.582068e-08              9.351446e-18 
## Dominance of   resistance                  Exposure             Male exposure 
##              6.528407e-38              6.660790e-23              1.426250e-05 
##        Starting frequency                  Coverage                 Dispersal 
##              9.760751e-81             1.531881e-125              9.664162e-29 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.154050e-15              2.993334e-32              5.412344e-13 
## Dominance of   resistance                  Exposure             Male exposure 
##              3.481640e-18              9.576041e-07              2.529119e-02 
##        Starting frequency                  Coverage                 Dispersal 
##              2.029387e-80              2.767899e-01              6.645464e-05
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-16.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              7.191772e-48              2.713605e-14              8.983791e-21 
## Dominance of   resistance                  Exposure             Male exposure 
##              4.919087e-67              2.272904e-13              1.101868e-03 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.097601e-54              2.038884e-08              1.484036e-09 
##                  Coverage                 Dispersal 
##              2.970703e-27              6.020441e-01 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.143110e-29              3.130298e-02              1.966945e-03 
## Dominance of   resistance                  Exposure             Male exposure 
##              2.879619e-05              1.735534e-01              2.238551e-01 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              4.868563e-15              8.983808e-07              6.133150e-01 
##                  Coverage                 Dispersal 
##              2.946694e-01              5.137393e-01
\end{verbatim}

\pagebreak

Classification trees

Classification trees were run on the same datasets, using the R package
Rpart. When evaluated on generations, those runs with a difference of 10
generations or less were regarded as ``equivalent''. A significant
subset of runs evaluated on mortality were also equivalent as they had
the same overall mortality under both policies (as might be expected if
both policies reached the same endpoint but by different routes). The
trees are interpreted as follows: • The numbers under each box give the
numbers of runs in the box which were equivalent, favoured rotations,
favours sequences, in that order. • The percentage given below each box
is the percentage of runs that fall into that box

As for PRCC, the only reasonable way to look for consistency seems to
require printing the graphs, spreading them on the floor in columns/rows
as in Figs 7\&8 in main ms and look for consistent patterns across row
and down columns. There was a tendency for low exposure to favour
rotations in Column A when evaluated on mean moratility but in general
there was no consistent patterns.

As noted above, scenario A3 and A4 reflect the most realistic deployment
scenarios but there was little consistence y between their trees.

Even when classification did occur they classes did not consistent
identify parameter spaces in which one policy was clearly favoured. For
example, Scenario D4 evaluated on mortality was relatively simple and
identified parameter spaces in which sequences, rotations or equivalence
was most likely. However, the numbers beneath each box show that
although one policy or equivalence tended to be favoured, there were a
large number of runs in the classification box which were contrary to
the favoured policy/equivalence. For example, rotations was the favoured
policy in the bottom right box of this figure with 134 run favouring
rotations but the classification box still contained 68 runs where the
policies were equivalent and 91 runs where sequences were favoured.

\pagebreak

\begin{verbatim}
[1] "Scenario A1 Generations tree cannot run"
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/tree-1.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-2.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-3.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-4.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-5.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-6.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-7.pdf}

\begin{verbatim}
[1] "Scenario B1 Generations tree cannot run"
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/tree-8.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-9.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-10.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-11.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-12.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-13.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-14.pdf}

\begin{verbatim}
[1] "Scenario C1 Generations tree cannot run"
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/tree-15.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-16.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-17.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-18.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-19.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-20.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-21.pdf}

\begin{verbatim}
[1] "Scenario D1 Generations tree cannot run"
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/tree-22.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-23.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-24.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-25.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-26.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-27.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-28.pdf}

\pagebreak

\end{document}

### 2020-12-14 inst/documents/paper_figs_rotations202010.Rmd
figure tweaks

--- 2020-03-23 inst/documents/paper_figs_rotations201911.Rmd
@@
 ---
-title: "paper_figs_rotations201911 : create paper figures from rotation runs"
+title: "paper_figs_rotations202010 : create paper figures from rotation runs"
 #output: pdf_document
@@
 
-```{r, eval=TRUE, include=FALSE}
+```{r setup, eval=TRUE, include=FALSE}
 #set to eval=TRUE to make figs for submission
@@
                fig.height=8.5)
-```
 
-```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
   library(rotations)
@@
 
-```{r fig2-examp-run-A1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig4-examp-run-A1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
-# TODO put into a reusable function to plot seq vs rot
-# TODO probably allow run_rot to return the ggplot object
+#2020-10 was fig2
 
@@
 
-```{r fig3-examp-run-A2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig5-examp-run-A2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
-#TODO replace later, quick pasted test to see effect of adding cost
+#2020-10 was fig3
 
@@
 
-```{r fig4-examp-run-C2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig6-examp-run-C2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
 # mortality switch
+
+#2020-10 was fig4
 
@@
 
-ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=TRUE, plot_refuge = FALSE, logy =FALSE  )
+ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=TRUE, plot_refuge = FALSE, logy =FALSE, legend_title = '' )
 ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=TRUE, plot_refuge = FALSE, logy =FALSE  )
@@
 
-```{r fig5-examp-run-A3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig7-examp-run-A3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 
-#TODO replace later, quick pasted test to see effect of adding dispersal
+#2020-10 was fig5
 
@@
 
-```{r fig6-examp-run-B2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig8-examp-run-B2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+#2020-10 was fig6
 
@@
 
-# calc percentage difference between rotation and sequence
+# calc percentage difference between rotation and sequence for resistance & mortality
 df_res_16$rms_percent <- 100 * df_res_16$gens_rot_minus_seq / df_res_16$gens_seq
+df_res_16$rms_mort_percent <- 100 * df_res_16$mort_rot_minus_seq / df_res_16$mort_seq
+#seems problem with above try Ians way
+df_res_16$rms_mort_percent <- 100 * (df_res_16$mort_rot / df_res_16$mort_seq -1)
+# Ian said this is equivalent and it does give nearly same result 10pow-14 difference
+#df_res_16$ian_rms_percent <- 100 * (df_res_16$gens_rot / df_res_16$gens_seq -1)
+#df_res_16$tst <- df_res_16$ian_rms_percent-df_res_16$rms_percent
 
@@
 
-```{r figx1-pcent-diff-rot-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig2-pcent-diff-rot-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# 2020-10 previously fig7
 
@@
 
+```{r fig3-mort-which-best-meanmort, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# 2020-10 previously fig8
+
+# ** using best strategy from generations calc above
+# ** so trying to see if being better in generations terms
+# ** has consequences for mean mortality
+
+#2020-10-08 Ian asked to change this to percent difference
+
+ggplot(df_res_16, aes(x=rms_mort_percent, fill=best_strategy)) + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
+#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
+     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     geom_histogram(binwidth=5, colour='darkgrey') + 
+     scale_fill_manual(values=colrs) +
+     theme_minimal() +
+     xlim(-40,40) +
+     xlab("percentage advantage of rotations over sequence in mean mortality") +  
+     ylab("count of randomly sampled runs") +  
+     theme(panel.grid.minor=element_blank(),
+           legend.position="bottom") +
+     facet_grid(rows=vars(id14), cols=vars(idAD),
+                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+#old version just mortality rather than %
+# ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
+#      geom_histogram(binwidth=0.02, colour='darkgrey') + 
+#      scale_fill_manual(values=colrs) +
+#      theme_minimal() +
+#      xlim(-0.2,0.2) +
+#      xlab("advantage of rotations over sequence in mean mortality") +
+#      ylab("count of randomly sampled runs") +  
+#      theme(panel.grid.minor=element_blank(),
+#            legend.position="bottom") +
+#      facet_grid(rows=vars(id14), cols=vars(idAD),
+#                 labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
+```
+
 # at each generations below threshold for rotations, which strategy was best
 
-```{r figx2-which-best-each-gens-rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig9-which-best-each-gens-rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+# 2020-10 fig9
 
@@
 #it shows what gens_rot the best strategies occur at
+
+# but difficult to see the ones where there is an advantage to either strategy because so few of them 
+# increasing binwidth to 50 made advantage run more visible
+
+#also good that it shows runs lasting longer when fitness costs and dispersal are added
+
 ggplot(df_res_16, aes(x=gens_rot, fill=best_strategy)) + 
-     geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     geom_histogram(binwidth=50, colour='darkgrey') + 
      scale_fill_manual(values=colrs) +
      theme_minimal() +
+     #scale_y_log10() + #log scale did make it easier to see lower values but misleading
      #xlim(-100,100) +
@@
 
+#Ian asked if could make the bars mean of rot & seq
+#but I think this is confusing & potentially wrong
+# ggplot(df_res_16, aes(x=0.5*(gens_rot+gens_seq), fill=best_strategy)) + 
+#      geom_histogram(binwidth=50, colour='darkgrey') + 
+#      scale_fill_manual(values=colrs) +
+#      theme_minimal() +
+#      #scale_y_log10() + #log scale did make it easier to see lower values but misleading
+#      #xlim(-100,100) +
+#      xlab("generations to reach thresholds for ROTATION") +
+#      ylab("count of randomly sampled runs") +    
+#      theme(panel.grid.minor=element_blank(),
+#            legend.position="bottom") +
+#      facet_grid(rows=vars(id14), cols=vars(idAD),
+#                 labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
+
 ```
@@
 
-```{r figx3-which-best-each-gens-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig9seq-which-best-each-gens-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+#same as previous but for seq rather than rot
+#maybe to go in SI
 
@@
 ggplot(df_res_16, aes(x=gens_seq, fill=best_strategy)) + 
-     geom_histogram(binwidth=widthbin, colour='darkgrey') + 
+     geom_histogram(binwidth=50, colour='darkgrey') + 
      scale_fill_manual(values=colrs) +
@@
 ```
-
-\pagebreak
 
@@
 ```
-
-```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
-
-# doesn't show much
-# Plots of how the difference between rotations and sequences respond to model inputs.
-
-# go through selected inputs and plot them on x with gens_rot_minus_seq on y 
-# TODO check whether this covers all the inputs we want to look at
-in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')
-
-y <- 'gens_rot_minus_seq'
-
-for(i in in_cols)
-{
-
-    #plotnum <- plotnum + 1
-
-    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
-    #plotlist[[plotnum]] <- 
-    print( ggplot(df_res_16, aes_string(x=i, y=y)) +
-
-             geom_point(shape=1, size=1, alpha=0.5, show.legend=FALSE) +
-             theme_minimal() +
-             geom_smooth(linetype='dashed',size=0.5) )
-
-}
-
-```
-
-# PRCC to show the sensitivity of the difference between rotation and sequence to inputs
-# this can only be used for the experiments where all insecticides are the same
-# (because in the ones where insecticides differ there are multiple values for each input)
-# start by running on just A expts
-
-```{r figxx-PRCC, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-
-  library(sensitivity)
-
-# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
-# 
-# Arguments
-# X a data frame containing model input variables
-# y a vector containing the responses 
-# rank logical. If TRUE, the analysis is done on the ranks.
-# nboot the number of bootstrap replicates.
-# conf the confidence level of the bootstrap confidence intervals.
-
-#same as above
-#TODO check that this covers all the inputs we want to look at
-in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')
-
-    x <- df_res_A[,in_cols]
-    y <- df_res_A['gens_rot_minus_seq']
-
-    pcc_res <- pcc(x, y, rank=TRUE)
-
-    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
-    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
-    #plot(pcc_res)
-
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
-
-    to_plot <- pcc_res$PRCC
-    #rename column 1 from 'original to PRCC
-    names(to_plot)[1] <- 'PRCC'
-    to_plot$inputs <- rownames(to_plot)
-
-    axis_text_x <- element_text(angle = 45,hjust = 1, vjust = 1)
-    height <- 0.9
-
-    gg <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
-             geom_point(shape=1, colour='red') +
-             theme(axis.text.x = axis_text_x) +
-             geom_hline(yintercept = 0, linetype=3) +
-             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
-             ylim(-1,1) +
-             #ggtitle(paste(strat_name[strategy_num])) +
-             ggtitle("Effect of inputs on benefit of rotations, all A expts.") +
-             xlab(NULL)
-
-    plot(gg)
-
-  #to put back in some ref gridlines that are removed by cowplot
-  #theme_set(theme_gray())
-  #theme_set(theme_bw())
-
-  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
-  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.6), labels='AUTO')
-
-```
-
-```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
-
-# Runs where rotations are >20% better (in red) are associated with higher dominance of cost and lower dominance of selection, but doesn't guarantee, must be interaction with other inputs too.
-# this was for earlier runs when there was some strange model behaviour, may not be the case now
-# and may not be very meaningful anyway
-
-#'cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs'
-
-#'cost','dom_cos','dom_sel'
-
-y <- 'gens_rot_minus_seq'
-
-ggplot(df_res_16, aes(x=dom_sel, y=dom_cos)) + 
-     geom_point(shape=3, size=1, alpha=0.2, show.legend=FALSE) +
-     theme_minimal() +
-     geom_point(data=filter(df_res_16, gens_rot_minus_seq/gens_seq >= 0.3 ), shape=1, size=1, col='red', alpha=0.8)
-
-```
-
-```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
-#eval=FALSE tmp because changes above will break this
-# summary plot of all runs, would show any differences between sequence & rotations as non horizontal lines
-
-# want to plot line between sequence and rotation value
-# any deviation from horizontal will be obvious
-# could use rotation interval as the x axis, so that longer intervals will be clearer
-
-# maybe not for publication
-# but useful for quick view
-
-#it was useful for me but now that I've moved to one row per scenario it won't work !!
-
-    gg <- ggplot( df_res_16, aes_string(x='rot_interval',
-                                         y='tot_gens_dep_under50',
-                                         group='scenario',
-                                         colour='scenario') ) +
-          geom_line( alpha=0.5, lwd=1 )
-
-    plot(gg)
-
-# plot gens againts exposure and effectiveness
-# as a quick test to see if I can set the starting values of these higher
-
-    gg <- ggplot( df_res_16, aes_string(x='eff',
-                                         y='tot_gens_dep_under50') ) +
-          geom_point( alpha=0.5 )
-    plot(gg)
-
-    gg <- ggplot( df_res_16, aes_string(x='expo_hi',
-                                         y='tot_gens_dep_under50') ) +
-          geom_point( alpha=0.5 )
-    plot(gg)
-
-    gg <- ggplot( df_res_16, aes_string(x='expo_hi*eff*male_expo_prop',
-                                         y='tot_gens_dep_under50') ) +
-          geom_point( alpha=0.5 )
-    plot(gg)
-
-```
-
-# new - mean mortality over whole sim
-# slightly surprising that not much variation
-# this one colours by best strategy in generation terms
-# B1 suggests that runs where rotation is better may have
-# lower mortality under rotation
-# try to find a different way of showing that 
-# (e.g. just plot gens_diff against mortality diff)
-
-```{r figxmort-rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-
-# ** mean mortality for rotations
-
-# have a go at classifying gens_rot so I can use it to colour the mean mort bars
-# cut(gens_rot, 5) this does that
-
-ggplot(df_res_16, aes(x=mort_rot, fill = cut(gens_rot, 5))) + 
-#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
-     geom_histogram(colour='darkgrey') + 
-     #geom_histogram(binwidth=0.02, colour='darkgrey') + 
-     #be interesting to color by generations to see if strategies that last longer have lower mortality ?
-     #scale_fill_manual(values=colrs) +
-     #scale_fill_continuous(aes(gens_rot)) +
-     scale_fill_brewer(palette = 'greens') +
-     theme_minimal() +
-     #xlim(-100,100) +
-     xlab("mean mortality over all generations for rotation") +
-     ylab("count of randomly sampled runs") +  
-     theme(panel.grid.minor=element_blank(),
-           legend.position="bottom") +
-     facet_grid(rows=vars(id14), cols=vars(idAD),
-                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
-
-```
-
-```{r figxgens_rot_v_mort_rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-
-# ** mean mortality for rotations
-
-ggplot(df_res_16, aes(x=gens_rot, fill = cut(mort_rot, 5))) + 
-#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
-     geom_histogram(colour='darkgrey') + 
-     #geom_histogram(binwidth=0.02, colour='darkgrey') + 
-     #be interesting to color by generations to see if strategies that last longer have lower mortality ?
-     #scale_fill_manual(values=colrs) +
-     #scale_fill_continuous(aes(gens_rot)) +
-     scale_fill_brewer(palette = 'greens') +
-     theme_minimal() +
-     #xlim(-100,100) +
-     #xlab("mean mortality over all generations for rotation") +
-     #ylab("count of randomly sampled runs") +  
-     theme(panel.grid.minor=element_blank(),
-           legend.position="bottom") +
-     facet_grid(rows=vars(id14), cols=vars(idAD),
-                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
-
-```
-
-```{r figxmort_rot_v_start_freqs, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-
-# i expect that mean mort may be largely influenced by startfreqs, but remember when diff. there are multiple
-
-# ** mean mortality for rotations
-
-ggplot(df_res_16, aes(x=start_freqs, y=mort_rot)) + 
-#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
-     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
-     geom_hline(yintercept = 0, linetype="dashed", color = 'darkgrey') +
-     geom_vline(xintercept = 0, linetype="dashed", color = 'darkgrey') +
-     geom_point(shape='.') +
-     geom_smooth() +
-     #geom_abline() +
-     # scale_fill_manual(values=colrs) +
-     theme_minimal() +
-     # #xlim(-100,100) +
-     # xlab("advantage of rotations over sequence in mean mortality") +
-     # ylab("count of randomly sampled runs") +  
-     theme(panel.grid.minor=element_blank()) +
-     facet_grid(rows=vars(id14), cols=vars(idAD),
-                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
-
-```
-
-```{r figxmort-rot-minus-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-
-# ** mean mortality difference between rot & seq
-# why do these seem not to sum to totals ?
-
-#ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
-ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
-     geom_histogram(binwidth=0.02, colour='darkgrey') + 
-     #scale_fill_manual(values=colrs) +
-     theme_minimal() +
-     #xlim(-100,100) +
-     xlab("advantage of rotations over sequence in mean mortality") +
-     ylab("count of randomly sampled runs") +  
-     theme(panel.grid.minor=element_blank(),
-           legend.position="bottom") +
-     facet_grid(rows=vars(id14), cols=vars(idAD),
-                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
-
-```
-
-```{r figxmort-which-best-meanmort, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-
-# ** using best strategy from generations calc above
-# ** so trying to see if being better in generations terms
-# ** has consequences for mean mortality
-
-ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
-#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
-     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
-     geom_histogram(binwidth=0.02, colour='darkgrey') + 
-     scale_fill_manual(values=colrs) +
-     theme_minimal() +
-     xlim(-0.2,0.2) +
-     xlab("advantage of rotations over sequence in mean mortality") +
-     ylab("count of randomly sampled runs") +  
-     theme(panel.grid.minor=element_blank(),
-           legend.position="bottom") +
-     facet_grid(rows=vars(id14), cols=vars(idAD),
-                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
-
-```
-
-#try to establish whether
-#runs that have higher generations for rotations
-#also have lower mean mortality
-
-```{r figxmort-v-gens, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-
-#gens_rot_minus_seq, facet by A-D and 1-4, better facet labels
-ggplot(df_res_16, aes(x=gens_rot_minus_seq, y=mort_rot_minus_seq)) + 
-#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
-     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
-     geom_hline(yintercept = 0, linetype="dashed", color = 'darkgrey') +
-     geom_vline(xintercept = 0, linetype="dashed", color = 'darkgrey') +
-     geom_point(shape='.') +
-     geom_smooth(method='lm') +
-     #geom_abline() +
-     # scale_fill_manual(values=colrs) +
-     theme_minimal() +
-     # #xlim(-100,100) +
-     # xlab("advantage of rotations over sequence in mean mortality") +
-     # ylab("count of randomly sampled runs") +  
-     theme(panel.grid.minor=element_blank()) +
-     facet_grid(rows=vars(id14), cols=vars(idAD),
-                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))
-
-```

### 2020-12-14 inst/documents/rotations_notes.Rmd
figure tweaks

@@
 
-# TODO
-
-2020-02-25 : probably get Sam to rerun the 16k runs
-
-first try to fix the issues above, or check they don't cause problems
+2020-02-25 : get Sam to rerun the 16k runs
+
+#2020-10-05 reviewing runs for rotations paper & Qs from Ian
+
+Andy….I assume the following description is how you transformed the data to plot Figure 7 ?
+We quantify the relative performance of sequential and rotations policy (see later discussion of Figures 7 and 8) using the metric, z.
+z=(L(r)/L(s) -1)*100
+
+this is my fig x1, the output is rms_percent
+
+df_res_16$rms_percent <- 100 * df_res_16$gens_rot_minus_seq / df_res_16$gens_seq
+
+Ian asks to change the mortality fig (8) to %difference.
+
+I think it is :
+figxmort-which-best-meanmort
+
+The rational for keeping it to absolute mortality differences is that the values are meaningful to those interested in insecticide resistance, and that mortality is already a %, so talking about percent of a percent gets really confusing.
+Currently the figure shows that most runs result in just a 1% difference in mortality - that readers may equate e.g. to the difference between 89 & 90% in a bioassay.
+
+# TODO 2020-20
+
+# TODO (from before)
 
@@
 
-Try some rotation_runs2019.Rmd with new outputs.
-
-a. create a function to compare seq and rot plots
-1. check random number issues
-2. look at paper3 folder
-
-Questions to answer re 16k runs :
-
-1. why with mortality switch are there more runs favouring sequence ?
-
 check on dispersal limits using new method. Seems that anything above 0.4 leads to near complete mixing
 run_rot(migration=0.4, coverage=0.5)
-#interesting how small levels of migration cause resistance freqs to decline even with no costs
-
+interesting how small levels of migration cause resistance freqs to decline even with no costs
 write word description of what dispersal does
 rename migration to dispersal ?
-
-migration limits ???
-I need to check on them more ! to make sure I know what they are doing.
+migration limits ??? I need to check on them more ! to make sure I know what they are doing.
 
@@
 CHECK ALL INPUT PARAMS
-
-~ create some plot functions to work on the output files (df_res_all) e.g. plot_rotseq_diff()
-
-~ try to get run progress message into console or terminal
-~ improve readme.md with details of how to run simulations
 
@@
 ~ check it works with diff insecticides i.e. eff=c(0.7,0.8,0.9) I suspect not because mort_from_resist() or something else not properly vectorised.
-
-~ TIDY UP CODE AND IMPROVE DOCUMENTATION e.g. function args
-
-whatever I do I need to make mort/survival thing more transparent
-
-options
-~ pass as mortality and then convert to survival for calculation
-~ pass as survival convert to mortality for plot
-
-~ ian interested in starting frequencies at 10e-4, can do that for the runs with different values per insecticide.
-~ we won't run sensitivity analysis on those runs, but just to check that rotations and sequence runs still similar
-
-rotations paper :
-~ think about what questions managers will want answered
-
-A) when do both strategies work (include looking at insecticides and rotation level)
- - maybe do PRCC
-B) if neither work, when are rotations better ?
-
-small changes :
-~ dominance of resistance
-~ how to term success. 'At least one insecticide still below resistance thresholds'
-
-Look into whether to use min_rwr_interval OR min_gens_switch_back 
-I'm thinking that I'll stick with min_rwr_interval
-
-I think takehome message maybe that rotations are much better than sequences only in rare circumstances.
-
-May also be something useful I can say about the number of insecticides too ? (Which is of interest to IVCC).
-
-~ can I create some simple figs to illustrate how an increasing rate of resistance evolution with frequency can lead to rotations being favoured
-
-~ modify sensi_an_rotations so that I have a reproduceable record of a) no cost or dispersal b)cost only c)dispersal only, d) cost & dispersal
-
-think about killer plot to display difference between rot & seq generations
-
-~ maybe come back to i_in_use ribbons not meeting issue
-~ add in some unit-tests ?
-
-~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ? I don't think so because it requires linkage.
-
-~ add error-check to run_rot if n.insecticides=1 & migration > 0
-
-Maybe explore idea from Rowland (1981) that resistance declines faster when it is mostly held by the RR (i.e. at higher allele frequencies)
-
-Look into data on changing allele frequencies over time (e.g. for Kdr)...
-
-It occurred to me that maybe a benefit of rotations is that if it is difficult to measure resistance allele frequencies they provide a lower risk of reaching resistance thresholds for any insecticide.
-i.e.
-are they a better strategy in light of uncertainty in resistance allele frequencies ?
-
-~ can we get cost to decrease over time at high frequency as suggested in GPIRM ?
-
-BEWARE active in active_or_refuge is potentially confusing, does not mean that insecticide is active/deployed, just means its is the active area (a different insecticide may be deployed)
 
@@
 
-~ Ian & I talked about using fitness to determine changes.
-and that fitness could be used as an index of population, but Ian was sceptical, he said to look up fitness determinants of population on wikipedia, about soft & hard selection.
-
-~ test fitness_single_locus() from run_rot() and add to Rmd
-
 ~ check in resistance::fitnessSingleLocus() on separation of dominance of cost & selection. Was probably OK for Levick et al because we didn't have cost in. 
@@
 # state of play
-~ sensi_an_rotations1.Rmd : starting on sensitivity analysis
+~ sensi_an_rotations1.Rmd
 ~~  produces an output dataframe containing both inputs and outputs

### 2020-12-14 R/rot_plots.r

@@ -35,6 +35,7 @@ rot_plot_resistance <- function(df_res2,
                                 df_resanother = NULL,
                                 lwd = 1.5,
                                 title = 'auto', #NULL
+                                legend_title = "areas connected\nby migration",
                                 plot = TRUE,
                                 plot_mort = TRUE) {
   
@@ -69,7 +70,7 @@ rot_plot_resistance <- function(df_res2,
     #legend for the lines, allows me to set title & labels  
     if (!plot_mort & mort_or_freq=='freq')
     {
-      gg <- gg + scale_colour_manual("areas connected\nby migration",values=c("red3","navy"), labels=c("treated","untreated refugia"))
+      gg <- gg + scale_colour_manual(legend_title,values=c("red3","navy"), labels=c("treated","untreated refugia"))
     }
   } else
   {
@@ -114,12 +115,20 @@ rot_plot_resistance <- function(df_res2,
       
       #trying & initially failing to get mortality line to appear in legend 
       #fiddled with the orders of values & labels to get to correspond to lines
-      # TODO sort when plot_refuge==FALSE
-      scale_colour_manual("areas connected\nby migration",values=c("red3",thresh_col,"darkgreen","navy"), 
-                           labels=c("resistance\nin treated",thresh_name,"mortality in treated","resistance in\nuntreated refugia"),
-                           guide = guide_legend(linetype=c(1,3,1,1)))
-                           #guide = guide_legend(override.aes = list(fill = NA)))
-    
+      if (plot_refuge) {
+        
+        scale_colour_manual(legend_title,values=c("red3",thresh_col,"darkgreen","navy"), 
+                            labels=c("resistance\nin treated",thresh_name,"mortality in treated","resistance in\nuntreated refugia"),
+                            guide = guide_legend(linetype=c(1,3,1,1)))
+        
+      } else {
+        
+        scale_colour_manual(legend_title,values=c("red3",thresh_col,"darkgreen"), 
+                            labels=c("resistance\nin treated",thresh_name,"mortality in treated"),
+                            guide = guide_legend(linetype=c(1,3,1)))
+        
+      }
+
       
       #legend for mortality, allows me to set title & labels  
       #but this overwrites other legend
